## Tuners {#ext-tuner}

The dummy tuner function in this tutorial is called `blackBoxFun`.

A new tuner consists of an **objective function** and settings.
The former is the heart of the `r ref("Tuner")`.
It must fulfill the following requirements:

- Run the `r ref("PerformanceEvaluator")` for the parameter value `x`.
- Evaluate the `r ref("Learner")` using the given `r ref("Resampling")` object and add the `r ref("ResampleResult")` to the `r ref("BenchmarkResult")` object of the `r ref("PerformanceEvaluator")`.

### Objective function

In the following we describe how a possible implementation of a new objective function could look like: 

```{r 90-extending-16 }
blackBoxFun = function (x, pe) {
  # set measure
  measure = pe$measures[[1L]]
  hashes = pe$bmr$data$hash
  # convert ParamSet to data.table
  x = setDT(as.list(x))
  pe$eval(x)
  new_hash = setdiff(pe$bmr$data$hash, hashes)
  # calculate performance
  perf = pe$bmr$resample_result(new_hash)$aggregate(measure)
  if (measure$minimize) perf else -perf
  return (perf)
}
```

With `pe` being the `r ref("PerformanceEvaluator")` object, `blackBoxFun()` should be able to do the following (you can also take a look at [TunerRandomSearch](https://github.com/mlr-org/mlr3tuning/blob/991d11ab70ce99f200f19e570f0d362f4910828a/R/TunerRandomSearch.R) or [TunerGenSA](https://github.com/mlr-org/mlr3tuning/blob/991d11ab70ce99f200f19e570f0d362f4910828a/R/TunerGenSA.R).

```{r 90-extending-17, eval = TRUE}
library("mlr3tuning")
task = mlr3::mlr_tasks$get("iris")
learner = mlr3::mlr_learners$get("classif.rpart")
resampling = mlr3::mlr_resamplings$get("cv")
resampling$param_set$values$folds = 2
measures = mlr3::mlr_measures$mget("classif.ce")
param_set = paradox::ParamSet$new(
  params = list(
    paradox::ParamDbl$new("cp", lower = 0.001, upper = 0.1)
  )
)
pe = PerformanceEvaluator$new(task, learner, resampling, measures, param_set)
blackBoxFun(c(cp = 0.05), pe)
pe$bmr$aggregate()
```

### Tuner class

To actually call the optimizer using a dedicated R6 Tuner class, we add `blackBoxFun()` as a private method.
This can either be done for an existing class or a new R6 Tuner class can be created.

In this example we replace the private `.$tune_step()` method from `r ref("mlr3tuning::TunerGenSA")` with our new objective function which we defined above.

```{r 90-extending-18 }
TunerGenSA = R6Class("TunerGenSA",
  inherit = Tuner,
  public = list(
    initialize = function(pe, terminator, ...) {
      if (any(pe$param_set$storage_type != "numeric")) {
        err_msg = "Parameter types needs to be numeric"
        lg$error(err_msg)
        stopf(err_msg)
      }
      
      # Default settings:
      settings = list(smooth = FALSE, acceptance.param = -15, 
        simple.function = FALSE, temperature = 250)
      super$initialize(id = "GenSA", pe = pe, terminator = terminator, 
        settings = insert_named(settings, list(...)))
    }
  ),
  private = list(
    tune_step = function() {
      blackBoxFun = function (x, pe) {
        # set measure
        measure = pe$measures[[1L]]
        hashes = pe$bmr$data$hash
        # convert ParamSet to data.table
        x = setDT(as.list(x))
        pe$eval(x)
        new_hash = setdiff(pe$bmr$data$hash, hashes)
        # calculate performance
        perf = pe$bmr$resample_result(new_hash)$aggregate(measure)
        if (measure$minimize) perf else -perf
        return (perf)
      }
      self$GenSA_res = GenSA(fn = blackBoxFun, lower = self$pe$param_set$lower, upper = self$pe$param_set$upper,
        control = self$settings, pe = self$pe)
    }
  )
)
```

Note that the private method needs always be called `.$tune_step()` as it will be called from the `.$tune()` method of the `Tuner` class.

### Example

Now that the "new" `r ref("mlr3tuning::TunerGenSA")` tuner has been defined, we can test it in a small use case:

```{r 90-extending-19}
# does not work currently
task = mlr3::mlr_tasks$get("spam")
learner = mlr3::mlr_learners$get("classif.rpart")
learner$predict_type = "prob"
resampling = mlr3::mlr_resamplings$get("holdout")
measures = mlr3::mlr_measures$mget(c("classif.auc", "classif.ce"))
param_set = paradox::ParamSet$new(
  params = list(
    paradox::ParamDbl$new("cp", lower = 0.001, upper = 0.1)
  )
)
pe = PerformanceEvaluator$new(task, learner, resampling, measures, param_set)
tuner = TunerGenSA$new(pe, TerminatorEvaluations$new(3L))
tuner$tune()
tuner$pe$bmr$aggregate()
tuner$tune_result()
str(tuner$GenSA_res)
```
