## Define Tuning Spaces** {#searchspace}

When running an optimization, it is important to inform the tuning algorithm about what hyperparameters are valid.
Here the names, types, and valid ranges of each hyperparameter are important.
All this information is communicated with objects of the class `r ref("ParamSet")`, which is defined in `r mlr_pkg("paradox")`.
While it is possible to create `r ref("ParamSet")`-objects using its `$new`-constructor, it is much shorter and readable to use the `r ref("ps")`-shortcut, which will be presented here.
For an in-depth description of `r mlr_pkg("paradox")` and its classes, see the [`r mlr_pkg("paradox")` chapter](#paradox).

Note, that `r ref("ParamSet")` objects exist in two contexts.
First, `r ref("ParamSet")`-objects are used to define the space of valid parameter settings for a learner (and other objects).
Second, they are used to define a search space for tuning.
We are mainly interested in the latter.
For example we can consider the `minsplit` parameter of the `r ref("mlr_learners_classif.rpart", "classif.rpart Learner")`.
The `r ref("ParamSet")` associated with the learner has a lower but *no* upper bound.
However, for tuning the value, a lower *and* upper bound must be given because tuning search spaces need to be bounded.
For `r ref("Learner")` or `r ref("PipeOp")` objects, typically "unbounded" `r ref("ParamSet", text = "ParamSets")` are used.
Here, however, we will mainly focus on creating "bounded" `r ref("ParamSet", text = "ParamSets")` that can be used for tuning.
See the [in-depth `r mlr_pkg("paradox")` chapter](#paradox) for more details on using `r ref("ParamSet", text = "ParamSets")` to  define parameter ranges for use-cases besides tuning.

### Creating `r ref("ParamSet")`s

An empty `"ParamSet")` -- not yet very useful -- can be constructed using just the `"ps")` call:

```{r optimization-027}
search_space = ps()
print(search_space)
```

`r ref("ps")` takes named `r ref("Domain")` arguments that are turned into parameters.
A possible search space for the `"classif.svm"` learner could for example be:

```{r optimization-028}
search_space = ps(
  cost = p_dbl(lower = 0.1, upper = 10),
  kernel = p_fct(levels = c("polynomial", "radial"))
)
print(search_space)
```

There are five domain constructors that produce a parameters when given to `r ref("ps")`:

| Constructor               | Description                          | Is bounded?                        | Underlying Class    |
| :-----------------------: | :----------------------------------: | :--------------------------------: | :-----------------: |
| `r ref("p_dbl")`                   | Real valued parameter ("double")     | When `upper` and `lower` are given | `r ref("ParamDbl")` |
| `r ref("p_int")`                  | Integer parameter                    | When `upper` and `lower` are given | `r ref("ParamInt")` |
| `r ref("p_fct")`                   | Discrete valued parameter ("factor") | Always                             | `r ref("ParamFct")` |
| `r ref("p_lgl")`                 | Logical / Boolean parameter          | Always                             | `r ref("ParamLgl")` |
| `r ref("p_uty")`                   | Untyped parameter                    | Never                              | `r ref("ParamUty")` |

These domain constructors each take some of the following arguments:

* **`lower`**, **`upper`**: lower and upper bound of numerical parameters (`r ref("p_dbl")` and `r ref("p_int")`). These need to be given to get bounded parameter spaces valid for tuning.
* **`levels`**: Allowed categorical values for `p_fct` parameters.
  Required argument for `r ref("p_fct")`.
  See [below](#autolevel) for more details on this parameter.
* **`trafo`**: transformation function, see [below](#searchspace-trafo).
* **`depends`**: dependencies, see [below](#searchspace-depends).
* **`tags`**: Further information about a parameter, used for example by the [`hyperband`](#hyperband) tuner.
* **`default`**: Value corresponding to default behavior when the parameter is not given.
  Not used for tuning search spaces.
* **`special_vals`**: Valid values besides the normally accepted values for a parameter.
  Not used for tuning search spaces.
* **`custom_check`**: Function that checks whether a value given to `r ref("p_uty")` is valid.
  Not used for tuning search spaces.

The `lower` and `upper` parameters are always in the first and second position respectively, except for `r ref("p_fct")` where `levels` is in the first position.
It is preferred to omit the labels (ex: upper = 0.1 becomes just 0.1). This way of defining a `r ref("ParamSet")` is more concise than the equivalent definition above.
Preferred:

```{r optimization-029}
search_space = ps(cost = p_dbl(0.1, 10), kernel = p_fct(c("polynomial", "radial")))
```

### Transformations (`trafo`) {#searchspace-trafo}

We can use the `r mlr_pkg("paradox")` function `r ref("generate_design_grid")` to look at the values that would be evaluated by grid search.
(We are using `r ref("rbindlist()")` here because the result of `$transpose()` is a list that is harder to read.
If we didn't use `$transpose()`, on the other hand, the transformations that we investigate here are not applied.) In `generate_design_grid(search_space, 3)`, `search_space` is the `r ref("ParamSet")` argument and 3 is the specified resolution in the parameter space.
The resolution for categorical parameters is ignored; these parameters always produce a grid over all of their valid levels.
For numerical parameters the endpoints of the params are always included in the grid, so if there were 3 levels for the kernel instead of 2 there would be 9 rows, or if the resolution was 4 in this example there would be 8 rows in the resulting table.

```{r optimization-030}
library("data.table")
rbindlist(generate_design_grid(search_space, 3)$transpose())
```

We notice that the `cost` parameter is taken on a linear scale.
We assume, however, that the difference of cost between `0.1` and `1` should have a similar effect as the difference between `1` and `10`.
Therefore it makes more sense to tune it on a *logarithmic scale*.
This is done by using a **transformation** (`trafo`).
This is a function that is applied to a parameter after it has been sampled by the tuner.
We can tune `cost` on a logarithmic scale by sampling on the linear scale `[-1, 1]` and computing `10^x` from that value.
```{r optimization-031}
search_space = ps(
  cost = p_dbl(-1, 1, trafo = function(x) 10^x),
  kernel = p_fct(c("polynomial", "radial"))
)
rbindlist(generate_design_grid(search_space, 3)$transpose())
```

It is even possible to attach another transformation to the `r ref("ParamSet")` as a whole that gets executed after individual parameter's transformations were performed.
It is given through the `.extra_trafo` argument and should be a function with parameters `x` and `param_set` that takes a list of parameter values in `x` and returns a modified list.
This transformation can access all parameter values of an evaluation and modify them with interactions.
It is even possible to add or remove parameters.
(The following is a bit of a silly example.)

```{r optimization-032}
search_space = ps(
  cost = p_dbl(-1, 1, trafo = function(x) 10^x),
  kernel = p_fct(c("polynomial", "radial")),
  .extra_trafo = function(x, param_set) {
    if (x$kernel == "polynomial") {
      x$cost = x$cost * 2
    }
    x
  }
)
rbindlist(generate_design_grid(search_space, 3)$transpose())
```

The available types of search space parameters are limited: continuous, integer, discrete, and logical scalars.
There are many machine learning algorithms, however, that take parameters of other types, for example vectors or functions.
These can not be defined in a search space `r ref("ParamSet")`, and they are often given as `r ref("ParamUty")` in the `r ref("Learner")`'s `r ref("ParamSet")`.
When trying to tune over these hyperparameters, it is necessary to perform a Transformation that changes the type of a parameter.

An example is the `class.weights` parameter of the [Support Vector Machine](https://machinelearningmastery.com/cost-sensitive-svm-for-imbalanced-classification/) (SVM), which takes a named vector of class weights with one entry for each target class.
The trafo that would tune `class.weights` for the `r ref("mlr_tasks_spam")`, `'tsk("spam")` dataset could be:

```{r optimization-033}
search_space = ps(
  class.weights = p_dbl(0.1, 0.9, trafo = function(x) c(spam = x, nonspam = 1 - x))
)
generate_design_grid(search_space, 3)$transpose()
```

(We are omitting `r ref("rbindlist()")` in this example because it breaks the vector valued return elements.)

### Automatic Factor Level Transformation {#autolevel}

A common use-case is the necessity to specify a list of values that should all be tried (or sampled from).
It may be the case that a hyperparameter accepts function objects as values and a certain list of functions should be tried.
Or it may be that a choice of special numeric values should be tried.
For this, the `r ref("p_fct")` constructor's `level` argument may be a value that is not a `character` vector, but something else.
If, for example, only the values `0.1`, `3`, and `10` should be tried for the `cost` parameter, even when doing random search, then the following search space would achieve that:

```{r optimization-034}
search_space = ps(
  cost = p_fct(c(0.1, 3, 10)),
  kernel = p_fct(c("polynomial", "radial"))
)
rbindlist(generate_design_grid(search_space, 3)$transpose())
```

This is equivalent to the following:
```{r optimization-035}
search_space = ps(
  cost = p_fct(c("0.1", "3", "10"),
    trafo = function(x) list(`0.1` = 0.1, `3` = 3, `10` = 10)[[x]]),
  kernel = p_fct(c("polynomial", "radial"))
)
rbindlist(generate_design_grid(search_space, 3)$transpose())
```

Note: Though the resolution is 3 here, in this case it doesn't matter because both `cost` and `kernel` are factors (the resolution for categorical variables is ignored, these parameters always produce a grid over all their valid levels).

This may seem silly, but makes sense when considering that factorial tuning parameters are always `character` values:

```{r optimization-036}
search_space = ps(
  cost = p_fct(c(0.1, 3, 10)),
  kernel = p_fct(c("polynomial", "radial"))
)
typeof(search_space$params$cost$levels)
```

Be aware that this results in an "unordered" hyperparameter, however.
Tuning algorithms that make use of ordering information of parameters, like genetic algorithms or model based optimization, will perform worse when this is done.
For these algorithms, it may make more sense to define a `r ref("p_dbl")` or `r ref("p_int")` with a more fitting trafo.

The `class.weights` case from above can also be implemented like this, if there are only a few candidates of `class.weights` vectors that should be tried.
Note that the `levels` argument of `r ref("p_fct")` must be named if there is no easy way for `as.character()` to create names:

```{r optimization-037}
search_space = ps(
  class.weights = p_fct(
    list(
      candidate_a = c(spam = 0.5, nonspam = 0.5),
      candidate_b = c(spam = 0.3, nonspam = 0.7)
    )
  )
)
generate_design_grid(search_space)$transpose()
```

### Parameter Dependencies (`depends`) {#searchspace-depends}

Some parameters are only relevant when another parameter has a certain value, or one of several values.
The [Support Vector Machine](https://machinelearningmastery.com/cost-sensitive-svm-for-imbalanced-classification/) (SVM), for example, has the `degree` parameter that is only valid when `kernel` is `"polynomial"`.
This can be specified using the `depends` argument.
It is an expression that must involve other parameters and be of the form `<param> == <scalar>`, `<param> %in% <vector>`, or multiple of these chained by `&&`.
To tune the `degree` parameter, one would need to do the following:

```{r optimization-038}
search_space = ps(
  cost = p_dbl(-1, 1, trafo = function(x) 10^x),
  kernel = p_fct(c("polynomial", "radial")),
  degree = p_int(1, 3, depends = kernel == "polynomial")
)
rbindlist(generate_design_grid(search_space, 3)$transpose(), fill = TRUE)
```

### Creating Tuning ParamSets from other ParamSets {#tune-token}

Having to define a tuning `r ref("ParamSet")` for a `r ref("Learner")` that already has parameter set information may seem unnecessarily tedious, and there is indeed a way to create tuning `r ref("ParamSet", "ParamSets")` from a `r ref("Learner")`'s `r ref("ParamSet")`, making use of as much information as already available.

This is done by setting values of a `r ref("Learner")`'s `r ref("ParamSet")` to so-called `r ref("TuneToken")`s, constructed with a `r ref("to_tune")` call.
This can be done in the same way that other hyperparameters are set to specific values.
It can be understood as the hyperparameters being tagged for later tuning.
The resulting `r ref("ParamSet")` used for tuning can be retrieved using the `$search_space()` method.

```{r optimization-039}
learner = lrn("classif.svm")
learner$param_set$values$kernel = "polynomial" # for example
learner$param_set$values$degree = to_tune(lower = 1, upper = 3)

print(learner$param_set$search_space())

rbindlist(generate_design_grid(
  learner$param_set$search_space(), 3)$transpose()
)
```

It is possible to omit `lower` here, because it can be inferred from the lower bound of the `degree` parameter itself.
For other parameters, that are already bounded, it is possible to not give any bounds at all, because their ranges are already bounded.
An example is the logical `shrinking` hyperparameter:
```{r optimization-040}
learner$param_set$values$shrinking = to_tune()

print(learner$param_set$search_space())

rbindlist(generate_design_grid(
  learner$param_set$search_space(), 3)$transpose()
)
```

`"to_tune")` can also be constructed with a `r ref("Domain")` object, i.e. something constructed with a `p_***` call.
This way it is possible to tune continuous parameters with discrete values, or to give trafos or dependencies.
One could, for example, tune the `cost` as above on three given special values, and introduce a dependency of `shrinking` on it.
Notice that a short form for `to_tune(<levels>)` is a short form of `to_tune(p_fct(<levels>))`.

:::{.callout-note}
When introducing the dependency, we need to use the `degree` value from *before* the implicit trafo, which is the name or `as.character()` of the respective value, here `"val2"`!
:::

```{r optimization-041}
learner$param_set$values$type = "C-classification" # needs to be set because of a bug in paradox
learner$param_set$values$cost = to_tune(c(val1 = 0.3, val2 = 0.7))
learner$param_set$values$shrinking = to_tune(p_lgl(depends = cost == "val2"))

print(learner$param_set$search_space())

rbindlist(generate_design_grid(learner$param_set$search_space(), 3)$transpose(), fill = TRUE)
```

The `"search_space()` picks up dependencies fromt the underlying `r ref("ParamSet")` automatically.
So if the `kernel` is tuned, then `degree` automatically gets the dependency on it, without us having to specify that.
(Here we reset `cost` and `shrinking` to `NULL` for the sake of clarity of the generated output.)

```{r optimization-042}
learner$param_set$values$cost = NULL
learner$param_set$values$shrinking = NULL
learner$param_set$values$kernel = to_tune(c("polynomial", "radial"))

print(learner$param_set$search_space())

rbindlist(generate_design_grid(learner$param_set$search_space(), 3)$transpose(), fill = TRUE)
```

It is even possible to define whole `r ref("ParamSet")`s that get tuned over for a single parameter.
This may be especially useful for vector hyperparameters that should be searched along multiple dimensions.
This `r ref("ParamSet")` must, however, have an `.extra_trafo` that returns a list with a single element, because it corresponds to a single hyperparameter that is being tuned.
Suppose the `class.weights` hyperparameter should be tuned along two dimensions:

```{r optimization-043}
learner$param_set$values$class.weights = to_tune(
  ps(spam = p_dbl(0.1, 0.9), nonspam = p_dbl(0.1, 0.9),
    .extra_trafo = function(x, param_set) list(c(spam = x$spam, nonspam = x$nonspam))
))
head(generate_design_grid(learner$param_set$search_space(), 3)$transpose(), 3)
```



