---
aliases:
  - "/introduction_and_overview.html"
---

# Introduction and Overview {#sec-introduction}

{{< include ../../common/_setup.qmd >}}

`r chapter = "Introduction and Overview"`
`r authors(chapter)`

Welcome to the **M**achine **L**earning in **R** universe.
In this book, we will guide you through the functionality offered by `mlr3` step by step.
If you want to contribute to our universe, ask any questions, read documentation, or just chat with the team, head to `r link("https://github.com/mlr-org/mlr3")` which has several useful links in the README.

The `r mlr3` [@mlr3] package and the wider `mlr3` ecosystem provide a generic, `r index("object-oriented", "object-oriented programming")`, and extensible framework for `r index("regression")` (@sec-tasks), `r index("classification")` (@sec-classif), and other machine learning `r index("tasks")` (@sec-special) for the R language [@R].
On the most basic level, the unified interface provides functionality to train, test, and evaluate many machine learning algorithms.
You can also take this a step further with hyperparameter optimization, computational pipelines, model interpretation, and much more.
`mlr3` has similar overall aims to `caret` and `tidymodels` for R, `scikit-learn` for Python, and `MLJ` for Julia.
In general, `r mlr3` is designed to provide more flexibility than other ML frameworks while still offering easy ways to use advanced functionality.
While `tidymodels` in particular makes it very easy to perform simple ML tasks, `r mlr3` is more geared towards advanced ML.

Before we can show you the full power of `mlr3`, we recommend installing the `r mlr3verse` package, which will install several, important packages in the `mlr3` ecosystem.

```{r introduction_and_overview-001, eval = FALSE}
install.packages("mlr3verse")
```

Chapters that were added after the release of the printed version of this book are marked with a '+'.

## Installation Guidelines {#installguide}

There are many packages in the `mlr3` ecosystem that you may want to use as you work through this book.
All our packages can be installed from GitHub and R-universe[^runiverse]; the majority (but not all) packages can also be installed from CRAN.
We recommend adding the mlr-org R-universe to your R options so you can install all packages with `install.packages()`, without having to worry which package repository it comes from.
To do this, install `r ref_pkg("usethis")` and run the following:

[^runiverse]: R-universe is an alternative package repository to CRAN. The bit of code below tells R to look at both R-universe and CRAN when trying to install packages. R will always install the latest version of a package.

```{r introduction_and_overview-002, eval = FALSE}
usethis::edit_r_profile()
```

In the file that opens add or change the `repos` argument in `options` so it looks something like the code below (you might need to add the full code block below or just edit the existing `options` function).

```{r introduction_and_overview-003, eval = FALSE}
options(repos = c(
  mlrorg = "https://mlr-org.r-universe.dev",
  CRAN = "https://cloud.r-project.org/"
))
```

Save the file, restart your R session, and you are ready to go!

If you want the latest development version of any of our packages, run

```{r introduction_and_overview-004, eval = FALSE}
remotes::install_github("mlr-org/{pkg}")
```

with `{pkg}` replaced with the name of the package you want to install.
You can see an up-to-date list of all our extension packages at `r link("https://github.com/mlr-org/mlr3/wiki/Extension-Packages")`.

### Docker Images {.unnumbered .unlisted}

As an alternative to installing packages in R, you can use pre-built Docker images from the `r link("https://github.com/mlr-org/mlr3docker")` repository.
These images provide ready-to-use environments with `mlr3` and its ecosystem already installed.
You can pull the images from `r link("https://hub.docker.com/u/mlrorg")`:

* `mlrorg/mlr3-slim:latest`: A lightweight image with `mlr3verse` installed, built on `rocker/r-ver:latest`
* `mlrorg/mlr3-full:latest`: A complete image with `mlr3verse` and `mlr3extralearners` with all dependencies, built on `rocker/rstudio:latest`

To pull and run one of these images, use:

```{r introduction_and_overview-004a, eval = FALSE}
docker pull mlrorg/mlr3-slim:latest
docker run -it mlrorg/mlr3-slim:latest R
```

The full image includes RStudio, which can be accessed via a web browser after running the container.
For more details, see the `r link("https://github.com/mlr-org/mlr3docker")` repository.

## How to Use This Book {#howtouse}

You could read this book cover to cover but you may benefit more from dipping in and out of chapters as suits your needs, we have provided a comprehensive index to help you find relevant pages and sections.
We do recommend reading the first part of the book in its entirety as this will provide you with a complete overview of our basic infrastructure and design, which is used throughout our ecosystem.

We have marked sections that are particularly complex with respect to either technical or methodological detail and could be skipped on a first read with the following information box:

{{< include ../../common/_optional.qmd >}}

Each chapter includes examples, API references, and explanations of methodologies.
At the end of each part of the book we have included exercises for you to test yourself on what you have learned; you can find the solutions to these exercises at `r link("https://mlr3book.mlr-org.com/solutions.html")`.
We have marked more challenging (and possibly time-consuming) exercises with an asterisk, '*'.

If you want more detail about any of the tasks used in this book or links to all the `mlr3` dictionaries, please see the appendices in the online version of the book at `r link("https://mlr3book.mlr-org.com/")`.

#### Reproducibility {.unnumbered .unlisted}

At the start of each chapter we run `set.seed(123)` and use `r ref_pkg("renv")` to manage package versions, you can find our lockfile at `r link("https://github.com/mlr-org/mlr3book/blob/main/book/renv.lock")`.


## mlr3book Code Style {#styleguide}

Throughout this book we will use the following code style:

1. We always use `=` instead of `<-` for assignment.
2. Class names are in `UpperCamelCase`
3. Function and method names are in `lower_snake_case`
4. When referencing functions, we will only include the package prefix (e.g., `pkg::function`) for functions outside the `mlr3` universe or when there may be ambiguity about in which package the function lives. Note you can use `environment(function)` to see which namespace a function is loaded from.
5. We denote packages, fields, methods, and functions as follows:

    * `package` (highlighted in the first instance)
    * `package::function()` or `function()` (see point 4)
    * `$field` for fields (data encapsulated in an R6 class)
    * `$method()` for methods (functions encapsulated in an R6 class)
    * `Class` (for R6 classes primarily, these can be distinguished from packages by context)

Now let us see this in practice with our first example.

## mlr3 by Example

The `mlr3` universe includes a wide range of tools taking you from basic ML to complex experiments.
To get started, here is an example of the simplest functionality -- training a model and making predictions.

```{r introduction_and_overview-005}
library(mlr3)
task = tsk("penguins")
split = partition(task)
learner = lrn("classif.rpart")

learner$train(task, row_ids = split$train)
learner$model

prediction = learner$predict(task, row_ids = split$test)
prediction

prediction$score(msr("classif.acc"))
```

In this example, we trained a decision tree on a subset of the `r ref("palmerpenguins::penguins")` dataset, made predictions on the rest of the data and then evaluated these with the accuracy measure.
In @sec-basics we will break this down in more detail.

The `mlr3` interface also lets you run more complicated experiments in just a few lines of code:

```{r introduction_and_overview-006, eval = FALSE}
library(mlr3verse)

tasks = tsks(c("breast_cancer", "sonar"))

glrn_rf_tuned = as_learner(ppl("robustify") %>>% auto_tuner(
    tnr("grid_search", resolution = 5),
    lrn("classif.ranger", num.trees = to_tune(200, 500)),
    rsmp("holdout")
))
glrn_rf_tuned$id = "RF"

glrn_stack = as_learner(ppl("robustify") %>>% ppl("stacking",
    lrns(c("classif.rpart", "classif.kknn")),
    lrn("classif.log_reg")
))
glrn_stack$id = "Stack"

learners = c(glrn_rf_tuned, glrn_stack)
bmr = benchmark(benchmark_grid(tasks, learners, rsmp("cv", folds = 3)))

bmr$aggregate(msr("classif.acc"))
```

```{r introduction_and_overview-007, output = FALSE, echo = FALSE}
library(mlr3verse)
library(mlr3pipelines)
library(mlr3benchmark)

tasks = tsks(c("breast_cancer", "sonar"))
glrn_rf_tuned = auto_tuner(
    tnr("grid_search", resolution = 5),
    lrn("classif.ranger", num.trees = to_tune(200, 500)),
    rsmp("holdout")
)
glrn_rf_tuned = pipeline_robustify(NULL, glrn_rf_tuned, TRUE) %>>%
    po("learner", glrn_rf_tuned)
glrn_stack = ppl(
    "stacking",
    base_learners = lrns(c("classif.rpart", "classif.kknn")),
    lrn("classif.log_reg"))
glrn_stack = pipeline_robustify(NULL, glrn_stack, TRUE) %>>%
    po("learner", glrn_stack)

learners = c(glrn_rf_tuned, glrn_stack)
bmr = benchmark(benchmark_grid(tasks, learners, rsmp("holdout")))
```

```{r introduction_and_overview-008, echo = FALSE}
aggr = bmr$aggregate(msr("classif.acc"))[, c("task_id", "learner_id",
  "classif.acc")]
aggr$learner_id = rep(c("RF", "Stack"), 2)
aggr
```

In this (much more complex!) example we chose two tasks and two learners and used automated tuning to optimize the number of trees in the random forest learner (@sec-optimization), and a machine learning pipeline that imputes missing data, collapses factor levels, and stacks models (@sec-pipelines and @sec-pipelines-nonseq).
We also showed basic features like loading learners (@sec-basics) and choosing resampling strategies for benchmarking (@sec-performance).
Finally, we compared the performance of the models using the mean accuracy with three-fold cross-validation.

You will learn how to do all this and more in this book.

## The `mlr3` Ecosystem

Throughout this book, we often refer to `mlr3`, which may refer to the single `r mlr3` base package but usually refers to all packages in our ecosystem, this should be clear from context.
The `r mlr3` *package* provides the base functionality that the rest of the ecosystem depends on for building more advanced machine learning tools.
@fig-mlr3verse shows the packages in our ecosystem that extend `r mlr3` with capabilities for preprocessing, pipelining, visualizations, additional learners, additional task types, and much more.

```{r introduction_and_overview-009, echo = FALSE, out.width = "100%"}
#| label: fig-mlr3verse
#| fig-cap: Overview of the `mlr3` ecosystem, the packages with gray dashed lines are still in development, all others have a stable interface.
#| fig-alt: "Mindmap showing the packages of the mlr3verse and their relationship. Center `mlr3`, immediately connected to that are 'Learners', 'Evaluation', 'Tuning', 'Feature Selection', 'Utilities', 'Special Tasks', 'Data', and 'Pipelines'. Within each group is: Learners: `mlr3learners`, `mlr3extralearners`, `mlr3torch`; Evaluation: `mlr3measures`, `mlr3benchmark`; Tuning: `mlr3tuning`, `miesmuschel`, `mlr3hyperband`, `mlr3mbo`, `bbotk`, `mlr3tuningspaces`; Feature Selection: `mlr3filters`, `mlr3fselect`; Utilities: `mlr3misc`, `mlr3viz`, `mlr3verse`, `mlr3batchmark`, `paradox`; Special Tasks: `mlr3spatiotempcv`, `mlr3spatial`, `mlr3proba`, `mlr3cluster`, `mlr3fda`, `mlr3fairness`; Data: `mlr3db`, `mlr3oml`, `mlr3data`; Pipelines: `mlr3pipelines`. `mlr3fda` and `mlr3torch` are connected by gray dashed lines."
include_multi_graphics("mlr3_ecosystem")
```

A complete and up-to-date list of extension packages can be found at `r link("https://mlr-org.com/ecosystem.html")`.

As well as packages within the `mlr3` ecosystem, software in the `mlr3verse` also depends on the following popular and well-established packages:

*   `r ref_pkg("R6")`: The class system predominantly used in `mlr3`.
*   `r ref_pkg("data.table")`: High-performance extension of R's `data.frame`.
*   `r ref_pkg("digest")`: Cryptographic hash functions.
*   `r ref_pkg("uuid")`: Generation of universally unique identifiers.
*   `r ref_pkg("lgr")`: Configurable logging library.
*   `r ref_pkg("mlbench")` and `r ref_pkg("palmerpenguins")`: Machine learning datasets.
*   `r ref_pkg("future")` / `r ref_pkg("future.apply")` / `r ref_pkg("parallelly")`: For parallelization (@sec-parallelization).
*   `r ref_pkg("evaluate")`: For capturing output, warnings, and exceptions (@sec-error-handling).

We build on `r ref_pkg("R6")` for object orientation and `r ref_pkg("data.table")` to store and operate on tabular data.
As both are core to `mlr3` we *briefly* introduce both packages for beginners; in-depth expertise with these packages is not necessary to work with `mlr3`.

### R6 for Beginners {#sec-r6}

`r ref_pkg("R6")` is one of R's more recent paradigms for `r index('object-oriented programming')`.
If you have experience with any (class) object-oriented programming then R6 should feel familiar.
We focus on the parts of R6 that you need to know to use `mlr3`.

*Objects* are created by constructing an instance of an `r ref("R6::R6Class")` variable using the `$new()` initialization method.
For example, say we have implemented a class called `Foo`, then  `foo = Foo$new(bar = 1)` would create a new object of class `Foo` and set the `bar` argument of the constructor to the value `1`.
In practice, we implement a lot of sugar functionality (@sec-mlr3-utilities) in `mlr3` that make construction and access a bit more convenient.

Some `R6` objects may have mutable states that are encapsulated in their *fields*, which can be accessed through the dollar, `$`, operator.
Continuing the previous example, we can access the `bar` value in the `foo` object by using `foo$bar` or we could give it a new value, e.g. `foo$bar = 2`.
These fields can also be 'active bindings', which perform additional computations when referenced or modified.

In addition to fields, *methods* allow users to inspect the object's state, retrieve information, or perform an action that changes the internal state of the object.
For example, in `mlr3`, the `$train()` method of a learner changes the internal state of the learner by building and storing a model.
Methods that modify the internal state of an object often return the object itself.
Other methods may return a new R6 object.
In both cases, it is possible to 'chain' methods by calling one immediately after the other using the `$`-operator; this is similar to the `%>%`-operator used in `tidyverse` packages.
For example, `Foo$bar()$hello_world()` would run the `$bar()` method of the object `Foo` and then the `$hello_world()` method of the object returned by `$bar()` (which may be `Foo` itself).

Fields and methods can be public or private.
The public fields and methods define the API to interact with the object.
In `mlr3`, you can safely ignore private methods unless you are looking to extend our universe by adding a new class (@sec-technical).

Finally, `R6` objects are `environments`, and as such have reference semantics.
This means that, for example, `foo2 = foo` does not create a new variable called `foo2` that is a copy of `foo`.
Instead, it creates a variable called `foo2` that references `foo`, and so setting `foo$bar = 3` will also change `foo2$bar` to `3` and vice versa.
To copy an object, use the `$clone(deep = TRUE)` method, so to copy `foo`: `foo2 = foo$clone(deep = TRUE)`[`$clone()`]{.aside}.

For a longer introduction, we recommend the `R6` vignettes found at `r link("https://r6.r-lib.org/")`; more detail can be found in `r link("https://adv-r.hadley.nz/r6.html")`.

### data.table for Beginners {#sec-data.table}

The package `r ref_pkg("data.table")` implements `r ref("data.table()")`, which is a popular alternative to R's `data.frame()`.
We use `r ref_pkg("data.table")` because it is blazingly fast and scales well to bigger data.

As with `data.frame`, `data.table`s can be constructed with `r ref("data.table()")` or `r ref("as.data.table()")`:

```{r introduction_and_overview-010.table-001.table-002}
library(data.table)
# converting a matrix with as.data.table
as.data.table(matrix(runif(4), 2, 2))

# using data.table
dt = data.table(x = 1:6, y = rep(letters[1:3], each = 2))
dt
```

`data.table`s can be used much like `data.frame`s, but they provide additional functionality that makes complex operations easier.
For example, data can be summarized by groups with a `by` argument in the `[` operator and they can be modified in-place with the `:=` operator.

```{r introduction_and_overview-011.table-003.table-004}
# mean of x column in groups given by y
dt[, mean(x), by = "y"]
# adding a new column with :=
dt[, z := x * 3]
dt
```

Finally `data.table` also uses reference semantics so you will need to use `r ref("data.table::copy()")` to clone a `data.table`.
For an in-depth introduction, we recommend the vignette @datatable.

## Essential mlr3 Utilities {#sec-mlr3-utilities}

`mlr3` includes a few important utilities that are essential to simplifying code in our ecosystem.

### Sugar Functions {.unnumbered .unlisted}

Most objects in `mlr3` can be created through convenience functions called helper functions or `r index("sugar functions")`.
They provide shortcuts for common code idioms, reducing the amount of code a user has to write.
For example `lrn("regr.rpart")` returns the learner without having to explicitly create a new R6 object.
We heavily use sugar functions throughout this book and provide the equivalent "full form" for complete detail at the end of each chapter.
The sugar functions are designed to cover the majority of use cases for most users, knowledge about the full `R6` backend is only required if you want to build custom objects or extensions.

Many object names in `mlr3` are standardized according to the convention: `mlr_<type>_<key>`, where `<type>` will be `tasks`, `learners`, `measures`, and other classes that will be covered in the book, and `<key>` refers to the ID of the object.
To simplify the process of constructing objects, you only need to know the object key and the sugar function for constructing the type.
For example: `mlr_tasks_mtcars` becomes `tsk("mtcars")`;`mlr_learners_regr.rpart` becomes `lrn("regr.rpart")`; and `mlr_measures_regr.mse` becomes `msr("regr.mse")`.
Throughout this book, we will refer to all objects using this abbreviated form.

### Dictionaries {.unnumbered .unlisted}

`mlr3` uses `r index('dictionaries')` to store R6 classes, which associate keys (unique identifiers) with objects (R6 objects).
Values in dictionaries are often accessed through sugar functions that retrieve objects from the relevant dictionary, for example `lrn("regr.rpart")` is a wrapper around `mlr_learners$get("regr.rpart")` and is thus a simpler way to load a decision tree learner from `r ref("mlr_learners")`.
We use dictionaries to group large collections of relevant objects so they can be listed and retrieved easily.
For example, you can see an overview of available learners (that are in loaded packages) and their properties with `as.data.table(mlr_learners)` or by calling the sugar function without any arguments, e.g. `lrn()`.

### mlr3viz {.unnumbered .unlisted}

`r mlr3viz` includes all plotting functionality in `mlr3` and uses `r ref_pkg("ggplot2")` under the hood.
We use `r ref("ggplot2::theme_minimal()")` in all our plots to unify our aesthetic, but as with all `ggplot` outputs, users can fully customize this.
`r mlr3viz` extends `fortify` and `autoplot` for use with common `r mlr3` outputs including `r ref("Prediction")`, `r ref("Learner")`, and `r ref("BenchmarkResult")` objects (which we will introduce and cover in the next chapters).
We will cover major plot types throughout the book.
The best way to learn about `r mlr3viz` is through experimentation; load the package and see what happens when you run `autoplot` on an `mlr3` object.
Plot types are documented in the respective manual page that can be accessed through `?autoplot.<class>`, for example, you can find different types of plots for regression tasks by running `?autoplot.TaskRegr`.

## Design Principles

{{< include ../../common/_optional.qmd >}}

Learning from over a decade of design and adaptation from `r ref_pkg("mlr")` to `r mlr3`, we now follow these design principles in the `r mlr3` ecosystem:

*   **Object-oriented programming**.
We embrace `r ref_pkg("R6")` for a clean, object-oriented design, object state changes, and reference semantics.
This means that the state of common objects (e.g. tasks (@sec-tasks) and learners (@sec-learners)) is encapsulated within the object, for example, to keep track of whether a model has been trained, without the user having to worry about this.
We also use inheritance to specialize objects, e.g. all learners are derived from a common base class that provides basic functionality.
*   **Tabular data**.
Embrace `r ref_pkg("data.table")` for its top-notch computational performance as well as tabular data as a structure that can be easily processed further.
*   **Unified tabular input and output data formats.**
This considerably simplifies the API and allows easy selection and "split-apply-combine" (aggregation) operations.
We combine `data.table` and `R6` to place references to non-atomic and compound objects in tables and make heavy use of list columns.
*   **Defensive programming and type safety**.
All user input is checked with `r ref_pkg("checkmate")` [@checkmate].
We use `data.table`, which has behavior that is more consistent than several base R methods (e.g., indexing `data.frame`s simplifies the result when the `drop` argument is omitted).
And we have extensive unit tests!
*   **Light on dependencies**.
One of the main maintenance burdens for `r ref_pkg("mlr")` was to keep up with changing learner interfaces and behavior of the many packages it depended on.
We require far fewer packages in `r mlr3`, which makes installation and maintenance easier.
We still provide the same functionality, but it is split into more packages that have fewer dependencies individually.
*   **Separation of computation and presentation**.
Most packages of the `r mlr3` ecosystem focus on processing and transforming data, applying ML algorithms, and computing results.
Our core packages do not provide visualizations because their dependencies would make installation unnecessarily complex, especially on headless servers (i.e., computers without a monitor where graphical libraries are not installed).
Hence, visualizations of data and results are provided in `r mlr3viz`.

::: {.content-visible when-format="html"}
`r citeas(chapter)`
:::
