\hypertarget{references}{%
\chapter*{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Apley2020}{}}%
Apley, Daniel W., and Jingyu Zhu. 2020. {``Visualizing the Effects of
Predictor Variables in Black Box Supervised Learning Models.''}
\emph{Journal of the Royal Statistical Society Series B: Statistical
Methodology} 82 (4): 1059--86. \url{https://doi.org/10.1111/rssb.12377}.

\leavevmode\vadjust pre{\hypertarget{ref-Au2022}{}}%
Au, Quay, Julia Herbinger, Clemens Stachl, Bernd Bischl, and Giuseppe
Casalicchio. 2022. {``Grouped Feature Importance and Combined Features
Effect Plot.''} \emph{Data Mining and Knowledge Discovery} 36 (4):
1401--50. \url{https://doi.org/10.1007/s10618-022-00840-5}.

\leavevmode\vadjust pre{\hypertarget{ref-bagnall2017great}{}}%
Bagnall, Anthony, Jason Lines, Aaron Bostrom, James Large, and Eamonn
Keogh. 2017. {``The Great Time Series Classification Bake Off: A Review
and Experimental Evaluation of Recent Algorithmic Advances.''}
\emph{Data Mining and Knowledge Discovery} 31: 606--60.
\url{https://doi.org/10.1007/s10618-016-0483-9}.

\leavevmode\vadjust pre{\hypertarget{ref-Baniecki2019}{}}%
Baniecki, Hubert, and Przemyslaw Biecek. 2019. {``{modelStudio}:
Interactive Studio with Explanations for ML Predictive Models.''}
\emph{Journal of Open Source Software} 4 (43): 1798.
\url{https://doi.org/10.21105/joss.01798}.

\leavevmode\vadjust pre{\hypertarget{ref-Baniecki2023}{}}%
Baniecki, Hubert, Dariusz Parzych, and Przemyslaw Biecek. 2023. {``The
Grammar of Interactive Explanatory Model Analysis.''} \emph{Data Mining
and Knowledge Discovery}, 1573--756X.
\url{https://doi.org/10.1007/s10618-023-00924-w}.

\leavevmode\vadjust pre{\hypertarget{ref-fairmlbook}{}}%
Barocas, Solon, Moritz Hardt, and Arvind Narayanan. 2019. \emph{Fairness
and Machine Learning: Limitations and Opportunities}. fairmlbook.org.

\leavevmode\vadjust pre{\hypertarget{ref-future119}{}}%
Bengtsson, Henrik. 2020. {``Future 1.19.1 - Making Sure Proper Random
Numbers Are Produced in Parallel Processing.''}
\url{https://www.jottr.org/2020/09/22/push-for-statistical-sound-rng/}.

\leavevmode\vadjust pre{\hypertarget{ref-avoiddetect}{}}%
---------. 2022. {``Please Avoid detectCores() in Your {R} Packages.''}
\url{https://www.jottr.org/2022/12/05/avoid-detectcores/}.

\leavevmode\vadjust pre{\hypertarget{ref-bergstra2012}{}}%
Bergstra, James, and Yoshua Bengio. 2012. {``Random Search for
Hyper-Parameter Optimization.''} \emph{Journal of Machine Learning
Research} 13: 281--305.
\url{https://jmlr.org/papers/v13/bergstra12a.html}.

\leavevmode\vadjust pre{\hypertarget{ref-Biecek2018}{}}%
Biecek, Przemyslaw. 2018. {``{DALEX}: Explainers for Complex Predictive
Models in {R}.''} \emph{Journal of Machine Learning Research} 19 (84):
1--5. \url{https://jmlr.org/papers/v19/18-416.html}.

\leavevmode\vadjust pre{\hypertarget{ref-biecek_burzykowski_2021}{}}%
Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory
Model Analysis}. Chapman; Hall/CRC, New York.
\url{https://ema.drwhy.ai/}.

\leavevmode\vadjust pre{\hypertarget{ref-binder2020}{}}%
Binder, Martin, Florian Pfisterer, and Bernd Bischl. 2020. {``Collecting
Empirical Data about Hyperparameters for Data Driven AutoML.''} In
\emph{Proceedings of the 7th ICML Workshop on Automated Machine Learning
(AutoML 2020)}.
\url{https://www.automl.org/wp-content/uploads/2020/07/AutoML_2020_paper_63.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-mlr3pipelines}{}}%
Binder, Martin, Florian Pfisterer, Michel Lang, Lennart Schneider, Lars
Kotthoff, and Bernd Bischl. 2021. {``{mlr3pipelines} - Flexible Machine
Learning Pipelines in {R}.''} \emph{Journal of Machine Learning
Research} 22 (184): 1--7.
\url{https://jmlr.org/papers/v22/21-0281.html}.

\leavevmode\vadjust pre{\hypertarget{ref-hpo_practical}{}}%
Bischl, Bernd, Martin Binder, Michel Lang, Tobias Pielok, Jakob Richter,
Stefan Coors, Janek Thomas, et al. 2023. {``Hyperparameter Optimization:
Foundations, Algorithms, Best Practices, and Open Challenges.''}
\emph{Wiley Interdisciplinary Reviews: Data Mining and Knowledge
Discovery}, e1484. \url{https://doi.org/10.1002/widm.1484}.

\leavevmode\vadjust pre{\hypertarget{ref-bischl2021openml}{}}%
Bischl, Bernd, Giuseppe Casalicchio, Matthias Feurer, Pieter Gijsbers,
Frank Hutter, Michel Lang, Rafael Gomes Mantovani, Jan N. van Rijn, and
Joaquin Vanschoren. 2021. {``Open{ML} Benchmarking Suites.''} In
\emph{Thirty-Fifth Conference on Neural Information Processing Systems
Datasets and Benchmarks Track (Round 2)}.
\url{https://openreview.net/forum?id=OCrD8ycKjG}.

\leavevmode\vadjust pre{\hypertarget{ref-mlr}{}}%
Bischl, Bernd, Michel Lang, Lars Kotthoff, Julia Schiffner, Jakob
Richter, Erich Studerus, Giuseppe Casalicchio, and Zachary M. Jones.
2016. {``{mlr}: {M}achine {L}earning in {R}.''} \emph{Journal of Machine
Learning Research} 17 (170): 1--5.
\url{https://jmlr.org/papers/v17/15-066.html}.

\leavevmode\vadjust pre{\hypertarget{ref-JSSv064i11}{}}%
Bischl, Bernd, Michel Lang, Olaf Mersmann, Jörg Rahnenführer, and Claus
Weihs. 2015. {``BatchJobs and BatchExperiments: Abstraction Mechanisms
for Using r in Batch Environments.''} \emph{Journal of Statistical
Software} 64 (11): 1--25. \url{https://doi.org/10.18637/jss.v064.i11}.

\leavevmode\vadjust pre{\hypertarget{ref-bischl2012resampling}{}}%
Bischl, Bernd, Olaf Mersmann, Heike Trautmann, and Claus Weihs. 2012.
{``Resampling Methods for Meta-Model Validation with Recommendations for
Evolutionary Computation.''} \emph{Evolutionary Computation} 20 (2):
249--75.
\href{https://doi.org/10.1162/EVCO_a_00069\%20}{https://doi.org/10.1162/EVCO\_a\_00069
}.

\leavevmode\vadjust pre{\hypertarget{ref-bishop_2006}{}}%
Bishop, Christopher M. 2006. \emph{Pattern Recognition and Machine
Learning}. Springer.

\leavevmode\vadjust pre{\hypertarget{ref-bommert2020}{}}%
Bommert, Andrea, Xudong Sun, Bernd Bischl, Jörg Rahnenführer, and Michel
Lang. 2020. {``Benchmark for Filter Methods for Feature Selection in
High-Dimensional Classification Data.''} \emph{Computational Statistics
\& Data Analysis} 143: 106839.
\url{https://doi.org/10.1016/j.csda.2019.106839}.

\leavevmode\vadjust pre{\hypertarget{ref-Breiman1996}{}}%
Breiman, Leo. 1996. {``Bagging Predictors.''} \emph{Machine Learning} 24
(2): 123--40. \url{https://doi.org/10.1007/BF00058655}.

\leavevmode\vadjust pre{\hypertarget{ref-breiman2001random}{}}%
---------. 2001a. {``Random Forests.''} \emph{Machine Learning} 45:
5--32. \url{https://doi.org/10.1023/A:1010933404324}.

\leavevmode\vadjust pre{\hypertarget{ref-Breiman2001}{}}%
---------. 2001b. {``Statistical Modeling: The Two Cultures (with
Comments and a Rejoinder by the Author).''} \emph{Statistical Science}
16 (3). \url{https://doi.org/10.1214/ss/1009213726}.

\leavevmode\vadjust pre{\hypertarget{ref-Bucker2022}{}}%
Bücker, Michael, Gero Szepannek, Alicja Gosiewska, and Przemyslaw
Biecek. 2022. {``Transparency, Auditability, and Explainability of
Machine Learning Models in Credit Scoring.''} \emph{Journal of the
Operational Research Society} 73 (1): 70--90.
\url{https://doi.org/10.1080/01605682.2021.1922098}.

\leavevmode\vadjust pre{\hypertarget{ref-byrd1995limited}{}}%
Byrd, Richard H., Peihuang Lu, Jorge Nocedal, and Ciyou Zhu. 1995. {``A
Limited Memory Algorithm for Bound Constrained Optimization.''}
\emph{SIAM Journal on Scientific Computing} 16 (5): 1190--1208.
\url{https://doi.org/10.1137/0916069}.

\leavevmode\vadjust pre{\hypertarget{ref-caton-arxiv20a}{}}%
Caton, S., and C. Haas. 2020. {``Fairness in Machine Learning: A
Survey.''} \emph{Arxiv} 2010.04053 {{[}cs.LG{]}}.
\url{https://doi.org/10.48550/arXiv.2010.04053}.

\leavevmode\vadjust pre{\hypertarget{ref-chandrashekar2014}{}}%
Chandrashekar, Girish, and Ferat Sahin. 2014. {``A Survey on Feature
Selection Methods.''} \emph{Computers and Electrical Engineering} 40
(1): 16--28. \url{https://doi.org/10.1016/j.compeleceng.2013.11.024}.

\leavevmode\vadjust pre{\hypertarget{ref-chen2016xgboost}{}}%
Chen, Tianqi, and Carlos Guestrin. 2016. {``{XGB}oost: A Scalable Tree
Boosting System.''} In \emph{Proceedings of the 22nd {ACM SIGKDD}
International Conference on Knowledge Discovery and Data Mining},
785--94. \url{https://doi.org/10.1145/2939672.2939785}.

\leavevmode\vadjust pre{\hypertarget{ref-Collett2014}{}}%
Collett, David. 2014. \emph{Modelling Survival Data in Medical
Research}. 3rd ed. CRC. \url{https://doi.org/10.1201/b18041}.

\leavevmode\vadjust pre{\hypertarget{ref-couronne2018random}{}}%
Couronné, Raphael, Philipp Probst, and Anne-Laure Boulesteix. 2018.
{``Random Forest Versus Logistic Regression: A Large-Scale Benchmark
Experiment.''} \emph{BMC Bioinformatics} 19: 1--14.
\url{https://doi.org/10.1186/s12859-018-2264-5}.

\leavevmode\vadjust pre{\hypertarget{ref-Dandl2020}{}}%
Dandl, Susanne, Christoph Molnar, Martin Binder, and Bernd Bischl. 2020.
{``Multi-Objective Counterfactual Explanations.''} In \emph{Parallel
Problem Solving from Nature {PPSN} {XVI}}, 448--69. Springer
International Publishing.
\url{https://doi.org/10.1007/978-3-030-58112-1_31}.

\leavevmode\vadjust pre{\hypertarget{ref-davis2006relationship}{}}%
Davis, Jesse, and Mark Goadrich. 2006. {``The Relationship Between
Precision-Recall and ROC Curves.''} In \emph{Proceedings of the 23rd
International Conference on Machine Learning}, 233--40.
\url{https://doi.org/10.1145/1143844.1143874}.

\leavevmode\vadjust pre{\hypertarget{ref-de2011ames}{}}%
De Cock, Dean. 2011. {``Ames, Iowa: Alternative to the Boston Housing
Data as an End of Semester Regression Project.''} \emph{Journal of
Statistics Education} 19 (3).
\url{https://doi.org/10.1080/10691898.2011.11889627}.

\leavevmode\vadjust pre{\hypertarget{ref-demsar2006}{}}%
Demšar, Janez. 2006. {``Statistical Comparisons of Classifiers over
Multiple Data Sets.''} \emph{Journal of Machine Learning Research} 7
(1): 1--30. \url{https://jmlr.org/papers/v7/demsar06a.html}.

\leavevmode\vadjust pre{\hypertarget{ref-ding2010investigation}{}}%
Ding, Yufeng, and Jeffrey S Simonoff. 2010. {``An Investigation of
Missing Data Methods for Classification Trees Applied to Binary Response
Data.''} \emph{Journal of Machine Learning Research} 11 (6): 131--70.
\url{https://www.jmlr.org/papers/v11/ding10a.html}.

\leavevmode\vadjust pre{\hypertarget{ref-dobbin2011}{}}%
Dobbin, Kevin K., and Richard M. Simon. 2011. {``Optimally Splitting
Cases for Training and Testing High Dimensional Classifiers.''}
\emph{BMC Medical Genomics} 4 (1): 31.
\url{https://doi.org/10.1186/1755-8794-4-31}.

\leavevmode\vadjust pre{\hypertarget{ref-uci}{}}%
Dua, Dheeru, and Casey Graff. 2017. {``{UCI} Machine Learning
Repository.''} University of California, Irvine, School of Information;
Computer Sciences. \url{https://archive.ics.uci.edu/ml}.

\leavevmode\vadjust pre{\hypertarget{ref-Eddelbuettel2020}{}}%
Eddelbuettel, Dirk. 2020. {``Parallel Computing with {R}: A Brief
Review.''} \emph{{WIREs} Computational Statistics} 13 (2).
\url{https://doi.org/10.1002/wics.1515}.

\leavevmode\vadjust pre{\hypertarget{ref-hpo_automl}{}}%
Feurer, Matthias, and Frank Hutter. 2019. {``Hyperparameter
Optimization.''} In \emph{Automated Machine Learning: Methods, Systems,
Challenges}, edited by Frank Hutter, Lars Kotthoff, and Joaquin
Vanschoren, 3--33. Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-05318-5_1}.

\leavevmode\vadjust pre{\hypertarget{ref-feurer2015initializing}{}}%
Feurer, Matthias, Jost Springenberg, and Frank Hutter. 2015.
{``Initializing {B}ayesian Hyperparameter Optimization via
Meta-Learning.''} In \emph{Proceedings of the AAAI Conference on
Artificial Intelligence}. Vol. 29. 1.
\url{https://doi.org/10.1609/aaai.v29i1.9354}.

\leavevmode\vadjust pre{\hypertarget{ref-Fisher2019pfi}{}}%
Fisher, Aaron, Cynthia Rudin, and Francesca Dominici. 2019. {``All
Models Are Wrong, but Many Are Useful: Learning a Variable's Importance
by Studying an Entire Class of Prediction Models Simultaneously.''}
\url{https://doi.org/10.48550/arxiv.1801.01489}.

\leavevmode\vadjust pre{\hypertarget{ref-Friedman2001pdp}{}}%
Friedman, Jerome H. 2001. {``Greedy Function Approximation: A Gradient
Boosting Machine.''} \emph{The Annals of Statistics} 29 (5).
\url{https://doi.org/10.1214/aos/1013203451}.

\leavevmode\vadjust pre{\hypertarget{ref-garnett_2022}{}}%
Garnett, Roman. 2022. \emph{Bayesian Optimization}. Cambridge University
Press. \url{https://bayesoptbook.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-amlb2022}{}}%
Gijsbers, Pieter, Marcos L. P. Bueno, Stefan Coors, Erin LeDell,
Sébastien Poirier, Janek Thomas, Bernd Bischl, and Joaquin Vanschoren.
2022. {``AMLB: An AutoML Benchmark.''} arXiv.
\url{https://doi.org/10.48550/ARXIV.2207.12560}.

\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2015ice}{}}%
Goldstein, Alex, Adam Kapelner, Justin Bleich, and Emil Pitkin. 2015.
{``Peeking Inside the Black Box: Visualizing Statistical Learning with
Plots of Individual Conditional Expectation.''} \emph{Journal of
Computational and Graphical Statistics} 24 (1): 44--65.
\url{https://doi.org/10.1080/10618600.2014.907095}.

\leavevmode\vadjust pre{\hypertarget{ref-gower1971general}{}}%
Gower, John C. 1971. {``A General Coefficient of Similarity and Some of
Its Properties.''} \emph{Biometrics}, 857--71.
\url{https://doi.org/10.2307/2528823}.

\leavevmode\vadjust pre{\hypertarget{ref-grinsztajn2022why}{}}%
Grinsztajn, Leo, Edouard Oyallon, and Gael Varoquaux. 2022. {``Why Do
Tree-Based Models Still Outperform Deep Learning on Typical Tabular
Data?''} In \emph{Thirty-Sixth Conference on Neural Information
Processing Systems Datasets and Benchmarks Track}.
\url{https://openreview.net/forum?id=Fp7__phQszn}.

\leavevmode\vadjust pre{\hypertarget{ref-guidotti2022counterfactual}{}}%
Guidotti, Riccardo. 2022. {``Counterfactual Explanations and How to Find
Them: Literature Review and Benchmarking.''} \emph{Data Mining and
Knowledge Discovery}, 1--55.
\url{https://doi.org/10.1007/s10618-022-00831-6}.

\leavevmode\vadjust pre{\hypertarget{ref-guidotti2018survey}{}}%
Guidotti, Riccardo, Anna Monreale, Salvatore Ruggieri, Franco Turini,
Fosca Giannotti, and Dino Pedreschi. 2018. {``A Survey of Methods for
Explaining Black Box Models.''} \emph{ACM Computing Surveys (CSUR)} 51
(5): 1--42. \url{https://doi.org/10.1145/3236009}.

\leavevmode\vadjust pre{\hypertarget{ref-guyon2003}{}}%
Guyon, Isabelle, and André Elisseeff. 2003. {``An Introduction to
Variable and Feature Selection.''} \emph{Journal of Machine Learning
Research} 3 (Mar): 1157--82.
\url{https://www.jmlr.org/papers/v3/guyon03a.html}.

\leavevmode\vadjust pre{\hypertarget{ref-hand2001simple}{}}%
Hand, David J, and Robert J Till. 2001. {``A Simple Generalisation of
the Area Under the ROC Curve for Multiple Class Classification
Problems.''} \emph{Machine Learning} 45: 171--86.
\url{https://doi.org/10.1023/A:1010920819831}.

\leavevmode\vadjust pre{\hypertarget{ref-hansen2011}{}}%
Hansen, Nikolaus, and Anne Auger. 2011. {``CMA-ES: Evolution Strategies
and Covariance Matrix Adaptation.''} In \emph{Proceedings of the 13th
Annual Conference Companion on Genetic and Evolutionary Computation},
991--1010. \url{https://doi.org/10.1145/2001858.2002123}.

\leavevmode\vadjust pre{\hypertarget{ref-hastie2001}{}}%
Hastie, Trevor, Jerome Friedman, and Robert Tibshirani. 2001. \emph{The
Elements of Statistical Learning}. Springer New York.
\url{https://doi.org/10.1007/978-0-387-21606-5}.

\leavevmode\vadjust pre{\hypertarget{ref-Hooker2019PleaseSP}{}}%
Hooker, Giles, and Lucas K. Mentch. 2019. {``Please Stop Permuting
Features: An Explanation and Alternatives.''}
\url{https://doi.org/10.48550/arxiv.1905.03151}.

\leavevmode\vadjust pre{\hypertarget{ref-Horn2015}{}}%
Horn, Daniel, Tobias Wagner, Dirk Biermann, Claus Weihs, and Bernd
Bischl. 2015. {``Model-Based Multi-Objective Optimization: Taxonomy,
Multi-Point Proposal, Toolbox and Benchmark.''} In \emph{Evolutionary
Multi-Criterion Optimization}, edited by António Gaspar-Cunha, Carlos
Henggeler Antunes, and Carlos Coello Coello, 64--78.
\url{https://doi.org/10.1007/978-3-319-15934-8_5}.

\leavevmode\vadjust pre{\hypertarget{ref-huang_2012}{}}%
Huang, D., T. T. Allen, W. I. Notz, and N. Zheng. 2012. {``Erratum to:
Global Optimization of Stochastic Black-Box Systems via Sequential
Kriging Meta-Models.''} \emph{Journal of Global Optimization} 54 (2):
431--31. \url{https://doi.org/10.1007/s10898-011-9821-z}.

\leavevmode\vadjust pre{\hypertarget{ref-Huang2022}{}}%
Huang, Jonathan, Galal Galal, Mozziyar Etemadi, and Mahesh Vaidyanathan.
2022. {``Evaluation and Mitigation of Racial Bias in Clinical Machine
Learning Models: Scoping Review.''} \emph{JMIR Med Inform} 10 (5).
\url{https://doi.org/10.2196/36388}.

\leavevmode\vadjust pre{\hypertarget{ref-hutter2019automated}{}}%
Hutter, Frank, Lars Kotthoff, and Joaquin Vanschoren, eds. 2019.
\emph{Automated Machine Learning - Methods, Systems, Challenges}.
Springer.

\leavevmode\vadjust pre{\hypertarget{ref-datatable}{}}%
{``Introduction to Data.table.''} 2023.
\url{https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html}.

\leavevmode\vadjust pre{\hypertarget{ref-james_introduction_2014}{}}%
James, Gareth, Daniela Witten, Trevor Hastie, and Robert Tibshirani.
2014. \emph{An Introduction to Statistical Learning: With Applications
in {R}}. Springer Publishing Company, Incorporated.
\url{https://doi.org/10.1007/978-1-4614-7138-7}.

\leavevmode\vadjust pre{\hypertarget{ref-jamieson_2016}{}}%
Jamieson, Kevin, and Ameet Talwalkar. 2016. {``Non-Stochastic Best Arm
Identification and Hyperparameter Optimization.''} In \emph{Proceedings
of the 19th International Conference on Artificial Intelligence and
Statistics}, edited by Arthur Gretton and Christian C. Robert,
51:240--48. Proceedings of Machine Learning Research. Cadiz, Spain:
PMLR. \url{https://proceedings.mlr.press/v51/jamieson16.html}.

\leavevmode\vadjust pre{\hypertarget{ref-japkowicz2011evaluating}{}}%
Japkowicz, Nathalie, and Mohak Shah. 2011. \emph{Evaluating Learning
Algorithms: A Classification Perspective}. Cambridge University Press.
\url{https://doi.org/10.1017/CBO9780511921803}.

\leavevmode\vadjust pre{\hypertarget{ref-jones_1993_lipschitzian}{}}%
Jones, Donald R., Cary D. Perttunen, and Bruce E. Stuckman. 1993.
{``Lipschitzian Optimization Without the {L}ipschitz Constant.''}
\emph{Journal of Optimization Theory and Applications} 79 (1): 157--81.
\url{https://doi.org/10.1007/BF00941892}.

\leavevmode\vadjust pre{\hypertarget{ref-jones_1998}{}}%
Jones, Donald R., Matthias Schonlau, and William J. Welch. 1998.
{``Efficient Global Optimization of Expensive Black-Box Functions.''}
\emph{Journal of Global Optimization} 13 (4): 455--92.
\url{https://doi.org/10.1023/A:1008306431147}.

\leavevmode\vadjust pre{\hypertarget{ref-Kalbfleisch2011}{}}%
Kalbfleisch, John D, and Ross L Prentice. 2011. \emph{The Statistical
Analysis of Failure Time Data}. Vol. 360. John Wiley \& Sons.
\url{https://doi.org/10.1002/9781118032985}.

\leavevmode\vadjust pre{\hypertarget{ref-karl2022}{}}%
Karl, Florian, Tobias Pielok, Julia Moosbauer, Florian Pfisterer, Stefan
Coors, Martin Binder, Lennart Schneider, et al. 2022. {``Multi-Objective
Hyperparameter Optimization--an Overview.''} \emph{arXiv Preprint
arXiv:2206.07438}. \url{https://doi.org/10.48550/arXiv.2206.07438}.

\leavevmode\vadjust pre{\hypertarget{ref-kim2009estimating}{}}%
Kim, Ji-Hyun. 2009. {``Estimating Classification Error Rate: Repeated
Cross-Validation, Repeated Hold-Out and Bootstrap.''}
\emph{Computational Statistics \& Data Analysis} 53 (11): 3735--45.
\url{https://doi.org/10.1016/j.csda.2009.04.009}.

\leavevmode\vadjust pre{\hypertarget{ref-kim_2021}{}}%
Kim, Jungtaek, and Seungjin Choi. 2021. {``On Local Optimizers of
Acquisition Functions in Bayesian Optimization.''} In \emph{Machine
Learning and Knowledge Discovery in Databases}, edited by Frank Hutter,
Kristian Kersting, Jefrey Lijffijt, and Isabel Valera, 675--90.
\url{https://doi.org/10.1007/978-3-030-67661-2_40}.

\leavevmode\vadjust pre{\hypertarget{ref-knowles_2006}{}}%
Knowles, Joshua. 2006. {``ParEGO: A Hybrid Algorithm with on-Line
Landscape Approximation for Expensive Multiobjective Optimization
Problems.''} \emph{IEEE Transactions on Evolutionary Computation} 10
(1): 50--66. \url{https://doi.org/10.1109/TEVC.2005.851274}.

\leavevmode\vadjust pre{\hypertarget{ref-kohavi1995}{}}%
Kohavi, Ron. 1995. {``A Study of Cross-Validation and Bootstrap for
Accuracy Estimation and Model Selection.''} In \emph{Proceedings of the
14th International Joint Conference on {Artificial} Intelligence -
{Volume} 2}, 1137--43. {IJCAI}'95. {San Francisco, CA, USA}: {Morgan
Kaufmann Publishers Inc.}

\leavevmode\vadjust pre{\hypertarget{ref-Kohavi1997}{}}%
Kohavi, Ron, and George H. John. 1997. {``Wrappers for Feature Subset
Selection.''} \emph{Artificial Intelligence} 97 (1): 273--324.
\url{https://doi.org/10.1016/S0004-3702(97)00043-X}.

\leavevmode\vadjust pre{\hypertarget{ref-Krzyzinski2023}{}}%
Krzyziński, Mateusz, Mikołaj Spytek, Hubert Baniecki, and Przemysław
Biecek. 2023. {``{SurvSHAP(t)}: Time-Dependent Explanations of Machine
Learning Survival Models.''} \emph{Knowledge-Based Systems} 262: 110234.
\url{https://doi.org/10.1016/j.knosys.2022.110234}.

\leavevmode\vadjust pre{\hypertarget{ref-kuehn_2018}{}}%
Kuehn, Daniel, Philipp Probst, Janek Thomas, and Bernd Bischl. 2018.
{``Automatic Exploration of Machine Learning Experiments on OpenML.''}
\url{https://arxiv.org/abs/1806.10961}.

\leavevmode\vadjust pre{\hypertarget{ref-checkmate}{}}%
Lang, Michel. 2017. {``{checkmate}: Fast Argument Checks for Defensive
{R} Programming.''} \emph{The {R} Journal} 9 (1): 437--45.
\url{https://doi.org/10.32614/RJ-2017-028}.

\leavevmode\vadjust pre{\hypertarget{ref-mlr3}{}}%
Lang, Michel, Martin Binder, Jakob Richter, Patrick Schratz, Florian
Pfisterer, Stefan Coors, Quay Au, Giuseppe Casalicchio, Lars Kotthoff,
and Bernd Bischl. 2019. {``{mlr3}: A Modern Object-Oriented Machine
Learning Framework in {R}.''} \emph{Journal of Open Source Software},
December. \url{https://doi.org/10.21105/joss.01903}.

\leavevmode\vadjust pre{\hypertarget{ref-batchtools}{}}%
Lang, Michel, Bernd Bischl, and Dirk Surmann. 2017. {``{batchtools}:
Tools for {R} to Work on Batch Systems.''} \emph{The Journal of Open
Source Software} 2 (10). \url{https://doi.org/10.21105/joss.00135}.

\leavevmode\vadjust pre{\hypertarget{ref-lecun1998gradient}{}}%
LeCun, Yann, Léon Bottou, Yoshua Bengio, and Patrick Haffner. 1998.
{``Gradient-Based Learning Applied to Document Recognition.''}
\emph{Proceedings of the IEEE} 86 (11): 2278--2324.
\url{https://doi.org/10.1109/5.726791}.

\leavevmode\vadjust pre{\hypertarget{ref-li_2018}{}}%
Li, Lisha, Kevin Jamieson, Giulia DeSalvo, Afshin Rostamizadeh, and
Ameet Talwalkar. 2018. {``Hyperband: A Novel Bandit-Based Approach to
Hyperparameter Optimization.''} \emph{Journal of Machine Learning
Research} 18 (185): 1--52.
\url{https://jmlr.org/papers/v18/16-558.html}.

\leavevmode\vadjust pre{\hypertarget{ref-Lindauer2022}{}}%
Lindauer, Marius, Katharina Eggensperger, Matthias Feurer, André
Biedenkapp, Difan Deng, Carolin Benjamins, Tim Ruhkopf, René Sass, and
Frank Hutter. 2022. {``{SMAC3}: A Versatile Bayesian Optimization
Package for Hyperparameter Optimization.''} \emph{Journal of Machine
Learning Research} 23 (54): 1--9.
\url{https://www.jmlr.org/papers/v23/21-0888.html}.

\leavevmode\vadjust pre{\hypertarget{ref-lipton2018mythos}{}}%
Lipton, Zachary C. 2018. {``The Mythos of Model Interpretability: In
Machine Learning, the Concept of Interpretability Is Both Important and
Slippery.''} \emph{Queue} 16 (3): 31--57.
\url{https://doi.org/10.1145/3236386.3241340}.

\leavevmode\vadjust pre{\hypertarget{ref-lopez2016}{}}%
López-Ibáñez, Manuel, Jérémie Dubois-Lacoste, Leslie Pérez Cáceres,
Mauro Birattari, and Thomas Stützle. 2016. {``The {irace} Package:
Iterated Racing for Automatic Algorithm Configuration.''}
\emph{Operations Research Perspectives} 3: 43--58.
\url{https://doi.org/10.1016/j.orp.2016.09.002}.

\leavevmode\vadjust pre{\hypertarget{ref-Lundberg2019}{}}%
Lundberg, Scott M., Gabriel G. Erion, and Su-In Lee. 2019. {``Consistent
Individualized Feature Attribution for Tree Ensembles.''} arXiv.
\url{https://doi.org/10.48550/arxiv.1802.03888}.

\leavevmode\vadjust pre{\hypertarget{ref-Mehrabi2021}{}}%
Mehrabi, Ninareh, Fred Morstatter, Nripsuta Saxena, Kristina Lerman, and
Aram Galstyan. 2021. {``A Survey on Bias and Fairness in Machine
Learning.''} \emph{ACM Comput. Surv.} 54 (6).
\url{https://doi.org/10.1145/3457607}.

\leavevmode\vadjust pre{\hypertarget{ref-MicciBarreca2001}{}}%
Micci-Barreca, Daniele. 2001. {``A Preprocessing Scheme for
High-Cardinality Categorical Attributes in Classification and Prediction
Problems.''} \emph{{ACM} {SIGKDD} Explorations Newsletter} 3 (1):
27--32. \url{https://doi.org/10.1145/507533.507538}.

\leavevmode\vadjust pre{\hypertarget{ref-mitchell21}{}}%
Mitchell, Shira, Eric Potash, Solon Barocas, Alexander D'Amour, and
Kristian Lum. 2021. {``Algorithmic Fairness: Choices, Assumptions, and
Definitions.''} \emph{Annual Review of Statistics and Its Application}
8: 141--63.
\url{https://doi.org/10.1146/annurev-statistics-042720-125902}.

\leavevmode\vadjust pre{\hypertarget{ref-molinaro2005prediction}{}}%
Molinaro, Annette M, Richard Simon, and Ruth M Pfeiffer. 2005.
{``Prediction Error Estimation: A Comparison of Resampling Methods.''}
\emph{Bioinformatics} 21 (15): 3301--7.
\url{https://doi.org/10.1093/bioinformatics/bti499}.

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2022}{}}%
Molnar, Christoph. 2022. \emph{Interpretable Machine Learning: A Guide
for Making Black Box Models Explainable}. 2nd ed.
\url{https://christophm.github.io/interpretable-ml-book}.

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2018}{}}%
Molnar, Christoph, Bernd Bischl, and Giuseppe Casalicchio. 2018.
{``{iml}: An {R} Package for Interpretable Machine Learning.''}
\emph{JOSS} 3 (26): 786. \url{https://doi.org/10.21105/joss.00786}.

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2022pitfalls}{}}%
Molnar, Christoph, Gunnar König, Julia Herbinger, Timo Freiesleben,
Susanne Dandl, Christian A. Scholbeck, Giuseppe Casalicchio, Moritz
Grosse-Wentrup, and Bernd Bischl. 2022. {``General Pitfalls
of~Model-Agnostic Interpretation Methods for~Machine Learning Models.''}
In \emph{xxAI - Beyond Explainable AI: International Workshop, Held in
Conjunction with ICML 2020, July 18, 2020, Vienna, Austria, Revised and
Extended Papers}, edited by Andreas Holzinger, Randy Goebel, Ruth Fong,
Taesup Moon, Klaus-Robert Müller, and Wojciech Samek, 39--68. Cham:
Springer International Publishing.
\url{https://doi.org/10.1007/978-3-031-04083-2_4}.

\leavevmode\vadjust pre{\hypertarget{ref-Morales2022}{}}%
Morales-Hernández, Alejandro, Inneke Van Nieuwenhuyse, and Sebastian
Rojas Gonzalez. 2022. {``A Survey on Multi-Objective Hyperparameter
Optimization Algorithms for Machine Learning.''} \emph{Artificial
Intelligence Review}, 1--51.
\url{https://doi.org/10.1007/s10462-022-10359-2}.

\leavevmode\vadjust pre{\hypertarget{ref-Niederreiter1988}{}}%
Niederreiter, Harald. 1988. {``Low-Discrepancy and Low-Dispersion
Sequences.''} \emph{Journal of Number Theory} 30 (1): 51--70.
\url{https://doi.org/10.1016/0022-314X(88)90025-X}.

\leavevmode\vadjust pre{\hypertarget{ref-pargent2022regularized}{}}%
Pargent, Florian, Florian Pfisterer, Janek Thomas, and Bernd Bischl.
2022. {``Regularized Target Encoding Outperforms Traditional Methods in
Supervised Machine Learning with High Cardinality Features.''}
\emph{Computational Statistics} 37 (5): 2671--92.
\url{https://doi.org/10.1007/s00180-022-01207-6}.

\leavevmode\vadjust pre{\hypertarget{ref-Poulos2018}{}}%
Poulos, Jason, and Rafael Valle. 2018. {``Missing Data Imputation for
Supervised Learning.''} \emph{Applied Artificial Intelligence} 32 (2):
186--96. \url{https://doi.org/10.1080/08839514.2018.1448143}.

\leavevmode\vadjust pre{\hypertarget{ref-provost2013}{}}%
Provost, Foster, and Tom Fawcett. 2013. \emph{Data Science for Business:
What You Need to Know about Data Mining and Data-Analytic Thinking}.
O'Reilly Media.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
R Core Team. 2019. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: {R} Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-Ribeiro2016lime}{}}%
Ribeiro, Marco, Sameer Singh, and Carlos Guestrin. 2016. {``{``}Why
Should {I} Trust You?{''}: Explaining the Predictions of Any
Classifier.''} In \emph{Proceedings of the 2016 Conference of the North
{A}merican Chapter of the Association for Computational Linguistics:
Demonstrations}, 97--101. San Diego, California: Association for
Computational Linguistics. \url{https://doi.org/10.18653/v1/N16-3020}.

\leavevmode\vadjust pre{\hypertarget{ref-Romaszko2019}{}}%
Romaszko, Kamil, Magda Tatarynowicz, Mateusz Urbański, and Przemysław
Biecek. 2019. {``modelDown: Automated Website Generator with
Interpretable Documentation for Predictive Machine Learning Models.''}
\emph{Journal of Open Source Software} 4 (38): 1444.
\url{https://doi.org/10.21105/joss.01444}.

\leavevmode\vadjust pre{\hypertarget{ref-Ruspini1970}{}}%
Ruspini, Enrique H. 1970. {``Numerical Methods for Fuzzy Clustering.''}
\emph{Information Sciences} 2 (3): 319--50.
\url{https://doi.org/10.1016/S0020-0255(70)80056-1}.

\leavevmode\vadjust pre{\hypertarget{ref-2018aequitas}{}}%
Saleiro, Pedro, Benedict Kuester, Abby Stevens, Ari Anisfeld, Loren
Hinkson, Jesse London, and Rayid Ghani. 2018. {``Aequitas: A Bias and
Fairness Audit Toolkit.''} \emph{arXiv Preprint arXiv:1811.05577}.
\url{https://doi.org/10.48550/arXiv.1811.05577}.

\leavevmode\vadjust pre{\hypertarget{ref-Schmidberger2009}{}}%
Schmidberger, Markus, Martin Morgan, Dirk Eddelbuettel, Hao Yu, Luke
Tierney, and Ulrich Mansmann. 2009. {``State of the Art in Parallel
Computing with {R}.''} \emph{Journal of Statistical Software} 31 (1).
\url{https://doi.org/10.18637/jss.v031.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-Schratz2021}{}}%
Schratz, Patrick, Marc Becker, Michel Lang, and Alexander Brenning.
2021. {``{mlr3spatiotempcv}: Spatiotemporal Resampling Methods for
Machine Learning in {R},''} October.
\url{https://arxiv.org/abs/2110.12674}.

\leavevmode\vadjust pre{\hypertarget{ref-Silverman1986}{}}%
Silverman, Bernard W. 1986. \emph{Density Estimation for Statistics and
Data Analysis}. Vol. 26. CRC press.

\leavevmode\vadjust pre{\hypertarget{ref-Simon2007}{}}%
Simon, Richard. 2007. {``Resampling Strategies for Model Assessment and
Selection.''} In \emph{Fundamentals of Data Mining in Genomics and
Proteomics}, edited by Werner Dubitzky, Martin Granzow, and Daniel
Berrar, 173--86. {Boston, MA}: {Springer US}.
\url{https://doi.org/10.1007/978-0-387-47509-7_8}.

\leavevmode\vadjust pre{\hypertarget{ref-Snoek2012}{}}%
Snoek, Jasper, Hugo Larochelle, and Ryan P Adams. 2012. {``Practical
Bayesian Optimization of Machine Learning Algorithms.''} In
\emph{Advances in Neural Information Processing Systems}, edited by F.
Pereira, C. J. Burges, L. Bottou, and K. Q. Weinberger. Vol. 25.
\url{https://proceedings.neurips.cc/paper_files/paper/2012/file/05311655a15b75fab86956663e1819cd-Paper.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-Sonabend2021b}{}}%
Sonabend, Raphael Edward Benjamin. 2021. {``A Theoretical and
Methodological Framework for Machine Learning in Survival Analysis:
Enabling Transparent and Accessible Predictive Modelling on
Right-Censored Time-to-Event Data.''} PhD, University College London
(UCL). \url{https://discovery.ucl.ac.uk/id/eprint/10129352/}.

\leavevmode\vadjust pre{\hypertarget{ref-MLSA}{}}%
Sonabend, Raphael, and Andreas Bender. 2023. \emph{Machine Learning in
Survival Analysis}. \url{https://www.mlsabook.com}.

\leavevmode\vadjust pre{\hypertarget{ref-Sonabend2022}{}}%
Sonabend, Raphael, Andreas Bender, and Sebastian Vollmer. 2022.
{``Avoiding {C}-Hacking When Evaluating Survival Distribution
Predictions with Discrimination Measures.''} Edited by Zhiyong Lu.
\emph{Bioinformatics} 38 (17): 4178--84.
\url{https://doi.org/10.1093/bioinformatics/btac451}.

\leavevmode\vadjust pre{\hypertarget{ref-mlr3proba}{}}%
Sonabend, Raphael, Franz J Király, Andreas Bender, Bernd Bischl, and
Michel Lang. 2021. {``{mlr3proba}: An {R} Package for Machine Learning
in Survival Analysis.''} \emph{Bioinformatics}, February.
\url{https://doi.org/10.1093/bioinformatics/btab039}.

\leavevmode\vadjust pre{\hypertarget{ref-Sonabend2022a}{}}%
Sonabend, Raphael, Florian Pfisterer, Alan Mishler, Moritz Schauer,
Lukas Burk, Sumantrak Mukherjee, and Sebastian Vollmer. 2022.
{``Flexible Group Fairness Metrics for Survival Analysis.''} In
\emph{DSHealth 2022 Workshop on Applied Data Science for Healthcare at
KDD2022}. \url{https://arxiv.org/abs/2206.03256}.

\leavevmode\vadjust pre{\hypertarget{ref-Stein1987}{}}%
Stein, Michael. 1987. {``Large Sample Properties of Simulations Using
Latin Hypercube Sampling.''} \emph{Technometrics} 29 (2): 143--51.
\url{https://doi.org/10.2307/1269769}.

\leavevmode\vadjust pre{\hypertarget{ref-Strobl2008}{}}%
Strobl, Carolin, Anne-Laure Boulesteix, Thomas Kneib, Thomas Augustin,
and Achim Zeileis. 2008. {``Conditional Variable Importance for Random
Forests.''} \emph{{BMC} Bioinformatics} 9 (1).
\url{https://doi.org/10.1186/1471-2105-9-307}.

\leavevmode\vadjust pre{\hypertarget{ref-Trumbelj2013Shapley}{}}%
Štrumbelj, Erik, and Igor Kononenko. 2013. {``Explaining Prediction
Models and Individual Predictions with Feature Contributions.''}
\emph{Knowledge and Information Systems} 41 (3): 647--65.
\url{https://doi.org/10.1007/s10115-013-0679-x}.

\leavevmode\vadjust pre{\hypertarget{ref-Thornton2013}{}}%
Thornton, Chris, Frank Hutter, Holger H. Hoos, and Kevin Leyton-Brown.
2013. {``Auto-{WEKA}.''} In \emph{Proceedings of the 19th {ACM} {SIGKDD}
International Conference on Knowledge Discovery and Data Mining}. {ACM}.
\url{https://doi.org/10.1145/2487575.2487629}.

\leavevmode\vadjust pre{\hypertarget{ref-tsallis1996}{}}%
Tsallis, Constantino, and Daniel A Stariolo. 1996. {``Generalized
Simulated Annealing.''} \emph{Physica A: Statistical Mechanics and Its
Applications} 233 (1-2): 395--406.
\url{https://doi.org/10.1016/S0378-4371(96)00271-3}.

\leavevmode\vadjust pre{\hypertarget{ref-openml2013}{}}%
Vanschoren, Joaquin, Jan N. van Rijn, Bernd Bischl, and Luis Torgo.
2013. {``OpenML: Networked Science in Machine Learning.''} \emph{SIGKDD
Explorations} 15 (2): 49--60.
\url{https://doi.org/10.1145/2641190.2641198}.

\leavevmode\vadjust pre{\hypertarget{ref-Wachter2017}{}}%
Wachter, Sandra, Brent Mittelstadt, and Chris Russell. 2017.
{``Counterfactual Explanations Without Opening the Black Box: Automated
Decisions and the {GDPR}.''} \emph{{SSRN} Electronic Journal}.
\url{https://doi.org/10.2139/ssrn.3063289}.

\leavevmode\vadjust pre{\hypertarget{ref-Wachter2021}{}}%
---------. 2021. {``Why Fairness Cannot Be Automated: Bridging the Gap
Between EU Non-Discrimination Law and AI.''} \emph{Computer Law \&
Security Review} 41: 105567.
https://doi.org/\url{https://doi.org/10.1016/j.clsr.2021.105567}.

\leavevmode\vadjust pre{\hypertarget{ref-Watson2021}{}}%
Watson, David S, and Marvin N Wright. 2021. {``Testing Conditional
Independence in Supervised Learning Algorithms.''} \emph{Machine
Learning} 110 (8): 2107--29.
\url{https://doi.org/10.1007/s10994-021-06030-6}.

\leavevmode\vadjust pre{\hypertarget{ref-Wexler2019}{}}%
Wexler, James, Mahima Pushkarna, Tolga Bolukbasi, Martin Wattenberg,
Fernanda Viégas, and Jimbo Wilson. 2019. {``The What-If Tool:
Interactive Probing of Machine Learning Models.''} \emph{IEEE
Transactions on Visualization and Computer Graphics} 26 (1): 56--65.
\url{https://doi.org/10.1109/TVCG.2019.2934619}.

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2017R}{}}%
Wickham, Hadley, and Garrett Grolemund. 2017. \emph{{R} for Data
Science: Import, Tidy, Transform, Visualize, and Model Data}. 1st ed.
O'Reilly Media. \url{https://r4ds.had.co.nz/}.

\leavevmode\vadjust pre{\hypertarget{ref-williams_2006}{}}%
Williams, Christopher KI, and Carl Edward Rasmussen. 2006.
\emph{Gaussian Processes for Machine Learning}. Vol. 2. 3. MIT press
Cambridge, MA.

\leavevmode\vadjust pre{\hypertarget{ref-Wisniewski2022}{}}%
Wiśniewski, Jakub, and Przemysław Biecek. 2022. {``The {R} Journal:
Fairmodels: A Flexible Tool for Bias Detection, Visualization, and
Mitigation in Binary Classification Models.''} \emph{The {R} Journal}
14: 227--43. \url{https://doi.org/10.32614/RJ-2022-019}.

\leavevmode\vadjust pre{\hypertarget{ref-Wolpert1992}{}}%
Wolpert, David H. 1992. {``Stacked Generalization.''} \emph{Neural
Networks} 5 (2): 241--59.
\url{https://doi.org/10.1016/S0893-6080(05)80023-1}.

\leavevmode\vadjust pre{\hypertarget{ref-xiang2013}{}}%
Xiang, Yang, Sylvain Gubian, Brian Suomela, and Julia Hoeng. 2013.
{``Generalized Simulated Annealing for Global Optimization: The GenSA
Package.''} \emph{{R} Journal} 5 (1): 13.
\url{https://doi.org/10.32614/RJ-2013-002}.

\end{CSLReferences}