## Logging and Verbosity

We use the [lgr](https://github.com/s-fleck/lgr) package for logging and progress output.
Because _lgr_ comes with its own exhaustive vignette, we will just briefly give examples how you can change the most important settings related to logging in `r mlr_pkg("mlr3")`.

### Available logging levels

_lgr_ comes with certain numeric thresholds which correspond to verbosity levels of the logging.
For `r mlr_pkg("mlr3")` the default is set to 400 which corresponds to level "info".
The following ones are available:

```{r 04-technical-062}
library("lgr")
getOption("lgr.log_levels")
```

### Global Setting

_lgr_ comes with a global option called `"lgr.default_threshold"` which can be set via `options()`.
You can set a specific level in your `.Rprofile` which is then used for all packages that use the _lgr_ package.
This approach may not be desirable if you want to only change the logging level for `r mlr_pkg("mlr3")`.

### Changing mlr3 logging levels

To change the setting for `r mlr_pkg("mlr3")` only, you need to change the threshold of the `r mlr_pkg("mlr3")` logger like this:

```{r 04-technical-063, eval = FALSE}
lgr::get_logger("mlr3")$set_threshold("<level>")`
```

Remember that this change only applies to the current R session.
