---
aliases:
  - "/solutions.html"
---

# Solutions to exercises {#sec-solutions}

::: {.content-visible when-format="html"}

{{< include ../../common/_setup.qmd >}}
<!-- ENSURE ONLINE ONLY AND REFERENCE IN PREFACE -->

## Solutions to @sec-basics

1. Set the seed to `124` then train a classification tree model with `lrn("classif.rpart")` and default hyperparameters on 80% of the data in the predefined `"sonar"` task. Evaluate the model's performance with the classification error measure on the remaining data. Also think about why we need to set the seed in this example.

Set the seed, load the `"sonar"` task, then the classification tree with `predict_type = "prob"` (needed for exercise 3), and the required measure.

```{r solutions-001}
set.seed(124)
task = tsk("sonar")
learner = lrn("classif.rpart", predict_type = "prob")
measure = msr("classif.ce")
```

Use `partition()` to split the dataset then train the model.
We set the seed because `partition()` introduces an element of randomness when splitting the data.

```{r}
splits = partition(task, ratio = 0.8)
learner$train(task, splits$train)
```

Once the model is trained, generate the predictions on the test set and score them.

```{r solutions-002}
prediction = learner$predict(task, splits$test)
prediction$score(measure)
```

1. Calculate the true positive, false positive, true negative, and false negative rates of the predictions made by the model in exercise 1.

Using `$confusion`, generate a confusion matrix and extract the required statistics,

```{r solutions-003}
prediction$confusion
```

```{r solutions-004, echo=FALSE}
confusion = as.data.frame(prediction$confusion)

TP = confusion$Freq[1]
FP = confusion$Freq[2]
TN = confusion$Freq[4]
FN = confusion$Freq[3]
```

Since the rows represent predictions (response) and the columns represent the ground truth values, the TP, FP, TN, and FN rates are as follows:

- True Positive (TP) = `r TP`

- False Positive (FP) = `r FP`

- True Negative (TN) = `r TN`

- False Positive (FN) = `r FN`

1. Since in this case we want the model to predict the negative class more often, we will raise the threshold (note the `predict_type` for the learner must be `prob` for this to work).

```{r solutions-005}
# raise threshold from 0.5 default to 0.6
prediction$set_threshold(0.6)

prediction$confusion
```
One reason we might want the false positive rate to be lower than the false negative rate is if we felt it was worse for a positive prediction to be incorrect (meaning the true label was the negative label) than it was for a negative prediction to be incorrect (meaning the true label was the positive label).

## Solutions to @sec-performance

1. Apply the "bootstrap" resampling strategy on `tsk("mtcars")` and evaluate the performance of `lrn("classif.rpart")`.
Use 100 replicates and a sampling ratio of 80%.
Calculate the MSE for each iteration and visualize the result.
Finally, calculate the aggregated performance score.

```{r solutions-006}
set.seed(3)
task = tsk("mtcars")
learner = lrn("regr.rpart")
resampling = rsmp("bootstrap", repeats = 100, ratio = 0.8)

rr = resample(task, learner, resampling)
rr$score(msr("regr.mse"))
autoplot(rr)
# Alternatively: Histogram
autoplot(rr, type = "histogram")

rr$aggregate(msr("regr.mse"))
```


2. Use `tsk("spam")` and five-fold CV to benchmark Random forest (`lrn("classif.ranger")`), Logistic Regression (`lrn("classif.log_reg")`), and XGBoost (`lrn("classif.xgboost")`) with respect to AUC.
Which learner appears to do best? How confident are you in your conclusion?
How would you improve upon this?

```{r solutions-007, warning=FALSE}
set.seed(3)
design = benchmark_grid(
  tasks = tsk("spam"),
  learners = lrns(c("classif.ranger", "classif.log_reg", "classif.xgboost"),
                  predict_type = "prob"),
  resamplings = rsmp("cv", folds = 5)
)

bmr = benchmark(design)

mlr3viz::autoplot(bmr, measure = msr("classif.auc"))
```

This is only a small example for a benchmark workflow, but without tuning (see @sec-optimization), the results are naturally not suitable to make any broader statements about the superiority of either learner for this task.


3. A colleague claims to have achieved a 93.1% classification accuracy using `lrn("classif.rpart")` on `tsk("penguins_simple")`.
You want to reproduce their results and ask them about their resampling strategy.
They said they used a custom three-fold CV with folds assigned as `factor(task$row_ids %% 3)`.
See if you can reproduce their results.

```{r solutions-008}
task = tsk("penguins_simple")
rsmp_cv = rsmp("custom_cv")

rsmp_cv$instantiate(task = task, f = factor(task$row_ids %% 3))

rr = resample(
  task = task,
  learner = lrn("classif.rpart"),
  resampling = rsmp_cv
)

rr$aggregate(msr("classif.acc"))
```


## Solutions to @sec-optimization

1. Tune the `mtry`, `sample.fraction`, ` num.trees` hyperparameters of a random forest model (`lrn("regr.ranger")`) on the `"mtcars"` task.
Use a simple random search with 50 evaluations and select a suitable batch size.
Evaluate with a three-fold CV and the root mean squared error.

```{r solutions-009}
set.seed(4)
learner = lrn("regr.ranger",
  mtry.ratio      = to_tune(0, 1),
  sample.fraction = to_tune(1e-1, 1),
  num.trees       = to_tune(1, 2000)
)

instance = ti(
  task = tsk("mtcars"),
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measures = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

tuner = tnr("random_search", batch_size = 10)

tuner$optimize(instance)
```

2. Evaluate the performance of the model created in Question 1 with nested resampling.
Use a holdout validation for the inner resampling and a three-fold CV for the outer resampling.
Print the unbiased performance estimate of the model.

```{r solutions-010}
set.seed(4)
learner = lrn("regr.ranger",
  mtry.ratio      = to_tune(0, 1),
  sample.fraction = to_tune(1e-1, 1),
  num.trees       = to_tune(1, 2000)
)

at = auto_tuner(
  tuner = tnr("random_search", batch_size = 10),
  learner = learner,
  resampling = rsmp("holdout"),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

task = tsk("mtcars")
outer_resampling = rsmp("cv", folds = 3)
rr = resample(task, at, outer_resampling, store_models = TRUE)

rr$aggregate()
```

1. Tune and benchmark an XGBoost model against a logistic regression model on the `"spam"` task and determine which has the best Brier score.
Use mlr3tuningspaces and nested resampling.

```{r solutions-011}
library(mlr3tuningspaces)

lrn_xgboost = lts(lrn("classif.xgboost", predict_type = "prob"))

at_xgboost = auto_tuner(
  tuner = tnr("random_search", batch_size = 1),
  learner = lrn_xgboost,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.bbrier"),
  term_evals = 2,
)

lrn_logreg = lrn("classif.log_reg", predict_type = "prob")

at_logreg = auto_tuner(
  tuner = tnr("random_search", batch_size = 1),
  learner = lrn_logreg,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.bbrier"),
  term_evals = 2,
)

task = tsk("spam")
outer_resampling = rsmp("cv", folds = 3)

design = benchmark_grid(
  tasks = task,
  learners = list(at_xgboost, at_logreg),
  resamplings = outer_resampling
)

bmr = benchmark(design, store_models = TRUE)

bmr
```



## Solutions to @sec-optimization-advanced

1. We first construct the objective function and optimization instance:

```{r}
library(bbotk)
library(mlr3mbo)

rastrigin = function(xdt) {
  D = ncol(xdt)
  y = 10 * D + rowSums(xdt^2 - (10 * cos(2 * pi * xdt)))
  data.table(y = y)
}

objective = ObjectiveRFunDt$new(
  fun = rastrigin,
  domain = ps(x1 = p_dbl(lower = -5.12, upper = 5.12),
    x2 = p_dbl(lower = -5.12, upper = 5.12)),
  codomain = ps(y = p_dbl(tags = "minimize")),
  id = "rastrigin2D")

instance = OptimInstanceSingleCrit$new(
  objective = objective,
  terminator = trm("evals", n_evals = 40))
```

Based on the different surrogate models, we can construct two optimizers:

```{r}
library(mlr3mbo)

surrogate_gp = srlrn(lrn("regr.km", covtype = "matern5_2",
  optim.method = "BFGS", control = list(trace = FALSE)))

surrogate_rf = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))

acq_function = acqf("cb", lambda = 1)

acq_optimizer = acqo(opt("nloptr", algorithm = "NLOPT_GN_ORIG_DIRECT"),
  terminator = trm("stagnation", iters = 100, threshold = 1e-5))

optimizer_gp = opt("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate_gp,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)

optimizer_rf = opt("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate_rf,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)
```

We then evaluate the given initial design on the instance and optimize it with the first BO algorithm using a Gaussian Process as surrogate model:

```{r, output=FALSE}
initial_design = data.table(
  x1 = c(-3.95, 1.16, 3.72, -1.39, -0.11, 5.00, -2.67, 2.44),
  x2 = c(1.18, -3.93, 3.74, -1.37, 5.02, -0.09, -2.65, 2.46))

instance$eval_batch(initial_design)

optimizer_gp$optimize(instance)

gp_data = instance$archive$data
gp_data[, y_min := cummin(y)]
gp_data[, nr_eval := seq_len(.N)]
gp_data[, surrogate := "Gaussian Process"]
```

Afterwards, we clear the instance, evaluate the initial design again and optimize the instance with the second BO algorithm using a random forest as surrogate model:

```{r, output=FALSE}
instance$archive$clear()

instance$eval_batch(initial_design)

optimizer_rf$optimize(instance)

rf_data = instance$archive$data
rf_data[, y_min := cummin(y)]
rf_data[, nr_eval := seq_len(.N)]
rf_data[, surrogate := "Random forest"]
```

We collect all data and visualize the anytime performance:

```{r}
library(ggplot2)
library(viridisLite)
all_data = rbind(gp_data, rf_data)
ggplot(aes(x = nr_eval, y = y_min, colour = surrogate), data = all_data) +
  geom_step() +
  scale_colour_manual(values = viridis(2, end = 0.8)) +
  labs(y = "Best Observed Function Value", x = "Number of Function Evaluations",
       colour = "Surrogate Model") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

2. We first construct the non-parallelized objective function and the optimization instance:

```{r}
schaffer1 = function(xss) {
  evaluations = lapply(xss, FUN = function(xs) {
    Sys.sleep(5)
    list(y1 = xs$x, y2 = (xs$x - 2)^2)
  })
  rbindlist(evaluations)
}

objective = ObjectiveRFunMany$new(
  fun = schaffer1,
  domain = ps(x = p_dbl(lower = -10, upper = 10)),
  codomain = ps(y1 = p_dbl(tags = "minimize"), y2 = p_dbl(tags = "minimize")),
  id = "schaffer1")

instance = OptimInstanceMultiCrit$new(
  objective = objective,
  terminator = trm("run_time", secs = 60))
```

Using the surrogate, acquisition function and acquisition function optimizer that are provided, we can proceed to optimize the instance via ParEGO:

```{r}
surrogate = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))

acq_function = acqf("ei")

acq_optimizer = acqo(opt("random_search", batch_size = 100),
  terminator = trm("evals", n_evals = 100))

optimizer = opt("mbo",
  loop_function = bayesopt_parego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer,
  args = list(q = 4))
```

```{r, output=FALSE}
optimizer$optimize(instance)
```

We observe that 12 points were evaluated in total (which makes sense as the objective function evaluation is not yet parallelized and the overhead of each function evaluation is given by 5 seconds).
While the points are appropriately evaluated in batches of size `q = 4` (with the initial design automatically constructed as the first batch), we do not experience any acceleration of the optimization process unless the function evaluation is explicitly parallelized.

```{r}
nrow(instance$archive$data)
instance$archive$data[, c("x", "timestamp", "batch_nr")]
```

We now parallelize the evaluation of the objective function and proceed to optimize the instance again:

```{r}
library(future)
library(future.apply)
plan("multisession", workers = 4)

schaffer1_parallel = function(xss) {
  evaluations = future_lapply(xss, FUN = function(xs) {
    Sys.sleep(5)
    list(y1 = xs$x, y2 = (xs$x - 2)^2)
  })
  rbindlist(evaluations)
}

objective_parallel = ObjectiveRFunMany$new(
  fun = schaffer1_parallel,
  domain = ps(x = p_dbl(lower = -10, upper = 10)),
  codomain = ps(y1 = p_dbl(tags = "minimize"), y2 = p_dbl(tags = "minimize")),
  id = "schaffer1_parallel")

instance_parallel = OptimInstanceMultiCrit$new(
  objective = objective_parallel,
  terminator = trm("run_time", secs = 60))
```

```{r, output=FALSE}
optimizer$optimize(instance_parallel)
```

By parallelizing the evaluation of the objective function we used our compute resources much more efficiently and evaluated many more points:

```{r}
nrow(instance_parallel$archive$data)
instance_parallel$archive$data[, c("x", "timestamp", "batch_nr")]
```




## Solutions to @sec-feature-selection

1. Calculate a correlation filter on the `"mtcars"` task.

```{r solutions-012}
library(mlr3verse)
filter = flt("correlation")

task = tsk("mtcars")
filter$calculate(task)

as.data.table(filter)
```

2. Use the filter from the first exercise to select the five best features in the `mtcars` dataset.

```{r solutions-013}
keep = names(head(filter$scores, 5))
task$select(keep)
task$feature_names
```

3. Apply a backward selection to `tsk("penguins")` with `lrn("classif.rpart")` and holdout resampling by the measure classification accuracy. Compare the results with those in @sec-fs-wrapper-example.

```{r solutions-014}
library(mlr3fselect)

instance = fselect(
  fselector = fs("sequential", strategy = "sbs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = rsmp("holdout"),
  measure = msr("classif.acc")
)
as.data.table(instance$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
instance$result_feature_set
```

Answer the following questions:

  a. Do the selected features differ?

Yes, the backward selection selects more features.

  b. Which feature selection method achieves a higher classification accuracy?

In this example, the backwards example performs slightly better, but this depends heavily on the random seed and could look different in another run.

  c. Are the accuracy values in b) directly comparable? If not, what has to be changed to make them comparable?

No, they are not comparable because the holdout sampling called with `rsmp("holdout")` creates a different holdout set for the two runs. A fair comparison would create a single resampling instance and use it for both feature selections (see @sec-performance for details):

```{r solutions-015}
resampling = rsmp("holdout")
resampling$instantiate(tsk("penguins"))

instance_sfs = fselect(
  fselector = fs("sequential", strategy = "sfs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = resampling,
  measure = msr("classif.acc")
)
instance_sbs = fselect(
  fselector = fs("sequential", strategy = "sbs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = resampling,
  measure = msr("classif.acc")
)
as.data.table(instance_sfs$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
as.data.table(instance_sbs$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
```

Alternatively, one could automate the feature selection and perform a benchmark between the two wrapped learners.

4. Automate the feature selection as in @sec-autofselect with `tsk("spam")` and `lrn("classif.log_reg")`.

```{r solutions-016, warning=FALSE}
library(mlr3fselect)
library(mlr3learners)

at = auto_fselector(
  fselector = fs("random_search"),
  learner = lrn("classif.log_reg"),
  resampling = rsmp("holdout"),
  measure = msr("classif.acc"),
  terminator = trm("evals", n_evals = 50)
)

design = benchmark_grid(
  task = tsk("spam"),
  learner = list(at, lrn("classif.log_reg")),
  resampling = rsmp("cv", folds = 3)
)

bmr = benchmark(design)

aggr = bmr$aggregate(msrs(c("classif.acc", "time_train")))
as.data.table(aggr)[, .(learner_id, classif.acc, time_train)]
```

## Solutions to @sec-pipelines

1. Concatenate the named PipeOps using `%>>%`.
  To get a `r ref("Learner")` object, use `r ref("as_learner()")`
```{r pipelines-001}
library(mlr3pipelines)
library(mlr3learners)

graph = po("imputeoor") %>>% po("scale") %>>% lrn("classif.log_reg")
graph_learner = as_learner(graph)
```

1. After training, the underlying `lrn("classif.log_reg")` can be accessed through the `$base_learner()` method.
  Alternatively, the learner can be accessed explicitly using `po("learner")`.
```{r pipelines-002}
graph_learner$train(tsk("pima"))

# access the learner through the $base_learner() method
model = graph_learner$base_learner()$model
coef(model)

# access the learner explicitly through the PipeOp
pipeop = graph_learner$graph_model$pipeops$classif.log_reg
model = pipeop$learner_model$model
coef(model)
```

3. Set the `$keep_results` flag of the Graph to `TRUE` to keep the results of the individual PipeOps.
  Afterwards, the input of the `lrn("classif.log_reg")` can be accessed through the `$.result` field of its predecessor, the `po("scale")`.
  Note that the `$.result` is a `list`, we want to access its only element, named `$output`.
```{r pipelines-003}
graph_learner$graph$keep_results = TRUE
graph_learner$train(tsk("pima"))

# access the input of the learner
scale_result = graph_learner$graph_model$pipeops$scale$.result

scale_output_task = scale_result$output

age_column = scale_output_task$data()$age

# check if the age column is standardized:
# 1. does it have mean 0? -- almost, up to numerical precision!
mean(age_column)
# 2. does it have standard deviation 1? -- yes!
sd(age_column)
```

## Solutions to @sec-pipelines-nonseq

1. To use `po("select")` to *remove*, instead of *keep*, a feature based on a pattern, use `r ref("selector_invert")` together with `r ref("selector_grep")`.
  To remove the "`R`" class columns in @sec-pipelines-stack, the following `po("select")` could be used:

```{r pipelines-004}
po("select", selector = selector_invert(selector_grep("\\.R")))
```
which would have the benefit that it would keep the columns pertaining to all other classes, even if the `"sonar"` task had more target classes.

2. A solution that does not need to specify the target classes at all is to use a custom `r ref("Selector")`, as was shown in @sec-pipelines-bagging:

```{r pipelines-005}
selector_remove_one_prob_column = function(task) {
  class_removing = task$class_names[[1]]
  selector_use = selector_invert(selector_grep(paste0("\\.", class_removing)))

selector_use(task)
}
```
Using this selector in @sec-pipelines-stack, one could use the resulting stacking learner on any classification task with arbitrary target classes.

3. As the first hint states, two `po("imputelearner")` objects are necessary: one to impute missing values in factor columns using a classification learner, and another to impute missing values in numeric columns using a regression learner.
  Additionally, `ppl("robustify")` is used along with the `r ref("ranger::ranger")`-based learners inside `po("imputelearner")` because the data passed to the imputation learners still contains missing values, which `ranger::ranger` cannot handle.

```{r pipelines-006}
gr_impute_factors = po("imputelearner", id = "impute_factors",
  learner = ppl("robustify", learner = lrn("classif.ranger")) %>>%
    lrn("classif.ranger"),
  affect_columns = selector_type("factor")
)
gr_impute_numerics = po("imputelearner", id = "impute_numerics",
  learner = ppl("robustify", learner = lrn("regr.ranger")) %>>%
    lrn("regr.ranger"),
  affect_columns = selector_type(c("numeric", "integer"))
)

gr_impute = gr_impute_numerics %>>% gr_impute_factors

imputed = gr_impute$train(tsk("penguins"))[[1]]

# e.g. see how row 4 was imputed
# original:
tsk("penguins")$data(rows = 4)
# imputed:
imputed$data(rows = 4)
```

## Solutions for @sec-preprocessing

We will consider a similar prediction problem as throughout this section, using the King County Housing data instead (available with `tsk("kc_housing")`).
To evaluate the models, we again use 10-fold cv and the mean absolute error.
The learner we want to use is a elastic-net regression by `lrn("regr.glmnet")`.
For now we will ignore the `date` column and simply remove it:

```{r, warning=FALSE, message=FALSE}
kc_housing = tsk("kc_housing")
kc_housing$select(setdiff(kc_housing$feature_names, "date"))
```

1. Have a look at the features, are there any features which might be problematic? If so, change or remove them.

```{r, warning=FALSE, message=FALSE}
summary(kc_housing)
```


`zipcode` should not really be interpreted as a numeric value, so we cast it to a factor.
We could argue to remove `lat` and `long` as handling them as linear effects is not necessarily a suitable, but we will keep them since `glmnet` performs internal feature selection anyways.

```{r, warning=FALSE, message=FALSE}
zipencode = po("mutate", mutation = list(zipcode = ~ as.factor(zipcode)), id = "zipencode")
```

2. Check the dataset and learner properties to understand which preprocessing steps you need to do.

```{r, warning=FALSE, message=FALSE}
print(kc_housing)
kc_housing$missings()
```

The data has missings and a categorical feature (since we are encoding the zipcode as a factor).

```{r, warning=FALSE, message=FALSE}
glmnet = lrn("regr.glmnet")
glmnet$properties
glmnet$feature_types
```

`glmnet` does not support factors or missing values. So our pipeline needs to handle both.

3. Build a suitable pipeline that allows glmnet to be trained on the dataset.

```{r, warning=FALSE, message=FALSE}
glmnet_preproc = GraphLearner$new(
  zipencode %>>%
  po("fixfactors") %>>%
    po("encodeimpact") %>>%
    list(
    po("missind",
        type = "integer",
        affect_columns = selector_type("integer")
    ),
    po("imputehist",
        affect_columns = selector_type("integer")
    )) %>>%
    po("featureunion") %>>%
    po("imputeoor",
        affect_columns = selector_type("factor")
    ) %>>%
    glmnet,
  id = "regr.glmnet_preproc")

log_glmnet_preproc = ppl("targettrafo", graph = glmnet_preproc)
log_glmnet_preproc$param_set$values$targetmutate.trafo = function(x) log(x)
log_glmnet_preproc$param_set$values$targetmutate.inverter = function(x) list(response = exp(x$response))
log_glmnet_preproc = GraphLearner$new(log_glmnet_preproc, id = "regr.log_glmnet_preproc")
```

First we fix the factor levels to ensure that all 70 zipcodes are fixed.
We can consider 70 levels high cardinality, so we use impact encoding.
We use the same imputation strategy as in @sec-preprocessing.
Since the target is highly skewed, we also apply a log-transformation of the target.


4. As a comparison, apply `pipeline_robustify` to glmnet and compare the results with your pipeline.

```{r, warning=FALSE, message=FALSE}
glmnet_robustify = GraphLearner$new(
    zipencode %>>%
    mlr3pipelines::pipeline_robustify() %>>%
      glmnet,
    id = "regr.glmnet_robustify"
)

log_glmnet_robustify = ppl("targettrafo", graph = glmnet_robustify)
log_glmnet_robustify$param_set$values$targetmutate.trafo = function(x) log(x)
log_glmnet_robustify$param_set$values$targetmutate.inverter = function(x) list(response = exp(x$response))
log_glmnet_robustify = GraphLearner$new(log_glmnet_robustify, id = "regr.log_glmnet_robustify")

learners = list(
  lrn("regr.featureless", robust = TRUE),
  glmnet_preproc,
  log_glmnet_preproc,
  glmnet_robustify,
  log_glmnet_robustify
)

set.seed(123L)
cv10 = rsmp("cv")
cv10$instantiate(kc_housing)

design = benchmark_grid(kc_housing, learners = learners, cv10)
bmr = benchmark(design)
bmr$aggregate(measure = msr("regr.mae"))[, .(learner_id, regr.mae)]
```

The log-transformed `glmnet` with impact encoding results in the best model, although only by a very small margin.


5. Consider the `date` feature: How can you extract information from this feature that `glmnet` can use? Check how much this improves your pipeline.
Also consider the spatial nature of the dataset: Can you extract an additional feature from the lat/long coordinates? (Hint: Downtown Seattle has lat/long coordinates `47.605`/`-122.334`).

```{r, warning=FALSE, message=FALSE}
extractors = po("mutate", mutation = list(
  date = ~ as.numeric(date),
  distance_downtown = ~ sqrt((lat - 47.605)^2 + (long  + 122.334)^2)))

kc_housing_full = extractors$train(list(tsk("kc_housing")))[[1]]
kc_housing_full$id = "kc_housing_feat_extr"


design_ext = benchmark_grid(kc_housing_full, learners = learners, cv10)
bmr_ext = benchmark(design_ext)
bmr$combine(bmr_ext)
bmr$aggregate(measure = msr("regr.mae"))[, .(learner_id, task_id, regr.mae)]
```

We simply convert the `date` feature into a numeric timestamp so that `glmnet` can handle the feature.
We create one additional feature as the distance to downtown Seattle.
This improves the average error of our model by a further 1600$.

## Solutions to @sec-technical

1. Consider the following example where you resample a learner (debug learner, sleeps for 3 seconds during train) on 4 workers using the multisession backend:
```{r technical-050, eval = FALSE}
task = tsk("penguins")
learner = lrn("classif.debug", sleep_train = function() 3)
resampling = rsmp("cv", folds = 6)

future::plan("multisession", workers = 4)
resample(task, learner, resampling)
```

i. Assuming that the learner would actually calculate something and not just sleep: Would all CPUs be busy?
ii. Prove your point by measuring the elapsed time, e.g., using `r ref("system.time()")`.
iii. What would you change in the setup and why?

Not all CPUs would be utilized in the example.
All 4 of them are occupied for the first 4 iterations of the cross validation.
The 5th iteration, however, only runs in parallel to the 6th fold, leaving 2 cores idle.
This is supported by the elapsed time of roughly 6 seconds for 6 jobs compared to also roughly 6 seconds for 8 jobs:

```{r solutions-022, eval = FALSE}
task = tsk("penguins")
learner = lrn("classif.debug", sleep_train = function() 3)

future::plan("multisession", workers = 4)

resampling = rsmp("cv", folds = 6)
system.time(resample(task, learner, resampling))

resampling = rsmp("cv", folds = 8)
system.time(resample(task, learner, resampling))
```

If possible, the number of resampling iterations should be an integer multiple of the number of workers.
Therefore, a simple adaptation either increases the number of folds for improved accuracy of the error estimate or reduces the number of folds for improved runtime.

2. Create a new custom classification measure (either using methods demonstrated in @sec-extending or with `msr("classif.costs")` which scores predictions using the mean over the following classification costs:

* If the learner predicted label "A" and the truth is "A", assign score 0
* If the learner predicted label "B" and the truth is "B", assign score 0
* If the learner predicted label "A" and the truth is "B", assign score 1
* If the learner predicted label "B" and the truth is "A", assign score 10

The rules can easily be translated to R code where we expect `truth` and `prediction` to be factor vectors of the same length with levels `"A"` and `"B"`:

```{r solutions-023}
costsens = function(truth, prediction) {
    score = numeric(length(truth))
    score[truth == "A" & prediction == "B"] = 10
    score[truth == "B" & prediction == "A"] = 1

    mean(score)
}
```

This function can be embedded in the `Measure` class accordingly.

```{r solutions-024}
MeasureCustom = R6::R6Class("MeasureCustom",
  inherit = mlr3::MeasureClassif, # classification measure
  public = list(
    initialize = function() { # initialize class
      super$initialize(
        id = "custom", # unique ID
        packages = character(), # no dependencies
        properties = character(), # no special properties
        predict_type = "response", # measures response prediction
        range = c(0, Inf), # results in values between (0, 1)
        minimize = TRUE # smaller values are better
      )
    }
  ),

  private = list(
    .score = function(prediction, ...) { # define score as private method
      # define loss
      costsens = function(truth, prediction) {
        score = numeric(length(truth))
        score[truth == "A" & prediction == "B"] = 10
        score[truth == "B" & prediction == "A"] = 1

        mean(score)
      }

      # call loss function
      costsens(prediction$truth, prediction$response)
    }
  )
)
```

An alternative (as pointed to by the hint) can be constructed by first translating the rules to a matrix of misclassification costs, and then feeding this matrix to the constructor of `msr("classif.costs")`:

```{r solutions-025}
# truth in columns, prediction in rows
C = matrix(c(0, 10, 1, 0), nrow = 2)
rownames(C) = colnames(C) = c("A", "B")
C

msr("classif.costs", costs = C)
```


## Solutions to @sec-large-benchmarking

{{< include ./solutions_large-scale_benchmarking.qmd >}}

## Solutions to @sec-interpretation

1. Prepare a `mlr3` regression task for `fifa` data. Select only variables describing the age and skills of footballers. Train any predictive model for this task, e.g. `lrn("regr.ranger")`.

```{r solutions-044, warning=FALSE, message=FALSE}
library(DALEX)
library(ggplot2)
data("fifa", package = "DALEX")
old_theme = set_theme_dalex("ema")

library(mlr3)
library(mlr3learners)
set.seed(1)

fifa20 = fifa[,5:42]
task_fifa = as_task_regr(fifa20, target = "value_eur", id = "fifa20")

learner = lrn("regr.ranger")
learner$train(task_fifa)
learner$model
```

2. Use the permutation importance method to calculate variable importance ranking. Which variable is the most important? Is it surprising?

**With `iml`**

```{r solutions-045, warning=FALSE, message=FALSE}
library(iml)
model = Predictor$new(learner,
                data = fifa20,
                y = fifa$value_eur)

effect = FeatureImp$new(model,
                loss = "rmse")
effect$plot()
```

**With `DALEX`**

```{r solutions-046, warning=FALSE, message=FALSE}
library(DALEX)
ranger_exp = DALEX::explain(learner,
  data = fifa20,
  y = fifa$value_eur,
  label = "Fifa 2020",
  verbose = FALSE)

ranger_effect = model_parts(ranger_exp, B = 5)
head(ranger_effect)
plot(ranger_effect)
```

3. Use the Partial Dependence profile to draw the global behavior of the model for this variable. Is it aligned with your expectations?

**With `iml`**

```{r solutions-047, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

effect = FeatureEffects$new(model)
plot(effect, features = num_features)
```

**With `DALEX`**

```{r solutions-048, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

ranger_profiles = model_profile(ranger_exp, variables = num_features)
plot(ranger_profiles)
```

4 Choose one of the football players. You can choose some well-known striker (e.g. Robert Lewandowski) or a well-known goalkeeper (e.g. Manuel Neuer). The following tasks are worth repeating for several different choices.

```{r solutions-049, warning=FALSE, message=FALSE}
player_1 = fifa["R. Lewandowski", 5:42]
```

5. For the selected footballer, calculate and plot the Shapley values. Which variable is locally the most important and has the strongest influence on the valuation of the footballer?

**With `iml`**

```{r solutions-050, warning=FALSE, message=FALSE}
shapley = Shapley$new(model, x.interest = player_1)
plot(shapley)
```

**With `DALEX`**

```{r solutions-051, warning=FALSE, message=FALSE}
ranger_shap = predict_parts(ranger_exp,
             new_observation = player_1,
             type = "shap", B = 1)
plot(ranger_shap, show_boxplots = FALSE)
```

6. For the selected footballer, calculate the Ceteris Paribus / Individual Conditional Expectation profiles to draw the local behavior of the model for this variable. Is it different from the global behavior?

**With `DALEX`**

```{r solutions-052, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

ranger_ceteris = predict_profile(ranger_exp, player_1)
plot(ranger_ceteris, variables = num_features) +
  ggtitle("Ceteris paribus for R. Lewandowski", " ")
```

## Solutions to @sec-special

1. Run a benchmark experiment on `tsk("german_credit")` with `lrn("classif.featureless")`, `lrn("classif.log_reg")`, and `lrn("classif.ranger")`. Tune the prediction thresholds of all learners by encapsulating them in a `po("learner_cv")` (with two-fold CV), followed by a `po("tunethreshold")`. Use `msr("classif.costs", costs = costs)`, where the `costs` matrix is as follows: true positive is `-10`, true negative is `-1`, false positive is `2`, and false negative is `3`. Use this measure in `po("tunethreshold")` and when evaluating your benchmark experiment.

```{r, message=FALSE, warning=FALSE}
set.seed(1)
# Load task and learners
tsk_german = tsk("german_credit")
learners = lrns(c("classif.featureless", "classif.log_reg",
  "classif.ranger"), predict_type = "prob")

# Create costs matrix
costs = matrix(c(-10, 3, 2, -1), nrow = 2,
  dimnames = list("Predicted Credit" = c("good", "bad"),
                    Truth = c("good", "bad")))
costs
# Create measure
meas_costs = msr("classif.costs", costs = costs)

# Create a function to wrap learners in internal cross-validation
#  to tune the threshold
pipe = function(l) {
  po("learner_cv", l, resampling.folds = 2) %>>%
    po("tunethreshold", measure = meas_costs)
}

# Benchmark
learners = lapply(learners, pipe)
design = benchmark_grid(tsk_german, learners, rsmp("holdout"))
bmr = benchmark(design)$aggregate(meas_costs)
bmr[, .(learner_id, classif.costs)]
```

Based on these results, the logistic regression performs the best with the greatest increase to costs, however the difference is only marginal compared to the other learners.

2. Train and test a survival forest using `lrn("surv.rfsrc")` (from `mlr3extralearners`). Run this experiment using `tsk("rats")` and `partition()`. Evaluate your model with the RCLL measure.

```{r}
# Get learners
library(mlr3extralearners)
# Get survival models
library(mlr3proba)
set.seed(1)
# Use partition to split data and test our model
tsk_rats = tsk("rats")
splits = partition(tsk_rats)
learner = lrn("surv.rfsrc")
prediction = learner$train(tsk_rats, splits$train)$predict(tsk_rats, splits$test)
prediction$score(msr("surv.rcll"))
```

The right-censored logloss provides a measure of predictive accuracy, but it is quite hard to interpret it without comparison to another model.
To yield a more informative value, we could either compute the RCLL for an uninformed baseline like the Kaplan-Meier estimator, or we could use the `ERV` (explained residual variation) parameter in the measure, which returns the RCLL as a percentage increase in performance compared to an uninformed baseline (in this case the Kaplan-Meier estimator):

```{r}
lrn("surv.kaplan")$
  train(tsk_rats, splits$train)$
  predict(tsk_rats, splits$test)$
  score(msr("surv.rcll"))

prediction$score(msr("surv.rcll", ERV = TRUE),
  task = tsk_rats, train_set = splits$train)
```

Now we can see that our model is only marginally better than the Kaplan-Meier baseline (a 2% performance increase).

3. Estimate the density of the "precip" task from the `mlr3proba` package using `lrn("dens.hist")`, evaluate your estimation with the logloss measure. As a stretch goal, look into the documentation of `distr6` to learn how to analyse your estimated distribution further.

```{r}
# Get density models
library(mlr3proba)
set.seed(1)
# Run experiment
tsk_precip = tsk("precip")
learner = lrn("dens.hist")
prediction = learner$train(tsk_precip)$predict(tsk_precip)
prediction
prediction$score(msr("dens.logloss"))
```

As before the logloss is not too informative by itself but as the Histogram is itself a baseline, we can use this value for comparison to more sophisticated models. To learn more about our predicted distribution, we could use `distr6` to summarise the distribution and to compute values such as the pdf and cdf:

```{r}
prediction$distr$summary()
# pdf evaluated at `50`
prediction$distr$pdf(50)
```

4. Run a benchmark clustering experiment on the "wine" dataset without a label column. Compare the performance of k-means learner with `k` equal to `2`, `3` and `4` using the silhouette measure and the insample resampling technique. What value of `k` would you choose based on the silhouette scores?

```{r, messages=FALSE, warnings=FALSE}
set.seed(1)
# Load clustering models and tasks
library(mlr3cluster)
# Create the clustering dataset by extracting only the features from the
#  wine task
tsk_wine = tsk("wine")
tsk_wine = as_task_clust(tsk_wine$data(cols = tsk_wine$feature_names))
# Create learners and ensure they have unique IDs
learners = c(
  lrn("clust.kmeans", centers = 2, id = "K=2"),
  lrn("clust.kmeans", centers = 3, id = "K=3"),
  lrn("clust.kmeans", centers = 4, id = "K=4")
)
# Benchmark
meas = msr("clust.silhouette")
design = benchmark_grid(tsk_wine, learners, rsmp("insample"))
benchmark(design)$aggregate(meas)[, .(learner_id, clust.silhouette)]
```

We can see that we get the silhouette closest to `1` with `K=2` so we might use this value for future experiments.

## Solutions to @sec-fairness

1. Load the `adult_train` task and try to build a first model.
  Train a simple model and evaluate it on the `adult_test` task that is also available with `mlr3fairness`.

For now we simply load the data and look at the data.

```{r}
library(mlr3)
library(mlr3fairness)
set.seed(8)

tsk_adult_train = tsk("adult_train")
tsk_adult_test = tsk("adult_test")
tsk_adult_train
```

We can now train a simple model, e.g., a decision tree and evaluate for accuracy.

```{r}
learner = lrn("classif.rpart")
learner$train(tsk_adult_train)
prediction = learner$predict(tsk_adult_test)
prediction$score()
```


2. Assume our goal is to achieve parity in *false omission rates*.
  Construct a fairness metric that encodes this and againg evaluate your model.
  Construct a fairness metric that encodes this and evaluate your model.
  In order to get a deeper understanding, look at the `groupwise_metrics` function to obtain performance in each group.

The metric is available via the key `"fairness.fomr"`.
Note, that evaluating our prediction now requires that we also provide the task.

```{r}
msr_1 = msr("fairness.fomr")
prediction$score(msr_1, tsk_adult_test)
```

The `groupwise_metrics` function creates a metric for each group specified in the `pta` column role:

```{r}
tsk_adult_test$col_roles$pta
```

```{r}
msr_2 = groupwise_metrics(base_measure = msr("classif.fomr"), task = tsk_adult_test)
```

We can then use this metric to evaluate our model again.
This gives us the false omission rates for male and female individuals separately.

```{r}
prediction$score(msr_2, tsk_adult_test)
```

1. Improve your model by employing pipelines that use pre- or post-processing methods for fairness.
  Evaluate your model along the two metrics and visualize the results.
  Compare the different models using an appropriate visualization.

First we can again construct the learners above.
```{r}
library(mlr3pipelines)
lrn_1 = po("reweighing_wts") %>>% lrn("classif.rpart")
lrn_2 = po("learner_cv", lrn("classif.rpart")) %>>%
  po("EOd")
```

And run the benchmark again. Note, that we use three-fold CV this time for comparison.

```{r}
learners = list(learner, lrn_1, lrn_2)
design = benchmark_grid(tsk_adult_train, learners, rsmp("cv", folds = 3L))
bmr = benchmark(design)
bmr$aggregate(msrs(c("classif.acc", "fairness.fomr")))
```

We can now again visualize the result.

```{r}
library(ggplot2)
fairness_accuracy_tradeoff(bmr, msr("fairness.fomr")) +
  scale_color_viridis_d("Learner") +
  theme_minimal()
```

1. Add `"race"` as a second sensitive attribute to your dataset.
  Add the information to your task and evaluate the initial model again. What changes?
  Again study the `groupwise_metrics`.

This can be achieved by adding "race" to the `"pta"` col_role.

```{r}
tsk_adult_train$set_col_roles("race", add_to = "pta")
tsk_adult_train
```

```{r}
tsk_adult_test$set_col_roles("race", add_to = "pta")
prediction$score(msr_1, tsk_adult_test)
```

If we now evaluate for the intersection, we obtain a large deviation from `0`.
Note, that the metric by default computes the maximum discrepancy between all metrics for the non-binary case.

If we now get the `groupwise_metrics`, we will get a metric for the intersection of each group.

```{r}
msr_3 = groupwise_metrics(msr("classif.fomr"),  tsk_adult_train)
unname(sapply(msr_3, function(x) x$id))
```

```{r}
prediction$score(msr_3, tsk_adult_test)
```

And we can see, that the reason might be, that the false omission rate for female Amer-Indian-Eskimo is at `1.0`!
We can investigate this further by looking at actual counts:

```{r}
table(tsk_adult_test$data(cols = c("race", "sex", "target")))
```

One of the reasons might be that there are only 3 individuals in the ">50k" category!
This is an often encountered problem, as error metrics have a large variance when samples are small.
Note, that the pre- and post-processing methods in general do not all support multiple protected attributes.

:::
