## Benchmarking {#benchmarking}

Comparing the performance of different learners on multiple tasks and/or different resampling schemes is a recurrent task.
This operation is usually referred to as "benchmarking" in the field of machine-learning.
`r mlr_pkg("mlr3")` offers the `r ref("benchmark()")` function for convenience.

### Design Creation {#bm-design}

In _mlr3_ we require you to supply a "design" of your benchmark experiment.
By "design" we essentially mean the matrix of settings you want to execute.
A "design" consists of `r ref("Task")`, `r ref("Learner")` and `r ref("Resampling")`.

Here, we call `r ref("benchmark()")` to perform a single holdout split on a single task and two learners.
We use the `r ref("benchmark_grid()")` function to create an exhaustive design and properly instantiate the resampling:

```{r 06-benchmarking-001}
library(data.table)
design = benchmark_grid(
  tasks = tsk("iris"),
  learners = list(lrn("classif.rpart"), lrn("classif.featureless")),
  resamplings = rsmp("holdout")
)
print(design)
bmr = benchmark(design)
```

Note that the holdout splits have been automatically instantiated for each row of the design.
As a result, the `rpart` learner used a different training set than the `featureless` learner.
However, for comparison of learners you usually want the learners to see the same splits into train and test sets.
To overcome this issue, the resampling strategy needs to be [**manually instantiated**](#resamp-inst) before creating the design.

While the interface of `r ref("benchmark()")` allows full flexibility, the creation of such design tables can be tedious.
Therefore, `r mlr_pkg("mlr3")` provides a convenience function to quickly generate design tables and instantiate resampling strategies in an exhaustive grid fashion: `r ref("benchmark_grid()")`.

```{r 06-benchmarking-002}
# get some example tasks
tasks = lapply(c("german_credit", "sonar"), tsk)

# get some learners and for all learners ...
# * predict probabilities
# * predict also on the training set
library(mlr3learners)
learners = c("classif.featureless", "classif.rpart", "classif.ranger", "classif.kknn")
learners = lapply(learners, lrn,
  predict_type = "prob", predict_sets = c("train", "test"))

# compare via 3-fold cross validation
resamplings = rsmp("cv", folds = 3)

# create a BenchmarkDesign object
design = benchmark_grid(tasks, learners, resamplings)
print(design)
```

### Execution and Aggregation of Results {#bm-exec}

After the [benchmark design](#bm-design) is ready, we can directly call `r ref("benchmark()")`

```{r 06-benchmarking-003}
# execute the benchmark
bmr = benchmark(design)
```

Note that we did not instantiate the resampling instance, but `r ref("benchmark_grid()")` took care of it for us:
each resampling strategy is instantiated for each task during the construction of the exhaustive grid.

After the benchmark, we can calculate and aggregate the performance with `.$aggregate()`:

```{r 06-benchmarking-004}
# measures:
# * area under the curve (auc) on training
# * area under the curve (auc) on test
measures = list(
  msr("classif.auc", id = "auc_train", predict_sets = "train"),
  msr("classif.auc", id = "auc_test")
)
bmr$aggregate(measures)
```

We can aggregate the results further.
For example, we might be interested which learner performed best over all tasks simultaneously.
Simply aggregating the performances with the mean is usually not statistically sound.
Instead, we calculate the rank statistic for each learner grouped by task, and then aggregate the calculated ranks grouped by learner.
Since the AUC needs to be maximized, we multiply with $-1$ so that the best learner gets a rank of 1.

```{r 06-benchmarking-005}
tab = bmr$aggregate(measures)
ranks = tab[, .(learner_id, rank_train = rank(-auc_train), rank_test = rank(-auc_test)), by = task_id]
print(ranks)

ranks[, .(mrank_train = mean(rank_train), mrank_test = mean(rank_test)), by = learner_id][order(mrank_test)]
```

Unsurprisingly, the featureless learner is outperformed.

### Converting specific benchmark objects to resample objects

A `r ref("BenchmarkResult")` object is essentially a collection of multiple `r ref("ResampleResult")` objects.
As these are stored in a column of the aggregated `data.table()`, we can easily extract them:

```{r 06-benchmarking-006}
tab = bmr$aggregate(measures)
rr = tab[task_id == "sonar" & learner_id == "classif.ranger"]$resample_result[[1]]
print(rr)
```

We can now investigate this resampling and even single resampling iterations using one of the approach shown in [the previous section](#bm-exec):

```{r 06-benchmarking-007}
measure = msr("classif.auc")
rr$aggregate(measure)

# get the iteration with worst AUC
perf = rr$score(measure)
i = which.min(perf$classif.auc)

# get the corresponding learner and train set
print(rr$learners[[i]])
head(rr$resampling$train_set(i))
```
