## Learners {#learners}

Objects of class `mlr3::Learner` provide a unified interface to many popular machine learning algorithms in R.
They consist of methods to train and predict a model for a `mlr3::Task` and provide meta-information about the learners, such as the hyperparameters you can set.

The package ships with a minimal set of classification and regression learners to avoid lots of dependencies;
some of the most popular learners are supported via the [mlr3learners](https://mlr3learners.mlr-org.com) package:

* (penalized) linear and logistic regression
* $k$-Nearest Neighbors regression and classification
* Linear and Quadratic Discriminant Analysis
* Naive Bayes
* Support-Vector machines
* Gradient Boosting
* Random Regression Forests and Random Classification Forests
* Kriging

The creation of custom learners is covered in Section \@ref(ext-learner).

### Predefined Learners

Similar to `r ref("mlr_tasks")`, the `r ref("Dictionary")` `r ref("mlr_learners")` can be queried for available learners:

```{r 01-learners-001}
library(mlr3learners)
mlr_learners
```

Each learner has the following information:

* `feature_types`: the type of features the learner can deal with.
* `packages`: the packages required to train a model with this learner and make predictions.
* `properties`: additional properties and capabilities.
  For example, a learner has the property "missings" if it is able to handle missing feature values, and "importance" if it computes and allows to extract data on the relative importance of the features.
* `predict_types`: possible prediction types. For example, a classification learner can predict labels ("response") or probabilities ("prob").

For a tabular overview of integrated learners, see Section \@ref(list-learners).


You can get a specific learner using its `id`, listed under `key` in the dictionary:

```{r 01-learners-002}
learner = mlr_learners$get("classif.rpart")
print(learner)
```

The field `param_set` stores a description of the hyperparameters the learner has, their ranges, defaults, and current values:

```{r 01-learners-003}
learner$param_set
```

The set of current hyperparameter values is stored in the `values` field of the `param_set` field.
You can change the current hyperparameter values by assigning a named list to this field:

```{r 01-learners-004}
learner$param_set$values = list(cp = 0.01, xval = 0)
learner
```
Note that this operation just overwrites all previously set parameters.
If you just want to add or update hyperparameters, you can use `r ref("mlr3misc::insert_named()")`:
```{r 01-learners-005}
learner$param_set$values = mlr3misc::insert_named(
  learner$param_set$values,
  list(cp = 0.02, minsplit = 2)
)
learner
```
This updates `cp` to `0.02`, sets `minsplit` to `2` and keeps the previously set parameter `xval`.

Again, there is an alternative to writing down the lengthy `mlr_learners$get()` part: `r ref("lrn()")`.
This function additionally allows to construct learners with specific hyperparameters or settings of a different identifier:
```{r 01-learners-006}
lrn("classif.rpart", id = "rp", cp = 0.001)
```
If you pass hyperparameters here, it is added to the default parameters in a `r ref("mlr3misc::insert_named()", text = "insert_named")`-fashion.
