## Resampling {#resampling}

### Settings {#resamp-settings}

In this example we use the _iris_ task and a simple classification tree (package `rpart`).

```{r 01-resampling-001}
task = tsk("iris")
learner = lrn("classif.rpart")
```

When performing resampling with a dataset, we first need to define which approach should be used.
The resampling strategies of _mlr3_ can be queried using the `.$keys()` method of the `r ref("mlr_resamplings")` dictionary.

```{r 01-resampling-002}
mlr_resamplings
```

Additional resampling methods for special use cases will be available via extension packages, such as `r gh_pkg("mlr-org/mlr3spatiotemporal")` for spatial data (still in development).

The model fit conducted in the [train/predict/score](#train-predict) chapter is equivalent to a "holdout", so let's consider this one first.
Again, we can retrieve elements via `$get()` or with a convenience function (`r ref("rsmp()")`):

```{r 01-resampling-003}
resampling = rsmp("holdout")
print(resampling)
```

Note that the `Instantiated` field is set to `FALSE`.
This means we did not actually apply the strategy on a dataset yet, but just performed a dry-run.
Applying the strategy on a dataset is done in section next [Instantation](#instantation).

By default we get a .66/.33 split of the data.
There are two ways in which the ratio can be changed:

1. Overwriting the slot in `.$param_set$values` using a named list.

```{r 01-resampling-004}
resampling$param_set$values = list(ratio = 0.8)
```

2. Specifying the resampling parameters directly during construction:

```{r 01-resampling-005}
rsmp("holdout", ratio = 0.8)
```

### Instantiation {#resamp-inst}

So far we just set the stage and selected the resampling strategy.
To actually perform the splitting, we need to apply the settings on a dataset.
This can be done in two ways:

1. Manually by calling the method `.$instantiate()` on a `r ref("Task")`

```{r 01-resampling-006}
resampling = rsmp("cv", folds = 3L)
resampling$instantiate(task)
resampling$iters
resampling$train_set(1)
```

2. Automatically by passing the resampling object to `resample()`.
   Here, the splitting is done within the `resample()` call based on the supplied `r ref("Task")`.

```{r 01-resampling-007}
learner1 = lrn("classif.rpart") # simple classification tree
learner2 = lrn("classif.featureless") # featureless learner, prediction majority class
rr1 = resample(task, learner1, resampling)
rr2 = resample(task, learner2, resampling)

setequal(rr1$resampling$train_set(1), rr2$resampling$train_set(1))
```

If you want to compare multiple learners, you should use the same resampling per task to reduce the variance of the performance estimation (**method 1**).
If you use **method 2** (and do not instantiate manually before), the resampling splits will differ between both runs.

If you aim is to compare different `r ref("Task")`, `r ref("Learner")` or `r ref("Resampling")`, you are better off using the `r ref("benchmark()")` function.
It is basically a wrapper around `r ref("resample()")` simplifying the handling of multiple settings.

If you discover this only after you've run multiple `r ref("resample()")` calls, don't worry - you can transform multiple single `r ref("ResampleResult")` objects into a `r ref("BenchmarkResult")` (explained later) using the `.$combine()` method.


### Execution {#resamp-exec}

With a `r ref("Task")`, a `r ref("Learner")` and `r ref("Resampling")` object we can call `r ref("resample()")` and create a `r ref("ResampleResult")` object.

Before we go into more detail, let's change the resampling to a "3-fold cross-validation" to better illustrate what operations are possible with a `r ref("ResampleResult")`.
Additionally, we tell `r ref("resample()")` to keep the fitted models via the flag `store_models`:

```{r 01-resampling-008}
task = tsk("iris")
learner = lrn("classif.rpart")
resampling = rsmp("cv", folds = 3L)

rr = resample(task, learner, resampling, store_models = TRUE)
print(rr)
```

The following operations are supported with `r ref("ResampleResult")` objects:

* Extract the performance for the individual resampling iterations:

```{r 01-resampling-009}
rr$score("classif.ce")
```

* Extract and inspect the resampling splits:

```{r 01-resampling-010}
rr$resampling
rr$resampling$iters
rr$resampling$test_set(1)
rr$resampling$train_set(3)
```

* Retrieve the learner of a specific iteration and inspect it:

```{r 01-resampling-011}
lrn = rr$learners[[1]]
lrn$model
```

### Custom resampling

Sometimes it is necessary to perform resampling with custom splits.
If you want to do that because you are coming from a specific modeling field, first take a look at the _mlr3_ extension packages to make sure your custom resampling method hasn't been implemented already.

If your custom resampling method is widely used in your field, feel welcome to integrate it into one of the existing _mlr3_ extension packages or create your own extension package.

A manual resampling instance can be created using the `"custom"` template from the `r ref("mlr_resamplings")` dictionary.

```{r 01-resampling-012}
resampling = rsmp("custom")
resampling$instantiate(task,
  list(c(1:10, 51:60, 101:110)),
  list(c(11:20, 61:70, 111:120))
)
resampling$iters
resampling$train_set(1)
resampling$test_set(1)
```
