# Large-Scale Benchmarking {#sec-large-benchmarking}

{{< include ../../common/_setup.qmd >}}

`r chapter = "Large-Scale Benchmarking"`
`r authors(chapter)`

```{r large_benchmarking-001}
#| include: false
#| cache: false
lgr::get_logger("mlr3oml")$set_threshold("off")
library("mlr3batchmark")
library("batchtools")
library("mlr3oml")
if (!dir.exists(file.path("openml", "manual"))) {
  dir.create(file.path("openml", "manual"), recursive = TRUE)
}
options(mlr3oml.cache = file.path("openml", "cache"))
```

In the world of ML, there are many methods that are hard to evaluate using mathematical analysis alone.
Even if formal analysis is successful, it is often an open question whether real-world datasets satisfy the necessary assumptions for the theorems to hold.
Consequently, researchers resort to conducting benchmark experiments to answer fundamental scientific questions.
Such studies evaluate different algorithms on a wide range of datasets, with the aim of identifying which method performs best.
These empirical investigations are essential for understanding the capabilities and limitations of existing methods and for developing new and improved approaches.
To carry them out effectively, researchers need access to datasets, spanning a wide range of domains and problem types.
This is because one can only draw conclusions about the kind of datasets on which the benchmark study was conducted.
Fortunately, there are several online resources available for acquiring such datasets, with `r link("https://www.openml.org/", "OpenML")` [@openml2013] being a popular choice.
In the first part of this chapter, we will show how to use OpenML through the connector package `r mlr3oml`.

However, evaluating algorithms on such a large scale requires significant computational resources.
For this reason, researchers often utilize High-Performance Computing (HPC) clusters which allow executing experiments massively parallel.
The R package `r ref_pkg("batchtools")` [@batchtools] is a tool for managing and executing experiments on such clusters.
In the second section we will show how to use batchtools and its mlr3 connector for benchmarking: `r mlr3batchmark`.
Once the experiments are complete, visualizations and statistical tests are used to draw conclusions from the benchmark experiments using the `r mlr3benchmark` package.

A common design for benchmarking is to compare a set of Learners $L_1, \ldots, L_n$ by evaluating their performance on a set of tasks $T_1, \ldots, T_k$ which each have an associated resampling $R_1, \ldots, R_k$.
We call such a task-resampling combination $C_j = (T_j, R_j)$ a `r index("resampled task")`.
Furthermore, we define $M = (M_1, \ldots, M_m)$ to be the evaluation measures.
We here focus on the most frequent case of a full experimental grid design, where each experiment is defined by a triple $E_{i, j} = (L_i, C_j, M)$, which evaluates a learner $L_i$ on a resampled task $C_j$ using performance measures $M$.
Running the benchmark consists of evaluating all these triples.
The execution of a single resample experiment can again be subdivided into computationally independent resampling iterations to achieve a finer granularity for parallelization (see @sec-parallel-resample).

As a guiding example throughout this chapter, we will compare the random forest implementation `lrn("classif.ranger")` with the logistic regression learner `lrn("classif.log_reg")`.
This question holds significant practical importance and has already been studied by, e.g., @couronne2018random.

The following example compares these two learners using a simple holdout resampling on three classification tasks that ship with `r mlr3`.
@sec-performance already covered how to conduct such benchmark experiments, and we recommend revisiting this chapter if anything is unclear.
The metric of choice is the classification accuracy.
Note that we "robustify" both learners to work on a wide range of tasks using the respective preprocessing pipeline, c.f. @sec-pipelines-nonseq.
The `ppl("robustify")` creates a preprocessing pipeline, i.e. a predefined `r ref("Graph")` consisting of common operations such as missing value imputation or feature encoding.
By specifying the learner during construction of the pipeline, some preprocessing operations can be omitted, e.g. when a learner can handle missing values itself.
Moreover, we specify the fallback learner to be a featureless learner (compare @sec-fallback) and we use `"try"` as encapsulation which will be discussed later in this chapter.
<!-- FIXME: reference to robustify specifically when preprocessing chapter is ready -->

```{r large_benchmarking-002}
#| warning: false
# create logistic regression pipeline
lrn_logreg = lrn("classif.log_reg")
lrn_logreg = as_learner(
  ppl("robustify", learner = lrn_logreg) %>>% lrn_logreg)
lrn_logreg$id = "logreg"
lrn_logreg$fallback = lrn("classif.featureless")
lrn_logreg$encapsulate = c(train = "try", predict = "try")

# create random forest pipeline
lrn_ranger = lrn("classif.ranger")
lrn_ranger = as_learner(
  ppl("robustify", learner = lrn_ranger) %>>% lrn_ranger)
lrn_ranger$id = "ranger"
lrn_ranger$fallback = lrn("classif.featureless")
lrn_ranger$encapsulate = c(train = "try", predict = "try")

# create full grid design with holdout resampling
set.seed(123)
design = benchmark_grid(
  tsks(c("german_credit", "sonar", "spam")),
  list(lrn_logreg, lrn_ranger),
  rsmp("holdout")
)

# run the benchmark
bmr = benchmark(design)

# retrieve results
aggr = bmr$aggregate(msr("classif.acc"))
aggr[, .(task_id, learner_id, classif.acc)]
```

Looking only at the aggregated performance measures, the random forest outperforms the logistic regression on all three datasets.
However, this analysis is not conclusive because

* only three tasks are considered,
* only a single holdout resampling is used, and
* the difference between performances might not be statistically significant.

In the subsequent sections, we will show the steps required to scale up this analysis using the tools mentioned earlier.

## Getting Data with OpenML {#sec-openml}

In order to be able to draw meaningful conclusions from benchmark experiments, a good choice of datasets and tasks is essential.
It is therefore helpful to

1. have convenient access to a large collection of datasets and be able to filter them for specific properties, and
1. be able to easily share datasets, tasks and collections with others, so that they can evaluate their methods on the same problems and thus allow cross-study comparisons.

OpenML is a platform that facilitates the sharing and dissemination of machine learning research data and satisfies these two desiderata.
Like `mlr3`, it is free and open source.
Unlike `mlr3`, it is not tied to a programming language and can for example also be used from its Python interface [@feurer2021openml] or from Java.
Its goal is to make it easier for researchers to find the data they need to answer the questions they have.
Its design was guided by the `r link("https://www.go-fair.org/fair-principles/", "FAIR")` principles, which stand for **F**indability, **A**ccessibility, **I**nteroperability and **R**eusability.
The purpose of these principles is to make scientific data more easily discoverable and reusable.
More concretely, OpenML is a repository for storing, organizing and retrieving datasets, algorithms and experimental results in a standardized way.
Entities have unique identifiers and standardized (meta) data.
Everything is accessible through a REST API or the web interface.

In this section we will cover some of the main features of OpenML and how to use them via the  `r mlr3oml` interface package.
OpenML supports different types of objects and we will cover the following:

*   `r index("OpenML **Dataset**")`: (Usually tabular) data with additional metadata.
    The latter includes for example a description of the data and a license.
    When accessed via `mlr3oml`, it can be converted to a `r ref("mlr3::DataBackend")`.
    As most OpenML datasets also have a designated target column, they can often directly be converted to an `r ref("mlr3::Task")`.
*   `r index("OpenML **Task**")`: A machine learning task, i.e. a concrete problem specification on an OpenML dataset.
    This includes splits into train and test sets, thereby differing from the `mlr3` task definition, and corresponds to the notion of a resampled task defined in the introduction.
    Thus, it can be converted to both a `r ref("mlr3::Task")` and corresponding instantiated `r ref("mlr3::Resampling")`.
*   `r index("OpenML **Task Collection**")`: A container object that allows to group tasks.
    This allows the creation of benchmark suites, such as the OpenML CC-18 [@bischl2021openml], which is a curated collection of classification tasks.

While OpenML also supports other objects such as representations of algorithms (flows) and experiment results (runs), they are not covered in this chapter.
For more information about these features, we refer to the OpenML `r link("htts://openml.org", "website")` or the documentation of the `mlr3oml` package.

### Dataset {#sec-openml-dataset}

To illustrate the OpenML dataset class, we will use the dataset with `r link("https://openml.org/d/1590", "ID 1590")` -- the well-known adult data.
Such an ID can either be found by searching for objects on the OpenML website or through the REST API.
This will be covered in more detail in @sec-openml-filtering.

We load the object into R using `r ref("mlr3oml::odt()")`, which returns an object of class `r ref("OMLData")`.

```{r large_benchmarking-003}
library("mlr3oml")
odata = odt(id = 1590)
odata
```

This dataset contains information about `r odata$nrow` adults -- such as their age or education -- and the goal is usually to predict the *class* variable, which indicates whether a person has an income above 50K dollars per year.
The `r ref("OMLData")` object not only contains the data itself, but comes with additional metadata that is accessible through its fields.

```{r large_benchmarking-004}
odata$license
```

The actual data can be accessed through `$data`.

```{r large_benchmarking-005}
odata$data
```

:::{.callout-tip}
When working with OpenML objects, these are downloaded in a piecemeal fashion from the OpenML server.
This way you can, e.g., access the metadata from the `OMLData` object without loading the dataset.
While accessing the `$data` slot in the above example, the download is automatically triggered, imported in R, and the `data.frame` gets stored in the `odata` object.
All subsequent accesses to `$data` will be transparently redirected to the in-memory `data.frame`.
Additionally, many objects can be permanently cached on the local file system.
This caching can be enabled by setting the option `mlr3oml.cache` to either `TRUE` or a specific path to be used as the cache folder.
:::

After we have loaded the data of interest into R, the next step is to convert it into a format usable with `mlr3`.
The class that comes closest to the OpenML dataset is the `r ref("mlr3::DataBackend")` (see @sec-backends) and it is possible to convert `OMLData` objects by calling `r ref("as_data_backend()")`.
This is a recurring theme throughout this section: OpenML and mlr3 objects are well interoperable.

```{r large_benchmarking-006}
backend = as_data_backend(odata)
backend
```

We can create an `mlr3` task from the backend via `r ref("as_task_classif()")`.

```{r large_benchmarking-007}
task = as_task_classif(backend, target = "class")
```

Some datasets on OpenML contain columns that should neither be used as a feature nor a target.
The column names that are usually included as features are accessible through the field `$feature_names` and we assign them to the `mlr3` task accordingly.
Note that for the dataset at hand this would not have been necessary, as all non-target columns are to be treated as predictors, but we include it for clarity.

```{r}
task$col_roles$feature = odata$feature_names
task
```

Alternatively, as the OpenML adult dataset comes with a default target, you can also directly convert it to a task with the appropriate type.
This will also set the features of the task appropriately.

```{r large_benchmarking-008}
task = as_task(odata)
```

### Task {#sec-openml-task}

OpenML tasks are built on top of OpenML datasets and additionally specify the target variable, the train-test splits to use for resampling, and more.
Similar to `mlr3`, OpenML has different types of tasks, such as regression or classification.
A task associated with the adult data from earlier has ID `r link("https://openml.org/t/359983", "359983")`.
We can load the object using the `r ref("mlr3oml::otsk()")` function, which returns an `r ref("OMLTask")` object.
Note that this task object is different from an `mlr3` `r ref("Task")` and cannot directly be used for ML.
However, it contains all required information and there is a convenient converter, as shown below.

```{r large_benchmarking-009}
otask = otsk(id = 359983)
otask
```

The `r ref("OMLData")` object associated with the underlying dataset can be accessed through the `$data` field.

```{r large_benchmarking-010}
otask$data
```

The data splits associated with the estimation procedure are accessible through the field `$task_splits`.
In `mlr3` terms, these are the instantiation of an `mlr3` `r ref("Resampling")` on a specific `Task`.

```{r large_benchmarking-011}
otask$task_splits
```

The OpenML task can be converted to both an `mlr3` `Task` and a `r ref("ResamplingCustom")` instantiated on the task.
To convert to the former we can use `r ref("as_task()")`.

```{r large_benchmarking-012}
task = as_task(otask)
task
```

The accompanying resampling can be created using `r ref("as_resampling()")`.

```{r large_benchmarking-013}
resampling = as_resampling(otask)
resampling
```

:::{.callout-tip}
As a shortcut, it is also possible to create the objects using the `"oml"` task or resampling using the `r ref("tsk()")` and `r ref("rsmp()")` constructors and pass the `data_id` or `task_id` to query, e.g. `tsk("oml", task_id = 359983)`.
:::


### Filtering of Data and Tasks {#sec-openml-filtering}

Besides working with objects with known IDs, another important question is how to find IDs of relevant datasets or tasks.
Because objects on OpenML have strict metadata, they can be filtered w.r.t these properties.
This is possible through either the website or the REST API.
In addition, the website also supports targeted text queries to search for specific datasets such as the "adult" data from earlier.

The `r ref("list_oml_data()")` function allows to filter datasets for specific properties.
As an example, we might only be interested in comparing the random forest and the logistic regression on datasets with less than 4 features and 100 to 1000 observations.
By setting `number_classes` to 2, we only receive datasets where the default target has two different values.
To keep the output readable, we only show the first 5 results from that query.

```{r}
#| include: false
path_odatasets = file.path("openml", "manual", "odatasets_filter.rds")
```

```{r large_benchmarking-014, eval = !file.exists(path_odatasets)}
odatasets = list_oml_data(
  limit = 5,
  number_features = c(1, 4),
  number_instances = c(100, 1000),
  number_classes = 2
)
```

```{r}
#| include: false
if (file.exists(path_odatasets)) {
  odatasets = readRDS(path_odatasets)
} else {
  saveRDS(odatasets, path_odatasets)
}
```


The table below confirms that indeed only datasets with the specified properties were returned.
We only show a subset of the columns for readability.

```{r large_benchmarking-016}
odatasets[,
  .(data_id, NumberOfClasses, NumberOfFeatures, NumberOfInstances)]
```

Besides datasets, it is also possible to filter tasks.
This can be done using `list_oml_tasks()` and works analogously to the previous example.

We could now start looking at the returned IDs in more detail in order to verify whether they are suitable for our purposes.
This process can be tedious, as some datasets have hard-to-detect quirks to look out for.
A solution to this problem is to use an existing curated task collection, which we will cover next.

### Task Collection {#sec-openml-collection}

The OpenML task collection is a container object bundling existing tasks.
This allows for the creation of `r index("benchmark suites")`, which are curated collections of tasks, satisfying certain quality criteria.
One example for such a benchmark suite is the `r link("https://www.openml.org/search?type=study&study_type=task&id=99", "OpenML CC-18")`, which contains curated classification tasks [@bischl2021openml].
Other collections available on OpenML include the `r link("https://www.openml.org/search?type=study&study_type=task&id=271", "AutoML benchmark")` [@amlb2022] or a `r link("https://www.openml.org/search?type=study&study_type=task&id=304", "benchmark for tabular deep learning")` [@grinsztajn2022why].


```{r}
#| include: false
path_otask_collection = file.path("openml", "manual", "otask_collection99.rds")
```

```{r large_benchmarking-017, eval = !file.exists(path_otask_collection)}
otask_collection = ocl(id = 99)
```

```{r}
#| include: false
if (file.exists(path_otask_collection)) {
  otask_collection = readRDS(path_otask_collection)
} else {
  # need to trigger the download
  otask_collection$task_ids
  saveRDS(otask_collection, path_otask_collection)
}
```

We can create an `r ref("OMLCollection")` using `r ref("mlr3oml::ocl()")`.
We see that the CC-18 contains 72 classification tasks on different datasets.

```{r large_benchmarking-019}
otask_collection
```

The contained tasks can be accessed through `$task_ids`.

```{r large_benchmarking-020}
otask_collection$task_ids
```

We will now define our experimental design using tasks from the CC-18.
If we wanted to get all tasks and resamplings, we could achieve this using the converters `r ref("as_tasks()")` and `r ref("as_resamplings()")`.
However, as the CC-18 contains not only binary classification tasks we use `r ref("list_oml_tasks()")` to subset the collection further.
We pass the task IDs from the CC-18 as argument `task_id` and request the number of classes to be 2.

```{r}
#| include: false
path_binary_cc18 = file.path("openml", "manual", "binary_cc18.rds")
```

```{r large_benchmarking-021, eval = !file.exists(path_binary_cc18)}
binary_cc18 = list_oml_tasks(
  task_id = otask_collection$task_ids, number_classes = 2)
```

```{r}
#| include: false
if (!file.exists(path_binary_cc18)) {
  saveRDS(binary_cc18, path_binary_cc18)
} else {
  binary_cc18 = readRDS(path_binary_cc18)
}
```


In order to keep the runtime reasonable in later examples, we only use the first six tasks.

```{r large_benchmarking-023}
binary_cc18[, .(task_id, name, NumberOfClasses)]
ids = binary_cc18$task_id[1:6]
```

We now define the learners, tasks, and resamplings for the experiment.
In addition to the random forest and the logistic regression, we also include a featureless learner as a baseline.

```{r large_benchmarking-024}
otasks = lapply(ids, otsk)

tasks = lapply(otasks, as_task)
resamplings = lapply(otasks, as_resampling)

lrn_featureless = lrn("classif.featureless", id = "featureless")
learners = list(lrn_logreg, lrn_ranger, learner_featureless)
```

To define the design table, we use `r ref("benchmark_grid()")` and set `paired` to `TRUE`.
This option can be used in situations where each resampling is instantiated on a corresponding task (therefore the `tasks` and `resamplings` below must have the same length) and each learner should be evaluated on every resampled task.

```{r large_benchmarking-025}
large_design = benchmark_grid(
  tasks, learners, resamplings, paired = TRUE
)
large_design
```

## Experiment Execution on HPC Clusters {#sec-hpc-exec}

Once an experimental design is finalized, the next step is to run it.
Parallelizing this step is conceptually straightforward, because we are facing an embarrassingly parallel problem (see @sec-parallelization).
Not only are the resample experiments independent, but even their individual iterations are.
However, if the experiment is large, parallelization on a local machine as shown in @sec-parallelization often is not enough and using a distributed computing system, such as an HPC cluster, is required.
While access to HPC clusters is widespread at universities and research-driven companies, the effort required to work on these systems is still considerable.
The R package `r ref_pkg("batchtools")` provides a framework to simplify running large batches of computational experiments in parallel from R.
It is highly flexible, making it suitable for a wide range of computational experiments, including ML, optimisation, simulation, and more.

:::{.callout-tip}
In @sec-parallel-resample we have already touched upon different parallelization backends.
The package `r ref_pkg("future")` also comes with a plan for `batchtools`.
However, for larger experiments, the additional control over the execution which `batchtools` offers is invaluable.
Therefore, we recommend future's `"batchtools"` plan only for moderately sized experiments which complete within a couple of hours.
To estimate the total runtime, a subset of the benchmark is usually executed and measured, and then the runtime is extrapolated.
:::

### HPC Basics {#sec-hpc-basics}

An HPC cluster is a collection of interconnected computers or servers providing computational power beyond what a single computer can achieve.
They are used for solving complex problems in chemistry, physics, engineering, ML and other fields that require a large amount of computational resources.
An HPC cluster typically consists of multiple compute nodes, each with multiple CPU/GPU cores, memory, and local storage.
These nodes are connected together by a high-speed network and network file system which enables the nodes to communicate and work together on a given task.
These clusters are also designed to run parallel applications that can be split into smaller tasks and distributed across multiple compute nodes to be executed simultaneously.
We will leverage this capacity to parallelize the execution of the benchmark experiment.
The most important difference between such a cluster and a personal computer, is that the nodes cannot be accessed directly, but instead computational jobs are queued by a `r index("scheduling system")` like `r link("https://slurm.schedmd.com", "Slurm")` (Simple Linux Utility for Resource Management).
A scheduling system is a software tool that orchestrates the allocation of computing resources to users or applications on the cluster.
It ensures that multiple users and applications can access the resources of the cluster in a fair and efficient manner, and also helps to maximize the utilization of the computing resources.

@fig-hpc contains a rough sketch of an HPC architecture.
Multiple users can log in into the head node (typically via `r link("https://en.wikipedia.org/wiki/Secure_Shell", "SSH")`) and add their computational workloads of the form "Execute Computation X using Resources Y for Z amount of time" to the queue.
One such instruction will be referred to as a `r index("computational job")`.
The scheduling system controls when these computational jobs are executed.

```{r large_benchmarking-026}
#| label: fig-hpc
#| fig-cap: "Illustration of an HPC cluster architecture."
#| fig-align: "center"
#| fig-alt: "A rough sketch of the architecture of an HPC cluster. Ann and Bob both have access to the cluster and can log in to the head node. There, they can submit jobs to the scheduling system, which adds them to its queue and determines when they are run."
#| echo: false
include_multi_graphics("Figures/mlr3book_figures-32.svg", "Figures/mlr3book_figures-32.png")
```

Common challenges when interoperating with a scheduling systems are that

1. the code to submit a job depends on the scheduling system,
1. code that runs locally has to be adapted to run on the cluster,
1. the burden is on the user to account for seeding to ensure reproducibility, and
1. it is cumbersome to query the status of jobs and debug failures.

In the following we will see how `batchtools` mitigates these problems.

### General Setup and Experiment Registry {#sec-registry}

Our goal in this section is to run the benchmark design `large_design` shown in @sec-openml-collection on an HPC cluster.
We use the packages `batchtools` and `r mlr3batchmark` for this.
The `mlr3batchmark` package assists in translating the machine learning problem defined with syntax and objects from `mlr3` to the more general "apply some algorithm A on some problem P" approach in `batchtools`, which is briefly outlined in @sec-custom-experiment-definition.

The central concept of `batchtools` is the experiment or `r index("job")`:
One replication of an experiment (or job) is defined by applying a (parameterized) algorithm to a (parameterized) problem.
A benchmark experiment in batchtools then consists of running many such experiments with different algorithms, algorithm parameters, problems, and problem parameters.
Each such experiment is computationally independent of all other experiments and constitutes the basic level of computation batchtools can parallelize.

In the introduction of this chapter, we have defined a benchmark experiment as evaluating a number of triples $E_{i, j} = (L_i, C_j, M)$ that are defined by a learner $L_i$, a resampled task $C_j = (T_j, R_j)$, and measures $M$.
While it might seem natural to define one such triplet as a job, each resampled task $C_j$ can be split up even further, namely into its resampling iterations $C^1_j, \ldots, C^{n_j}_j$, where $n_j$ is the number of iterations of resampling $R_j$.
This makes the computation more granular and gives you the opportunity to utilize more CPUs in parallel (c.f., @sec-parallelization).
Moreover, because the evaluation of the measures $M$ is (usually) computationally negligible, it is common to only parallelize the execution of the resample experiments and evaluate the measures $M$ afterwards.
For these reasons, we will understand one pair $(L_i, C^l_j)$ as a single batchtools experiment.^[Note that such a job does not have to coincide with the notion of a computational job defined earlier, more on that in @sec-batchtools-submission.]
The custom definition of problems and algorithm which yields a different granularity is demonstrated in @sec-custom-experiment-definition.

The first step is always to create an (or load an existing) experiment registry using the function `r ref("batchtools::makeExperimentRegistry()")` (or `r ref("batchtools::loadRegistry()")`, respectively).
This function constructs the inter-communication object for all functions in `batchtools` and corresponds to a folder on the file system.
Among other things, the experiment registry stores the

* algorithms, problems, and job definitions
* log outputs and status of submitted, running, and finished jobs
* job results
* `r index("cluster function", "Batchtools object that allows to configure the interaction with the scheduling system.")`, which defines the interaction with the scheduling system

While the first three bullet points should be relatively clear, the fourth needs some explanation.
By configuring the scheduling system through a cluster function, it is possible to make the interaction with the scheduling system independent of the scheduling software.
We will come back to this later and show how to change it to work on a Slurm cluster.

We create a registry in a subdirectory of our working directory - on a real cluster, make sure that this folder is stored on a shared network filesystem, otherwise the nodes cannot access it.
Furthermore, we set the registry's `seed` to 1 and the `packages` to `r mlr3verse`, which will make the package available in all our experiments.

```{r include = FALSE}
#| cache: false
if (dir.exists("experiments"))
    unlink("experiments", recursive = TRUE)
```

```{r large_benchmarking-027}
#| cache: false
library("batchtools")

# create registry
reg = makeExperimentRegistry(
  file.dir = "./experiments",
  seed = 1,
  packages = "mlr3verse"
)
```

When printing our newly created registry, we see that there are 0 problems, algorithms or jobs registered.
Among other things, we are informed that the "Interactive" cluster function (see `r ref("batchtools::makeClusterFunctionsInteractive()")`) is used and about the working directory for the experiments.

```{r large_benchmarking-028}
reg
```


### Experiment Definition Using mlr3batchmark {#sec-mlr3batchmark}

The next step is to populate the registry with problems and algorithms, which we will then use to define the jobs, i.e., the resampling iterations.
This is the first step where `mlr3batchmark` comes into play.
Doing this step with `batchtools` is also possible and gives you more flexibility and is demonstrated in @sec-custom-experiment-definition.
By calling `r ref("batchmark()")`, `mlr3` tasks and `mlr3` resamplings will be translated to batchtools problems, and `mlr3` learners are mapped to batchtools algorithms.
Then, jobs for all resampling iterations are created.

```{r large_benchmarking-029}
#| cache: false
#| output: false
library("mlr3batchmark")
batchmark(large_design, reg = reg)
```

::: {.callout-tip}
All batchtools functions that interoperate with a registry take a registry as an argument.
By default, this argument is set to the last created registry, which is currently the `reg` object defined earlier.
We nonetheless pass it explicitly in this section for clarity, but would not have to do so.
:::

When printing the registry, we confirm that six problems (one for each resampled task) but only a single algorithm is registered.
While a 1-on-1 mapping of the 3 algorithms in the design would have also been possible, the approach in `mlr3batchmark` uses a single algorithm parametrized with the learner identifiers instead for efficiency.
Furthermore, $180 = 3 \times 6 \times 10$ jobs, i.e. one for each resampling iteration, are registered.

```{r large_benchmarking-030}
reg
```

We can summarize the defined experiments using `r ref("batchtools::summarizeExperiments()")`.
There are 10 jobs for each combination of a learner and resampled task, as 10-fold cross-valdation is used as the resampling procedure.

```{r large_benchmarking-031}
summarizeExperiments(
  by = c("task_id", "learner_id"), reg = reg)
```

The function `r ref("batchtools::getJobTable()")` can be used to get more detailed information about the jobs.
Here, we only show a few selected columns for readability and unpack the list columns `algo.pars` and `prob.pars` using `r ref("batchtools::unwrap()")`.
Among other things, we see that each job has a unique `job.id`.
Each row in this job table represents one iteration (column `repl`) of a resample experiment.

```{r large_benchmarking-032}
job_table = getJobTable(reg = reg)
job_table = unwrap(job_table)
job_table = job_table[,
  .(job.id, learner_id, task_id, resampling_id, repl)
]

job_table
```


### Job Submission {#sec-batchtools-submission}

Once the experiments are defined, the next step is to submit them.
Before doing so, it is recommended to test each algorithm individually using `r ref("batchtools::testJob()")`.
We test the job with `job.id = 1` exemplary and specify `external = TRUE` to run the test in an external R session.
The return of the `testJob()` function (which is the return value of the `"run_learner"` algorithm) is a bit technical, because here not the complete objects but only the essential parts are returned to reduce the communication overhead (a named list `learner_state` and a named list `prediction`) - we do not give a detailed description of the list elements here.
We can for example access the training time through the `learner_state` as shown below.

<!-- ```{r large_benchmarking-033}
#| output: true
result = testJob(1, external = TRUE, reg = reg)
result$learner_state$train_time
``` -->

In case something goes wrong, `batchtools` comes with a bunch of useful debugging utilities covered in @sec-batchtools-monitoring.
Once we are confident that the jobs are defined correctly, we can proceed with their submission, which requires

1. specifying resource requirements for each computational job, and
1. (optionally) grouping multiple jobs into one computational job

Which resources can be configured depends on the cluster function that is set in the registry.
We earlier left it at its default value, which is the "Interactive" cluster function.
In the following we assume that we are working on a Slurm cluster.
Accordingly, we initialize the cluster function with `r ref("batchtools::makeClusterFunctionsSlurm()")` and a predefined `r link("https://github.com/mllg/batchtools/blob/master/inst/templates/slurm-simple.tmpl", "slurm-simple template")`.
A template file is a shell script with placeholders filled in by `batchtools` and contains

1. the command to start the computation via `Rscript` or `R CMD batch`, usually in the last line, and
1. comments which serve as annotations for the scheduler, e.g. to communicate resources or paths on the file system.

The exemplary template should work on many Slurm installations out-of-the-box, but can also easily be customized to also work with more advanced configurations.

```{r large_benchmarking-034}
cf = makeClusterFunctionsSlurm(template = "slurm-simple")
```

To proceed with the examples on a local machine, set the cluster functions to a Socket backend.

```{r large_benchmarking-035}
cf = makeClusterFunctionsSocket()
```

```{r}
#| include: false
cf = makeClusterFunctionsInteractive()
```

:::{.callout-tip}
It is possible to customize the cluster function.
More information is available in the documentation of the `batchtools` package.
:::

We update the value of the `$cluster.functions` and save the registry.
This only has to be done manually when modifying the fields explicitly.
Functions like `r ref("batchmark()")` internally save the registry when required.

```{r large_benchmarking-036}
#| eval: true
#| cache: false
reg$cluster.functions = cf
saveRegistry(reg = reg)
```

The jobs are submitted to the scheduler via `r ref("batchtools::submitJobs()")`.
The most important arguments of this function besides the registry are:

* `ids`, which are either a vector of job IDs to submit, or a data frame with columns `job.id` and `chunk`, which allows grouping multiple jobs into one larger computational job and thereby controlling the granularity.
  This often makes sense on HPCs, as submitting and running computational jobs comes with a considerable overhead and there are often hard limits for the maximum runtime (walltime).
* `resources`, which is a named list specifying the resource requirements for the submitted jobs.

We will `r ref("batchtools::chunk()")` the IDs in such a way that 5 iterations of one resample experiment are run sequentially in one computational job.
The optimal grouping depends on the concrete experiment and scheduling system.

```{r large_benchmarking-037}
ids = job_table$job.id
chunks = data.table(
  job.id = ids, chunk = chunk(ids, chunk.size = 5, shuffle = FALSE)
)
chunks
```

Furthermore, we specify the number of CPUs per computational job to 1, the walltime to 1 hour (3600 seconds), and the RAM limit to 8 GB.
The set of resources depends on your cluster and the corresponding template file.
For a list of resource names that are standardized across most implementations, see `r ref("batchtools::submitJobs()")`.
If you are unsure about the resource requirements, you can start a subset of jobs with conservative resource constraints, e.g. the maximum runtime allowed for your computing site.
Measured runtimes and memory usage can be queried with `r ref("batchtools::getJobTable()")` and finally used to better estimate the required resources for the remaining jobs.

<!-- output to false otherwise we see the interactive cluster function -->
```{r large_benchmarking-038}
#| cache: false
#| output: false
submitJobs(
  ids = chunks,
  resources = list(ncpus = 1, walltime = 3600, memory = 8000),
  reg = reg
)

# wait for all jobs to terminate
waitForJobs(reg = reg)
```


:::{.callout-tip}
A good approach to submit computational jobs is to do this from an R session that is running persistently.
One option is to use `r link("https://github.com/tmux/tmux/wiki", "TMUX (Terminal Multiplexer)")` on the head node to continue job submission (or computation, depending on the cluster functions) in the background.
:::


###  Job Monitoring, Error Handling, and Result Collection {#sec-batchtools-monitoring}



After submitting the jobs, the next phase is to wait for them to finish.
In case you terminated your running R session after job submission, you can load the experiment registry using `r ref("loadRegistry()")` to continue where you left off.

In any large scale experiment many things can and will go wrong, even if we test our jobs beforehand using `r ref("batchtools::testJob()")` as recommended earlier.
The cluster might have an outage, jobs may run into resource limits or crash, subtle bugs in your code could be triggered or any other error condition might arise.
In these situations it is important to quickly determine what went wrong and to recompute only the minimal number of required jobs.

The current status of the computation can be queried with `r ref("getStatus()")`, which lists the number of jobs categorized in multiple status groups:

```{r large_benchmarking-041}
getStatus(reg = reg)
```

To query the ids of jobs in the respective categories, see `r ref("findJobs()")` and, e.g., `findNotSubmitted()` or `findDone()`.

In our case, all experiments finished and none expired or crashed - albeit we took some countermeasures by extending our base learners with fallback learners and integrating them in the pipeline "robustify".
Nonetheless, it makes sense to check the logs for suspicious messages and warnings with `r ref("grepLogs()")` before proceeding with the analysis of the results.
For this purpose, we have set the encapsulation before to `"try"`: in contrast to `"evaluate"` or `"callr"` encapsulation, `"try"` does not capture the output from messages, warnings or errors and store them in the learner's log.
All output is printed to the console and redirected to a log file so that `batchtools` can operate on it with functions like `r ref("getLog()")` or `r ref("grepLogs()")`.

In the following, we extend the design with the debug learner (see @sec-error-handling) erring with a 50% probability.
By just calling `r ref("batchmark()")` with the new design again, the new experiments will be added to the registry on top of the existing jobs.
The `tasks` and `resamplings` below are once again those from the `large_design` from earlier.

```{r}
#| cache: false
#| output: false
extra_design = benchmark_grid(
  learners = lrns("classif.debug", error_train = 0.5),
  tasks = tasks,
  resampling = resamplings,
  paired = TRUE
)

batchmark(extra_design, reg = reg)
```

We can check the new state and submit the jobs which have not been submitted yet (i.e., the newly created jobs):

```{r}
#| cache: false
getStatus(reg = reg)
ids = findNotSubmitted(reg = reg)
```

We queue these jobs as usual by passing their IDs to `submitJobs()`:

```{r}
#| cache: false
#| output: false
submitJobs(ids, reg = reg)
waitForJobs(reg = reg)
```

After these jobs have terminated, we can get a summary of those that failed:

```{r}
error_ids = findErrors(reg = reg)
summarizeExperiments(
  error_ids, by = c("task_id", "learner_id"), reg = reg)
```

Unsurprisingly, all failed jobs have one thing in common: the debug learner.

Finally, it is time to collect the experiment output.
The results for benchmark experiments defined with `batchmark()` can be collected with `r ref("reduceResultsBatchmark()")`, which constructs a regular `r ref("BenchmarkResult")`.
It is also possible to retrieve single results with `r ref("loadResult()")`, but the returned values are hard to work with, as they are not optimized for usability, but efficiency.
Here, we only collect results which were not produced by the debug learner.

```{r}
ids = findExperiments(
  algo.pars = learner_id != "classif.debug", reg = reg)
bmr = reduceResultsBatchmark(ids, reg = reg)
bmr$aggregate()
```



### Custom Experiment Definition {#sec-custom-experiment-definition}

{{< include ../../common/_optional.qmd >}}

While `mlr3batchmark` excels for conducting benchmarks on HPCs, there can be situations in which more fine-grained control over the experiment definition is beneficial or even required.
Here, we will show how to define batchtools jobs that execute an `mlr3` benchmark experiment without the help of `mlr3batchmark`.
There is not one single way how to achieve this, and we here show a solution that sacrifices efficiency for simplicity.
Unless you have a specific reason to customize your experiment definition, we recommend using `mlr3batchmark`.
Like before, the first step is to create an experiment registry.

```{r include = FALSE}
#| cache: false
if (dir.exists("experiments-custom"))
  unlink("experiments-custom", recursive = TRUE)
```

```{r large_benchmarking-046}
#| cache: false
reg = makeExperimentRegistry(
  file.dir = "./experiments-custom",
  seed = 1,
  packages = "mlr3verse"
)
```

We can register a problem by calling `r ref("batchtools::addProblem()")`, whose main arguments beside the registry are:

* `name` to uniquely identify the problem,
* `data` to represent the static data part of a problem, and
* `fun`, which takes in the problem `data`, the problem parameters, and the `job` definition, see `r ref("batchtools::makeJob()")` and returns a problem `instance`.

We register all task-resampling combinations of the `large_design` using the task ID as the name.^[The `mlr3` task ID is not the same as the OpenML task ID.]
The problem `fun` takes in the static problem `data` and returns it as the problem `instance` as is.
If we were using problem parameters, we could modify the problem instance depending on their values.
In the code below, recall that the `tasks` and `resamplings` were originally used to define the `large_design`.

```{r large_benchmarking-047}
#| output: false
#| cache: false
for (i in seq_along(tasks)) {
  addProblem(
    name = tasks[[i]]$id,
    data = list(task = tasks[[i]], resampling = resamplings[[i]]),
    fun = function(data, job, ...) data,
    reg = reg
  )
}

```

When calling `batchtools::addProblem()`, not only is the problem added to the registry object from the active R session, but this information is also synced with the registry folder.

The next step is to register the algorithm we want to run, which we achieve by calling `r ref("batchtools::addAlgorithm()")`.
Besides the registry, it takes in the arguments:

* `name` to uniquely identify the algorithm and
* `fun`, which takes in the problem instance, the algorithm parameters, and the `job` definition.
  It defines the computational steps of an experiment and its return value is the experiment result, i.e. what can later be retrieved using `r ref("loadResult()")`.

The algorithm function receives a list containing the task and resampling as the problem `instance`, the `learner` as the algorithm parameter, and the `job` object.
It then executes the resample experiment defined by these three objects using `r ref("resample()")` and returns a `r ref("ResampleResult")`.
This differs from the `mlr3batchmark`, where one resampling iteration corresponds to one job in `batchtools`.
Here, one `batchtools` job represents a complete resample experiment.

```{r large_benchmarking-048}
#| cache: false
addAlgorithm(
  "run_learner",
  fun = function(instance, learner, job, ...) {
    resample(instance$task, learner, instance$resampling)
  },
  reg = reg
)

reg$algorithms
```

As we have now defined the problems and the algorithm, we can define concrete experiments using `r ref("batchtools::addExperiments()")`.
This function has arguments

* `prob.designs`, a named list of data frames. The name must match the problem name while the column names correspond to parameters of the problem.

* `algo.designs`, a named list of data frames. The name must match the algorithm name while the column names correspond to parameters of the algorithm.

In the code below, we add all resampling iterations for the six tasks as experiments.
By leaving `prob.designs` unspecified, experiments for all existing problems are created per default.
We set the algorithm parameters to all possible `learners`, i.e. the logistic regression, random forest, and featureless learner from `large_design`.
Note that whenever an experiment is added, the current seed is assigned to the experiment and then incremented.

```{r large_benchmarking-049}
#| cache: false
#| output: false
library("data.table")

algorithm_design = list(run_learner = data.table(learner = learners))
print(algorithm_design$run_learner)

addExperiments(algo.designs = algorithm_design, reg = reg)
```

We confirm that the algorithm, problems, and experiments (jobs) were added successfully.

```{r large_benchmarking-050}
summarizeExperiments()
```

@fig-batchtools-illustration summarizes the interplay between the batchtools problems, algorithms, and experiments.

```{r large_benchmarking-051}
#| label: fig-batchtools-illustration
#| fig-cap: "Illustration the batchtools problem, algorithm, and experiment. "
#| fig-align: "center"
#| fig-alt: "A problem consists of a static data part and applies the problem function to this data part (and potentially problem parameters) to return a problem instance. The algorithm function takes in a problem instance (and potentially algorithm parameters), executes one job and returns its result."
#| echo: false
include_multi_graphics("Figures/mlr3book_figures-31.svg", "Figures/mlr3book_figures-31.png")
```


We are now ready to submit the jobs to the cluster.
By specifying no job IDs, all experiments are submitted as independent jobs, i.e. one computational job executes one resample experiment.

```{r}
#| output: false
#| cache: false
submitJobs(reg = reg)
waitForJobs(reg = reg)
```

When a cluster jobs finishes, it stores its return value in the registry folder.
We can retrieve the job results using `r ref("batchtools::loadResult()")`.
It outputs the objects returned by the algorithm function, which in our case is a `ResampleResult`.

```{r large_benchmarking-054}
rr = loadResult(1, reg = reg)
rr
```

In order to use mlr3's post-processing tools, we need to convert all results into a `BenchmarkResult`.
We can do this, by combining all resample results using `r ref("batchtools::reduceResults()")`.

```{r large_benchmarking-055}
bmr = reduceResults(c, reg = reg)
bmr$aggregate()
```

While we took a different route than in @sec-mlr3batchmark to define the experiments and ran them at a different granularity, we arrived at the same result.


## Statistical Analysis {#sec-benchmark-analysis}

Once we successfully executed the benchmark experiment, we can proceed with its analysis.
The package `r ref("mlr3benchmark")` provides infrastructure for applying statistical significance tests on `r ref("BenchmarkResult")` objects.
Currently, Friedman tests and pairwise Friedman-Nemenyi tests [@demsar2006] are supported to analyze benchmark experiments with at least two independent tasks and at least two learners.
Before we can use these methods, we have to convert the benchmark result to a `r ref("mlr3benchmark::BenchmarkAggr")` using `r ref("as_benchmark_aggr()")`.
We can then perform a pairwise comparison using `$friedman_posthoc()`.
This method will first perform a global friedman test and only conduct the post-hoc tests if the former is significant.

```{r large_benchmarking-056}
library("mlr3benchmark")
bma = as_benchmark_aggr(bmr, measures = msr("classif.ce"))
bma$friedman_posthoc()
```

These results would indicate a statistically significant difference between the `"featureless"` learner and `"ranger"`, assuming a 95% confidence level.
This table can be summarized in a critical difference plot, which typically shows the mean rank of a learning algorithm on the x-axis along with a thick horizontal line that connects learners which are pairwise not significantly different (while correcting for multiple tests):

```{r large_benchmarking-057}
autoplot(bma, type = "cd")
```

While our experiment did now show a significant difference between the random forest and the logistic regression (they are connected in the plot), the former has a lower rank on average.
This is in line with the large benchmark study conducted by @couronne2018random, where the random forest outperformed the logistic regression in 69% of 243 real world datasets.

As a final note, it is important to be careful when interpreting such test results.
Because our datasets are not an iid sample from a population of datasets, we can only make inference about the data generating processes at hand, i.e. those that generated the datasets we used in the benchmark.

## Conclusion

In this chapter, we have explored how to conduct large scale machine learning experiments using `mlr3`.
We have shown how to acquire diverse datasets from OpenML through the `r mlr3oml` interface package.
Furthermore, we have learned how to execute large-scale experiments using the `r ref_pkg("batchtools")` package and its `r mlr3batchmark` integration.
Finally, we have demonstrated how to analyze the results using the `r mlr3benchmark` package, thereby extracting meaningful insights from the experiments.

The most important functions and classes we learned about are in @tbl-api-large-benchmarking alongside their R6 classes (if applicable).

| S3 function                         | R6 Class                   | Summary                                                             |
| ------------------------------------| ---------------------------| --------------------------------------------------------------------|
| `r ref("odt()")`                    | `r ref("OMLData")`         | Retrieve an OpenML Dataset                                          |
| `r ref("otsk()")`                   | `r ref("OMLTask")`         | Retrieve an OpenML Task                                             |
| `r ref("ocl()")`                    | `r ref("OMLCollection")`   | Retrieve an OpenML Collection                                       |
| `r ref("list_oml_data()")`          |-                           | Filter OpenML Datasets                                              |
| `r ref("list_oml_tasks()")`         |-                           | Filter OpenML Tasks                                                 |
| `r ref("makeExperimentRegistry()")` |-                           | Create a new registry                                               |
| `r ref("loadRegistry()")`           |-                           | Load an existing registry                                           |
| `r ref("saveRegistry()")`           |-                           | Save an existing registry                                           |
| `r ref("batchmark()")`              |-                           | Register problems, algorithms, and experiments from a design        |
| `r ref("addProblem()")`             |-                           | Register a new Problem                                              |
| `r ref("addAlgorithm()")`           |-                           | Register a  new algorithm                                           |
| `r ref("addExperiments()")`         |-                           | Register experiments using existing algorithms and problems         |
| `r ref("submitJobs()")`             |-                           | Submit jobs to the scheduler                                        |
| `r ref("getJobTable()")`            |-                           | Get an overview of all job definitions                              |
| `r ref("unwrap()")`                 |-                           | Get an overview of all job definitions                              |
| `r ref("getStatus()")`              |-                           | Get the status of the computation                                   |
| `r ref("reduceResultsBatchmark()")` |-                           | Load finished jobs as a benchmark result                            |
| `r ref("reduceResults()")`          |-                           | Combine experiment results                                          |
| `r ref("findExperiments()")`        |-                           | Find specific experiments                                           |
| `r ref("grepLogs()")`               |-                           | Search the log files                                                |
| `r ref("summarizeExperiments()")`   |-                           | Summarize defined experiments                                       |
| `r ref("getLog()")`                 |-                           | Get a specific log file                                             |
| `r ref("findErrors()")`             |-                           | Find ids of failed jobs                                             |

:Core functions for Open in `mlr3` with the underlying R6 class that are constructed when these functions are called (if applicable) and a summary of the purpose of the functions. {#tbl-api-large-benchmarking}

### Resources{.unnumbered .unlisted}

- Look at the short `r link("https://doi.org/10.21105/joss.00135", "batchtools paper")`, and please cite this if you use the package.
- Read the `r link("https://www.jstatsoft.org/v64/i11", "Paper on BatchJobs / BatchExperiments")` which are the batchtools predecessors, but the concepts still hold, and most examples work analogously.
- Learn more in the `r link("https://mllg.github.io/batchtools/articles/batchtools.html", "batchtools vignette")`.
- Explore the `r link("https://docs.openml.org/", "OpenML documentation")`.

## Exercises

In this exercise we will conduct an empirical study that compares two machine learning algorithms.
Our null hypothesis is that a single regression tree performs equally well than a random forest.

### Getting Data from OpenML {.unnumbered .unlisted}

1. Load the OpenML collection with ID `r link("269", "https://www.openml.org/search?type=study&study_type=task&id=269")`. It contains regression tasks from the AutoML benchmark [@amlb2022].
2. Find all tasks with less than 4000 observations and convert them to `mlr3` tasks.
3. Create an experimental design that compares the random forest in `r ref_pkg("ranger")` with the regression tree from `r ref_pkg("rpart")` on those tasks.
   You can use 3-fold CV instead of the OpenML resamplings to save time.

### Executing the Experiments using batchtools {.unnumbered .unlisted}

1. Create a registry and populate it with the experiments.
1. (Optional) Change the cluster function to either "Socket" or "Multicore" (the latter does not work on Windows).
1. Submit the jobs and once they are finished, collect the results.

### Analyzing the Results {.unnumbered .unlisted}

1. Conduct a global Friedman test and interpret the results.
   As an evaluation measure, use the mean-square error.
1. Inspect the ranks of the results.

::: {.content-visible when-format="html"}
`r citeas(chapter)`
:::
