## Tasks {#tasks}

Tasks are objects that contain the (usually tabular) data and additional meta-data to define a machine learning problem.
The meta-data is, for example, the name of the target variable for supervised machine learning problems, or the type of the dataset (e.g. a _spatial_ or _survival_).
This information is used for specific operations that can be performed on a task.

### Task Types {#tasks-types}

To create a task from a `r ref("data.frame()")`, `r ref("data.table()")` or `r ref("Matrix::Matrix()", text = "Matrix()")`, you first need to select the right task type:

* **Classification Task**: The target is a label (stored as `character()`or`factor()`) with only few distinct values.
  <br/>→ `r ref("TaskClassif")`.

* **Regression Task**: The target is a numeric quantity (stored as `integer()` or `double()`).
  <br/>→ `r ref("TaskRegr")`.

* **Survival Task**: The target is the (right-censored) time to an event. More censoring types are currently in development.
  <br/>→ `r ref("mlr3proba::TaskSurv")` in add-on package `r mlr_pkg("mlr3proba")`.

* **Density Task**: An unsupervised task to estimate the density.
  <br/>→ `r ref("mlr3proba::TaskDens")` in add-on package `r mlr_pkg("mlr3proba")`.

* **Cluster Task**: An unsupervised task type; there is no target and the aim is to identify similar groups within the feature space.
  <br/>→ `r ref("mlr3cluster::TaskClust")` in add-on package `r mlr_pkg("mlr3cluster")`.

* **Spatial Task**: Observations in the task have spatio-temporal information (e.g. coordinates).
  <br/>→ `r ref("mlr3spatiotempcv::TaskRegrST")` or `r ref("mlr3spatiotempcv::TaskClassifST")` in add-on package `r mlr_pkg("mlr3spatiotempcv")`.

* **Ordinal Regression Task**: The target is ordinal.
  <br/>→ `TaskOrdinal` in add-on package `r gh_pkg("mlr-org/mlr3ordinal")` (still in development).

### Task Creation {#tasks-creation}

As an example, we will create a regression task using the `r ref("datasets::mtcars", text = "mtcars")` data set from the package `datasets` and predict the numeric target variable `"mpg"` (miles per gallon).
We only consider the first two features in the dataset for brevity.

First, we load and prepare the data.

```{r 02-basics-tasks-001}
data("mtcars", package = "datasets")
data = mtcars[, 1:3]
str(data)
```

Next, we create a regression task, i.e. we construct a new instance of the R6 class `r ref("TaskRegr")`.
Usually, this is done by calling the constructor`TaskRegr$new()`.
Instead, we are calling the converter `r ref("as_task_regr()")` to convert our `data.frame()` stored as `data` to a task and provide the following information:

2. `x`: Object to convert. Works for `data.frame()`/`data.table()`/`tibble()` abstract data backends implemented in the class `r ref("DataBackendDataTable")`.
   The latter allows to connect to out-of-memory storage systems like SQL servers via the extension package `r mlr_pkg("mlr3db")`.
3. `target`: The name of the target column for the regression problem.
1. `id` (optional): An arbitrary identifier for the task, used in plots and summaries.
   If not provided, the deparsed and substituted name of `x` will be used.

```{r 02-basics-tasks-002}
library("mlr3")

task_mtcars = as_task_regr(data, target = "mpg", id = "cars")
print(task_mtcars)
```

The `print()` method gives a short summary of the task:
It has `r task_mtcars$nrow` observations and `r task_mtcars$ncol` columns, of which `r length(task_mtcars$feature_names)` are features.

We can also plot the task using the `r mlr_pkg("mlr3viz")` package, which gives a graphical summary of its properties:

```{r 02-basics-tasks-003, message=FALSE}
library("mlr3viz")
autoplot(task_mtcars, type = "pairs")
```

Note that instead of loading all the extension packages individually, it is often more convenient to load the `r mlr_pkg("mlr3verse")` package instead.
`mlr3verse` imports most `mlr3` packages and re-exports functions which are used for regular machine learning and data science tasks.


### Predefined tasks {#tasks-predefined}

`r mlr_pkg("mlr3")` ships with a few predefined machine learning tasks.
All tasks are stored in an R6 `r ref("Dictionary")` (a key-value store) named `r ref("mlr_tasks")`.
Printing it gives the keys (the names of the datasets):

```{r 02-basics-tasks-004}
mlr_tasks
```

We can get a more informative summary of the example tasks by converting the dictionary to a `r ref("data.table()")` object:

```{r 02-basics-tasks-005}
as.data.table(mlr_tasks)
```
In the above display, the columns `"lgl"` (`r ref("logical")`), `"int"` (`r ref("integer")`), `"dbl"` (`r ref("double")`), `"chr"` (`r ref("character")`), `"fct"` (`r ref("factor")`), `"ord"` (`r ref("ordered", text = "ordered factor")`) and `"pxc"` (`r ref("POSIXct")` time) display the number of features in the dataset with the corresponding storage type.

To get a task from the dictionary, one can use the `$get()` method from the `mlr_tasks` class and assign the return value to a new object.
Since `mlr3` arranges most of its object instances in dictionaries and extraction is such a common task, there is a shortcut for this: the function `r ref("tsk()")`.
Here, we retrieve the `r ref("mlr_tasks_penguins", text = "palmer penguins task")` originating from the package `r cran_pkg("palmerpenguins")`:

```{r 02-basics-tasks-006}
task_penguins = tsk("penguins")
print(task_penguins)
```

Note that dictionaries such as `r ref("mlr_tasks")` can get populated by extension packages.
E.g., `r mlr_pkg("mlr3data")` comes with some more example and toy tasks for regression and classification, and `r mlr_pkg("mlr3proba")` ships with additional survival and density estimation tasks.
Both packages will get loaded once we load the `r mlr_pkg("mlr3verse")` package, so we do it here and have a look at the available tasks again:
```{r 02-basics-tasks-007}
library(mlr3verse)
as.data.table(mlr_tasks)[, 1:4]
```
To get more information about the respective task, the corresponding man page can be found under `mlr_tasks_[id]`, e.g. `r ref("mlr_tasks_german_credit")`.



### Task API {#tasks-api}

All task properties and characteristics can be queried using the task's public fields and methods (see `r ref("Task")`).
Methods can also be used to change the stored data and the behavior of the task.

#### Retrieving Data {#tasks-retrieving}

The data stored in a task can be retrieved directly from fields, for example:

```{r 02-basics-tasks-008}
task_mtcars
task_mtcars$nrow
task_mtcars$ncol
```

More information can be obtained through methods of the object, for example:

```{r 02-basics-tasks-009}
task_mtcars$data()
```

In `r mlr_pkg("mlr3")`, each row (observation) has a unique identifier, stored as an `integer()`.
These can be passed as arguments to the `$data()` method to select specific rows:

```{r 02-basics-tasks-010}
head(task_mtcars$row_ids)

# retrieve data for rows with ids 1, 5, and 10
task_mtcars$data(rows = c(1, 5, 10))
```

Note that although the row ids are typically just the sequence from `1` to `nrow(data)`, they are only guaranteed to be unique natural numbers.
Keep that in mind, especially if you work with data stored in a real data base management system (see [backends](#backends)).

Similarly to row ids, target and feature columns also have unique identifiers, i.e. names (stored as `character()`).
Their names can be accessed via the public slots `$feature_names` and `$target_names`.
Here, "target" refers to the variable we want to predict and "feature" to the predictors for the task.

```{r 02-basics-tasks-011}
task_mtcars$feature_names
task_mtcars$target_names
```

The `row_ids` and column names can be combined when selecting a subset of the data:

```{r 02-basics-tasks-012}
# retrieve data for rows 1, 5, and 10 and only select column "mpg"
task_mtcars$data(rows = c(1, 5, 10), cols = "mpg")
```

To extract the complete data from the task, one can also simply convert it to a `data.table`:

```{r 02-basics-tasks-013}
summary(as.data.table(task_mtcars))
```

#### Roles (Rows and Columns) {#tasks-roles}

It is possible to assign different roles to rows and columns.
These roles affect the behavior of the task for different operations.
We already seen this for the target and feature columns which serve a different purpose.

For example, the previously-constructed _mtcars_ task has the following column roles:

```{r 02-basics-tasks-014}
print(task_mtcars$col_roles)
```
Columns can also have no role (they are ignored) or have multiple roles.
To add the row names of `mtcars` as an additional feature, we first add them to the data table as regular column and then recreate the task with the new column.

```{r 02-basics-tasks-015}
# with `keep.rownames`, data.table stores the row names in an extra column "rn"
data = as.data.table(datasets::mtcars[, 1:3], keep.rownames = TRUE)
task_mtcars = as_task_regr(data, target = "mpg", id = "cars")

# there is a new feature called "rn"
task_mtcars$feature_names
```

The row names are now a feature whose values are stored in the column `"rn"`.
We include this column here for educational purposes only.
Generally speaking, there is no point in having a feature that uniquely identifies each row.
Furthermore, the character data type will cause problems with many types of machine learning algorithms.

On the other hand, the identifier may be useful to label points in plots, for example to identify and label outliers.
Therefore we will change the role of the `rn` column by removing it from the list of features and assign the new role `"name"`.
There are two ways to do this:

1. Use the `r ref("Task")` method `$set_col_roles()` (recommended).
2. Simply modify the field `$col_roles`, which is a named list of vectors of column names.
   Each vector in this list corresponds to a column role, and the column names contained in that vector are designated as having that role.

Supported column roles can be found in the manual of `r ref("Task")`, or just by printing the names of the field `$col_roles`:.

```{r 02-basics-tasks-016}
# supported column roles, see ?Task
names(task_mtcars$col_roles)

# assign column "rn" the role "name", remove from other roles
task_mtcars$set_col_roles("rn", roles = "name")

# note that "rn" not listed as feature anymore
task_mtcars$feature_names

# "rn" also does not appear anymore when we access the data
task_mtcars$data(rows = 1:2)
task_mtcars$head(2)
```

Changing the role does not change the underlying data, it just updates the view on it.
The data is not copied in the code above.
The view is changed in-place though, i.e. the task object itself is modified.

Just like columns, it is also possible to assign different roles to rows.

Rows can have two different roles:

1. Role `use`:
  Rows that are generally available for model fitting (although they may also be used as test set in resampling).
  This role is the default role.

2. Role `validation`:
  Rows that are not used for training.
  Rows that have missing values in the target column during task creation are automatically set to the validation role.

There are several reasons to hold some observations back or treat them differently:

1. It is often good practice to validate the final model on an external validation set to identify possible overfitting.
2. Some observations may be unlabeled, e.g. in competitions like [Kaggle](https://www.kaggle.com/).

These observations cannot be used for training a model, but can be used to get predictions.

#### Task Mutators {#tasks-mutators}

As shown above, modifying `$col_roles` or `$row_roles` (either via `set_col_roles()`/`set_row_roles()` or directly by modifying the named list) changes the view on the data.
The additional convenience method `$filter()` subsets the current view based on row ids and `$select()` subsets the view based on feature names.

```{r 02-basics-tasks-017}
task_penguins = tsk("penguins")
task_penguins$select(c("body_mass", "flipper_length")) # keep only these features
task_penguins$filter(1:3) # keep only these rows
task_penguins$head()
```

While the methods discussed above allow to subset the data, the methods `$rbind()` and `$cbind()` allow to add extra rows and columns to a task.
Again, the original data is not changed.
The additional rows or columns are only added to the view of the data.

```{r 02-basics-tasks-018}
task_penguins$cbind(data.frame(letters = letters[1:3])) # add column foo
task_penguins$head()
```

### Plotting Tasks {#autoplot-task}

The `r mlr_pkg("mlr3viz")` package provides plotting facilities for many classes implemented in `r mlr_pkg("mlr3")`.
The available plot types depend on the inherited class, but all plots are returned as `r cran_pkg("ggplot2")` objects which can be easily customized.

For classification tasks (inheriting from `r ref("TaskClassif")`), see the documentation of `r ref("mlr3viz::autoplot.TaskClassif")` for the implemented plot types.
Here are some examples to get an impression:

```{r 02-basics-tasks-019, warning = FALSE, message = FALSE}
library("mlr3viz")

# get the pima indians task
task = tsk("pima")

# subset task to only use the 3 first features
task$select(head(task$feature_names, 3))

# default plot: class frequencies
autoplot(task)

# pairs plot (requires package GGally)
autoplot(task, type = "pairs")

# duo plot (requires package GGally)
autoplot(task, type = "duo")
```

Of course, you can do the same for regression tasks (inheriting from `r ref("TaskRegr")`) as documented in `r ref("mlr3viz::autoplot.TaskRegr")`:

```{r 02-basics-tasks-020, warning = FALSE, message = FALSE}
library("mlr3viz")

# get the complete mtcars task
task = tsk("mtcars")

# subset task to only use the 3 first features
task$select(head(task$feature_names, 3))

# default plot: boxplot of target variable
autoplot(task)

# pairs plot (requires package GGally)
autoplot(task, type = "pairs")
```
