## Benchmarking {#benchmarking}

Comparing the performance of different learners on multiple tasks and/or different resampling schemes is a common task.
This operation is usually referred to as "benchmarking" in the field of machine-learning.
The `r mlr_pkg("mlr3")` package offers the `r ref("benchmark()")` convenience function.

### Design Creation {#bm-design}

In `mlr3` we require you to supply a "design" of your benchmark experiment.
Such a design is essentially a table of settings you want to execute.
It consists of unique combinations of `r ref("Task")`, `r ref("Learner")` and `r ref("Resampling")` triplets.

We use the `r ref("benchmark_grid()")` function to create an exhaustive design and instantiate the resampling properly, so that all learners are executed on the same train/test split for each tasks.
We set the learners to predict probabilities and also tell them to predict the observations of the training set (by setting `predict_sets` to `c("train", "test")`).
Additionally, we use `r ref("tsks()")`, `r ref("lrns()")`, and `r ref("rsmps()")` to retrieve lists of `r ref("Task")`, `r ref("Learner")` and `r ref("Resampling")` in the same fashion as `r ref("tsk()")`, `r ref("lrn()")` and `r ref("rsmp()")`.

```{r 02-basics-benchmarking-001}
library(mlr3verse)

design = benchmark_grid(
  tasks = tsks(c("spam", "german_credit", "sonar")),
  learners = lrns(c("classif.ranger", "classif.rpart", "classif.featureless"),
    predict_type = "prob", predict_sets = c("train", "test")),
  resamplings = rsmps("cv", folds = 3)
)
print(design)
```

The created `design` can be passed to `r ref("benchmark()")` to start the computation.
It is also possible to create a custom design manually.
However, if you create a custom task with `data.table()`, the train/test splits will be different for each row of the design if you do not [**manually instantiate**](#resampling-inst) the resampling before creating the design.
See the help page on `r ref("benchmark_grid()")` for an example.


### Execution and Aggregation of Results {#bm-exec}

After the [benchmark design](#bm-design) is ready, we can directly call `r ref("benchmark()")`:

```{r 02-basics-benchmarking-003}
# execute the benchmark
bmr = benchmark(design)
```

Note that we did not instantiate the resampling instance manually.
`r ref("benchmark_grid()")` took care of it for us:
Each resampling strategy is instantiated once for each task during the construction of the exhaustive grid.

Once the benchmarking is done, we can aggregate the performance with `$aggregate()`.
We create two measures to calculate the AUC for the training set and for the predict set:

```{r 02-basics-benchmarking-004}
measures = list(
  msr("classif.auc", predict_sets = "train", id = "auc_train"),
  msr("classif.auc", id = "auc_test")
)

tab = bmr$aggregate(measures)
print(tab)
```

We can aggregate the results even further.
For example, we might be interested to know which learner performed best over all tasks simultaneously.
Simply aggregating the performances with the mean is usually not statistically sound.
Instead, we calculate the rank statistic for each learner grouped by task.
Then the calculated ranks grouped by learner are aggregated with `r cran_pkg("data.table")`.
Since the AUC needs to be maximized, we multiply the values by $-1$ so that the best learner has a rank of $1$.

```{r 02-basics-benchmarking-005}
library(data.table)
# group by levels of task_id, return columns:
# - learner_id
# - rank of col '-auc_train' (per level of learner_id)
# - rank of col '-auc_test' (per level of learner_id)
ranks = tab[, .(learner_id, rank_train = rank(-auc_train), rank_test = rank(-auc_test)), by = task_id]
print(ranks)

# group by levels of learner_id, return columns:
# - mean rank of col 'rank_train' (per level of learner_id)
# - mean rank of col 'rank_test' (per level of learner_id)
ranks = ranks[, .(mrank_train = mean(rank_train), mrank_test = mean(rank_test)), by = learner_id]

# print the final table, ordered by mean rank of AUC test
ranks[order(mrank_test)]
```

Unsurprisingly, the featureless learner is outperformed on both training and test set.
The classification forest also outperforms a single classification tree.


### Plotting Benchmark Results {#autoplot-benchmarkresult}

Analogously to plotting [tasks](#autoplot-task), [predictions](#autoplot-prediction) or [resample results](#autoplot-resampleresult), `r mlr_pkg("mlr3viz")` also provides a `r ref("ggplot2::autoplot()", text = "autoplot()")` method for benchmark results.

```{r 02-basics-benchmarking-006}
autoplot(bmr) + ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

We can also plot ROC curves.
To do so, we first need to filter the `r ref("BenchmarkResult")` to only contain a single `r ref("Task")`:

```{r 02-basics-benchmarking-007}
bmr_small = bmr$clone()$filter(task_id = "german_credit")
autoplot(bmr_small, type = "roc")
```

All available plot types are listed on the manual page of `r ref("autoplot.BenchmarkResult()")`.

### Extracting ResampleResults {#bm-resamp}

A `r ref("BenchmarkResult")` object is essentially a collection of multiple `r ref("ResampleResult")` objects.
As these are stored in a column of the aggregated `data.table()`, we can easily extract them:

```{r 02-basics-benchmarking-008}
tab = bmr$aggregate(measures)
rr = tab[task_id == "german_credit" & learner_id == "classif.ranger"]$resample_result[[1]]
print(rr)
```

We can now investigate this resampling and even single resampling iterations using one of the approaches shown in [the previous section](#bm-exec):

```{r 02-basics-benchmarking-009}
measure = msr("classif.auc")
rr$aggregate(measure)

# get the iteration with worst AUC
perf = rr$score(measure)
i = which.min(perf$classif.auc)

# get the corresponding learner and train set
print(rr$learners[[i]])
head(rr$resampling$train_set(i))
```

### Converting and Merging

A `r ref("ResampleResult")` can be casted to a `r ref("BenchmarkResult")` using the converter `r ref("as_benchmark_result()")`.
Additionally, two `r ref("BenchmarkResult", text = "BenchmarkResults")` can be merged into a larger result object.


```{r 02-basics-benchmarking-010}
task = tsk("iris")
resampling = rsmp("holdout")$instantiate(task)

rr1 = resample(task, lrn("classif.rpart"), resampling)
rr2 = resample(task, lrn("classif.featureless"), resampling)

# Cast both ResampleResults to BenchmarkResults
bmr1 = as_benchmark_result(rr1)
bmr2 = as_benchmark_result(rr2)

# Merge 2nd BMR into the first BMR
bmr1$combine(bmr2)

bmr1
```
