# Pipelines {#pipelines}

```{r vis_setup, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(visNetwork)
library(mlr3pipelines)
```

`r mlr_pkg("mlr3pipelines")` is a dataflow programming toolkit for machine learning in R utilizing the `r mlr_pkg("mlr3pipelines")` package. 
Machine learning workflows can be written as directed “Graphs” that represent data flows between preprocessing, model fitting, and ensemble learning units in an expressive and intuitive language. 
Using methods from the `r mlr_pkg("mlr3tuning")`, it is even possible to simultaneously optimize parameters of multiple processing units.

## The Building Blocks: POs

The building blocks of `r mlr_pkg("mlr3pipelines")` are **PipeOp**-objects (PO). 
They can be constructed directly using `PipeOp<NAME>$new()`, but the recommended way is to retrieve them from the **`mlr_pipeops`** dictionary. 
See `as.data.table(mlr_pipeops)` for a list and `?mlr_pipeops` for more details.

```{r 03-pipelines-1}
library("mlr3pipelines")
as.data.table(mlr_pipeops)
```

## The Pipeline Operator {#pipe-operator}

Single POs can be created using `mlr_pipeops$get(<name>)`.

```{r 03-pipelines-2}
pca = mlr_pipeops$get("pca")
```

Some POs require additional arguments for construction.  

```{r eval = FALSE}
learner = mlr_pipeops$get("learner")

# Error in cast_from_dict(learner, "Learner", mlr_learners, clone, FALSE) : argument "learner" is missing, with no default
```

```{r 03-pipelines-3}
learner = mlr_pipeops$get("learner",
  learner = mlr3learners::LearnerRegrRanger$new())
```

Hyperparameters of POs can be set through the `param_vals` argument. 
Here we set the fraction of features for a filter.

```{r 03-pipelines-4}
filter = mlr_pipeops$get("filter",
  filter = mlr3featsel::FilterVariance$new(),
  param_vals = list(filter.frac = 0.5))
```

## Nodes, Edges and Graphs {#pipe-nodes-edges-graphs}

POs are combined into `Graph`s. 
The manual way (= hard way) to construct a `Graph` is to create an empty one first, fill it with POs, and then connect **edges** between the POs.
POs are identified by their **`$id`**. 
Note that the operations all modify the object in-place and return the object itself, so multiple modifications can be chained.

For this example we use the "pca" PO defined above and a new PO named "mutate".
The latter creates a new feature from existing variables.

```{r 03-pipelines-5}
mutate = mlr_pipeops$get("mutate")
```

```{r 03-pipelines-6}
graph = Graph$new()$
  add_pipeop(mutate)$
  add_pipeop(filter)$
  add_edge("mutate", "variance")  # add connection mutate -> filter
```

The much quicker way is to use the `%>>%` operator to chain POs or `Graph`s. 
The same result as above can be achieved by doing

```{r 03-pipelines-7}
graph = mutate %>>% filter
```

Now the `Graph` can be inspected using its `$plot()` function:

```{r 03-pipelines-8}
#graph$plot(html = TRUE)
```

**Chaining multiple POs of the same kind**

If multiple POs of the same kind should be chained it is necessary to change the id to avoid name clashes.
This can be done by either accessing the `$id` slot or during construction.

Here we tr is used on data both for processing and to build an internal model of the data (which can be inspected using `graph$pipeops$<PipeOp Name>$state`). 
This model is used by the `$predict()` functions. 
For example when doing feature filtering during model training, it is necessary to remove the same features from new prediction data. 
Most POs handle data in the `mlr3` `Task` format, so the argument to `$train()` and `$predict()` is often a `Task`. 
The output is always a list with as many elements as there are terminal nodes in the `Graph` (usually just one).

```{r 03-pipelines-9}
task = mlr_tasks$get("iris")  # the "iris" task
task_processed = graph$train(task)[[1]]
```

### Imputation

## Modelling {#pipe-modelling}

The main purpose of `Graph`s is to build combined preprocessing and model fitting pipelines that can be used as `mlr3` `Learner`s. 
In the following we chain two preprocessing tasks 

- mutate (creation of a new feature)
- filter (filtering the dataset)

and then chain a PO learner to train and predict on the modified dataset.

```{r 03-pipelines-10}
graph = mutate %>>% filter %>>%
  mlr_pipeops$get("learner",
    learner = mlr_learners$get("classif.rpart"))
```

Until here we defined the main pipeline stored in `graph`.
Now we can train and predict the pipeline.

```{r 03-pipelines-11}
graph$train(task)
graph$predict(task)
```

Rather than calling `$train()` and `$predict()` manually, we can put the pipeline `graph` into a `GraphLearner` object.
A `GraphLearner` encapsulates the whole pipeline (including the preprocessing steps) and can be put into `resample()` or `benchmark()`.
If you are familiar with the old _mlr_ package, this is the equivalent of all the `make*Wrapper()` functions.
The pipeline being encapsulated (here `graph`) must always produce a `Prediction` with its `$predict()` call, so it will probably contain at least one `PipeOpLearner`.

```{r 03-pipelines-12}
glrn = GraphLearner$new(graph)
```

This learner can be used for model fitting, resampling, benchmarking, and tuning.

```{r 03-pipelines-13}
cv10 = mlr_resamplings$get("cv")
resample(task, glrn, cv10)
```

### Setting hyperparameters {#pipe-hyperpars}

Individual POs offer hyperparameters because they contain `$param_set` slots that can be read and written from `$param_set$values` (via the paradox package). 
The parameters get passed down to the `Graph`, and finally to the `GraphLearner`.
This makes it not only possible to easily change change the behavior of a `Graph` / `GraphLearner` and try different settings manually, but also to perform tuning using the mlr3tuning package.

```{r 03-pipelines-14}
glrn$param_set$values$variance.filter.frac = 0.25
resample(task, glrn, cv10)
```

If you are unfamiliar with tuning in mlr3 yet, we recommend to take a look at the section about [tuning](#tuning) first.
Here we define a ParamSet for the "rpart" learner and the "variance" filter which should be optimized during tuning.

```{r 03-pipelines-15}
library("paradox")
ps = ParamSet$new(list(
  ParamDbl$new("classif.rpart.cp", lower = 0, upper = 0.05),
  ParamDbl$new("variance.filter.frac", lower = 0.25, upper = 1)
))
```

After having defined the `PerformanceEvaluator`, a random search with 10 iterations is created.

```{r 03-pipelines-16}
library("mlr3tuning")
pe = PerformanceEvaluator$new(task, glrn, cv10, "classif.ce", ps)
tuner = TunerRandomSearch$new(pe, TerminatorEvaluations$new(10))
tuner$tune()
```

The tuning result can be inspected using the `$tune_result()` method.

```{r 03-pipelines-18}
tuner$tune_result()
```

## Non-linear Graphs {#pipe-nonlinear}

The Graphs seen so far all have a linear structure. 
Some POs may have multiple input or output channels.
These make it possible to create non-linear Graphs with alternative paths taken by the data.

Possible types are 

- [Branching](#pipe-model-ensembles-branching): 
  Splitting of a node into several paths, useful for example when comparing multiple feature-selection methods (pca, filters). 
  Only one path will be executed.
- [Copying](#pipe-model-ensembles-copying): 
  Splitting of a node into several paths, all paths will be executed (sequentially). 
  Parallel execution is not yet supported.
- [Stacking](#pipe-model-ensembles-stacking): 
  Single graphs are stacked onto each other, i.e. the output of one Graph is the input for another. 
  In machine learning this means that the prediction of one Graph is used as input for another Graph.
 
### Branching & Copying {#pipe-model-ensembles-branching-copying}

The `PipeOpBranch` and `PipeOpUnbranch` POs make it possible to specify multiple alternative paths.
Only one is actually executed, the others are ignored. 
The active path is determined by a hyperparameter. 
This concept makes it possible to tune alternative preprocessing paths (or learner models).

`PipeOp(Un)Branch` is initialized either with the number of branches, or with a `character`-vector indicating the names of the branches. 
If names are given, the "branch-choosing" hyperparameter becomes more readable.
In the following, we set three options

1. Doing nothing ("null")
1. Applying a PCA
1. Scaling the data

It is important to "unbranch" again after "branching", so that the outputs are merged into one result objects.

In the following we first create the branched graph and then show what happens if the "unbranching" is not applied.

```{r}
graph = mlr_pipeops$get("branch", c("null", "pca", "scale")) %>>%
  gunion(list(
      mlr_pipeops$get("null", id = "null1"),
      mlr_pipeops$get("pca"),
      mlr_pipeops$get("scale")
  ))
```

Without "unbranching":

```{r}
graph$plot()
```

With "unbranching":

```{r}
(graph %>>%
  mlr_pipeops$get("unbranch", c("null", "pca", "scale")))$plot()
```

### Model Ensembles {#pipe-model-ensembles}

We can leverage the different operations presented to connect POs to form powerful graphs.

Before we go into details, we split the task into train
and test indices.

```{r}
task = mlr_tasks$get("iris")
train.idx = sample(seq_len(task$nrow), 120)
test.idx = setdiff(seq_len(task$nrow), train.idx)
```

#### Bagging {#pipe-model-ensembles-bagging}

We first examine Bagging introduced by Breiman (1994).
The basic idea is to create multiple predictors and then aggregate those to a single, more powerful predictor.

> "... multiple versions are formed
> by making bootstrap replicates of the learning set
> and using these as new learning sets" (Breimann 1994)

Bagging then aggregates a set of predictors by averaging (regression) or majority vote (classification).
The idea behind bagging is, that a set of weak, but different predictors can be combined in order to arrive at a single, better predictor.

We can achieve this by downsampling our data before training a learner, repeating this for say $10$ times and then performing a majority vote on the predictions.

First, we create a simple pipeline, that uses `PipeOpSubsample` before a `PipeOpLearner` is trained:

```{r}
single_pred = PipeOpSubsample$new(param_vals = list(frac = 0.7)) %>>%
  PipeOpLearner$new(mlr_learners$get("classif.rpart"))
```

We can now copy this operation $10$ times using `greplicate`.

```{r}
pred_set = greplicate(single_pred, 10L)
```

Afterwards we need to aggregate the 10 pipelines
to form a single model:

```{r}
bagging = pred_set %>>%
  PipeOpMajorityVote$new(innum = 10L)
```

and plot again to see what happens:

```{r, fig.width=7.5}
bagging$plot(html = TRUE)
```

This pipeline can again be used in conjunction with
`GraphLearner` in order for Bagging to be used like a
[mlr3::Learner].

```{r}
baglrn = GraphLearner$new(bagging)
baglrn$train(task, train.idx)
baglrn$predict(task, test.idx)
```

In conjunction with different `Backends`, this can be a very powerful tool.
In cases when the data does not fully fit in memory, we can obtain a fraction of the data for each learner from a `DataBaseBackend` and then aggregate predictions over all learners.

#### Stacking {#pipe-model-ensembles-stacking}

Stacking (Wolpert, 1992) is another technique that can improve model performance. 
The basic idea behind stacking is the use of predictions from one model as features for a subsequent model to possibly improve performance. 

As an example we can train a decision tree and use the predictions from this model in conjunction with the original features in order to train an additional model on top.

In order to limit overfitting, we additionally do not predict on the original predictions of the learner, but instead on out-of-bag predictions. 
To do all this, we can use `PipeOpLearnerCV`.

`PipeOpLearnerCV` performs nested cross-validation on the training data, fitting a model in each fold. 
Each of the models is then used to predict on the out-of-fold data. 
As a result, we obtain predictions for every data point in our input data.

We first create a "level 0" learner, which is used to extract a lower level prediction.
We additionally `clone()` the learner object to obtain a copy of the learner, and set a custom id for the `PipeOp`.

```{r}
lrn = mlr_learners$get("classif.rpart")
lrn_0 = PipeOpLearnerCV$new(lrn$clone())
lrn_0$id = "rpart_cv"
```

Additionally, we use a `PipeOpNULL` in parallel to the "level 0" learner, in order to send the unchanged Task to the next level, where it is then combined with the predictions from our decision tree learner.

```{r}
level_0 = gunion(list(lrn_0, PipeOpNULL$new()))
```

Afterwards, we want to concatenate the predictions from `PipeOpLearnerCV` and the original Task using `PipeOpFeatureUnion`.

```{r}
combined = level_0 %>>% PipeOpFeatureUnion$new(2)
```

We can now train another learner on top of the combined features.

```{r, fig.width=7.5}
stack = combined %>>% PipeOpLearner$new(lrn$clone())
stack$plot(html = TRUE)
```

```{r}
stacklrn = GraphLearner$new(stack)
stacklrn$train(task, train.idx)
stacklrn$predict(task, test.idx)
```

In this vignette, we showed a very simple usecase for stacking.
In many real-world applications, stacking is done for multiple levels and on multiple representations of the dataset. 
On a lower level, different preprocessing methods can be defined in conjunction with several learners. 
On a higher level, we can then combine those predictions in order to form a very powerful model.

#### Multilevel Stacking

In order to showcase the power of `mlr3pipelines`, we will show a more complicated stacking example.

In this case, we train a _glmnet_ and 2 different _rpart_ models (some transform its inputs using `PipeOpPCA`) on our task in the "level 0" and concatenate them with the original features (via `PipeOpNull`).
The result is then passed on to "level 1", where we copy the concatenated features 3 times and put this task into a _rpart_ and a _glmnet_ model.
Additionally, we keep a version of the "level 0" output (via `PipeOpNull`) and pass this on to "level 2". 
In "level 2" we simply concatenate all "level 1" outputs and train a final decision tree.

```{r, eval = FALSE}
rprt = mlr_learners$get("classif.rpart")
rprt$predict_type = "prob"
glmn = mlr_learners$get("classif.glmnet")
glmn$predict_type = "prob"

#  Create Learner CV Operators
lrn_0 = PipeOpLearnerCV$new(rprt, id = "rpart_cv_1")
lrn_0$values$maxdepth = 5L
lrn_1 = PipeOpPCA$new(id = "pca1") %>>% PipeOpLearnerCV$new(rprt, id = "rpart_cv_2")
lrn_1$values$maxdepth = 1L
lrn_2 = PipeOpPCA$new(id = "pca2") %>>% PipeOpLearnerCV$new(glmn)

# Union them with a PipeOpNULL to keep original features
level_0 = gunion(list(lrn_0, lrn_1,lrn_2, PipeOpNULL$new(id = "NULL1")))

# Cbind the output 3 times, train 2 learners but also keep level
# 0 predictions
level_1 = level_0 %>>%
  PipeOpFeatureUnion$new(4) %>>%
  PipeOpCopy$new(3) %>>%
  gunion(list(
    PipeOpLearnerCV$new(rprt, id = "rpart_cv_l1"),
    PipeOpLearnerCV$new(glmn, id = "glmnt_cv_l1"),
    PipeOpNULL$new(id = "NULL_l1")
  ))

# Cbind predicitions, train a final learner.
level_2 = level_1 %>>%
  PipeOpFeatureUnion$new(3, id = "u2") %>>%
  PipeOpLearner$new(rprt,
    id = "rpart_l2")

# Plot the resulting graph
level_2$plot(html = TRUE)

task = mlr_tasks$get("iris"),
lrn = GraphLearner$new(level_2)

lrn$
  train(task, train.idx)$
  predict(task, test.idx)$
  score()
```

## Tuning {#pipe-tuning}
