# Train & Predict {#train-predict}

In this chapter, we explain how [tasks and learners](#tasks-and-learners) can be used to train a model and predict to a new dataset.

The concept is demonstrated on a supervised classification using the iris dataset and the **rpart** learner (classification tree).

Additionally, this chapter includes the following use-cases

- Functional Data Analysis using <model name> (WIP)
- Regression Analysis using <model name> (WIP)
- Survival Analysis using <model name> (WIP)
- Spatial Analysis using <model name> (WIP)

## Basic concept

### Creating Task and Learner Objects

The first step is to generate the following `mlr3` objects from the [task dictionary](#tasks) and the [learner dictionary](#learners), respectively:

1. The classification task
    ```{r 03-train-predict-001}
    task = mlr_tasks$get("iris")
    ```
2. A learner for the classification tree
    ```{r 03-train-predict-002}
    learner = mlr_learners$get("classif.rpart")
    ```

### Setting up the train/test splits of the data (#split-data)

It is common to train on a majority of the data.
Here we use 80% of all available observations and predict on the remaining 20% observations.
For this purpose, we create two index vectors:

```{r 03-train-predict-003}
train_set = sample(task$nrow, 0.8 * task$nrow)
test_set = setdiff(seq_len(task$nrow), train_set)
```

### Defining the "Experiment"

The process of fitting a machine-learning model, predicting on test data and scoring the predictions by comparing predicted and true labels is what we call an "experiment" in `mlr`.
We start by initializing a new `Experiment` object by passing the created `r ref("TaskClassif")` and `r ref("LearnerClassif")`:

```{r 03-train-predict-004}
e = Experiment$new(task = task, learner = learner)
print(e)
```

The output shows a summary of the state of the experiment.
We can see that the current state is "defined".
This means that the task and the learner have been stored, but nothing else happened so far.  
If we query the `state` slot specifically, the (ordered) factor levels of the output show us all other possible states of an experiment:

```{r 03-train-predict-005}
e$state
```

### Training

Now we can finally train the learner on the task by calling the `.$train()` function of the experiment `e` we created:

```{r 03-train-predict-006}
e$train(row_ids = train_set)
print(e)
```

The output indicates that the `Experiment` object was modified (its state is now `[trained]`).
Additionally, it was extended by a slot `model`:

```{r 03-train-predict-007}
print(e$model)
```

### Predicting

After the model was trained, we use the remaining part of the data for prediction.
Remember that we [initially split the data](#split-data) in `train_set` and `test_set`.

(Rather than using a subset of the initial data we could also pass a completetly new `data.frame` here.)

```{r 03-train-predict-008}
e$predict(row_ids = test_set)
print(e)
```

Now we gained a new slot named `Predictions` that we holds the prediction results.
Since we are using R6 objects in `mlr3`, the result is again directly stored in object `e`.

A preview of the generated predictions can be retrieved by querying the `prediction` slot.

```{r 03-train-predict-009}
e$prediction
```

If desired you can also extract the results to a normal `data.table` object:

```{r 03-train-predict-010}
e_dt = as.data.table(e$prediction)
head(e_dt)
```

#### Performance assessment

The last step of the experiment is quantifying the performance of the model by comparing the predicted labels with the true labels using a performance measure.
The default measure for the classification tasks is the "mean misclassification error".

```{r 03-train-predict-011}
task$measures
```

To conduct the performance assessment, we call the function `.$score()`.
We now get a new slot called `"performance"` in our object:

```{r 03-train-predict-012}
e$score()
```

If we now print the `Experiment` object `e` again, we see that it consists of all six slots which store all important information of the experiment:

```{r 03-train-predict-013}
print(e)
```

#### Chaining methods

The underlying R6 class structure makes it possible to chain all of the operations above (`$train()`, `$predict()` and `$score()`) in a single call:

```{r 03-train-predict-014}
Experiment$new(task = task, learner = learner)$train(train_set)$predict(test_set)$score()
```

## Use cases
