## Logging {#logging}

We use the `r cran_pkg("lgr")` package for logging and progress output.

### Changing mlr3 logging levels

To change the setting for `r mlr_pkg("mlr3")` for the current session, you need to retrieve the logger (which is a `r cran_pkg("R6")` object) from `r cran_pkg("lgr")`, and then change the threshold of the like this:

```{r 05-technical-logging-001, eval = FALSE}
requireNamespace("lgr")

logger = lgr::get_logger("mlr3")
logger$set_threshold("<level>")
```

The default log level is `"info"`.
All available levels can be listed as follows:

```{r 05-technical-logging-002}
getOption("lgr.log_levels")
```

To increase verbosity, set the log level to a higher value, e.g. to `"debug"` with:
```{r 05-technical-logging-003, eval = FALSE}
lgr::get_logger("mlr3")$set_threshold("debug")
```

To reduce the verbosity, reduce the log level to warn:

```{r 05-technical-logging-004, eval = FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
```

`r cran_pkg("lgr")` comes with a global option called `"lgr.default_threshold"` which can be set via `options()` to make your choice permanent across sessions.

Also note that the optimization packages such as `r mlr_pkg("mlr3tuning")`  `r mlr_pkg("mlr3fselect")` use the logger of their base package `r mlr_pkg("bbotk")`.
To disable the output from `r mlr_pkg("mlr3")`, but keep the output from `r mlr_pkg("mlr3tuning")`, reduce the verbosity for the logger `r mlr_pkg("mlr3")`
and optionally change the logger `r mlr_pkg("bbotk")` to the desired level.

```{r 05-technical-logging-005, eval=FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("info")
```

### Redirecting output

Redirecting output is already extensively covered in the documentation and vignette of `r cran_pkg("lgr")`.
Here is just a short example which adds an additional appender to log events into a temporary file in [JSON](https://en.wikipedia.org/wiki/JSON) format:
```{r 05-technical-logging-006, eval = knitr::is_html_output()}
tf = tempfile("mlr3log_", fileext = ".json")

# get the logger as R6 object
logger = lgr::get_logger("mlr")

# add Json appender
logger$add_appender(lgr::AppenderJson$new(tf), name = "json")

# signal a warning
logger$warn("this is a warning from mlr3")

# print the contents of the file
cat(readLines(tf))

# remove the appender again
logger$remove_appender("json")
```
