# Use Cases {#use-cases}

This chapter is a collection of use cases to showcase `r mlr_pkg("mlr3")`.

## House Price Prediction in King County {#use-case-regr-houses}

We use the `house_sales_prediction` dataset contained in this book in order to provide a use-case for the application of `mlr3` on real-world data.

```{r 08-use-cases-001}
library(mlr3book)
data("house_sales_prediction", package = "mlr3book")
```

### Exploratory Data Analysis

In order to get a quick impression of our data, we perform some initial *Exploratory Data Analysis*.
This helps us to get a first impression of our data and might help us arrive at additional features that can help with the prediction of the house prices.

We can get a quick overview using R's summary function:

```{r 08-use-cases-002}
summary(house_sales_prediction)
dim(house_sales_prediction)
```

Our dataset has `r nrow(house_sales_prediction)` observations and `r ncol(house_sales_prediction)` columns.
The variable we want to predict is `price`.
In addition to the price column, we have several other columns:

- `id:` A unique identifier for every house.

- `date`: A date column, indicating when the house was sold.
  This column is currently not encoded as a `date` and requires some preprocessing.

- `zipcode`: A column indicating the ZIP code. This is a categorical variable with many factor levels.

- `long, lat` The longitude and latitude of the house

- `...` several other numeric columns providing information about the house, such as number of rooms, square feet etc.

Before we continue with the analysis,  we preprocess some features so that they are stored in the correct format.

First we convert the `date` column to `numeric` to be able to treat it as a numeric feature:

```{r 08-use-cases-003, message = FALSE}
library(lubridate)
house_sales_prediction$date = ymd(substr(house_sales_prediction$date, 1, 8))
house_sales_prediction$date = as.numeric(as.Date(house_sales_prediction$date, origin = "1900-01-01"))
house_sales_prediction$date = house_sales_prediction$date
```

Afterwards, we convert the zipcode to a factor:

```{r 08-use-cases-004}
house_sales_prediction$zipcode = as.factor(house_sales_prediction$zipcode)
```

And add a new column **renovated** indicating whether a house was renovated at some point.

```{r 08-use-cases-005}
house_sales_prediction$renovated = as.numeric(house_sales_prediction$yr_renovated > 0)
# And drop the id column:
house_sales_prediction$id = NULL
```

Additionally we convert the price from Dollar to units of 1000 Dollar to improve readability.

```{r 08-use-cases-006}
house_sales_prediction$price = house_sales_prediction$price / 1000
```

We can now plot the density of the **price** to get a first impression on its distribution.

```{r 08-use-cases-007}
library(ggplot2)
ggplot(house_sales_prediction, aes(x = price)) + geom_density()
```

We can see that the prices for most houses lie between 75.000 and 1.5 million dollars.
There are few extreme values of up to 7.7 million dollars.

Feature engineering often allows us to incorporate additional knowledge about the data and underlying processes.
This can often greatly enhance predictive performance.
A simple example: A house which has `yr_renovated == 0` means that is has not been renovated yet.
Additionally we want to drop features which should not have any influence (`id column`).

After those initial manipulations, we load all required packages and create a Task containing our data.

```{r 08-use-cases-008}
library(mlr3)
library(mlr3viz)
tsk = TaskRegr$new("sales", house_sales_prediction, target = "price")
```

We can inspect associations between variables using `mlr3viz`'s `autoplot` function in order to get some good first impressions for our data.
Note, that this does in no way prevent us from using other powerful plot functions of our choice on the original data.

#### Distribution of the price:

The outcome we want to predict is the **price** variable.
The `autoplot` function provides a good first glimpse on our data.
As the resulting object is a `ggplot2` object, we can use `faceting` and other functions from **ggplot2** in order to enhance plots.

```{r 08-use-cases-009}
library(ggplot2)
autoplot(tsk) + facet_wrap(~renovated)
```

We can observe that renovated flats seem to achieve higher sales values, and this might thus be a relevant feature.

Additionally, we can for example look at the condition of the house.
Again, we clearly can see that the price rises with increasing condition.

```{r 08-use-cases-010}
autoplot(tsk) + facet_wrap(~condition)
```

#### Association between variables

In addition to the association with the target variable, the association between the features can also lead to interesting insights.
We investigate using variables associated with the quality and size of the house.
Note that we use `$clone()` and `$select()` to clone the task and select only a subset of the features for the `autoplot` function, as `autoplot` per default uses all features.
The task is cloned before we select features in order to keep the original task intact.

```{r 08-use-cases-011}
# Variables associated with quality
autoplot(tsk$clone()$select(tsk$feature_names[c(3, 17)]),
  type = "pairs")
```

```{r 08-use-cases-012}
autoplot(tsk$clone()$select(tsk$feature_names[c(9:12)]),
  type = "pairs")
```

### Splitting into train and test data

In `mlr3`, we do not create `train` and `test` data sets, but instead keep only a vector of train and test indices.

```{r 08-use-cases-013}
set.seed(4411)
train.idx = sample(seq_len(tsk$nrow), 0.7 * tsk$nrow)
test.idx = setdiff(seq_len(tsk$nrow), train.idx)
```

### A first model: Decision Tree

Decision tree's cannot only be used as a powerful tool for predictive models but for exploratory data analysis.
In order to fit a decision tree, we first get the `regr.rpart` learner from the `mlr_learners` dictionary.

For now we leave out the  `zipcode` variable, as we also have the `latitude` and `longitude` of each house.

```{r 08-use-cases-014}
tsk_nozip = tsk$clone()$select(tsk$feature_names[1:19])
# Get the learner
lrn = mlr_learners$get("regr.rpart")
# And train on the task
lrn$train(tsk_nozip, row_ids = train.idx)
```

```{r 08-use-cases-015, width = 10, height = 10}
plot(lrn$model)
text(lrn$model)
```

The learned tree relies on several variables in order to distinguish between cheaper and pricier houses.
The features we split along are **grade**, **sqft_living**, but also some features related to the area (longitude and latitude).

We can visualize the price across different regions in order to get more info:

```{r 08-use-cases-016, message = FALSE}
# Load the ggmap package in order to visualize on a map
library(ggmap)

# And create a quick plot for the price
qmplot(long, lat, maptype = "watercolor", color = log(price),
  data = house_sales_prediction[train.idx[1:3000],]) +
  scale_colour_viridis_c()

# And the zipcode
qmplot(long, lat, maptype = "watercolor", color = zipcode,
  data = house_sales_prediction[train.idx[1:3000],]) + guides(color = FALSE)
```

We can see that the price is clearly associated with the zipcode when comparing the
two plots.
As a result, we might want to indeed use the **zipcode** column in our future endeavours.

### A first baseline: Decision Tree

After getting an initial idea for our data, we might want to construct a first baseline, in order to see what a simple model already can achieve.

We use `resample` with `3-fold cross-validation` on our training data in order to get a reliable estimate of the algorithm's performance on future data.
Before we start with defining and training learners, we create a `r ref("Resampling")` in order to make sure that we always compare on exactly the same data.

```{r 08-use-cases-017}
library(mlr3learners)
cv3 = mlr_resamplings$get("cv3")
cv3$instantiate(tsk$clone()$filter(train.idx))
```

For the cross-validation we only use the **training data** by cloning the task and selecting only observations from the training set.

```{r 08-use-cases-018}
lrn_rpart = mlr_learners$get("regr.rpart")
res = resample(task = tsk$clone()$filter(train.idx), lrn_rpart, cv3)
res$performance("regr.mse")
sprintf("RMSE of the simple rpart: %s", round(sqrt(res$aggregate()), 2))
```

### Many Trees: Random Forest

We might be able to improve upon the **RMSE** using more powerful learners.
We first load the `mlr3learners` package, which contains the **ranger** learner (a package which implements the "Random Forest" algorithm).

```{r 08-use-cases-019}
lrn_ranger = mlr_learners$get("regr.ranger", param_vals = list(num.trees = 15L))
res = resample(task = tsk$clone()$filter(train.idx), lrn_ranger, cv3)
res$performance("regr.mse")
sprintf("RMSE of the simple ranger: %s", round(sqrt(res$aggregate()), 2))
```

Often tuning **RandomForest** methods does not increase predictive performances substantially.
If time permits, it can nonetheless lead to improvements and should thus be performed.
In this case, we resort to tune a different kind of model: **Gradient Boosted Decision Trees** from the package `r cran_pkg("xgboost")`.

### A better baseline: `AutoTuner`

Tuning can often further improve the performance.
In this case, we *tune* the xgboost learner in order to see whether this can improve performance.
For the `AutoTuner` we have to specify a **Termination Criterion** (how long the tuning should run) a **Tuner** (which tuning method to use) and a **ParamSet** (which space we might want to search through).
For now we do not use the **zipcode** column, as `r cran_pkg("xgboost")` cannot naturally
deal with categorical features.
The **AutoTuner** automatically performs nested cross-validation.

```{r 08-use-cases-020}
set.seed(444L)
library(mlr3tuning)
library(paradox)
lrn_xgb = mlr_learners$get("regr.xgboost")

# Define the ParamSet
ps = ParamSet$new(
  params = list(
    ParamDbl$new(id = "eta", lower = 0.2, upper = .4),
    ParamDbl$new(id = "min_child_weight", lower = 1, upper = 20),
    ParamDbl$new(id = "subsample", lower = .7, upper = .8),
    ParamDbl$new(id = "colsample_bytree",  lower = .9, upper = 1),
    ParamDbl$new(id = "colsample_bylevel", lower = .5, upper = .7),
    ParamInt$new(id = "nrounds", lower = 1L, upper = 25)
))

# Define the Terminator
terminator = TerminatorEvaluations$new(10)
at = AutoTuner$new(lrn_xgb, "cv3", measures = "regr.mse", ps,
  terminator, tuner = TunerRandomSearch, tuner_settings = list())
```

```{r 08-use-cases-021, echo = FALSE, results = 'hide'}
# And resample the AutoTuner
res = resample(tsk_nozip$clone()$filter(train.idx), at, cv3)
```

```{r 08-use-cases-022}
res$performance("regr.mse")
sprintf("RMSE of the tuned xgboost: %s", round(sqrt(res$aggregate()), 2))
```

We can obtain the resulting params in the respective splits by accessing the `r ref("ResampleResult")`.

```{r 08-use-cases-023}
sapply(res$learners, function(x) x$param_set$values)
```

**NOTE:** To keep runtime low, we only tune parts of the hyperparameter space of `r cran_pkg("xgboost")` in this example.
Additionally, we only allow for $10$ random search iterations, which is usually to little for real-world applications.
Nonetheless, we are able to obtain an improved performance when comparing to the `r cran_pkg("ranger")` model.

In order to further improve our results we have several options:

- Find or engineer better features
- Remove Features to avoid overfitting
- Obtain additional data (often prohibitive)
- Try more models
- Improve the tuning
   - Increase the tuning budget
   - Enlarge the tuning search space
   - Use a more efficient tuning algorithm
- Stacking and Ensembling (see [Pipelines](#pipelines))

Below we will investigate some of those possibilities and investigate whether this improves performance.

### Engineering Features: Mutating ZIP-Codes

In order to better cluster the zip codes, we compute a new feature: **med_price**:
It computes the median price in each zip-code.
This might help our model to improve the prediction.

```{r 08-use-cases-024}
# Create a new feature and append it to the task
zip_price = house_sales_prediction[, .(med_price = median(price)),by = zipcode]

# Join on the original data to match with original columns, then cbind to the task
tsk$cbind(house_sales_prediction[zip_price, on="zipcode"][,"med_price"])
```

Again, we run `resample` and compute the **RMSE**.

```{r 08-use-cases-025}
lrn_ranger = mlr_learners$get("regr.ranger")
res = resample(task = tsk$clone()$filter(train.idx), lrn_ranger, cv3)
res$performance("regr.mse")
sprintf("RMSE of ranger with med_price: %s", round(sqrt(res$aggregate()), 2))
```

### Obtaining a sparser model

In many cases, we might want to have a sparse model.
For this purpose we can use a `r ref("mlr3featse::Filter")` implemented in `r mlr_pkg("mlr3featsel")`.
This can prevent our learner from overfitting make it easier for humans to interpret models as fewer variables influence the resulting prediction.

```{r 08-use-cases-026}
library(mlr3featsel)
filter = FilterMRMR$new()
tsk_ftsel = filter$filter_nfeat(tsk$clone(), nfeat = 12)
```

The resulting **RMSE** is slightly higher, and at the same time we only use $12$ features.

```{r 08-use-cases-027}
lrn_ranger = mlr_learners$get("regr.ranger")
res = resample(task = tsk_ftsel$clone()$filter(train.idx), lrn_ranger, cv3)
res$performance("regr.mse")
sprintf("RMSE of ranger with filtering: %s", round(sqrt(res$aggregate()), 2))
```
