# Survival Analysis {#survival}

The package `r mlr_pkg("mlr3survival")` extends `r mlr_pkg("mlr3")` with the following objects for survival analysis:

* `r ref("mlr3survival::TaskSurv", text = "TaskSurv")` to define (right-censored) survival tasks
* `r ref("mlr3survival::LearnerSurv", text = "LearnerSurv")` as base class for survival learners
* `r ref("mlr3survival::PredictionSurv", text = "PredictionSurv")` as specialized class for `r ref("Prediction")` objects
* `r ref("mlr3survival::MeasureSurv", text = "MeasureSurv")` as specialized class for performance measures

## Example {#survival-example}

This example uses the `r ref("survival::rats", text = "rats")` data from `r cran_pkg("survival")` package which comes as predefined `r ref("TaskSurv")`.

```{r}
library(mlr3survival)
task = mlr_tasks$get("rats")

# the target column is a survival object:
head(task$truth())

# kaplan meier estimator of survival times
task$survfit()

# kaplan-meier estimator of survival times
# stratify on binarized age variable
sf = task$survfit("sex")
print(sf)

# plot using ggfortify
library("ggfortify")
autoplot(sf)
```

```{r survival-benchmark}
keys = mlr_learners$keys("^surv")
print(keys)
learners = mlr_learners$mget(keys)

keys = mlr_measures$keys("^surv")
print(keys)
measures = mlr_measures$mget(keys)

set.seed(1)
bmr = benchmark(expand_grid(task, learners, "cv"), measures = measures)
print(bmr)
```
