## Integrated Learners {#list-learners}

```{r include=FALSE}
requireNamespace("mlr3", quietly = TRUE)
requireNamespace("mlr3learners", quietly = TRUE)
requireNamespace("mlr3proba", quietly = TRUE)
requireNamespace("mlr3cluster", quietly = TRUE)
requireNamespace("mlr3extralearners", quietly = TRUE)
```


Listed are the `r length(mlr3::mlr_learners$keys())` learners whose implementations are available in one of the following packages:

* `r mlr_pkg("mlr3")`: debug learner and `r cran_pkg("rpart")` learners.
* `r mlr_pkg("mlr3learners")`: opinionated selection of some default learners.
* `r mlr_pkg("mlr3proba")`: base learners for survival and probabilistic regression.
* `r mlr_pkg("mlr3cluster")`: learners for unsupervised clustering.
* `r mlr_pkg("mlr3extralearners")`: more experimental learners for regression, classification and survival.

```{r 98-appendix-learners-001, echo=FALSE, message=FALSE}
d = suppressWarnings(data.table::as.data.table(mlr3::mlr_learners))


d$key = mlr3misc::map_chr(d$key, function(x) {
  txt = try(ref(sprintf("mlr_learners_%s", x), x), silent = TRUE)
  if (inherits(txt, "try-error"))
    return(x)
  else txt
})

d$feature_types = mlr3misc::map_chr(d$feature_types, function(x) {
  paste0(mlr3:::translate_types(x), collapse = ", ")
})

d$packages = mlr3misc::map_chr(d$packages, function(x) {
  ifelse(length(x) == 0, "", paste0(sapply(sort(x), cran_pkg), collapse = ", "))
  })


for (col in c("properties", "predict_types")) {
  d[[col]] = mlr3misc::map_chr(d[[col]], paste0, collapse = ", ")
}

d = d[, c("key", "packages", "feature_types", "properties", "predict_types"), with = FALSE]
knitr::kable(d, col.names = c("Id", "Packages", "Feature Types", "Properties", "Predict Types"))
```
