---
aliases:
  - "/predsets_valid_inttune.html"
---

# Predict Sets, Validation and Internal Tuning (+) {#sec-predsets-valid-inttune}

{{< include ../../common/_setup.qmd >}}

`r chapter = "Predict Sets, Validation and Internal Tuning (+)"` `r authors(chapter)`

## Predict Sets and Training Error Estimation {#sec-predict-sets}

In the @sec-performance chapter we have already studied in detail how to train, predict and evaluate many different learners. Evaluating a fully trained model usually requires making predictions on unseen test observations. When we predict directly with a trained learner, we can explicitly control which observations are used:

```{r}
tsk_sonar = tsk("sonar")
lrn_rf = lrn("classif.ranger")
lrn_rf$train(tsk_sonar, row_ids = 4:208)
pred1 = lrn_rf$predict(tsk_sonar, row_ids = 1:3)
pred2 = lrn_rf$predict_newdata(tsk_sonar$data(1:3))
```

But when using `resample()` or `benchmark()`, the default behavior is to predict on the *test* set of the resampling. It is also possible to make predictions on the *train* and *internal_valid* data, by configuring the `$predict_sets` of a learner. The *internal_valid* option (see next sections) is only possible if validation data is predefined in the task, or if a learner with property 'validation' has its `$validate` field set.

We now configure our learner to make predictions on the train and test sets during resampling. The former may sometimes be of interest for further analysis or to study overfitting. Or maybe we are simply curious.

```{r}
lrn_rf$predict_sets = c("train", "test")
rr = resample(tsk_sonar, lrn_rf, rsmp("cv", folds = 3))
```

The learner, during resampling, will now after having been trained for the current iteration, produce predictions on all requested sets. To access predictions in our case, we can either ask for a list of 3 prediction objects, one for each cross-validation fold, or we can ask for a combined prediction object for the whole CV -- which for a CV contains as many prediction rows as there are observations in the task.

```{r}
str(rr$predictions("test")) # or str(rr$predictions("train"))
rr$prediction("test") # or rr$prediction("train")
```

We can also apply measures to specific sets of the resample result:

```{r}
rr$aggregate(list(
  msr("classif.ce", predict_sets = "train", id = "ce_train"),
  msr("classif.ce", predict_sets = "test", id = "ce_test")
))
```

The default predict set for a measure is usually the test set. But we can request other sets here. If multiple predict sets are requested for the measure, their predictions are joined, before they are passed into the measure, which then usually calculates an aggregated score over all predicted rows of the set. In our case, unsurprisingly, the train error is lower than the test error.

If we only want to access information that is computed during training, we can even configure the learner not to make any predictions at all. This is useful, for example, for learners that already produce an estimate of their generalization error during training, e.g. using out-of-bag error estimates or validation scores. The former, which is only available to learners with the 'oob_error' property, can be accessed via `r ref("MeasureOOBError")`. The latter is available to learners with the 'validation' property and is implemented as `r ref("MeasureInternalValidScore")`. Below we evaluate a random forest using its out-of-bag error. Since we do not need any predict sets, we can use `r ref("ResamplingInsample")`, which will use the entire dataset for training.

```{r}
tsk_sonar = tsk("sonar")
lrn_rf$predict_sets = NULL
rsmp_in = rsmp("insample")
rr = resample(tsk_sonar, lrn_rf, rsmp_in, store_models = TRUE)
msr_oob = msr("oob_error")
rr$aggregate(msr_oob)
```

All this works in exactly the same way for benchmarking, tuning, nested resampling, and any other procedure where resampling is internally involved and we can either ask for predictions or apply measures at the end. Below we illustrate this by tuning the `mtry.ratio` parameter of the random forest using a simple grid search with 10 evaluations. Instead of explicitly making predictions on some test data and evaluating them, we use the out-of-bag error to evaluate the different values of `mtry.ratio`.
This can speed up the tuning process considerably.

```{r}
lrn_rf$param_set$set_values(
  mtry.ratio = to_tune(0.1, 1)
)

ti = tune(
  task = tsk_sonar,
  tuner = tnr("grid_search"),
  learner = lrn_rf,
  resampling = rsmp_in,
  measure = msr_oob,
  term_evals = 10,
  store_models = TRUE
)
```

## Validation {#sec-validation}

For iterative training (which many learners use) it can be interesting to track performance *during* training on *validation* data. One can use this for simple logging or posthoc analysis, but the major use case is early stopping. If the model’s performance on the training data keeps improving but the performance on the validation data plateaus or degrades, this indicates overfitting and we should stop iterative training. Handling this in an online fashion during training is much more efficient than configuring the number of iterations from the outside via traditional, offline hyperparameter tuning, where we would fit the model again and again with different iteration numbers.

In `mlr3`, learners can have the 'validation' and 'internal_tuning' properties to indicate whether they can make use of a validation set and whether they can internally optimize hyperparameters, for example by stopping early. To check if a given learner supports this, we can simply access its `$properties` field. Examples of such learners are boosting algorithms like XGBoost, LightGBM, or CatBoost, as well as deep learning models from `r ref_pkg("mlr3torch")`. In this section we will train XGBoost on sonar and keep track of its performance on a validation set.

```{r}
tsk_sonar = tsk("sonar")
lrn_xgb = lrn("classif.xgboost")
lrn_xgb
```

To enable validation, we need to configure how the validation data is constructed. For XGBoost there is a special `watchlist` parameter, but `mlr3` also provides a standardized -- and as we will see later, more powerful -- interface via the learner's `$validate` field. This field can be set to:

-   `NULL` to use no validation data (default),
-   a ratio indicating the proportion of training data to be used as the validation set,
-   `"predefined"` to use the validation data specified in the task (we will see shortly how to configure this), and
-   `"test"` to use the test set as validation data, which only works in combination with resampling and tuning.

::: callout-note
## Test Data Leakage

If a learner's `$validate` field is set to 'test', we will leak the resampling test set during training. This will lead to biased performance estimates if the validation scores are used for early stopping. Whether this is desireable depends on the context: if the test set is used to evaluate parameter configurations during HPO (i.e. it acts as a validation set), then this is usually OK; However, if the purpose of the test set is to provide an unbiased estimate of performance, e.g. to compare different learners, then this is not OK.
:::

Below, we configure the XGBoost learner to use $1/3$ of its training data for validation:

```{r}
lrn_xgb$validate = 1/3
```

Next, we set the number of iterations (`nrounds`) and which metric to track (`eval_metric`) and train the learner. Here, $1/3$ of the observations from the training task will be solely used for validation and the remaining $2/3$ for training. If stratification or grouping is enabled in the task, this will also be respected. This has already been covered in @sec-performance.

```{r}
lrn_xgb$param_set$set_values(
  nrounds = 100,
  eval_metric = "logloss"
)
lrn_xgb$train(tsk_sonar)
```

Because the XGBoost learner kept a log of the validation performance, we can now access this through the `$model` slot. Where exactly in the model this information is stored, depends on the specific learning algorithm. For XGBoost, the history is stored in `$evaluation_log`:

```{r}
tail(lrn_xgb$model$evaluation_log)
```

The validation loss over time is visualized in the figure below, with the iterations on the x-axis and the validation logloss on the y-axis:

```{r, out.width = "70%", echo = FALSE, warning = FALSE}
library(ggplot2)
set.seed(1)
ggplot(lrn_xgb$model$evaluation_log, aes(x = iter, y = test_logloss)) +
  geom_line() +
  labs(
    x = "Boosting Iteration",
    y = "Validation Logloss"
  ) + theme_minimal()
```

`mlr3` also provides a standardized acccessor for the final validation performance. We can access this via the `$internal_valid_scores` field, which is a named list containing possibly more than one validation metric.

```{r}
lrn_xgb$internal_valid_scores
```

In some cases one might want to have more control over the construction of the validation data. This can be useful, for example, if there is a predefined validation split to be used with a task. Such fine-grained control over the validation data is possible by setting the `validate` field to `"predefined"`.

```{r}
lrn_xgb$validate = "predefined"
```

This allows us to use the `$internal_valid_task` defined in the training task. Below, we set the validation task to use 60 randomly sampled ids and remove them from the primary task.

```{r}
valid_ids = sample(tsk_sonar$nrow, 60)
tsk_valid = tsk_sonar$clone(deep = TRUE)
tsk_valid$filter(valid_ids)
tsk_sonar$filter(setdiff(tsk_sonar$row_ids, valid_ids))
tsk_sonar$internal_valid_task = tsk_valid
```

Note that we could have achieved the same by simply setting `tsk_valid$internal_valid_task = valid_ids`, but we have chosen to show the more explicit way for the sake of clarity. The associated validation task now has 60 observations and the primary task 148:

```{r}
c(tsk_sonar$internal_valid_task$nrow, tsk_sonar$nrow)
```

When we now train the XGBoost learner on the task, the data from the internal validation task will be used for validation.
Note that the `$internal_valid_task` slot is always used internally, even if you set a ratio value in `learner$validate`, it is simply automatically constructed (and then passed down).

```{r}
lrn_xgb$train(tsk_sonar)
```

In many cases, however, one does not only train an individual learner, but combines it with other (preprocessing) steps in a `r ref("GraphLearner")`, see @sec-preprocessing. Validation in a `GraphLearner` is still possible, because preprocessing `PipeOp`s also handle the validation task. While the *train* logic of the `PipeOp`s is applied to the primary task, the *predict* logic is applied to the validation data. This ensures that there is no data leakage when the XGBoost learner evaluates its performance on the validation data. Below, we construct a `PipeOpPCA` and apply it to the sonar task with a validation task.

```{r}
po_pca = po("pca")
taskout = po_pca$train(list(tsk_sonar))[[1]]
taskout$internal_valid_task
```

The preprocessing that is applied to the `$internal_valid_task` during `$train()` is equivalent to making a prediction for it:

```{r}
po_pca$predict(list(tsk_sonar$internal_valid_task))[[1L]]
```

This means that tracking validation performance works even in complex graph learners, which would not be possible when setting the `watchlist` parameter of XGBoost. Below, we chain the PCA operator to XGBoost and convert it to a learner.

```{r}
glrn = as_learner(po_pca %>>% lrn_xgb)
```

While this almost 'just works', we now need to specify the `$validate` field on two levels:

1.  For the `GraphLearner` itself, i.e. how the validation data is created before the `Task` enters the graph.
2.  Which `PipeOp`s that have the property `"validation"` should actually use it.

This configuration can be simplified by using `set_validate()`. When applied to a `GraphLearner`, we can specify the arguments `validate` which determines *how* to create the validation data and optionally the argument `ids` which specifies *which* `PipeOp`s should use it. By default, the latter is set to the `$base_learner()` of the `Graph`, which is the last learner. This means that both calls below are equivalent:

```{r}
set_validate(glrn, validate = "predefined")
set_validate(glrn, validate = "predefined", ids = "classif.xgboost")
```

We can now train the graph learner just as before and inspect the final validation metric, which is now prefixed with the ID of the corresponding `PipeOp`.

```{r}
glrn$validate = "predefined"
glrn$train(tsk_sonar)
glrn$internal_valid_scores
```

::: callout-note
## Field `$validate` for `PipeOp`s

Since individual `PipeOp`s cannot control how the validation data is generated, only whether to use it, their `$validate` field can only be set to `NULL` or `"predefined"`. This is why we get an error when running `as_pipeop(lrn("classif.xgboost", validate = 0.3))`. When using validation in a GraphLearner, it is best to first construct the learner without specifying the validation data and then use `set_validate()`.
:::

## Internal Tuning {#sec-internal-tuning}

Not only can XGBoost log its validation performance, it can also monitor it to *early stop* its training, i.e. perform internal tuning of the `nrounds` hyperparameter during training. This is marked by the `"internal_tuning"` property:

```{r}
"internal_tuning" %in% lrn_xgb$properties
```

Early stopping for XGBoost can be enabled by specifying the `early_stopping_rounds` parameter. This is also known as *patience* and specifies for how many iterations the validation loss must not improve for the training to terminate. The metric that is used for early stopping is the first value that we passed to `eval_metric`, which was the logloss.

```{r}
lrn_xgb$param_set$set_values(
  early_stopping_rounds = 10,
  nrounds = 100
)
```

When we now train the learner, we can access the internally optimized `nrounds` through the `$internal_tuned_values` field.

```{r}
lrn_xgb$train(tsk_sonar)
lrn_xgb$internal_tuned_values
```

By using early stopping, we were able to already terminate training after `r lrn_xgb$internal_tuned_values$nrounds + lrn_xgb$param_set$values$early_stopping_rounds` iterations. Below, we visualize the validation loss over time and the optimal nrounds is marked red. We can see that the logloss plateaus after `r lrn_xgb$internal_tuned_values$nrounds` rounds, but training continues for a while afterwards due to the patience.

```{r, echo = FALSE, out.width = "70%"}
theme_set(theme_minimal())
data = lrn_xgb$model$evaluation_log
ggplot(data, aes(x = iter, y = test_logloss)) +
  geom_line() +
  geom_point(data = data.table(x = lrn_xgb$internal_tuned_values$nrounds,
    y = lrn_xgb$internal_valid_scores$logloss), aes(x = x, y = y, color = "red")) +
  labs(
    x = "Iteration", y = "Validation Logloss", 
    color = NULL
  )
```

So far we have only used the early stopping implementation of XGBoost to optimize `nrounds`, but have not tuned any other hyperparameters. This is where `r mlr3` comes in, as it allows us to combine the internal tuning of a learner with (non-internal) hyperparameter tuning via `r ref_pkg("mlr3tuning")`. To do this, we set both parameters to `to_tune()`, but mark `nrounds` to be tuned internally.

```{r}
lrn_xgb$param_set$set_values(
  eta = to_tune(0.001, 0.1, logscale = TRUE),
  nrounds = to_tune(upper = 500, internal = TRUE)
)
```

In such scenarios, one might often want to use the same validation data to optimize `eta` and `nrounds`. This is possible by specifying the `"test"` option of the `validate` field. This means that in each resampling iteration the validation data will be set to the test set, i.e. the same data that will also be used to evaluate the parameter configuration (to tune `eta`).

```{r}
lrn_xgb$validate = "test"
```

We will now continue to tune XGBoost on the sonar task using a simple grid search with 10 evaluations and 3-fold cross-validation. Internally, this will train XGBoost with 10 different values of `eta` and the `nrounds` parameter fixed at 500, i.e. the upper bound from above. For each value of `eta` a 3-fold cross-validation with early stopping will be performed, yielding 3 (possibly different) early stopped values for `nrounds` for each value of `eta`. These are combined into a single value according to an aggregation rule, which by default is set to averaging, but which can be overridden when creating the internal tune token, see `r ref("to_tune()")` for more information.

When combining internal tuning with hyperparameter optimization via `r ref_pkg("mlr3tuning")` we need to specify two performance metrics: one for the internal tuning and one for the `Tuner`. For this reason, `mlr3` requires the internal tuning metric to be set explicitly, even if a default value exists. There are two ways to use the same metric for both types of hyperparameter optimization:

1.  Use `msr("internal_valid_scores", select = <id>)`, i.e. the final validation score, as the tuning measure. As a learner can have multiple internal valid scores, the measure allows us to select one by specifying the `select` argument. We also need to specify whether the measure should be minimized.
2.  Set both, the `eval_metric` and the tuning measure to the same metric, e.g. `eval_metric = "error"` and `measure = msr("classif.ce")`. Some learners even allow to set the validation metric to an `mlr3::Measure`. You can find out which ones support this feature by checking their corresponding documentation. One example for this is XGBoost.

The advantage of using the first option is that the predict step can be skipped because the internal validation scores are already computed during training. This is similar to evaluating the random forest from @sec-predict-sets using the out-of-bag error.

```{r}
tsk_sonar = tsk("sonar")
lrn_xgb$predict_sets = NULL

ti = tune(
  tuner = tnr("grid_search"),
  learner = lrn_xgb,
  task = tsk_sonar,
  resampling = rsmp("cv", folds = 3),
  measure = msr("internal_valid_score",
    select = "logloss", minimize = TRUE),
  term_evals = 10L
)
```

The tuning result contains the best found configuration for both `eta` and `nrounds`.

```{r}
ti$result_learner_param_vals[c("eta", "nrounds")]
```

We now show how to extract the different parameter configurations from the tuning archive. All internally tuned parameters are accessible via the `$internal_tuned_values`. This is a list column, because it is possible to tune more than one parameter internally, e.g. in a `GraphLearner`. Below we extract the values for `eta` (transformed back from its log scale), `nrounds` (internally tuned) and the logloss. The latter was evaluated on the internal validation tasks, which corresponded to the `Resampling`'s test sets as we specified `validate = "test"`. By visualizing the results we can see an inverse relationship between the two tuning parameters: a smaller step size (eta) requires more boosting iterations (nrounds).

```{r, out.width = "70%"}
d = ti$archive$data

d = data.table(
  eta = exp(d$eta),
  nrounds = unlist(d$internal_tuned_values),
  logloss = d$logloss
)

ggplot(data = d, aes(x = eta, y = nrounds, color = logloss)) +
  geom_point() + theme_minimal()
```

This also works with an `r ref("AutoTuner")`, which will use the internally optimized `nrounds`, as well as the offline tuned `eta` for the final model fit.
This means that there is no validation or early stopping when training the final model, and we use all available data.

```{r}
at = auto_tuner(
  tuner = tnr("grid_search"),
  learner = lrn_xgb,
  resampling = rsmp("cv", folds = 3),
  measure = msr("internal_valid_score",
    select = "logloss", minimize = TRUE),
  term_evals = 10L
)
at$train(tsk_sonar)
```

If we were to resample the `AutoTuner` from above, we would still get valid performance estimates.
This is because the test set of the outer resampling is *never* used as validation data, since the final model fit does not perform any validation.
The validation data generated during the hyperparameter tuning uses the test set of the inner resampling, which is a subset of the training set of the outer resampling.

However, care must be taken when using the test set of a resampling for validation. Whether this is OK depends on the context and purpose of the resampling.
If the purpose of resampling is to get an unbiased performance estimate of algorithms, some of which stop early and some of which don't, this is not OK.
In such a situation, the former would have an unfair advantage over the latter. The example below illustrates such a case where this would not be a fair comparison between the two learners.

```{r}
lrn_xgb$param_set$set_values(
  eta = 0.1, nrounds = 500, early_stopping_rounds = 10
)
lrn_xgb$predict_sets = "test"

design = benchmark_grid(
  tsk_sonar, list(lrn_xgb, lrn("classif.rpart")), rsmp("cv", folds = 3)
)
bmr = benchmark(design)
bmr$aggregate(msr("classif.ce"))
```

At last, we will cover how to enable internal tuning when manually specifying a search space with the `ps()` function instead of the `to_tune()`-mechanism.
While the latter is more convenient and therefore usually recommended, manually defining a search space gives you for more flexibility with respect to parameter transformations, see e.g. @sec-tune-trafo.
There are two ways to do this:
1. We can include the internally tuned parameters in the primary `search_space`, specify an aggregation function and tag them with `"internal_tuning"`.
2. We define a seperate `internal_search_space` where we also define the aggregation function but don't include the `"internal_tuning"` tag.

We start by illustrating the first approach, which is more compact as the search space is defined as a single object.

```{r}
search_space = ps(
  eta = p_dbl(0.001, 0.1, logscale = TRUE),
  nrounds = p_int(upper = 500, tags = "internal_tuning",
    aggr = function(x) as.integer(mean(unlist(x))))
)
```

This search space can be passed to the `AutoTuner` and the optimization will then proceed as before.

```{r}
at = auto_tuner(
  tuner = tnr("grid_search"),
  learner = lrn_xgb,
  resampling = rsmp("cv", folds = 3),
  measure = msr("internal_valid_score",
    select = "logloss", minimize = TRUE),
  search_space = search_space,
  term_evals = 10L
)
at$train(tsk_sonar)
```

One reason to use the second approach is when a global parameter transformation (`.extra_trafo`) is applied to the primary search space.
Univariate transformation, such as logarithmic scaling, are no problem.
Below, we demonstrate this using an illustrative example where we sample set the `max_depth` of the trees depending on the value of `eta`:

```{r}
search_space = ps(eta = p_dbl(0.001, 0.1, logscale = TRUE),
  .extra_trafo = function(x, param_set) {
    x$max_depth = if (exp(x$eta) < 0.01) 10 else 20
    x
  })

internal_search_space = ps(
  nrounds = p_dbl(upper = 500, tags = "internal_tuning",
    aggr = function(x) as.integer(mean(unlist(x))))
)

at = auto_tuner(
  tuner = tnr("grid_search"),
  learner = lrn_xgb,
  resampling = rsmp("cv", folds = 3),
  measure = msr("internal_valid_score",
    select = "logloss", minimize = TRUE),
  search_space = search_space,
  internal_search_space = internal_search_space,
  term_evals = 10L
)
at$train(tsk_sonar)
```

## Conclusion

In this chapter we first learned how to evaluate machine learning methods on different prediction sets, namely *train*, *internal_valid* and *test*. Then we learned how to track the performance of an iterative learning procedure on a validation set. This technique also works seamlessly in a graphlearner, the only difference being that you have to specify not only how to create the validation data, but also which PipeOps should use it. Furthermore, mlr3's *internal tuning* mechanism allows you to combine hyperparameter tuning via `r mlr3tuning` with internal tuning of the learning algorithm, such as early stopping of XGBoost.

## Exercises

1.  Manually `$train()` a LightGBM classifier from `r ref_pkg("mlr3extralearners")` on the pima task using $1/3$ of the training data for validation. As the pima task has missing values, select a method from `r ref_pkg("mlr3pipelines")` to impute them. Explicitly set the evaluation metric to logloss (`"binary_logloss"`), the maximum number of boosting iterations to 1000, the patience parameter to 10, and the step size to 0.01. After training the learner, inspect the final validation scores as well as the early stopped number of iterations.

2.  Wrap the learner from exercise 1) in an `AutoTuner` using a three-fold CV for the tuning. Also change the rule for aggregating the different boosting iterations from averaging to taking the maximum across the folds. Don't tune any parameters other than `nrounds`, which can be done using `tnr("internal")`. Use the internal validation metric as the tuning measure. Compare this learner with a `lrn("classif.rpart")` using a 10-fold outer cross-validation with respect to classification accuracy.

3.  Consider the code below:

    ```{r}
    branch_lrn = as_learner(
      ppl("branch", list(
        lrn("classif.ranger"),
        lrn("classif.xgboost",
          early_stopping_rounds = 10,
          eval_metric = "error",
          eta = to_tune(0.001, 0.1, logscale = TRUE),
          nrounds = to_tune(upper = 1000, internal = TRUE)))))

    set_validate(branch_lrn, validate = "test", ids = "classif.xgboost")
    branch_lrn$param_set$set_values(branch.selection = to_tune())

    at = auto_tuner(
      tuner = tnr("grid_search"),
      learner = branch_lrn,
      resampling = rsmp("holdout", ratio = 0.8),
      # cannot use internal validation score because ranger does not have one
      measure = msr("classif.ce"),
      term_evals = 10L,
      store_models = TRUE
    )

    tsk_sonar = tsk("sonar")$filter(1:100)

    rr = resample(
      tsk_sonar, at, rsmp("holdout", ratio = 0.8), store_models = TRUE
    )
    ```

    Answer the following questions (ideally without running the code):

    3.1 During the hyperparameter optimization, how many observations are used to train the XGBoost algorithm (excluding validation data) and how many for the random forest? Hint: learners that cannot make use of validation data ignore it. 3.2 How many observations would be used to train the final model if XGBoost was selected? What if the random forest was chosen? 3.3 How would the answers to the last two questions change if we had set the `$validate` field of the graphlearner to `0.25` instead of `"test"`?

4.  Look at the (failing) code below:

    ```{r, eval = FALSE}
    tsk_sonar = tsk("sonar")
    glrn = as_learner(
      po("pca") %>>% lrn("classif.xgboost", validate = 0.3)
    )
    ```

    Can you explain *why* the code fails? Hint: Should the data that xgboost uses for validation be preprocessed according to the *train* or *predict* logic?

::: {.content-visible when-format="html"}
`r citeas(chapter)`
:::
