# Quick data.table Intro for Beginners {#data.table}

```{r 02-basics-R6-001, include = F}
library(mlr3)
library(mlr3book)
```

The package `r mlr3book::cran_pkg("data.table")` essentially implements the eponymous alternative to R's `data.frame()`, i.e. an object to store tabular data.

:::{.callout-note}
We decided to use `data.table()` because it is blazingly fast and scales quite well on bigger data.
Many functions of `r mlr_pkg("mlr3")` return `data.table`s which can conveniently be subsetted or combined with other outputs.
If you happen to not like the syntax or are feeling more comfortable with other tools, base `data.frame`s or `r cran_pkg("tibble")`/`r cran_pkg("dplyr")` is just a single `as.data.frame()` or `as_tibble()` away.
:::

Data tables can be constructed using the `data.table()` function (whose interface is similar to  to `data.frame()`) or by converting an object with `as.data.table()`.
```{r}
library("data.table")
df = data.frame(x = 1:12, y = rep(letters[1:3], each = 4))
dt = as.data.table(df)
```
Although both objects store the data identically in memory, they are considerably different in their operation.
First, the index operator `[` works slightly different.
For both objects, the first argument (`i`) is used to select rows and the second argument (`j`) is used to select columns.
However, column names are in the search path, and thus can be used directly:
```{r}
df[df$y == "a", ]
dt[y == "a", ]
```

Second, there is no optional `drop` argument (`drop` is always `FALSE` for `data.table()`), but instead multiple additional arguments to query the data from the `data.table()` in a very concise way.
Most importantly, you can group the data with argument `by` and combine this with aggregating functions provided via the second argument (`j`):
```{r}
dt[, mean(x), by = "y"]
```
There is also extensive support to perform all kinds of data base join operations (see, e.g., [here](https://rstudio-pubs-static.s3.amazonaws.com/52230_5ae0d25125b544caab32f75f0360e775.html)).

For an in-depth introduction, we refer to the [excellent vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html).
Also don't miss the other vignettes linked on the CRAN page of `r mlr3book::ref_pkg("data.table")`!
