# Advanced Tuning Methods and Black Box Optimization {#sec-optimization-advanced}

{{< include _setup.qmd >}}

`r chapter = "Advanced Tuning Methods and Black Box Optimization"`
`r authors(chapter)`

Having looked at the basic usage of `mlr3tuning`, we will now turn to more advanced methods.
We will begin in @sec-tuning-errors by continuing to look at single-objective tuning but will consider what happens when experiments go wrong and how to prevent fatal errors.
We will then extend the methodology from @sec-optimization to enable multi-objective tuning, where learners are optimized to multiple measures simultaneously, in @sec-multi-metrics-tuning we will consider important theory behind this and demonstrate how this is handled relatively simply in `mlr3` by making use of the same classes and methods we have already used.
The final two sections focus on specific optimization methods.
@sec-hyperband looks in detail at multi-fidelity tuning and the Hyperband tuner, consider some theory behind this method and them demonstrating it in practice with `r mlr3hyperband`.
Finally, @sec-bayesian-optimization takes a deep dive into black box Bayesian Optimization.
This is a more theory-heavy section to motivate the design of the classes and methods in `r mlr3mbo`.

## Error Handling and Memory Management {#sec-tuning-errors}

In this section we will look at how to use `mlr3` to ensure that tuning workflows are efficient and robust.
In particular, we will consider how to enable features that prevent fatal errors leading to irrecoverable data loss in the middle of an experiment, and then how to manage tuning experiments that may use up a lot of computer memory.

### Encapsulation and Fallback Learner {#sec-encapsulation-fallback}

Error handling is discussed in detail in @sec-error-handling, however it is very important in the context of tuning so here we will just practically demonstrate how to make use of encapsulation and fallback learners and explain why they are essential during HPO.

Even in simple ML problems, there is a lot of potential for things to go wrong.
For example when learners do not converge, run out of memory, or terminate with an error due to issues in the underlying data or bugs in the code.
As a common example, learners can fail if there are factor levels present in the test data that were not in the training data, models fail in this case as there have been no weights/coefficients trained for these new factor levels:

```{r, error = TRUE, warning = FALSE}
task_bh = tsk("boston_housing")
cv10 = rsmp("cv", folds = 10)
msr_rmse = msr("regr.rmse")
rnd_srch = tnr("random_search")
learner = lrn("regr.lm", df = to_tune(1, 2))

instance = tune(
  tuner = rnd_srch,
  task = task_bh,
  learner = learner,
  resampling = cv10,
  measures = msr_rmse,
  term_evals = 10
)
```

In the above example, we can see the tuning process breaks and we lose all information about the HPO process as the `instance` cannot be saved.
This is even worse in nested resampling or benchmarking, when errors could cause us to lose all progress across multiple configurations or even learners and tasks.

`r index('Encapsulation')` (@sec-encapsulation) allows errors to be isolated and handled, without disrupting the tuning process.
We can tell a learner to encapsulate an error by setting the `$encapsulate` field as follows:

```{r optimization-035}
learner$encapsulate = c(train = "evaluate", predict = "evaluate")
```

Note by passing `"evaluate"` to both `train` and `predict`, we are telling the learner to setup encapsulation in both the training and predicting stages, however we could have only set it for one stage.

Another common issue that cannot be easily solved during HPO is learners not converging and the process running indefinitely.
We can prevent this happening by setting the `timeout` field in a learner, which signals the learner to stop if it has been running for that much time, again this can be set for training and predicting individually:

```{r optimization-036}
learner$timeout = c(train = 30, predict = 30)
```

Now if either an error occurs, or the model timeout threshold is reached, then instead of breaking, the learner will simply not make predictions when errors are found.
Unfortunately, if predictions are not made, then our HPO experiment will still fail as for any resampling iteration with errors, the result will be `NA`, and so are unable to aggregate results across resampling iterations.
Therefore it is essential to also select a fallback learner (@sec-fallback), which is a learner that will be fitted if the learner of interest fails.
A common approach is to use a featureless baseline, `classif/regr.featureless`.
Below we set `regr.featureless`, which always predicts the mean `response`, by passing this learner to the `$fallback` field.

```{r optimization-037}
learner$fallback = lrn("regr.featureless")
```

We can now run our experiment and see errors that occurred during tuning in the archive.

```{r optimization-038}
instance = tune(
  tuner = rnd_srch,
  task = task_bh,
  learner = learner,
  resampling = cv10,
  measures = msr_rmse,
  term_evals = 10
)

as.data.table(instance$archive)[, .(df, regr.rmse, errors)]

# Reading the error in the first resample result
instance$archive$resample_result(1)$errors
```

The learner was tuned without breaking because the errors were encapsulated and logged before the fallback learners were used for fitting and predicting:

```{r}
instance$result
```

### Memory Management {#sec-memory-management}

Running a large tuning experiment requires a lot of working memory, especially when using nested resampling.
Most of the memory is consumed by the models since each resampling iteration creates one new model.
The option `store_models` in the functions `r ref("ti()")` and `r ref("auto_tuner()")` allows us to enable the storage of the models.
Storing the models is disabled by default and in most cases is not required.

The archive stores a `r ref("ResampleResult")` for each evaluated hyperparameter configuration.
The contained `r ref("Prediction")` objects can take up a lot of memory, especially with large datasets and many resampling iterations.
We can disable the storage of the resample results by setting `store_benchmark_result = FALSE` in the functions `r ref("ti()")` and `r ref("auto_tuner()")`.
Note that without the resample results, it is no longer possible to score the configurations on another measure.

When we run nested resampling with many outer resampling iterations, additional memory can be saved if we set `store_tuning_instance = FALSE` in the `r ref("auto_tuner()")` function.
However, the functions `r ref("extract_inner_tuning_results()")` and `r ref("extract_inner_tuning_archives()")` would then no longer work.

The option `store_models = TRUE` sets `store_benchmark_result` and `store_tuning_instance` to `TRUE` because the models are stored in the benchmark results which in turn is part of the instance.
This also means that `store_benchmark_result = TRUE` sets  `store_tuning_instance` to `TRUE`.

Finally, we can set `store_models = FALSE` in the `r ref("resample()")` or `r ref("benchmark()")` functions to disable the storage of the auto tuners when running nested resampling.
This way we can still access the aggregated performance (`rr$aggregate()`) but lose information about the inner resampling.


## Multi-Objective Tuning {#sec-multi-metrics-tuning}

So far we have considered optimizing a model with respect to one metric, but multi-criteria, or `r index("multi-objective", aside = TRUE)` optimization, is also possible.
A simple example of multi-objective optimization might be optimizing a classifier to minimize false positive *and* false negative predictions.
In another example, consider the single-objective problem of tuning a deep neural network to minimize classification error.
The best performing model is likely to be quite complex, possibly with many layers that will take a long time to train, which would not be appropriate when you have limited resources.
In this case we might want to simultaneously minimize the classification error and model complexity.

By definition, optimization of multiple metrics means these will be in competition (otherwise we would only optimize one of them) and therefore in general no *single* configuration exists that optimizes all metrics.
We therefore instead focus on the concept of `r index("Pareto optimality")`.
One hyperparameter configuration is said to `r index("Pareto-dominate")` another if the resulting model is equal or better in all metrics and strictly better in at least one metric.
For example say we are minimizing classification error, CE, and complexity, CP, for configurations A and B with CE of $CE_A$ and $CE_B$ respectively and CP of $CP_A$ and $CP_B$ respectively.
Then, A pareto-dominates B if: 1) $CE_A \leq CE_B$ and $CP_A < CP_B$ or; 2) $CE_A < CE_B$ and $CP_A \leq CP_B$.
All configurations that are not Pareto-dominated by any other configuration are called `r index('Pareto-efficient')` and the set of all these configurations is the `r index("Pareto set", aside = TRUE)`.
The metric values corresponding to these Pareto set are referred to as the `r index("Pareto front", aside = TRUE)`.

The goal of multi-objective HPO is to approximate the Pareto front.
We will now demonstrate multi-objective HPO by tuning a decision tree on the `sonar` task with respect to the classification error, as a measure of model performance, and the number of selected features, as a measure of model complexity (in a decision tree the number of selected features is straightforward to obtain by simply counting the number of unique splitting variables).
Methodological details on multi-objective HPO can be found in @hpo_multi.

We will tune `cp`, `minsplit`, and `maxdepth`:

```{r optimization-082}
learner = lrn("classif.rpart",
  cp = to_tune(1e-04, 1e-1),
  minsplit = to_tune(2, 64),
  maxdepth = to_tune(1, 30)
)

measures = msrs(c("classif.ce", "selected_features"))
```

Note that as we tune with respect to multiple measures, the function `ti` creates a `r ref("TuningInstanceMultiCrit")` instead of a `r ref("TuningInstanceSingleCrit")`.
We also have to set `store_models = TRUE` as this is required by the selected features measure.

```{r optimization-083}
instance = ti(
  task = tsk("sonar"),
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measures = measures,
  terminator = trm("evals", n_evals = 30),
  store_models = TRUE
)
instance
```

As before we will select and run a tuning algorithm, in this example we will use random search:

```{r optimization-084,output=FALSE}
tuner = tnr("random_search")
tuner$optimize(instance)
```

Finally, we inspect the best-performing configurations, i.e., the Pareto set, and visualize the corresponding estimated Pareto front (@fig-pareto).

```{r optimization-085}
instance$archive$best()[, .(cp, minsplit, maxdepth, classif.ce, selected_features)]
```

```{r optimization-086,message=FALSE}
#| label: fig-pareto
#| fig-cap: Pareto front of selected features and classification error. Purple dots represent tested configurations, each blue dot individually represents a Pareto-optimal configuration and all blue dots together represent the Pareto front.
#| fig-alt: Scatter plot with selected_features on x-axis and classif.ce on y-axis. Plot shows around 15 purple dots and four blue dots at roughly (1, 0.301), (4, 0.299), (6, 0.285), (8, 0.28)representing the pareto front, blue dots are joined by a line.
#| echo: false
library(ggplot2)
library(viridisLite)

ggplot(as.data.table(instance$archive), aes(x = selected_features, y = classif.ce)) +
  geom_point(
    data = ,
    shape = 21,
    size = 3,
    fill = viridis(3, end = 0.8)[1],
    alpha = 0.8,
    stroke = 0.5) +
  geom_step(
    data = instance$archive$best(),
    direction = "hv",
    colour = viridis(3, end = 0.8)[2],
    linewidth = 1) +
  geom_point(
    data = instance$archive$best(),
    shape = 21,
    size = 3,
    fill = viridis(3, end = 0.8)[2],
    alpha = 0.8,
    stroke = 0.5) +
  theme_minimal()
```

Determining which configuration to use from the Pareto front is up to you.
By definition there is no optimal configuration so this may depend on your use-case, for example if you would prefer lower complexity at the cost of higher error than you might prefer a configuration where `selected_features = 1`.
You can select one configuration and pass it to a learner for training using `$result_learner_param_vals`, so if we want to select the second configuration we would run:

```{r}
learner = lrn("classif.rpart")
learner$param_set$values = instance$result_learner_param_vals[[2]]
learner$param_set
```

As multi-objective tuning requires manual intervention to select a configuration, it is currently not possible to use `auto_tuner`.
Furthermore, it can also be quite difficult to compare multiple models over multiple measures.

## Multi-Fidelity Tuning via Hyperband {#sec-hyperband}

Increasingly large datasets and search spaces and costly to train models make hyperparameter optimization a time-consuming task.
Recent HPO methods often also make use of evaluating a configuration at multiple fidelity levels.
For example, a neural network can be trained for an increasing number of epochs, gradient boosting can be performed for an increasing number of boosting iterations and training data can always be subsampled to a smaller fraction of all available data.
The general idea of *multi-fidelity* HPO is that the performance of a model obtained by using computationally cheap lower fidelity evaluation (few numbers of epochs or boosting iterations, only using a small sample of all available data for training) is predictive of the performance of the model obtained using computationally expensive full model evaluation and this concept can be leveraged to make HPO more efficient (e.g., only continuing to evaluate those configurations on higher fidelities that appear to be promising with respect to their performance).
The fidelity parameter is part of the search space and controls the trade-off between the runtime and preciseness of the performance approximation.

A popular multi-fidelity HPO algorithm is *Hyperband* [@li_2018].
After having evaluated randomly sampled configurations on low fidelities, Hyperband iteratively allocates more resources to promising configurations and terminates low-performing ones.
In the following example, we will optimize XGBoost and use the number of boosting iterations as the fidelity parameter.
This means Hyperband will allocate increasingly more boosting iterations to well-performing hyperparameter configurations.
Increasing the number of boosting iterations increases the time to train a model but generally also the performance.
It is therefore a suitable fidelity parameter.
However, as already mentioned, Hyperband is not limited to machine learning algorithms that are trained iteratively.
In the second example, we will tune a support vector machine and use the size of the training data as the fidelity parameter.
Some prior knowledge about pipelines (@sec-pipelines) is beneficial but not necessary to fully understand the examples.
In the following, the terms fidelity and budget are often used interchangeably.

### Hyperband Tuner {#sec-hyperband-tuner}

Hyperband [@li_2018] builds upon the Successive Halving algorithm by @jamieson_2016.
Successive Halving is initialized with the number of starting configurations $n$, the proportion of configurations discarded in each stage $\eta$, and the minimum $r{_{min}}$ and maximum $r{_{max}}$ budget of a single evaluation.
The algorithm starts by sampling $n$ random configurations and allocating the minimum budget $r{_{min}}$ to them.
The configurations are evaluated and $\frac{1}{\eta}$ of the worst-performing configurations are discarded.
The remaining configurations are promoted to the next stage and evaluated on a larger budget.
This continues until one or more configurations are evaluated on the maximum budget $r{_{max}}$ and the best-performing configuration is selected.
The total number of stages is calculated so that each stage consumes approximately the same overall budget.
Successive Halving has the disadvantage that is not clear whether we should choose a large $n$ and try many configurations on a small budget or choose a small $n$ and train more configurations on the full budget.

Hyperband solves this problem by running Successive Halving with different numbers of stating configurations starting on different budget levels.
The algorithm is initialized with the same parameters as Successive Halving except for $n$.
Each run of Successive Halving is called a bracket and starts with a different budget $r{_{0}}$.
A smaller starting budget means that more configurations can be evaluated.
The most exploratory bracket is allocated the minimum budget $r{_{min}}$.
The next bracket increases the starting budget by a factor of $\eta$.
In each bracket, the starting budget increases further until the last bracket $s = 0$ essentially performs a random search with the full budget $r{_{max}}$.
The number of brackets $s{_{max}} + 1$ is calculated with $s{_{max}} = {\log_\eta \frac{r{_{max}} }{r{_{min}}}}$.
Under the condition that $r{_{0}}$ increases by $\eta$ with each bracket, $r{_{min}}$ sometimes has to be adjusted slightly in order not to use more than $r{_{max}}$ resources in the last bracket.
The number of configurations in the base stages is calculated so that each bracket uses approximately the same amount of budget.
@tbl-hyperband shows a full run of the Hyperband algorithm.
The bracket $s = 3$ is the most exploratory bracket and $s = 0$ essentially performs a random search using the full budget.

+-----+-------------------+-------------------+-------------------+-------------------+
|     | $s = 3$           | $s = 2$           | $s = 1$           | $s = 0$           |
+-----+---------+---------+---------+---------+---------+---------+---------+---------+
| $i$ | $n_{i}$ | $r_{i}$ | $n_{i}$ | $r_{i}$ | $n_{i}$ | $r_{i}$ | $n_{i}$ | $r_{i}$ |
+=====+=========+=========+=========+=========+=========+=========+=========+=========+
| 0   | 8       | 1       | 6       | 2       | 4       | 4       | 4       | 8       |
+-----+---------+---------+---------+---------+---------+---------+---------+---------+
| 1   | 4       | 2       | 3       | 4       | 2       | 8       |         |         |
+-----+---------+---------+---------+---------+---------+---------+---------+---------+
| 2   | 2       | 4       | 1       | 8       |         |         |         |         |
+-----+---------+---------+---------+---------+---------+---------+---------+---------+
| 3   | 1       | 8       |         |         |         |         |         |         |
+-----+---------+---------+---------+---------+---------+---------+---------+---------+

: Hyperband schedule with the number of configurations $n_{i}$ and resources $r_{i}$ for each bracket $s$ and stage $i$, when $\eta = 2$ , $r{_{min}} = 1$ and $r{_{max}} = 8$ {#tbl-hyperband}

### Example XGBoost {#sec-hyperband-example-xgboost}

In this practical example, we will optimize the hyperparameters of XGBoost on the `spam` dataset.
We begin by constructing the learner.

```{r optimization-062}
library(mlr3hyperband)

learner = lrn("classif.xgboost")
```

As the next step we define the search space.
The `nrounds` parameter controls the number of boosting iterations.
We specify a range from 16 to 128 boosting iterations.
This is used as $r{_{min}}$ and $r{_{max}}$ within Hyperband.
We need to tag the parameter with `"budget"`to identify it as a fidelity parameter.
For the other hyperparameters, we take the search space for XGBoost from @hpo_practical.
This search space usually work well for a wide range of datasets.

```{r optimization-063}
learner$param_set$set_values(
  nrounds           = to_tune(p_int(16, 128, tags = "budget")),
  eta               = to_tune(1e-4, 1, logscale = TRUE),
  max_depth         = to_tune(1, 20),
  colsample_bytree  = to_tune(1e-1, 1),
  colsample_bylevel = to_tune(1e-1, 1),
  lambda            = to_tune(1e-3, 1e3, logscale = TRUE),
  alpha             = to_tune(1e-3, 1e3, logscale = TRUE),
  subsample         = to_tune(1e-1, 1)
)
```

We proceed to construct the tuning instance.
Note that we use `trm("none")` because Hyperband terminates itself after all brackets have been evaluated.

```{r optimization-064}
instance = ti(
  task = tsk("spam"),
  learner = learner,
  resampling = rsmp("holdout"),
  measures = msr("classif.ce"),
  terminator = trm("none")
)
```

We then construct the Hyperband tuner and specify `eta = 2`.
In general, Hyperband can start all over from the beginning once the last bracket is evaluated.
We control the number of Hyperband runs with the `repetition` argument.
The setting `repetition = Inf` is useful when a terminator should stop the optimization, for example based on runtime.

```{r optimization-065}
tuner = tnr("hyperband", eta = 2, repetitions = 1)
```

Using `eta = 2` and 16 to 128 boosting iterations results in the following schedule.
This only prints a data table with the schedule and does not modify the tuner.

```{r optimization-066}
hyperband_schedule(r_min = 16, r_max = 128, eta = 2)
```

We can now proceed with the tuning:

```{r optimization-067}
#| output: false
tuner$optimize(instance)
```

The result is the configuration with the best performance.

```{r optimization-068}
instance$result[, .(classif.ce, nrounds)]
```

Note that the archive resulting of a Hyperband run contains the additional columns `bracket` and `stage`:

```{r optimization-069}
as.data.table(instance$archive)[,
  .(bracket, stage, classif.ce, eta, max_depth, colsample_bytree)]
```

### Example Support Vector Machine {#sec-hyperband-example-svm}

In this example, we will optimize the hyperparameters of a support vector machine on the `sonar` dataset.
We begin by constructing the learner and setting `type` to `"C-classification"`.

```{r optimization-070}
learner = lrn("classif.svm", id = "svm", type = "C-classification")
```

The `r mlr3pipelines` package features a `r ref("PipeOp")` for subsampling data.
This will be helpful when using the size of the training data as a fidelity parameter.

```{r optimization-071}
po("subsample")
```

This pipeline operator controls the size of the training dataset with the `frac` parameter.
We connect the `po("subsample")`  with the learner and get a `r ref("GraphLearner")`.

```{r optimization-072}
graph_learner = as_learner(
  po("subsample") %>>%
  learner
)
```

The graph learner subsamples and then fits a support vector machine on the data subset.
The parameter set of the graph learner is a combination of the parameter sets of the pipeline operator and learner.

```{r optimization-073}
as.data.table(graph_learner$param_set)[, .(id, lower, upper, levels)]
```

Next, we create the search space.
We have to prefix the hyperparameters with the id of the pipeline operators, because this reflects the way how they are represented in the parameter set of the graph learner.
The `subsample.frac` is the fidelity parameter that must be tagged with `"budget"` in the search space.
In the following, the dataset size is increased from 3.7% to 100%.
For the other hyperparameters, we take the search space for support vector machines from @binder2020.
This search space usually work well for a wide range of datasets.

```{r optimization-074}
graph_learner$param_set$set_values(
  subsample.frac  = to_tune(p_dbl(3^-3, 1, tags = "budget")),
  svm.kernel      = to_tune(c("linear", "polynomial", "radial")),
  svm.cost        = to_tune(1e-4, 1e3, logscale = TRUE),
  svm.gamma       = to_tune(1e-4, 1e3, logscale = TRUE),
  svm.tolerance   = to_tune(1e-4, 2, logscale = TRUE),
  svm.degree      = to_tune(2, 5)
)
```

Support vector machines can often crash during training or take an extensive time to train given certain hyperparameters.
We therefore set a timeout of 30 seconds and specify a fallback learner (@sec-encapsulation-fallback) to handle these cases.

```{r optimization-075}
graph_learner$encapsulate = c(train = "evaluate", predict = "evaluate")
graph_learner$timeout = c(train = 30, predict = 30)
graph_learner$fallback = lrn("classif.featureless")
```

Let us create the tuning instance.
Again, we use `trm("none")` because Hyperband controls the termination itself.

```{r optimization-076}
instance = ti(
  task = tsk("sonar"),
  learner = graph_learner,
  resampling = rsmp("cv", folds = 3),
  measures = msr("classif.ce"),
  terminator = trm("none")
)
```

We create the tuner and set `eta = 3`.

```{r optimization-077}
tuner = tnr("hyperband", eta = 3)
```

Using `eta = 3` and a lower bound of 3.7% for the dataset size results in the following Hyperband schedule.

```{r optimization-078}
hyperband_schedule(r_min = 3^-3, r_max = 1, eta = 3)
```

We can now start the tuning.

```{r optimization-079}
#| output: false
tuner$optimize(instance)
```

We observe that the best model is a support vector machine with a polynomial kernel.

```{r optimization-080}
instance$result[, .(classif.ce, subsample.frac, svm.kernel)]
```

The archive contains all evaluated configurations.
We can proceed to further investigate the 8 configurations that were evaluated on the full dataset.
The configuration with the best classification error on the full dataset was sampled in the second bracket.
The classification error was estimated to be 30% on 33% of the dataset and decreased to 14% on the full dataset (see bright purple line in @fig-hyperband).

```{r optimization-081, eval=FALSE}
#| echo: false
#| label: fig-hyperband
#| fig-cap: "Optimization paths of the 8 configurations evaluated on the complete dataset."
#| fig-alt: "Image showing the performance of 8 configurations evaluated on different training dataset sizes. The classification error decreases on larger training set sizes."
library(ggplot2)

data = as.data.table(instance$archive)[, i := factor(.GRP), by = "svm.cost"]
top = data[subsample.frac == 1, i]
data = data[list(top), , on = "i"]

ggplot(data, aes(x = subsample.frac, y = classif.ce, group = i)) +
  geom_vline(xintercept = 0.037, color = "grey85") +
  geom_vline(xintercept = 0.111, color = "grey85") +
  geom_vline(xintercept = 0.333, color = "grey85") +
  geom_vline(xintercept = 1, color = "grey85") +
  geom_line(aes(color=i), show.legend = FALSE) +
  geom_point(aes(color=i), size = 3, show.legend = FALSE, position=position_jitter(height = 0.003, width = 0)) +
  scale_x_continuous(breaks = c(0.034, 0.11, 0.33, 1), labels = function(x) paste0(as.character(x *100), "%")) +
  scale_color_viridis_d(alpha = 0.8) +
  xlab("Training dataset Size") +
  ylab("Classification Error") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

## Bayesian Optimization {#sec-bayesian-optimization}

In this section we will take a deep dive into `r index('Bayesian Optimization')` (BO), also known as Model Based Optimization (MBO) \index{Model Based Optimization|see{Bayesian Optimization}}.
BO is implemented via the `r mlr3mbo` package.
The design of BO is slightly different from what we have seen so far in other tuning methods so to help motivate this we will spend a little more time in this section on theory and methodology.

In hyperparameter optimization @sec-optimization), learners are passed a hyperparameter configuration and evaluated on a given task via a resampling technique to estimate its generalization performance with the goal to find the optimal hyperparameter configuration.
In general, no analytical description for the mapping from hyperparameter configuration to performance exists and gradient information is also not available.
HPO is therefore a prime example for `r index('black-box optimization', aside = TRUE)`, which considers the optimization of a function whose mathematical structure and analytical description is unknown or unexploitable.
As a result, the only observable information is the output value (i.e., generalization performance) of the function given an input value (i.e., hyperparameter configuration).
In fact, as evaluating the performance of a learner can take a substantial amount of time, HPO is quite an expensive black-box optimization problem.
In the real-world, black-box optimization problems are encountered all the time, for example modelling real-world experiments like crash tests or chemical reactions.

Many optimization algorithm classes exist that can be used for black-box optimization, which differ in how they tackle this problem; for example we saw in @sec-optimization methods including grid/random search and briefly discussed evolutionary strategies.
Bayesian optimization refers to a class of sample-efficient iterative global black-box optimization algorithms that rely on a 'surrogate model' trained on observed data to model the black-box function.
This surrogate model is typically a non-linear regression model that tries to capture the unknown function using limited observed data.
During each iteration, BO algorithms employ an 'acquisition function' to determine the next candidate point for evaluation.
This function measures the expected 'utility' of each point within the search space based on the prediction of the surrogate model.
The algorithm then selects the candidate point with the best acquisition function value, and evaluates the black-box function at that point to then update the surrogate model.
This iterative process continues until a termination criterion is met, such as reaching a pre-specified maximum number of evaluations or achieving a desired level of performance.
BO is a powerful method that often results in good optimization performance, especially if the cost of the black-box evaluation becomes expensive and optimization budget is tight.

In the rest of this section we will first provide an introduction to black-box optimization with the `r bbotk` package and then introduce the building blocks of BO algorithms and examine their interplay and interaction during the optimization process before we assemble these building blocks in a ready to use black-box optimizer with `r mlr3mbo`.
Readers who are primarily interested in how to utilize BO for hyperparameter optimization without delving deep into the underlying building blocks may want to skip to @sec-bayesian-tuning.
Detailed introductions to black-box optimization and BO are given in @hpo_practical, @hpo_automl and @garnett_2022.

As a running example throughout this section we will optimize the sinusoidal function $f: [0, 1] \rightarrow \mathbb{R}, x \mapsto 2x + \sin(14x)$ (@fig-bayesian-optimization-sinusoidal), which is characterized by two local minima and one global minimum.

### Black-Box Optimization {#sec-black-box-optimization}

The `r bbotk` (black-box optimization toolkit) package is the workhorse package for general black-box optimization within the `mlr3` ecosystem.
At the heart of the package are the R6 classes:

* `r ref("OptimInstanceSingleCrit")` and `r ref("OptimInstanceMultiCrit")`, which are used to construct an `r index('optimization instance', aside = TRUE)` that describes the optimization problem and stores the results
* `r ref("Optimizer")` which is used to construct and configure optimization algorithms.

These classes might look familiar after reading @sec-optimization, and in fact `r ref("TuningInstanceSingleCrit")` and `r ref("TuningInstanceMultiCrit")` inherit from `OptimInstanceSingle/MultiCrit` and `r ref("Tuner")` is closely based on `Optimizer`.

`OptimInstanceSingleCrit` relies on an `r ref("Objective", aside = TRUE)` function that wraps the actual mapping from a domain (all possible function inputs) to a codomain (all possible function outputs).

Objective functions can be created using different classes, all of which inherit from `r ref("Objective")`.
These classes provide different ways to define and evaluate objective functions and picking the right one will reduce type conversion overhead:

* `r ref("ObjectiveRFun")` wraps a function that takes a list describing a *single configuration* as input where elements can be of any type. It is suitable when the underlying function evaluation mechanism is given by evaluating a single configuration at a time.
* `r ref("ObjectiveRFunMany")` wraps a function that takes a list of *multiple configurations* as input where elements can be of any type and even mixed types. It is useful when the function evaluation of multiple configurations can be parallelized.
* `r ref("ObjectiveRFunDt")` wraps a function that operates on a `data.table`. It allows for efficient vectorized or batched evaluations directly on the `data.table` object, avoiding unnecessary data type conversions.

To start translating our problem to code we will use the `r ref("ObjectiveRFun")` class to take a single configuration as input.
The `Objective` requires specification of the function to optimize and its domain and codomain.
By tagging the codomain with `"minimize"` or `"maximize"` we specify the optimization direction.
Note how below our optimization function takes a `list` as an input with one element called `x`.

```{r}
library(bbotk)
sinus_1D = function(xs) 2 * xs$x * sin(14 * xs$x)

domain = ps(x = p_dbl(lower = 0, upper = 1))
codomain = ps(y = p_dbl(tags = "minimize"))
objective = ObjectiveRFun$new(sinus_1D,
  domain = domain, codomain = codomain)
```

We can visualize are objective by generating a grid of points on which we evaluate the function (@fig-bayesian-optimization-sinusoidal), this will help us identify its local minima and global minimum.

```{r}
#| label: fig-bayesian-optimization-sinusoidal
#| fig-cap: Visualization of the sinusoidal function. Local minima in blue triangles and global minimum in the red circle.
#| fig-alt: Line graph from (0,1) on the x-axis to (-2,2) on the y-axis; labelled 'x' and 'y' respectively. The line starts with a local minimum at (0,0), increases and then has a local minimum at around (0.35,-0.69), the function then increases and then decreases to the global minimum at around (0.79, -1.56).
library(ggplot2)
library(viridisLite)

xydt = generate_design_grid(domain, resolution = 1001)$data
xydt[, y := objective$eval_dt(xydt)$y]
optima = data.table(x = c(0, 0.3509406, 0.7918238))
optima[, y := objective$eval_dt(optima)$y]
optima[, type := c("local", "local", "global")]

ggplot(aes(x = x, y = y), data = xydt) +
  geom_line() +
  geom_point(aes(pch = type, color = type), size = 4, data = optima) +
  theme_minimal() +
  theme(legend.position = "none")
```

The global minimum, `r xydt[y == min(y), x]`, corresponds to the point of the domain with the lowest function value:

```{r}
xydt[y == min(y), ]
```

With the objective function defined, we can proceed to optimize it using `r ref("OptimInstanceSingleCrit")`.
This class allows us to wrap the objective function and explicitly specify a search space.
The search space defines the set of input values we want to optimize over, and it is typically a subset or transformation of the domain, though by default the entire domain is taken as the search space.
In black-box optimization, it is common for the domain, and hence also the search space, to have finite box constraints.
Similarly to HPO, transformations can sometimes be used to more efficiently search the space (@sec-logarithmic-transformations).

In the following, we use a simple random search to optimize the sinusoidal function over the whole domain and inspect the result from the `instance` in the usual way (@sec-tuner).
Analogously to tuners, `Optimizer`s in `r ref_pkg("bbotk")` are stored in the `r ref('mlr_optimizers')` dictionary and can be constructed with `r ref('opt()', aside = TRUE)`.

```{r, output=FALSE}
instance = OptimInstanceSingleCrit$new(objective,
  search_space = domain,
  terminator = trm("evals", n_evals = 20))
optimizer = opt("random_search", batch_size = 20)
optimizer$optimize(instance)
```

Similarly to how we can use `r ref("tune()")` to construct a tuning instance, here we can use `r ref("bb_optimize()", aside = TRUE)`:

```{r}
instance = bb_optimize(objective, method = "random_search",
  max_evals = 20)
instance$result
```

Now we have introduced the basic black-box optimization setup, we can introduce the building blocks of any Bayesian optimization algorithm.

### Building Blocks of Bayesian Optimization {#sec-bayesian-optimization-blocks}

Bayesian optimization (BO) is a global optimization algorithm that usually follows the following process:

1. Generate and evaluate an `r index('initial design', aside = TRUE)`
2. Loop:
    a. Fit a `r index('surrogate model', aside = TRUE)` on the archive of all observations made so far to model the unknown black-box function.
    b. Optimize an `r index('acquisition function', aside = TRUE)` to determine which points of the search space are promising candidate(s) that should be evaluated next.
    c. Evaluate the next candidate(s) and update the archive of all observations made so far.
    d. Check if a given termination criterion is met, if not go back to (a).

The acquisition function relies on the mean and standard deviation prediction of the surrogate model and requires no evaluation of the true black-box function, making it comparably cheap to optimize.
A good acquisition function will balance *exploiting* knowledge about regions where we observed that performance is good and the surrogate model has low uncertainty, with *exploring* regions where it has not yet evaluated points and as a result the uncertainty of the surrogate model is high.

We refer to these elements as the 'building blocks' of BO as it is a highly modular algorithm; as long as the above structure is in place, then the surrogate models, acquisition functions, and optimizers are all interchangeable.
The design of `r mlr3mbo` reflects this modularity, with the base class for `r ref("OptimizerMbo")` holding all the key elements: the BO algorithm loop structure (`r ref("loop_function")`), *surrogate* model (`r ref("Surrogate")`),  *acquisition function* (`r ref("AcqFunction")`), and *acquisition function optimizer* (`r ref("AcqOptimizer")`).
In this section, we will provide a more detailed explanation of these building blocks and explore their interplay and interaction during optimization.

#### The Initial Design {#sec-bayesian-optimization-initial}

Before we can fit a surrogate model to model the unknown black-box function, we need data.
The initial set of points that is evaluated before a surrogate model can be fit is referred to as the `r index('initial design')`.

`r mlr3mbo` allows you to either construct the initial design manually, or let `r ref("loop_function")` construct and evaluate this for you.
In this section we will demonstrate the first method, which requires more user-input but therefore allows more control over the initial design.

You could create an initial design manually but a more straightforward method might be to use one of the four design generators in `r paradox`:

* `r ref("generate_design_random()")`: Generate points uniformly at random
* `r ref("generate_design_grid()")`: Generate points in a uniform sized grid
* `r ref("generate_design_lhs()")`: Latin hypercube sampling [@Stein1987]
* `r ref("generate_design_sobol()")`: Sobol sequence [@Niederreiter1988]

@fig-bayesian-optimization-designs illustrates the difference in generated designs from these four methods assuming an initial design of size nine and a domain of two numeric variables from 0 to 1.
We observe that a random design does not necessarily cover the search well and in this example simply due to bad luck samples points close to each other leaving large areas unexplored.
The grid design results in points being equidistant from their nearest neighbor but does not cover the search space well as areas between points are unexplored.
In contrast, the LHS design provides a good space-filling property, as it ensures that each interval of each input variable (spanned by the horizontal and vertical dotted lines) is represented by exactly one sample point, which usually results in a more even coverage of the search space and a better representation of the distribution of the input variables (as seen in the marginal distributions).
The Sobol design works similarly to LHS but does not guarantee even coverage for a small number of samples.
However, constructing a Sobol design is more efficient than LHS, especially as the number of samples and dimensions grows and in fact the coverage of Sobol is better than LHS when the number of dimensions is large.

```{r, echo = FALSE}
#| label: fig-bayesian-optimization-designs
#| fig-cap: Comparing different samplers for constructing an initial design of 9 points on a domain of two numeric variables ranging from 0 to 1. Dotted horizontal and vertical lines partition the domain into equally sized bins. Histograms on the top and right visualize the marginal distributions of the generated sample.
#| fig-alt: "Plot shows four grids with x_1 on x-axis ranging from 0 to 1 and x_2 on y-axis ranging from 0 to 1. Each grid has bars above them and to the right representing marginal distributions. Top left: 'Random Design' nine points are scattered randomly across the grid with poor coverage. Marginal distributions are also random. Top right: 'Grid Design', points are uniformly scattered across the grid on lines x_1=0,x_1=0.5,x_1=1 and same for x_2. Marginal distributions show three long bars at each of the corresponding lines. Bottom left: 'LHS Design', points appear randomly scattered however marginal distributions are completely equal with equal sized bars along each axis. Bottom right: 'Sobol Design', very similar to 'LHS Design' however one of the bars in the marginal distribution is slightly longer than the others."

library(ggExtra)
library(gridExtra)

sample_domain = ps(x1 = p_dbl(lower = 0, upper = 1),
  x2 = p_dbl(lower = 0, upper = 1))
random_design = generate_design_random(sample_domain, n = 9)$data
grid_design = generate_design_grid(sample_domain, resolution = 3)$data
lhs_design = generate_design_lhs(sample_domain, n = 9)$data
sobol_design = generate_design_sobol(sample_domain, n = 9)$data

qs = seq(from = 0, to = 1, length.out = 10)

g_random = ggplot(aes(x = x1, y = x2), data = random_design) +
  geom_point(size = 3L) +
  geom_vline(xintercept = qs, linetype = 2) +
  geom_hline(yintercept = qs, linetype = 2) +
  labs(title = "Random Design", x = expression(x[1]), y = expression(x[2])) +
  theme_minimal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1))
g_random = ggMarginal(g_random, type = "histogram", bins = 10)

g_grid = ggplot(aes(x = x1, y = x2), data = grid_design) +
  geom_point(size = 3L) +
  geom_vline(xintercept = qs, linetype = 2) +
  geom_hline(yintercept = qs, linetype = 2) +
  labs(title = "Grid Design", x = expression(x[1]), y = expression(x[2])) +
  theme_minimal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1))
g_grid = ggMarginal(g_grid, type = "histogram", bins = 10)

g_lhs = ggplot(aes(x = x1, y = x2), data = lhs_design) +
  geom_point(size = 3L) +
  geom_vline(xintercept = qs, linetype = 2) +
  geom_hline(yintercept = qs, linetype = 2) +
  labs(title = "LHS Design", x = expression(x[1]), y = expression(x[2])) +
  theme_minimal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1))
g_lhs = ggMarginal(g_lhs, type = "histogram", bins = 10)

g_sobol = ggplot(aes(x = x1, y = x2), data = sobol_design) +
  geom_point(size = 3L) +
  geom_vline(xintercept = qs, linetype = 2) +
  geom_hline(yintercept = qs, linetype = 2) +
  labs(title = "Sobol Design", x = expression(x[1]), y = expression(x[2])) +
  theme_minimal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1))
g_sobol = ggMarginal(g_sobol, type = "histogram", bins = 10)

grid.arrange(g_random, g_grid, g_lhs, g_sobol, nrow = 2, ncol = 2)
```

Whichever of these methods you choose, the result is simply a `data.table`:

```{r}
generate_design_random(sample_domain, n = 3)$data
```

Therefore you could also specify a completely custom initial design by defining you own `data.table`.
Either way, when manually constructing an initial design (as opposed to letting `loop_function` automate this), it needs to be evaluated on the `r ref("OptimInstance")` before optimizing it.
Returning to our running example of minimizing the sinusoidal function, we will evaluate a custom initial design:

```{r}
instance = OptimInstanceSingleCrit$new(objective,
  terminator = trm("evals", n_evals = 20))
design = data.table(x = c(0.1, 0.34, 0.65, 1))
instance$eval_batch(design)
instance$archive$data
```

We can see how each point in our design was evaluated by the sinusoidal function, giving us data we can now use to start the iterative BO algorithm by fitting the surrogate model on that data.

#### Surrogate Model {#sec-bayesian-optimization-surrogate}

A `r index('surrogate model')` wraps a regression learner that models the unknown black-box function based on observed data.
In `r mlr3mbo`, the `r ref("SurrogateLearner", aside = TRUE)` is a higher-level R6 class inheriting from the base `r ref("Surrogate")` class, designed to construct and manage the surrogate model, including automatic construction of the `TaskRegr` that the learner should be trained on at each iteration of the BO loop.

Any regression learner in `mlr3` can be used, however, most acquisition functions depend on both mean and standard deviation predictions from the surrogate model and therefore not all learners are suitable for all scenarios.
Typical choices of regression learners used as surrogate models include Gaussian processes (`lrn("regr.km")`) for low dimensional numeric search spaces and random forests (e.g., `lrn("regr.ranger")`) for higher dimensional mixed (and / or hierarchical) search spaces.
A detailed introduction to Gaussian processes can be found in @williams_2006 and in-depth focus to Gaussian processes in the context of surrogate models in BO is given in @garnett_2022.
In this example we use a Gaussian process with Matérn 5/2 kernel, which uses `BFGS` as an optimizer to find the optimal kernel parameters and set `trace = FALSE` to prevent too much output during fitting.

```{r}
lrn_gp = lrn("regr.km", covtype = "matern5_2", optim.method = "BFGS",
  control = list(trace = FALSE))
```

A `r ref("SurrogateLearner")` can be constructed by passing a `LearnerRegr` object to the sugar function `r index('srlrn()', aside = TRUE, code = TRUE)`', alongside the `archive` of the instance:

```{r}
library(mlr3mbo)
surrogate = srlrn(lrn_gp, archive = instance$archive)
```

Internally, the regression learner is fit on a `TaskRegr` where features are the variables of the domain and the target is the codomain, the data is from the `r ref("bbotk::Archive")` of the `r ref("bbotk::OptimInstance")` that is to be optimized.

In our running example we have already initialized our archive with the initial design, so we can update our surrogate model, which essentially fits the Gaussian process, note how we use `$learner` to access the wrapped model:

```{r}
surrogate$update()
surrogate$learner$model
```

Having introduced the concept of a surrogate model, we can now move on to the acquisition function, which makes use of the surrogate model predictions to decide which candidate to evaluate next.

#### Acquisition Function {#sec-bayesian-optimization-acquisition}

Roughly speaking, an `r index('acquisition function')` relies on the prediction of a surrogate model and quantifies the expected 'utility' of each point of the search space if it were to be evaluated in the next iteration.

A popular example is the expected improvement [@jones_1998], which tells us how much we can expect a candidate point to improve over the best function value observed so far (the 'incumbent'), given the performance prediction of the surrogate model:

$$
\alpha_{\mathrm{EI}}(\mathbf{x}) = \mathbb{E} \left[ \max \left( f_{\mathrm{min}} - Y(\mathbf{x}), 0 \right) \right]
$$

Here, $Y(\mathbf{x)}$ is the surrogate model prediction (a random variable) for a given point $\mathbf{x}$ (which when using a Gaussian process follows a normal distribution) and $f_{\mathrm{min}}$ is the best function value observed so far (assuming minimization).
Optimizing the expected utility requires mean and standard deviation predictions from the model.

In `r mlr3mbo`, acquisition functions (of class `r ref( "AcqFunction")`) are stored in the `r ref("mlr_acqfunctions")` dictionary and can be constructed with `r ref('acqf()', aside = TRUE)`, passing the key of the method you want to use and our surrogate learner.
In our running example, we will use the expected improvement (`acqf("ei")`) to choose the next candidate for evaluation, which updates (`$update()`) the incumbent to ensure it is still the best value observed so far.

```{r}
acq_function = acqf("ei", surrogate = surrogate)
acq_function$update()
acq_function$y_best
```

You can use `$eval_dt` to evaluate the acquisition function for the domain given as `data.table`.
In @fig-bayesian-optimization-ei we evaluated the expected improvement on a uniform grid of points between 0 and 1 using the predicted mean and standard deviation from the Gaussian process.
We can see that the expected improvement is high in regions where the mean prediction (black dots) of the Gaussian process is low.

```{r}
#| label: fig-bayesian-optimization-ei
#| fig-cap: Expected improvement (darkred) based on the mean and uncertainty prediction (lightblue) of the Gaussian process surrogate model trained on an initial design of four points (black). Ribbons represent the mean plus minus the standard deviation prediction.
#| fig-alt: Expected improvement based on the mean and uncertainty prediction of the Gaussian process surrogate model. The expected improvement is high where the mean prediction is low but the standard deviation prediction still suggests some uncertainty.

xydt = generate_design_grid(domain, resolution = 1001)$data
# evaluate our sinusoidal function
xydt[, y := objective$eval_dt(xydt)$y]
# evluate expected improvement
xydt[, ei :=  acq_function$eval_dt(xydt[, "x"])]
# make predictions from our data
xydt[, c("mean", "se") :=  surrogate$predict(xydt[, "x"])]
xydt[1:3]

ggplot(xydt, mapping = aes(x = x, y = y)) +
  geom_point(size = 2, data = instance$archive$data) +
  geom_line() +
  geom_line(aes(y = mean), colour = "steelblue", linetype = 2) +
  geom_ribbon(aes(min = mean - se, max = mean + se),
    fill = "steelblue", alpha = 0.1) +
  geom_line(aes(y = ei * 40), linewidth = 1, colour = "darkred") +
  scale_y_continuous("y",
    sec.axis = sec_axis(~ . * 0.025, name = "EI",
      breaks = c(0, 0.025, 0.05))) +
  theme_minimal()
```

We will now proceed to optimize the acquisition function itself to find the candidate with the largest expected improvement.

#### Acquisition Function Optimizer {#sec-bayesian-optimization-acquisitionopt}

An `r index('acquisition function optimizer')` of class `r ref("AcqOptimizer", aside = TRUE)` is used to optimize the acquisition function by efficiently searching the space of potential candidates within a limited computational budget.

Due to the non-convex nature of most commonly used acquisition functions [@garnett_2022] it is typical to employ global optimization techniques for acquisition function optimization.
Widely used approaches for optimizing acquisition functions include derivative-free global optimization methods like branch and bound algorithms, such as the DIRECT algorithm [@jones_1993_lipschitzian], as well as multi-start local optimization methods, such as running the L-BFGS-B algorithm [@byrd1995limited] or a local search multiple times from various starting points [@kim_2021].
Assuming a purely numeric search space, the utilization of gradient information in acquisition function optimization is theoretically possible, but its feasibility relies on the specific surrogate model chosen.
However, even when employing a Gaussian process as surrogate model, allowing for the calculation of gradients of the acquisition function with respect to the input variables, the issue of vanishing gradients in the mean and standard deviation predictions concerning the input variables presents a significant challenge [@garnett_2022].
Consequently, in most cases, the optimization problem of the acquisition function can be regarded as a black box optimization problem itself, but a much cheaper one than the original.

`AcqOptimizer` objects are constructed with `r ref("mlr3mbo::acqo()", aside = TRUE)`, which takes as input a `r ref('botk::Optimizer')`, a `r ref("bbotk::Terminator")`, and the acquisition function.
Optimizers are stored in the `ref("mlr_optimizers")` dictionary and can be constructed with the sugar function `r ref('opt()', aside = TRUE)`.
The terminators are the same as those introduced in @sec-terminator.

Below we use a non-linear optimization with the DIRECT algorithm and we terminate the acquisition function if there is no improvement of at leas `1e-5` for `100` iterations.
The `$optimize()` method optimizes the acquisition function and returns the next candidate.

```{r}
acq_optimizer = acqo(
  optimizer = opt("nloptr", algorithm = "NLOPT_GN_ORIG_DIRECT"),
  terminator = trm("stagnation", iters = 100, threshold = 1e-5),
  acq_function = acq_function
)

candidate = acq_optimizer$optimize()
candidate
```

We have now manually shown how to run a single loop of the BO algorithm.
In practice one would use `r ref("OptimizerMbo")` to put all these pieces together to automate the process.
Before demonstrating this class we will first take a step back and introduce the `loop_function` which tells the algorithm how it should be run.

#### Using and Building Loop Functions {#sec-bayesian-optimization-loop}

The `r ref("loop_function", index = TRUE)` determines the behavior of the BO algorithm on a global level, i.e., how the subroutine should look like that is performed at each iteration to generate new candidates for evaluation.
Loop functions are stored in the `r ref("mlr_loop_functions")` dictionary.
As these are `S3` (not `R6`) classes, they can be simply loaded by just referencing the `key` (i.e., there is no constructor required).

```{r}
as.data.table(mlr_loop_functions)[, .(key, label, instance)]
```

Loop functions are relatively simple functions that take as input the classes that we have just discussed and define the BO loop.
You could pick and use one of the loop functions included in the dictionary above, or you can write your own for finer control over the BO process.
For example, a common choice of loop function is the Efficient Global Optimization (EGO) algorithm [@jones_1998] (`r ref('bayesopt_ego()')`), a simplified version of this code is shown at the end of this section to help demonstrate the algorithm and its purpose.
In short, the code sets up the relevant components discussed above and then loops the steps above:  1) update the surrogate model 2) update the acquisition function 3) optimize the acquisition function to yield a new candidate 4) evaluate the candidate and add it to the archive.
Two additional points to note are that: firstly the function checks to see if the initial grid has been manually constructed (as we did above), if it has not then it generates a grid using the Sobol method of size `init_design_size` or if that is null then of size four times the dimensionality of the search space; and secondly if there is an error during the loop then a fallback is used where the next candidate is proposed uniformly at random, ensuring that the process continues even in the presence of potential issues, we will return to this in @sec-practical-bayesian-optimization.

```{r, eval=FALSE}
bayesopt_ego = function(
    instance,
    surrogate,
    acq_function,
    acq_optimizer,
    init_design_size = NULL
  ) {

  # setting up the building blocks
  surrogate$archive = instance$archive # archive
  acq_function$surrogate = surrogate # surrogate model
  acq_optimizer$acq_function = acq_function # acquisition function

  # initial design
  search_space = instance$search_space
  # if initial design has not been manually constructed then use Sobol method
  if (is.null(init_design_size) && instance$archive$n_evals == 0L) {
    init_design_size = 4L * search_space$length
  }
  if (!is.null(init_design_size) && instance$archive$n_evals == 0L) {
    design = generate_design_sobol(search_space, n = init_design_size)$data
    instance$eval_batch(design)
  }

  # MBO loop
  repeat {
    candidate = tryCatch({
      # update the surrogate model
      acq_function$surrogate$update()
      # update the acquisition function
      acq_function$update()
      # optimize the acquisition function to yield a new candidate
      acq_optimizer$optimize()
    }, mbo_error = function(mbo_error_condition) {
      generate_design_random(search_space, n = 1L)$data
    })

    # evaluate the candidate and add it to the archive
    instance$eval_batch(candidate)
    if (instance$is_terminated) break
  }

  return(instance)
}
```

We are now ready to put everything together to automate the BO process.

### Automating BO with OptimizerMbo {#sec-bayesian-black-box-optimization}

`r ref("OptimizerMbo")` can be used to assemble the building blocks described above into a single object that can then be optimized.
The other benefit of this method is that you do not need to pass any of these building blocks to each other as the `r ref('opt()', aside = TRUE)` constructor will do this for you:

```{r}
surrogate = srlrn(lrn("regr.km", covtype = "matern5_2",
  optim.method = "BFGS", control = list(trace = FALSE)))
acq_function = acqf("ei")
acq_optimizer = acqo(opt("nloptr", algorithm = "NLOPT_GN_ORIG_DIRECT"),
  terminator = trm("stagnation", iters = 100, threshold = 1e-5))

optimizer = opt("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)
```

::: {.callout-tip}
Additional arguments for customizing certain loop functions can be passed through with the `args` parameter of `opt()`.
:::

In this example we will use the same initial design that we created before and will optimize our algorithm using `$optimize()`:

```{r}
instance = OptimInstanceSingleCrit$new(objective,
  terminator = trm("evals", n_evals = 20))
design = data.table(x = c(0.1, 0.34, 0.65, 1))
instance$eval_batch(design)
optimizer$optimize(instance)

instance$archive$best()
```

Using only a few evaluations, BO comes close to the true global optimum (0.792).
@fig-bayesian-optimization-sampling shows the sampling trajectory of candidates as the algorithm progressed, we can see that focus is increasingly given to more regions around the global optimum.
However, even in later optimization stages, the algorithm still explores new areas, illustrating that the expected improvement acquisition function indeed balances exploration and exploitation as we required.

```{r, echo = FALSE}
#| label: fig-bayesian-optimization-sampling
#| fig-cap: Sampling trajectory of the BO algorithm. Points of the initial design in black triangles. Sampled points are in dots with colour progressing from dark red to light red as the algorithm progresses.
#| fig-alt: Line graph of the same sinusoidal function as before but now there are dots from white to red along the line. There are more dots around the global minimum in later stages but still a spread of dots throughout the line.
ggplot() +
  geom_line(aes(x = x, y = y), data = xydt) +
  geom_point(aes(x = x, y = y), colour = "black", fill = "black", size = 2, pch = 24,
    data = instance$archive$data[batch_nr == 1]) +
  geom_point(aes(x = x, y = y, colour = batch_nr), size = 2,
    data = instance$archive$data[batch_nr > 1]) +
  scale_color_gradient(low = "darkred", high = "lavenderblush") +
  labs(colour = "Evaluation Number") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### Bayesian Optimization for HPO {#sec-bayesian-tuning}

`r mlr3mbo` can be used for HPO by making use of `r ref("TunerMbo", aside = TRUE)`, which is a wrapper around `r ref("OptimizerMbo")` and therefore works in the exact same way.
As an example, below we will tune the `cost` and `gamma` parameters of a classification SVM with a radial kernel on the `sonar` task with 3-fold CV for the inner and outer resamplings.
We set up `tnr("mbo")` using the same objects constructed above and then run our tuning experiment as usual:

```{r}
tuner = tnr("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)

lrn_svm = lrn("classif.svm", kernel = "radial", type = "C-classification",
  cost  = to_tune(1e-5, 1e5, logscale = TRUE),
  gamma = to_tune(1e-5, 1e5, logscale = TRUE))

instance = tune(
  tuner,
  task = tsk("sonar"),
  lrn_svm,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.ce"),
  term_evals = 25
)
```

Multi-objective tuning is also possible with BO with algorithms using many different design choices, for example whether they use a scalarization approach of objectives and only rely on a single surrogate model, or fit a surrogate model for each objective.
More details on multi-objective BO can for example be found in @Horn2015 or @Morales2022.

Below we will illustrate multi-objective tuning using the ParEGO [@knowles_2006] loop function.
ParEGO (`r ref("bayesopt_parego()")`) tackles multi-objective BO via a scalarization approach and models a single scalarized objective function via a single surrogate model and then proceeds to find the next candidate for evaluation making use of a standard single-objective acquisition function such as the expected improvement.
Other compatible loop functions can be found by looking at the `"instance"` column of `r ref('mlr_loop_functions')`.
We will tune three parameters of a decision tree with respect to true positive and true negative rate, the Pareto front is visualized in @fig-pareto-bayesopt.

```{r}
tuner = tnr("mbo",
  loop_function = bayesopt_parego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)

lrn_rpart = lrn("classif.rpart",
  cp = to_tune(1e-04, 1e-1),
  minsplit = to_tune(2, 64),
  maxdepth = to_tune(1, 30)
)

instance = tune(
  tuner,
  task = tsk("sonar"),
  lrn_svm,
  resampling = rsmp("cv", folds = 3),
  measure = msrs(c("classif.tpr", "classif.fpr")),
  term_evals = 25
)
```

```{r, echo = FALSE}
#| label: fig-pareto-bayesopt
#| fig-cap: Pareto front of TPR and FPR obtained via ParEGO. Purple dots represent tested configurations, each blue dot individually represents a Pareto-optimal configuration and all blue dots together represent the Pareto front.
#| fig-alt: Scatter plot with classif.tpr on x-axis (between 0.75 and 1.00) and classif.fpr on y-axis (between 0.2 and 1.0). The Pareto front is shown as the set of points at roughly (0.2, 0.9), (0.4, 0.93), (0.87, 1.0).
ggplot(aes(x = classif.tpr, y = classif.fpr),
  data = as.data.table(instance$archive)) +
  geom_point(
    data = ,
    shape = 21,
    size = 3,
    fill = viridis(3, end = 0.8)[1],
    alpha = 0.8,
    stroke = 0.5) +
  geom_step(
    data = instance$archive$best(),
    direction = "hv",
    colour = viridis(3, end = 0.8)[2],
    linewidth = 1) +
  geom_point(
    data = instance$archive$best(),
    shape = 21,
    size = 3,
    fill = viridis(3, end = 0.8)[2],
    alpha = 0.8,
    stroke = 0.5) +
  theme_minimal()
```

### Noisy Bayesian Optimization {#sec-noisy-bayesian-optimization}

So far, we implicitly assumed that the black-box function we are trying to optimize is deterministic, i.e., repeatedly evaluating the same point will always return the same objective function value.
However, real world black-box functions are often noisy, which means that repeatedly evaluating the same point will return different objective function values due to background noise on top of the black-box function.
For example, if you were modelling a machine in a factory to estimate the rate of production, even if all parameters of the machine were controlled, we would still expect different performance at different times due to uncontrollable background facts such as environmental conditions.

In `r bbotk`, you can mark an `r ref("Objective")` object as noisy by passing the `"noisy"` tag to the `properties` parameter, which allows us to use methods that can treat such objectives differently.

```{r}
sinus_1D_noisy = function(xs) {
  y = 2 * xs$x * sin(14 * xs$x) + rnorm(1, mean = 0, sd = 0.1)
  y
}
domain = ps(x = p_dbl(lower = 0, upper = 1))
codomain = ps(y = p_dbl(tags = "minimize"))
objective_noisy = ObjectiveRFun$new(sinus_1D_noisy,
  domain = domain, codomain = codomain, properties = "noisy")
```

Noisy objectives can be treated in different ways:

1. A surrogate model can be used to incorporate the noise
2. An acquisition function can be used that respects noisiness
3. The final best point(s) after optimization (i.e., the `$result` field of the instance) can be chosen in a way to reflect noisiness

In the first case, instead of using an interpolating Gaussian process, we could instead use Gaussian process regression that estimates the measurement error by setting `nuggest.estim = TRUE`:

```{r, output=FALSE}
srlrn(lrn("regr.km", nugget.estim = TRUE))
```

This will result in the Gaussian process not perfectly interpolating training data and the standard deviation prediction associated with the training data will be non-zero, reflecting the uncertainty in the observed function values due to the measurement error.
A more in-depth discussion of noise free vs. noisy observations in the context of Gaussian processes can be found in Chapter 2 of @williams_2006.

For the second option, one example of an acquisition function that properly respects noisiness is the Augmented expected improvement [@huang_2012] (`acqf("aei")`) which essentially rescales the expected improvement, taking measurement error into account.

Finally, `r mlr3mbo` allows for explicitly specifying how the final result after optimization is assigned to the instance (i.e., what will be saved in `instance$result`) with a `r index('result assigner', aside = TRUE)`, which can be specified during the construction of a `OptimizerMbo` or `TunerMbo`.
`r ref('ResultAssignerSurrogate')` uses a surrogate model to predict the mean of all evaluated points and proceeds to choose the point with the best mean prediction as the final optimization result.
In contrast, the default method, `r ref('ResultAssignerArchive')`, just picks the best point according to the evaluations logged in `archive`.
Result assigners are stored in the `r ref('mlr_result_assigners')` dictionary and can be constructed with `r ref('ras()')`.

```{r, eval = FALSE}
tnr("mbo",
  loop_function = bayesopt_parego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer,
  result_assigner = ras("surrogate")
)
```

### Practical Considerations in Bayesian Optimization {#sec-practical-bayesian-optimization}

`r mlr3mbo` tries to use 'intelligent' defaults regarding the choice of surrogate model, acquisition function, acquisition function optimizer and even the loop function.
For example, in the case of a purely numeric search space, `r mlr3mbo` will by default use a Gaussian process as surrogate model and a random forest as fallback learner and additionally encapsulates (@sec-encapsulation-fallback) the learner via the `r ref_pkg("evaluate")` package.
In the case of a mixed or hierarchical search space, `r mlr3mbo` will use a random forest as surrogate model.
As a result of defaults existing for all building blocks, users can perform BO without specifying any building blocks and can still expect decent optimization performance.
To see an up-to-date overview of these defaults, take a look at the help page `r ref('mbo_defaults')`.
We will finish this section with some practical considerations to think about when using BO.

#### Error handling {.unnumbered .unlisted}

In the context of BO, there is plenty of room for potential failure of building blocks which could break the whole process.
For example, if two points in the training data are too close to each other, fitting the Gaussian process surrogate model can fail.

`r mlr3mbo` has several built-in safety nets to catch errors.
`r ref("Surrogate")` includes the `catch_errors` configuration control parameter, which, if set to `TRUE`, catches all errors that occur during training or updating of the surrogate model.
`r ref("AcqOptimizer")` also has the `catch_errors` configuration control parameter, which can be used to catch all errors that occur during the acquisition function optimization, either due to the surrogate model failing to predict or the acquisition function optimizer erroring.
If errors are caught in any of these steps, the standard behavior of any `r ref("loop_function")` is to trigger a fallback, which proposes the next candidate uniformly at random.
Note, when setting `catch_errors = TRUE` for the `r ref("AcqOptimizer")`, it is usually not necessary to also explicitly set `catch_errors = TRUE` for the `r ref("Surrogate")`, though this may be useful when debugging.

In the worst-case scenario, if all iterations errored, the BO algorithm will simply perform a random search.
Ideally, fallback learners (@sec-encapsulation-fallback) could also be used, which will be employed before proposing the next candidate randomly.
The value of the acquisition function is also always logged into the archive of the optimization instance so inspecting this is a good idea to ensure the algorithm behaved as expected and that the acquisition function column is populated as expected (`instance$archive$data`).

#### Surrogate models {.unnumbered .unlisted}

In practice, users may prefer a more robust BO variant over a potentially better performing but unstable variant.
Even if the `catch_errors` parameters are turned on and are never triggered, that does not guarantee that the BO algorithm ran as intended.
For instance, Gaussian processes are sensitive to the choice of kernel and kernel parameters, typically estimated through maximum likelihood estimation, suboptimal parameter values can result in white noise models with a constant mean and standard deviation prediction (except for the interpolation of training data).
In this case, the surrogate model will not provide useful mean and standard deviation predictions resulting in poor overall performance of the BO algorithm.
Another practical consideration regarding the choice of surrogate model can be overhead.
Fitting a vanilla Gaussian process scales cubicly in the number of data points and therefore the overhead of the BO algorithm grows with the number of iterations.
Furthermore, vanilla Gaussian processes natively cannot handle categorical input variables or dependencies in the search space (recall that in HPO we often deal with mixed hierarchical spaces).
In contrast, a random forest -- popularly used as a surrogate model in the *SMAC* package (@Lindauer2022) -- is cheap to train, quite robust in the sense that it is not as sensitive to its hyperparameters as a Gaussian process, and can easily handle mixed hierarchical spaces.
On the downside, a random forest is not really Bayesian (i.e., there is no posterior predictive distribution) and suffers from poor uncertainty estimates and poor extrapolation.

#### Warmstarting {.unnumbered .unlisted}

Warmstarting is a technique in optimization where previous optimization runs are used to improve the convergence rate and final solution of a new, related optimization run.
In BO, warmstarting can be achieved by providing a set of likely well-performing configurations as part of the initial design.
This approach can be particularly advantageous because it allows the surrogate model to start with prior knowledge of the optimization landscape in relevant regions.
In `r mlr3mbo`, warmstarting is straightforward by specifying a custom initial design.
Furthermore, a convenient feature of `r mlr3mbo` is the ability to continue optimization in an online fashion even after an optimization run has been terminated.
Both `r ref("OptimizerMbo")` and `r ref("TunerMbo")` support this feature, allowing optimization to resume on a given instance even if the optimization was previously interrupted or terminated.

#### Termination {.unnumbered .unlisted}

Common termination criteria include stopping after a fixed number of evaluations, once a given walltime budget has been reached, when performance reaches a certain level, or when performance improvement stagnates.
In the context of BO, it can also be sensible to stop the optimization if the best acquisition function value falls below a certain threshold.
For instance, terminating the optimization if the expected improvement of the next candidate(s) is negligible can be a reasonable approach.
At the time of publishing, terminators based on acquisition functions have not been implemented but this feature will be coming soon.

#### Parallelization {.unnumbered .unlisted}

The standard behavior of most BO algorithms is to sequentially propose a single candidate that should be evaluated next.
Users may want to use parallelization to compute candidates more efficiently.
If you are using BO for HPO, then the most efficient method is to parallelize the nested resampling, see @sec-nested-resampling-parallelization.
Alternatively, if the loop function supports candidates being proposed in batches (e.g., `bayesopt_parego()`) then the `q` argument to the loop function can be set to propose `q` candiates in each iteration that will be evaluated in parallel.

## Conclusion

In this chapter, we looked at advanced tuning methods.
We started by thinking about the types of errors that can occur during tuning and how to handle these to ensure your HPO process does not crash.
We then looked at multi-fidelity tuning, in which the Hyberband tuner can be used to efficiently tune algorithms by making use of fidelity parameters that control learner complexity.
We will return to Hyperband in @sec-pipelines-nonseq where we will learn how to make use of pipelines in order to tune algorithm with Hyperband.
<!-- FIXME: UPDATE REFERENCE ABOVE ONCE MERGED -->
Finally, we took a deep dive into Bayesian Optimization (BO) to look at how `r bbotk`, `r mlr3mbo`, and `r mlr3tuning` can be used together to implement complex BO tuning algorithms in `mlr3`, allowing for highly flexible and sample-efficient algorithms.
In the next chapter we will look at feature selection and see how `r mlr3filters` and `r mlr3fselect` use a very similar design interface to `r mlr3tuning`.

@tbl-api-advanced-tuning summarizes the most important functions and methods seen in this chapter.

| Underlying R6 Class | Constructor (if applicable) | Important methods |
| --------------------------- | --------------------- | -------------------------------------------- |
| `r ref("Learner")` | `r ref("lrn")`| `$encapsulate`/`$fallback` |
| `r ref("TunerHyperband")` | `tnr("hyperband")`| - |
| `r ref("bbotk::Objective")` | `r ref("srlrn()")`|  |
| `r ref("mlr3mbo::SurrogateLearner")` | `r ref("mlr3mbo::srlrn()")`|  |
| `r ref("mlr3mbo::AcqFunction")` | `r ref("mlr3mbo::acqf()")` |  |
| `r ref("mlr3mbo::AcqOptimizer")` | `r ref("mlr3mbo::acqo()")` |  |
| - | `r ref("mlr3mbo::loop_function")` | - |
| `r ref('mlr3mbo::OptimizerMbo')` | `r ref("mlr3mbo::opt()")` | |
| `r ref('mlr3mbo::TunerMbo')` | `tnr("mbo")` | |

: Important classes and functions covered in this chapter with underlying `R6` class (if applicable), constructor to create an object of the class, and important class methods. {#tbl-api-advanced-tuning}

## Exercises

<!-- FIXME - FIX BELOW USING ALL SECTIONS ABOVE -->

1. Minimize the 2D Rastrigin function $f: [-5.12, 5.12] \times [-5.12, 5.12] \rightarrow \mathbb{R}$, $\mathbf{x} \mapsto 10 D+\sum_{i=1}^D\left[x_i^2-10 \cos \left(2 \pi x_i\right)\right]$, $D = 2$ via BO (standard sequential single-objective BO via `bayesopt_ego()`) using the lower confidence bound with `lambda = 1` as acquisition function and `"NLOPT_GN_ORIG_DIRECT"` via `opt("nloptr")` as acquisition function optimizer (similarly as above).
Specify a budget of 40 function evaluations.
Use either a Gaussian process with Matérn 5/2 kernel (`"regr.km"`, similarly as above) or a random forest (`"regr.ranger"`) as surrogate model and compare the anytime performance (similarly as in @fig-bayesian-sinusoidal_bo_rs) of these two BO algorithms.
As an initial design, use the following points:
```{r}
initial_design = data.table(
  x1 = c(-3.95, 1.16, 3.72, -1.39, -0.11, 5.00, -2.67, 2.44),
  x2 = c(1.18, -3.93, 3.74, -1.37, 5.02, -0.09, -2.65, 2.46))
```
You can use the following function skeleton as a starting point to construct the objective function (using the `r ref("ObjectiveRFunDt")` class):
```{r}
rastrigin = function(xdt) {
  D = ncol(xdt)
  y = 10 * D + rowSums(xdt^2 - (10 * cos(2 * pi * xdt)))
  data.table(y = y)
}
```
The different surrogate models should for example look like the following:
```{r}
surrogate_gp = srlrn(lrn("regr.km", covtype = "matern5_2",
  optim.method = "BFGS", control = list(trace = FALSE)))

surrogate_rf = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))
```

2. Minimize the following function: $f: [-10, 10] \rightarrow \mathbb{R}^2, x \mapsto \left(x^2, (x - 2)^2\right)$.
Use the ParEGO algorithm (`r ref("mlr_loop_functions_parego")`) in a batch setting of four candidates (`q = 4`) and parallelize the actual objective function evaluation using the `r ref_pkg("future")` package (using four workers in a `multisession` plan).
Construct the objective function using the `r ref("ObjectiveRFunMany")` class.
For illustrative reasons, suspend the execution for five seconds every time a point is to be evaluated (making use of the `Sys.sleep()` function).
Use the following surrogate model, acquisition function and acquisition function optimizer (recall that ParEGO uses a scalarization approach to multi-objective optimization):
```{r}
surrogate = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))

acq_function = acqf("ei")

acq_optimizer = acqo(opt("random_search", batch_size = 100),
  terminator = trm("evals", n_evals = 100))
```
Terminate the optimization after a runtime of 60 seconds.
How many points did the BO algorithm evaluate (including the initial design) when properly parallelizing the evaluation of the objective function?
Compare this to the number of points the BO algorithm evaluated when *not* parallelizing the evaluation (but still using a batch of size `q = 4`).
Note that `q = 4` must be passed to the `r ref("OptimizerMbo")` via the `args` parameter.
You can use the following (non-parallelized) function skeleton as a starting point to construct the objective function (note that `r ref_pkg("future.apply")` might be useful to implement the parallelization):
```{r}
# non-parallelized version of the Schaffer function N.1
schaffer1 = function(xss) {
  evaluations = lapply(xss, FUN = function(xs) {
    Sys.sleep(5)
    list(y1 = xs$x, y2 = (xs$x - 2)^2)
  })
  rbindlist(evaluations)
}
```

::: {.content-visible when-format="html"}
`r citeas(chapter)`
:::
