# Quickstart {-}

{{< include _setup.qmd >}}

```{r index-002, eval = FALSE}
install.packages("mlr3")
```

As a 30-second introductory example, we will train a decision tree model on the first 120 rows of iris data set and make predictions on the final 30, measuring the accuracy of the trained model.

```{r index-003}
library("mlr3")
task = tsk("iris")
learner = lrn("classif.rpart")

# train a model of this learner for a subset of the task
learner$train(task, row_ids = 1:120)
# this is what the decision tree looks like
learner$model

predictions = learner$predict(task, row_ids = 121:150)
predictions
# accuracy of our model on the test set of the final 30 rows
predictions$score(msr("classif.acc"))
```

:::{.callout-tip}
A collection of use cases and examples can be found in the [mlr3gallery](https://mlr-org.com/gallery) on our [website](https://mlr-org.com).

We highly recommend to take a look at our `r mlr3book::mlr_pkg("cheatsheets")` while diving deeper into `r mlr3book::mlr_pkg("mlr3")`.
:::
