---
params:
    rebuild_cache: true
---

# Solutions to exercises {#sec-solutions}

::: {.content-visible when-format="html"}

{{< include common/_setup.qmd >}}
<!-- ENSURE ONLINE ONLY AND REFERENCE IN PREFACE -->

## Solutions to @sec-basics

1. Set the seed to `124` then train a classification tree model with `classif.rpart` and default hyperparameters on 80% of the data in the predefined `sonar` task. Evaluate the model's performance with the classification error measure on the remaining data. Also think about why we need to set the seed in this example.

Set the seed, load the `sonar` task, then the classification tree with `predict_type = "prob"` (needed for exercise 3), and the required measure.

```{r solutions-001}
set.seed(124)
task = tsk("sonar")
learner = lrn("classif.rpart", predict_type = "prob")
measure = msr("classif.ce")
```

Use `partition` to split the dataset then train the model.
We set the seed because `partition` introduces an element of randomness when splitting the data.

```{r}
splits = partition(task, ratio = 0.8)
learner$train(task, splits$train)
```

Once the model is trained, generate the predictions on the test set and score them.

```{r solutions-002}
preds = learner$predict(task, splits$test)
preds$score(measure)
```

2. Calculate the true positive, false positive, true negative, and false negative rates of the predictions made by the model in exercise 1.

Using `$confusion`, generate a confusion matrix and extract the required statistics,

```{r solutions-003}
preds$confusion
```

```{r solutions-004, echo=FALSE}
confusion = as.data.frame(preds$confusion)

TP = confusion$Freq[1]
FP = confusion$Freq[2]
TN = confusion$Freq[4]
FN = confusion$Freq[3]
```

Since the rows represent predictions (response) and the columns represent the ground truth values, the TP, FP, TN, and FN rates are as follows:

- True Positive (TP) = `r TP`

- False Positive (FP) = `r FP`

- True Negative (TN) = `r TN`

- False Positive (FN) = `r FN`

1. Since in this case we want the model to predict the negative class more often, we will raise the threshold (note the `predict_type` for the learner must be `prob` for this to work).

```{r solutions-005}
# raise threshold from 0.5 default to 0.6
preds$set_threshold(0.6)

preds$confusion
```
One reason we might want the false positive rate to be lower than the false negative rate is if we felt it was worse for a positive prediction to be incorrect (meaning the true label was the negative label) than it was for a negative prediction to be incorrect (meaning the true label was the positive label).

## Solutions to @sec-performance

1. Apply the "bootstrap" resampling strategy on the `mtcars` task and evaluate the performance of the `classif.rpart` decision tree learner.
Use 100 replicates and an a sampling ratio of 80%.
Calculate the MSE for each iteration and visualize the result.
Finally, calculate the aggregated performance score.

```{r solutions-006}
set.seed(3)
task = tsk("mtcars")
learner = lrn("regr.rpart")
resampling = rsmp("bootstrap", repeats = 100, ratio = 0.8)

rr = resample(task, learner, resampling)
rr$score(msr("regr.mse"))
autoplot(rr)
# Alternatively: Histogram
autoplot(rr, type = "histogram")

rr$aggregate(msr("regr.mse"))
```


2. Use the `spam` task and 5-fold cross-validation to benchmark Random Forest (`classif.ranger`), Logistic Regression (`classif.log_reg`), and XGBoost (`classif.xgboost`) with regards to AUC.
Which learner appears to do best? How confident are you in your conclusion?
How would you improve upon this?

```{r solutions-007, warning=FALSE}
set.seed(3)
grid = benchmark_grid(
  tasks = tsk("spam"),
  learners = lrns(c("classif.ranger", "classif.log_reg", "classif.xgboost"),
                  predict_type = "prob"),
  resamplings = rsmp("cv", folds = 5)
)

bmr = benchmark(grid)

mlr3viz::autoplot(bmr, measure = msr("classif.auc"))
```

This is only a small example for a benchmark workflow, but without tuning (see @sec-optimization), the results are naturally not suitable to make any broader statements about the superiority of either learner for this task.


3. A colleague claims to have achieved a 93.1% classification accuracy using the `classif.rpart` learner on the `penguins_simple` task.
You want to reproduce their results and ask them about their resampling strategy.
They said they used a custom 3-fold cross-validation with folds assigned as `factor(task$row_ids %% 3)`.
See if you can reproduce their results.

```{r solutions-008}
task = tsk("penguins_simple")
r = rsmp("custom_cv")

r$instantiate(task = task, f = factor(task$row_ids %% 3))

rr = resample(
  task = task,
  learner = lrn("classif.rpart"),
  resampling = r
)

rr$aggregate(msr("classif.acc"))
```


## Solutions to @sec-optimization

1. Tune the `mtry`, `sample.fraction`, ` num.trees` hyperparameters of a random forest model (`regr.ranger`) on the `r ref("mlr_tasks_mtcars", text = "Motor Trend")` dataset (`mtcars`).
Use a simple random search with 50 evaluations and select a suitable batch size.
Evaluate with a 3-fold cross-validation and the root mean squared error.

```{r solutions-009}
set.seed(4)
learner = lrn("regr.ranger",
  mtry.ratio      = to_tune(0, 1),
  sample.fraction = to_tune(1e-1, 1),
  num.trees       = to_tune(1, 2000)
)

instance = ti(
  task = tsk("mtcars"),
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measures = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

tuner = tnr("random_search", batch_size = 10)

tuner$optimize(instance)
```

2. Evaluate the performance of the model created in Question 1 with nested resampling.
Use a holdout validation for the inner resampling and a 3-fold cross-validation for the outer resampling.
Print the unbiased performance estimate of the model.

```{r solutions-010}
set.seed(4)
learner = lrn("regr.ranger",
  mtry.ratio      = to_tune(0, 1),
  sample.fraction = to_tune(1e-1, 1),
  num.trees       = to_tune(1, 2000)
)

at = auto_tuner(
  tuner = tnr("random_search", batch_size = 10),
  learner = learner,
  resampling = rsmp("holdout"),
  measure = msr("regr.rmse"),
  terminator = trm("evals", n_evals = 50)
)

task = tsk("mtcars")
outer_resampling = rsmp("cv", folds = 3)
rr = resample(task, at, outer_resampling, store_models = TRUE)

rr$aggregate()
```

3. Tune and benchmark an XGBoost model against a logistic regression `r ref("mlr_tasks_spam", text = "Spam")` dataset and determine which has the best Brier score.
Use mlr3tuningspaces and nested resampling.

```{r solutions-011}
library(mlr3tuningspaces)

learner_xgboost = lts(lrn("classif.xgboost", predict_type = "prob"))

at_xgboost = auto_tuner(
  tuner = tnr("random_search", batch_size = 1),
  learner = learner_xgboost,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.bbrier"),
  term_evals = 2,
)

learner_logreg = lrn("classif.log_reg", predict_type = "prob")

at_logreg = auto_tuner(
  tuner = tnr("random_search", batch_size = 1),
  learner = learner_logreg,
  resampling = rsmp("cv", folds = 3),
  measure = msr("classif.bbrier"),
  term_evals = 2,
)

task = tsk("spam")
outer_resampling = rsmp("cv", folds = 3)

design = benchmark_grid(
  tasks = task,
  learners = list(at_xgboost, at_logreg),
  resamplings = outer_resampling
)

bmr = benchmark(design, store_models = TRUE)

bmr
```



## Solutions to @sec-feature-selection

1. Calculate a correlation filter on the `r ref("mlr_tasks_mtcars", text = "Motor Trend")` dataset (`mtcars`).

```{r solutions-012}
library("mlr3verse")
filter = flt("correlation")

task = tsk("mtcars")
filter$calculate(task)

as.data.table(filter)
```

2. Use the filter from the first exercise to select the five best features in the `mtcars` dataset.

```{r solutions-013}
keep = names(head(filter$scores, 5))
task$select(keep)
task$feature_names
```

3. Apply a backward selection to the `r ref("mlr_tasks_penguins", text = "penguins")` dataset with a classification tree learner `"classif.rpart"` and holdout resampling by the measure classification accuracy. Compare the results with those in @sec-fs-wrapper-example.

```{r solutions-014}
library("mlr3fselect")

instance = fselect(
  fselector = fs("sequential", strategy = "sbs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = rsmp("holdout"),
  measure = msr("classif.acc")
)
as.data.table(instance$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
instance$result_feature_set
```

Answer the following questions:

  a. Do the selected features differ?

Yes, the backward selection selects more features.

  b. Which feature selection method achieves a higher classification accuracy?

In this example, the backwards example performs slightly better, but this depends heavily on the random seed and could look different in another run.

  c. Are the accuracy values in b) directly comparable? If not, what has to be changed to make them comparable?

No, they are not comparable because the holdout sampling called with `rsmp("holdout")` creates a different holdout set for the two runs. A fair comparison would create a single resampling instance and use it for both feature selections (see @sec-performance for details):

```{r solutions-015}
resampling = rsmp("holdout")
resampling$instantiate(tsk("penguins"))

sfs = fselect(
  fselector = fs("sequential", strategy = "sfs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = resampling,
  measure = msr("classif.acc")
)
sbs = fselect(
  fselector = fs("sequential", strategy = "sbs"),
  task =  tsk("penguins"),
  learner = lrn("classif.rpart"),
  resampling = resampling,
  measure = msr("classif.acc")
)
as.data.table(sfs$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
as.data.table(sbs$result)[, .(bill_depth, bill_length, body_mass, classif.acc)]
```

Alternatively, one could automate the feature selection and perform a benchmark between the two wrapped learners.

4. Automate the feature selection as in @sec-autofselect with the `r ref("mlr_tasks_spam", text = "spam")` dataset and a logistic regression learner (`"classif.log_reg"`). Hint: Remember to call `library("mlr3learners")` for the logistic regression learner.

```{r solutions-016, warning=FALSE}
library("mlr3fselect")
library("mlr3learners")

at = auto_fselector(
  fselector = fs("random_search"),
  learner = lrn("classif.log_reg"),
  resampling = rsmp("holdout"),
  measure = msr("classif.acc"),
  terminator = trm("evals", n_evals = 50)
)

grid = benchmark_grid(
  task = tsk("spam"),
  learner = list(at, lrn("classif.log_reg")),
  resampling = rsmp("cv", folds = 3)
)

bmr = benchmark(grid)

aggr = bmr$aggregate(msrs(c("classif.acc", "time_train")))
as.data.table(aggr)[, .(learner_id, classif.acc, time_train)]
```

## Solutions to @sec-pipelines

1. Concatenate the named PipeOps using the `r ref("concat_graphs", "%>>%")` operator.
  To get a `r ref("Learner")` object, use `r ref("as_learner", "as_learner()")`
```{r pipelines-001}
library("mlr3pipelines")
library("mlr3learners")

graph = po("imputeoor") %>>% po("scale") %>>% lrn("classif.log_reg")
graph_learner = as_learner(graph)
```

2. After training, the underlying `r ref("LearnerClassifLogReg", "lrn(\"classif.log_reg\")")` can be accessed through the `$base_learner()` method.
  Alternatively, the learner can be accessed explicitly using `r ref("PipeOpLearner", "po(\"learner\")")`.
```{r pipelines-002}
graph_learner$train(tsk("pima"))

# access the learner through the $base_learner() method
model = graph_learner$base_learner()$model
coef(model)

# acces the learner explicitly through the PipeOp
pipeop = graph_learner$graph_model$pipeops$classif.log_reg
model = pipeop$learner_model$model
coef(model)
```

3. Set the `$keep_results` flag of the Graph to `TRUE` to keep the results of the individual PipeOps.
  Afterwards, the input of the `r ref("LearnerClassifLogReg", "lrn(\"classif.log_reg\")")` can be accessed through the `$.result` field of its predecessor, the `r ref("PipeOpScale", "po(\"scale\")")`.
  Note that the `$.result` is a `list`, we want to access its only element, named `$output`.
```{r pipelines-003}
graph_learner$graph$keep_results = TRUE
graph_learner$train(tsk("pima"))

# access the input of the learner
scale_result = graph_learner$graph_model$pipeops$scale$.result

scale_output_task = scale_result$output

age_column = scale_output_task$data()$age

# check if the age column is standardized:
# 1. does it have mean 0? -- almost, up to numerical precision!
mean(age_column)
# 2. does it have standard deviation 1? -- yes!
sd(age_column)
```

4. To use `r ref("PipeOpSelect", "po(\"select\")")` to *remove*, instead of *keep*, a feature based on a pattern, use `r ref("selector_invert")` together with `r ref("selector_grep")`.
  To remove the "`R`" class columns in @sec-pipelines-stack, the following `r ref("PipeOpSelect", "po(\"select\")")` could be used:
```{r pipelines-004}
po("select", selector = selector_invert(selector_grep("\\.R")))
```
which would have the benefit that it would keep the columns pertaining to all other classes, even if the `r ref("mlr_tasks_sonar", "tsk(\"sonar\")")` data set had more target classes.

5. A solution that does not need to specify the target classes at all is to use a custom `r ref("Selector")`, as was shown in @sec-pipelines-bagging:
```{r pipelines-005}
selector_remove_one_prob_column = function(task) {
  class_removing = task$class_names[[1]]
  selector_use = selector_invert(selector_grep(paste0("\\.", class_removing)))

selector_use(task)
}
```
Using this selector in @sec-pipelines-stack, one could use the resulting stacking learner on any classification task with arbitrary target classes.

6. As the first hint states, two `r ref("PipeOpImputeLearner", "po(\"imputelearner\")")` objects are necessary: one to impute missing values in factor columns using a classification learner, and another to impute missing values in numeric columns using a regression learner.
  Additionally, `r ref("pipeline_robustify", "ppl(\"robustify\")")` is used along with the `r ref("ranger::ranger")`-based learners inside `r ref("PipeOpImputeLearner", "po(\"imputelearner\")")` because the data passed to the imputation learners still contains missing values, which `r ref("ranger::ranger")` cannot handle.
```{r pipelines-006}
impute_factors = po("imputelearner", id = "impute_factors",
  learner = ppl("robustify", learner = lrn("classif.ranger")) %>>%
    lrn("classif.ranger"),
  affect_columns = selector_type("factor")
)
impute_numerics = po("imputelearner", id = "impute_numerics",
  learner = ppl("robustify", learner = lrn("regr.ranger")) %>>%
    lrn("regr.ranger"),
  affect_columns = selector_type(c("numeric", "integer"))
)

impute_graph = impute_numerics %>>% impute_factors

imputed = impute_graph$train(tsk("penguins"))[[1]]

# e.g. see how row 4 was imputed
# original:
tsk("penguins")$data(rows = 4)
# imputed:
imputed$data(rows = 4)
```

## Solutions to @sec-special

1. Run a benchmark experiment on the `german_credit` task with algorithms: `featureless`, `log_reg`, `ranger`. Tune the `featureless` model using `tunetreshold` and `learner_cv`. Use 2-fold CV and evaluate with `msr("classif.costs", costs = costs)` where you should make the parameter `costs` so that the cost of a true positive is -10, the cost of a true negative is -1, the cost of a false positive is 2, and the cost of a false negative is 3. Use `set.seed(11)` to make sure you get the same results as us. Are your results surprising?

```{r solutions-017}
library(mlr3verse)
set.seed(11)

costs = matrix(c(-10, 3, 2, -1), nrow = 2, dimnames =
  list("Predicted Credit" = c("good", "bad"),
    Truth = c("good", "bad")))
cost_measure = msr("classif.costs", costs = costs)

gr = po("learner_cv", lrn("classif.featureless", predict_type = "prob")) %>>%
  po("tunethreshold", measure = cost_measure)
task = tsk("german_credit")
learners = list(as_learner(gr), lrn("classif.log_reg"), lrn("classif.ranger"))
bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 2)))
bmr$aggregate(cost_measure)[, c(4, 7)]
```

2. Train and predict a survival forest using `rfsrc` (from `mlr3extralearners`). Run this experiment using `task = tsk("rats"); split = partition(task)`. Evaluate your model with the RCLL measure.
```{r solutions-018}
library(mlr3verse)
library(mlr3proba)
library(mlr3extralearners)
set.seed(11)

task = tsk("rats")
split = partition(task)

lrn("surv.rfsrc")$
  train(task, split$train)$
  predict(task, split$test)$
  score(msr("surv.rcll"))
```

3. Estimate the density of the `tsk("precip")` data using `logspline` (from `mlr3extralearners`). Run this experiment using `task = tsk("precip"); split = partition(task)`. Evaluate your model with the logloss measure.
```{r solutions-019}
library(mlr3verse)
library(mlr3proba)
library(mlr3extralearners)
set.seed(11)

task = tsk("precip")
split = partition(task)

lrn("dens.logspline")$
  train(task, split$train)$
  predict(task, split$test)$
  score(msr("dens.logloss"))
```

4. Run a benchmark clustering experiment on the `wine` dataset without a label column. Compare the performance of `k-means` learner with `k` equal to 2, 3 and 4 using the silhouette measure. Use insample resampling technique. What value of `k` would you choose based on the silhouette scores?

```{r solutions-020}
library(mlr3)
library(mlr3cluster)
set.seed(11)
learners = list(
  lrn("clust.kmeans", centers = 2L, id = "k-means, k=2"),
  lrn("clust.kmeans", centers = 3L, id = "k-means, k=3"),
  lrn("clust.kmeans", centers = 4L, id = "k-means, k=4")
)

task = as_task_clust(tsk("wine")$data()[, -1])
measure = msr("clust.silhouette")
bmr = benchmark(benchmark_grid(task, learners, rsmp("insample")))
bmr$aggregate(measure)[, c(4, 7)]
```

Based on the silhouette score, we can choose `k = 2`.

5. Run a (spatially) unbiased classification benchmark experiment on the `ecuador` task with a featureless learner and xgboost, evaluate with the binary Brier score.

You can use any resampling method from `r mlr3spatiotempcv`, in this solution we use 4-fold spatial environmental blocking.

```{r solutions-021}
library(mlr3verse)
library(mlr3spatial)
library(mlr3spatiotempcv)
set.seed(11)
learners = lrns(paste0("classif.", c("xgboost", "featureless")),
  predict_type = "prob")
rsmp_sp = rsmp("spcv_env", folds = 4)
design = benchmark_grid(tsk("ecuador"), learners, rsmp_sp)
bmr = benchmark(design)
bmr$aggregate(msr("classif.bbrier"))[, c(4, 7)]
```


## Solutions to @sec-technical

1. Consider the following example where you resample a learner (debug learner, sleeps for 3 seconds during train) on 4 workers using the multisession backend:
```{r technical-050, eval = FALSE}
task = tsk("penguins")
learner = lrn("classif.debug", sleep_train = function() 3)
resampling = rsmp("cv", folds = 6)

future::plan("multisession", workers = 4)
resample(task, learner, resampling)
```

i. Assuming that the learner would actually calculate something and not just sleep: Would all CPUs be busy?
ii. Prove your point by measuring the elapsed time, e.g., using `r ref("system.time()")`.
iii. What would you change in the setup and why?

Not all CPUs would be utilized in the example.
All 4 of them are occupied for the first 4 iterations of the cross validation.
The 5th iteration, however, only runs in parallel to the 6th fold, leaving 2 cores ilde.
This is supported by the elapsed time of roughly 6 seconds for 6 jobs compared to also roughly 6 seconds for 8 jobs:

```{r solutions-022, eval = FALSE}
task = tsk("penguins")
learner = lrn("classif.debug", sleep_train = function() 3)

future::plan("multisession", workers = 4)

resampling = rsmp("cv", folds = 6)
system.time(resample(task, learner, resampling))

resampling = rsmp("cv", folds = 8)
system.time(resample(task, learner, resampling))
```

If possible, the number of resampling iterations should be an integer multiple of the number of workers.
Therefore, a simple adaptation either increases the number of folds for improved accuracy of the error estimate or reduces the number of folds for improved runtime.

2. Create a new custom classification measure (either using methods demonstrated in @sec-extending or with `r ref("mlr_measures_classif.costs")`) which scores predictions using the mean over the following classification costs:

* If the learner predicted label "A" and the truth is "A", assign score 0
* If the learner predicted label "B" and the truth is "B", assign score 0
* If the learner predicted label "A" and the truth is "B", assign score 1
* If the learner predicted label "B" and the truth is "A", assign score 10

The rules can easily be translated to R code where we expect `truth` and `prediction` to be factor vectors of the same length with levels `"A"` and `"B"`:

```{r solutions-023}
costsens = function(truth, prediction) {
    score = numeric(length(truth))
    score[truth == "A" & prediction == "B"] = 10
    score[truth == "B" & prediction == "A"] = 1

    mean(score)
}
```

This function can be embedded in the `Measure` class accordingly.

```{r solutions-024}
MeasureCustom = R6::R6Class("MeasureCustom",
  inherit = mlr3::MeasureClassif, # classification measure
  public = list(
    initialize = function() { # initialize class
      super$initialize(
        id = "custom", # unique ID
        packages = character(), # no dependencies
        properties = character(), # no special properties
        predict_type = "response", # measures response prediction
        range = c(0, Inf), # results in values between (0, 1)
        minimize = TRUE # smaller values are better
      )
    }
  ),

  private = list(
    .score = function(prediction, ...) { # define score as private method
      # define loss
      costsens = function(truth, prediction) {
        score = numeric(length(truth))
        score[truth == "A" & prediction == "B"] = 10
        score[truth == "B" & prediction == "A"] = 1

        mean(score)
      }

      # call loss function
      costsens(prediction$truth, prediction$response)
    }
  )
)
```

An alternative (as pointed to by the hint) can be constructed by first translating the rules to a matrix of misclassification costs, and then feeding this matrix to the
 constructor of the `r ref("mlr_measures_classif.costs")` measure:

```{r solutions-025}
# truth in columns, prediction in rows
C = matrix(c(0, 10, 1, 0), nrow = 2)
rownames(C) = colnames(C) = c("A", "B")
print(C)

msr("classif.costs", costs = C)
```

## Solutions to @sec-large-benchmarking


### Getting Data from OpenML {.unnumbered .unlisted}

```{r solutions-026}
#| cache: false
#| include: false
library("mlr3oml")
lgr::get_logger("mlr3oml")$set_threshold("off")

if (!dir.exists("openml")) {
  dir.create(file.path("openml", "manual"), recursive = TRUE)
}
options(mlr3oml.cache = file.path("openml", "cache"))
```

We access the AutoML benchmark suite with ID 269 using the `r ref("mlr3oml::ocl()")` function.

```{r}
#| include: false
path_automl_suite = file.path("openml", "manual", "automl_suite.rds")
```

```{r solutions-028, eval = !file.exists(path_automl_suite)}
library("mlr3oml")
automl_suite = ocl(id = 269)
automl_suite$task_ids
```

```{r}
#| include: false
if (file.exists(path_automl_suite)) {
  automl_suite = readRDS(path_automl_suite)
} else {
  # need to access the ids to trigger the download
  automl_suite$task_ids
  saveRDS(automl_suite, path_automl_suite)
}
```

We can find all tasks with less than 4000 observations by specifying this constraint in `r ref("mlr3oml::list_oml_tasks()")`.

```{r}
#| include: false
path_automl_table = file.path("openml", "manual", "automl_table.rds")
```


```{r solutions-030, eval = !file.exists(path_automl_table)}
tbl = list_oml_tasks(
  task_id = automl_suite$task_ids, number_instances = c(0, 4000)
)
```

```{r}
#| include: false
if (file.exists(path_automl_table)) {
  tbl = readRDS(path_automl_table)
} else {
  saveRDS(tbl, path_automl_table)
}
```

This returns a table which only contains matching tasks from the AutoML benchmark.

```{r solutions-032}
tbl[, .(task_id, data_id, name, NumberOfInstances)]
```

We can create mlr3 tasks from these OpenML IDs using `tsk("oml")`.

```{r solutions-033}
tasks = lapply(tbl$task_id, function(id) tsk("oml", task_id = id))
```

Below, we define the robustified learners with a featureless fallback learner.

```{r solutions-034}
learner_ranger = as_learner(
  ppl("robustify", learner = lrn("regr.ranger")) %>>%
    po("learner", lrn("regr.ranger"))
)
learner_ranger$id = "ranger"
learner_ranger$fallback = lrn("regr.featureless")

learner_rpart = as_learner(
  ppl("robustify", learner = lrn("regr.rpart")) %>>%
    po("learner", lrn("regr.rpart"))
)
learner_rpart$id = "rpart"
learner_rpart$fallback = lrn("regr.featureless")

learners = list(learner_ranger, learner_rpart)
```

Finally, we create the experimental design using `r ref("benchmark_grid()")`.

```{r solutions-035}
# we set the seed, as benchmark_grid instantiates the resamplings
set.seed(123)
resampling = rsmp("cv", folds = 3)
design = benchmark_grid(tasks, learners, resampling)
design
```

### Executing the Experiments using batchtools {.unnumbered .unlisted}

We start with loading the relevant libraries and creating a registry.
By specifying the registry's `file.dir` to `NA` we are using a temporary directory.

```{r solutions-036}
#| cache: false
library("mlr3batchmark")
library("batchtools")

reg = makeExperimentRegistry(
  file.dir = NA,
  seed = 1,
  packages = "mlr3verse"
)
```

Then, we change the cluster function to "Multicore" (or "Socket" if you are on Windows).

```{r solutions-037}
#| cache: false
# Mac and Linux
reg$cluster.functions = makeClusterFunctionsMulticore()

# Windows:
reg$cluster.functions = makeClusterFunctionsSocket()

saveRegistry(reg)
```

```{r}
#| include: false
#| cache: false
reg$cluster.functions = makeClusterFunctionsInteractive()
saveRegistry(reg)
```

The next two steps are to populate the registry with the experiments using `batchmark()` and to submit them.
By specifying no IDs in `submitJobs()`, all jobs returned by `findNotSubmitted()` are queued, which in this case are all existing jobs.

```{r solutions-038}
#| output: false
batchmark(design, reg = reg)
submitJobs(reg = reg)
waitForJobs(reg = reg)
```

After the execution of the experiment finished, we can collect the results like below.

```{r solutions-040}
bmr = reduceResultsBatchmark(reg = reg)
bmr
```

### Analyzing the Results {.unnumbered .unlisted}

As a first step, we load `r ref_pkg("mlr3benchmark")` and create a benchmark aggregate using `msr("regr.mse")`.

```{r solutions-041}
library("mlr3benchmark")
bma = as_benchmark_aggr(bmr, measures = msr("regr.mse"))
bma
```

We can now use the `$friedman_test()` method of the benchmark aggregate to apply said test to the experiment results.
Note that we don't need a post-hoc test, because we are only comparing two algorithms.

```{r solutions-042}
bma$friedman_test()
```

This experimental design was not able to detect a significant difference on the 5% level.

Next, we inspect the ranks using the `$rank_data()` method of the `r ref("BenchmarkAggr")`, where we specify `minimize` to `TRUE`, because a lower mean square error is better.

```{r solutions-043}
bma$rank_data(minimize = TRUE)
```

The random forest is better on 7 of the 10 tasks.

## Solutions to @sec-interpretation

1. Prepare a `mlr3` regression task for `fifa` data. Select only variables describing the age and skills of footballers. Train any predictive model for this task, e.g. `regr.ranger`.

```{r solutions-044, warning=FALSE, message=FALSE}
library("DALEX")
library("ggplot2")
data("fifa", package = "DALEX")
old_theme = set_theme_dalex("ema")

library("mlr3")
library("mlr3learners")
set.seed(1)

fifa20 <- fifa[,5:42]
task_fifa = as_task_regr(fifa20, target = "value_eur", id = "fifa20")

learner = lrn("regr.ranger")
learner$train(task_fifa)
learner$model
```

2. Use the permutation importance method to calculate variable importance ranking. Which variable is the most important? Is it surprising?

**With `iml`**

```{r solutions-045, warning=FALSE, message=FALSE}
library(iml)
model = Predictor$new(learner,
                data = fifa20,
                y = fifa$value_eur)

effect = FeatureImp$new(model,
                loss = "rmse")
effect$plot()
```

**With `DALEX`**

```{r solutions-046, warning=FALSE, message=FALSE}
library("DALEX")
ranger_exp = DALEX::explain(learner,
  data = fifa20,
  y = fifa$value_eur,
  label = "Fifa 2020",
  verbose = FALSE)

ranger_effect = model_parts(ranger_exp, B = 5)
head(ranger_effect)
plot(ranger_effect)
```

3. Use the Partial Dependence profile to draw the global behavior of the model for this variable. Is it aligned with your expectations?

**With `iml`**

```{r solutions-047, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

effect = FeatureEffects$new(model)
plot(effect, features = num_features)
```

**With `DALEX`**

```{r solutions-048, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

ranger_profiles = model_profile(ranger_exp, variables = num_features)
plot(ranger_profiles)
```

4 Choose one of the football players. You can choose some well-known striker (e.g. Robert Lewandowski) or a well-known goalkeeper (e.g. Manuel Neuer). The following tasks are worth repeating for several different choices.

```{r solutions-049, warning=FALSE, message=FALSE}
player_1 <- fifa["R. Lewandowski", 5:42]
```

5. For the selected footballer, calculate and plot the Shapley values. Which variable is locally the most important and has the strongest influence on the valuation of the footballer?

**With `iml`**

```{r solutions-050, warning=FALSE, message=FALSE}
shapley = Shapley$new(model, x.interest = player_1)
plot(shapley)
```

**With `DALEX`**

```{r solutions-051, warning=FALSE, message=FALSE}
ranger_shap = predict_parts(ranger_exp,
             new_observation = player_1,
             type = "shap", B = 1)
plot(ranger_shap, show_boxplots = FALSE)
```

6. For the selected footballer, calculate the Ceteris Paribus / Individual Conditional Expectation profiles to draw the local behavior of the model for this variable. Is it different from the global behavior?

**With `DALEX`**

```{r solutions-052, warning=FALSE, message=FALSE}
num_features = c("movement_reactions", "skill_ball_control", "age")

ranger_ceteris = predict_profile(ranger_exp, player_1)
plot(ranger_ceteris, variables = num_features) +
  ggtitle("Ceteris paribus for R. Lewandowski", " ")
```

## Solutions to @sec-bayesian-optimization

1. We first construct the objective function and optimization instance:

```{r}
rastrigin = function(xdt) {
  D = ncol(xdt)
  y = 10 * D + rowSums(xdt^2 - (10 * cos(2 * pi * xdt)))
  data.table(y = y)
}

objective = ObjectiveRFunDt$new(
  fun = rastrigin,
  domain = ps(x1 = p_dbl(lower = -5.12, upper = 5.12),
    x2 = p_dbl(lower = -5.12, upper = 5.12)),
  codomain = ps(y = p_dbl(tags = "minimize")),
  id = "rastrigin2D")

instance = OptimInstanceSingleCrit$new(
  objective = objective,
  terminator = trm("evals", n_evals = 40))
```

Based on the different surrogate models, we can construct two optimizers:

```{r}
surrogate_gp = srlrn(lrn("regr.km", covtype = "matern5_2",
  optim.method = "BFGS", control = list(trace = FALSE)))

surrogate_rf = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))

acq_function = acqf("cb", lambda = 1)

acq_optimizer = acqo(opt("nloptr", algorithm = "NLOPT_GN_ORIG_DIRECT"),
  terminator = trm("stagnation", iters = 100, threshold = 1e-5))

optimizer_gp = opt("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate_gp,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)

optimizer_rf = opt("mbo",
  loop_function = bayesopt_ego,
  surrogate = surrogate_rf,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer)
```

We then evaluate the given initial design on the instance and optimize it with the first BO algorithm using a Gaussian Process as surrogate model:

```{r, output=FALSE}
initial_design = data.table(
  x1 = c(-3.95, 1.16, 3.72, -1.39, -0.11, 5.00, -2.67, 2.44),
  x2 = c(1.18, -3.93, 3.74, -1.37, 5.02, -0.09, -2.65, 2.46))

instance$eval_batch(initial_design)

optimizer_gp$optimize(instance)

gp_data = instance$archive$data
gp_data[, y_min := cummin(y)]
gp_data[, nr_eval := seq_len(.N)]
gp_data[, surrogate := "Gaussian Process"]
```

Afterwards, we clear the instance, evaluate the initial design again and optimize the instance with the second BO algorithm using a random forest as surrogate model:

```{r, output=FALSE}
instance$archive$clear()

instance$eval_batch(initial_design)

optimizer_rf$optimize(instance)

rf_data = instance$archive$data
rf_data[, y_min := cummin(y)]
rf_data[, nr_eval := seq_len(.N)]
rf_data[, surrogate := "Random Forest"]
```

We collect all data and visualize the anytime performance:

```{r}
library(ggplot2)
library(viridisLite)
all_data = rbind(gp_data, rf_data)
ggplot(aes(x = nr_eval, y = y_min, colour = surrogate), data = all_data) +
  geom_step() +
  scale_colour_manual(values = viridis(2, end = 0.8)) +
  labs(y = "Best Observed Function Value", x = "Number of Function Evaluations",
       colour = "Surrogate Model") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

2. We first construct the non-parallelized objective function and the optimization instance:

```{r}
schaffer1 = function(xss) {
  evaluations = lapply(xss, FUN = function(xs) {
    Sys.sleep(5)
    list(y1 = xs$x, y2 = (xs$x - 2)^2)
  })
  rbindlist(evaluations)
}

objective = ObjectiveRFunMany$new(
  fun = schaffer1,
  domain = ps(x = p_dbl(lower = -10, upper = 10)),
  codomain = ps(y1 = p_dbl(tags = "minimize"), y2 = p_dbl(tags = "minimize")),
  id = "schaffer1")

instance = OptimInstanceMultiCrit$new(
  objective = objective,
  terminator = trm("run_time", secs = 60))
```

Using the surrogate, acquisition function and acquisition function optimizer that are provided, we can proceed to optimize the instance via ParEGO:

```{r}
surrogate = srlrn(lrn("regr.ranger", num.trees = 10L, keep.inbag = TRUE,
  se.method = "jack"))

acq_function = acqf("ei")

acq_optimizer = acqo(opt("random_search", batch_size = 100),
  terminator = trm("evals", n_evals = 100))

optimizer = opt("mbo",
  loop_function = bayesopt_parego,
  surrogate = surrogate,
  acq_function = acq_function,
  acq_optimizer = acq_optimizer,
  args = list(q = 4))
```

```{r, output=FALSE}
optimizer$optimize(instance)
```

We observe that 12 points were evaluated in total (which makes sense as the objective function evaluation is not yet parallelized and the overhead of each function evaluation is given by 5 seconds).
While the points are appropriately evaluated in batches of size `q = 4` (with the initial design automatically constructed as the first batch), we don't experience any acceleration of the optimization process unless the function evaluation is explicitly parallelized.

```{r}
nrow(instance$archive$data)
instance$archive$data[, c("x", "timestamp", "batch_nr")]
```

We now parallelize the evaluation of the objective function and proceed to optimize the instance again:

```{r}
library(future)
library(future.apply)
plan("multisession", workers = 4)

schaffer1_parallel = function(xss) {
  evaluations = future_lapply(xss, FUN = function(xs) {
    Sys.sleep(5)
    list(y1 = xs$x, y2 = (xs$x - 2)^2)
  })
  rbindlist(evaluations)
}

objective_parallel = ObjectiveRFunMany$new(
  fun = schaffer1_parallel,
  domain = ps(x = p_dbl(lower = -10, upper = 10)),
  codomain = ps(y1 = p_dbl(tags = "minimize"), y2 = p_dbl(tags = "minimize")),
  id = "schaffer1_parallel")

instance_parallel = OptimInstanceMultiCrit$new(
  objective = objective_parallel,
  terminator = trm("run_time", secs = 60))
```

```{r, output=FALSE}
optimizer$optimize(instance_parallel)
```

By parallelizing the evaluation of the objective function we used our compute resources much more efficiently and evaluated many more points:

```{r}
nrow(instance_parallel$archive$data)
instance_parallel$archive$data[, c("x", "timestamp", "batch_nr")]
```

## Solutions to @sec-fairness

1. Load the `"adult_train"` task and try to build a first model.
  Train a simple model and evaluate it on the `"adult_test"` task that is also available with `mlr3fairness`.

For now we simply load the data and look at the data.

```{r}
library("mlr3")
library("mlr3fairness")
set.seed(8)

trn = tsk("adult_train")
tst = tsk("adult_test")
trn
```

We can now train a simple model, e.g., a decision tree and evaluate for accuracy.

```{r}
l = lrn("classif.rpart")
l$train(trn)
prd = l$predict(tst)
prd$score()
```


2. Assume our goal is to achieve parity in *false omission rates*.
  Construct a fairness metric that encodes this and againg evaluate your model.
  In order to get a deeper understanding, look at the `groupwise_metrics` function to obtain performance in each group.

The metric is available via the key `"fairness.fomr"`.
Note, that evaluating our prediction now requires that we also provide the task.

```{r}
m = msr("fairness.fomr")
prd$score(m, tst)
```

The `groupwise_metrics` function creates a metric for each group specified in the `pta` column role:

```{r}
tst$col_roles$pta
```

```{r}
m2 = groupwise_metrics(base_measure = msr("classif.fomr"), task = tst)
```

We can then use this metric to evaluate our model again.
This gives us the false omission rates for male and female individuals separately.

```{r}
prd$score(m2, tst)
```

3. Now try to improve your model, e.g. by employing pipelines that use pre- or post-processing methods for fairness.
  Evaluate your model along the two metrics and visualize the results.
  Compare the different models using an appropriate visualization.

First we can again construct the learners above.
```{r}
library("mlr3pipelines")
l1 = po("reweighing_wts") %>>% lrn("classif.rpart")
l2 = po("learner_cv", lrn("classif.rpart")) %>>%
  po("EOd")
```

And run the benchmark again. Note, that we use 3-fold cross-validation this time for comparison.

```{r}
lrns = list(l, l1, l2)
grd = benchmark_grid(trn, lrns, rsmp("cv", folds = 3L))
bmr = benchmark(grd)
bmr$aggregate(msrs(c("classif.acc", "fairness.fomr")))
```

We can now again visualize the result.

```{r}
library("ggplot2")
fairness_accuracy_tradeoff(bmr, msr("fairness.fomr")) +
scale_color_viridis_d("Learner") +
theme_minimal()
```

4. Now we want to add a second sensitive attribute to your dataset, in this case `"race"`.
  Add the information to your task and evaluate for the initial model again. What changes?
  Again study the `groupwise_metrics`.

This can be achieved by adding "race" to the `"pta"` col_role.

```{r}
trn$set_col_roles("race", add_to = "pta")
trn
```

```{r}
tst$set_col_roles("race", add_to = "pta")
prd$score(m, tst)
```

If we now evaluate for the intersection, we obtain a large deviation from `0`.
Note, that the metric by default computes the maximum discrepancy between all metrics for the non-binary case.

If we now get the `groupwise_metrics`, we will get a metric for the intersection of each group.

```{r}
m3 = groupwise_metrics(msr("classif.fomr"),  trn)
unname(sapply(m3, function(x) x$id))
```

```{r}
prd$score(m3, tst)
```

And we can see, that the reason might be, that the false omission rate for female Amer-Indian-Eskimo is at `1.0`!
We can investigate this further by looking at actual counts:

```{r}
table(tst$data(cols = c("race", "sex", "target")))
```

One of the reasons might be that there are only 3 individuals in the ">50k" category!
This is an often encountered problem, as error metrics have a large variance when samples are small.
Note, that the pre- and post-processing methods in general do not all support multiple protected attributes.

:::