---
author:
  - name: Raphael Sonabend
    orcid: 0000-0001-9225-4654
    email: raphaelsonabend@gmail.com
    affiliations:
      - name: Imperial College London
  - name: Patrick Schratz
    orcid: 0000-0003-0748-6624
  - name: Damir Pulatov
    orcid: 0000-0003-4901-7201
    affiliations:
      - name: University of Wyoming
abstract: Previous chapters have considered the core functionality of functions and features in the mlr3 universe and other helper packages, which altogether create a modular ecosystem. This chapter highlights the power of that modularity by introducing implemented machine learning tasks beyond regression and classification. The chapter begins with an overview to mlr3proba, which is an extension for probabilistic supervised learning. mlr3proba currently includes survival analysis, density estimation, and some support for probabilistic regression. Next, the chapter looks at the packages mlr3spatial and mlr3spatiotempcv, which combined implement spatial and spatiotemporal analysis. This section also includes detailed consideration to custom cross-validation required for spatiotemporal modelling. Finally, mlr3cluster is introduced, which includes support for unsupervised clustering. As each of these tasks may be new to readers, we include an overview to the theory behind each, before looking at the extensions to common mlr3 objects (learners, predictions, tasks, and measures), and then finally consider task-specific extensions.
---

```{r, include = FALSE}
set.seed(8)
```

# Beyond Deterministic Regression and Single-Label Classification {#sec-special}

{{< include _setup.qmd >}}

This chapter title is a mouthful, but it highlights the limited scope we have considered up until now.
So far, this book has focused on detailed explanation of the `r mlr3` universe of packages, which is abstracted in such a way that allows users to choose a level of complexity for your project that suits you.
For example, you could limit your usage of these packages to just `r mlr3` to train a very small subset of regression and classification learners on a set of data and then use your model to make predictions for new unseen data.
You could extend this to many regression and classification learners with `r mlr3learners` and `r mlr3extralearners`.
By making use of `r mlr3tuning` and `r mlr3pipelines` you could optimise your model further including with pre- and post-processing of data and algorithms.
However, the real power of the `r mlr3` universe, is in its ability to take all of these features, and extend them to any machine learning task.
In @sec-basics we introduced deterministic regression and deterministic & probabilistic single-label classification (@tbl-alltasks).
But our infrastructure also works well for survival analysis, density estimation, clustering, spatiotemporal analysis, time-series forecasting, deep learning architectures, image analysis, multi-label classification, probabilistic regression, and many many more.
Some of these tasks are implemented in extension packages as part of the `r mlr3` universe (see @fig-mlr3verse), some are available by creating pipelines with `r mlr3pipelines`, some are in development, and some are waiting to be developed (maybe by some readers of this book).
In this chapter, we will take you through just a subset of these new tasks, focusing on the ones that have a stable API.
As we work through this chapter we will refer to the 'building blocks' of `r mlr3`, this refers to the base classes that must be extended to create new tasks, these are `r ref("Prediction")`, `r ref("Learner")`, `r ref("Measure")`, and `r ref("Task")`.
In many cases, to implement a new machine learning paradigm, all you need to do is create a class that extends the above and then implement new concrete learners that inherit from these (read @sec-extending if you are interested in extending new tasks or adding new learners and/or measures).

@tbl-alltasks summarises available extension tasks, including the package they are implemented in, the task status, and the suffix used in the building block functions.
Only tasks with stable APIs are discussed in this chapter, the table also includes tasks in development.

| Task | Package | Status | Description |
| ---- | ------- | ------ | --------------- |
| Deterministic regression | `r mlr3` | Stable | Point prediction of a continuous variable. |
| Probabilistic regression | `r mlr3proba` | Stable | Probability distribution prediction of a continuous variable. |
| Deterministic single-label classification | `r mlr3` | Stable | Prediction of a single category an observation falls into. |
| Probabilistic single-label classification | `r mlr3` | Stable | Prediction of the probability of an observation falling into one or more mutually exclusive categories. |
| Cost-sensitive classification | `r mlr3` and `r mlr3pipelines` | Stable | Classification predictions with unequal costs associated with misclassifications. |
| Survival analysis | `r mlr3proba` | Stable | Time-to-event predictions with possible 'censoring'. |
| Density estimation | `r mlr3proba` | Stable | Unsupervised estimation of probability density functions. |
| Spatiotemporal analysis | `r mlr3spatiotempcv` and `r mlr3spatial` | Stable | Supervised prediction of data with spatial (e.g., coordinates) and/or temporal outcomes. |
| Cluster analysis | `r mlr3cluster` | Stable | Unsupervised estimation of homogeneous clusters of data points. |
| Time series forecasting | `mlr3temporal` | Developing | Supervised predictions of target over time (e.g., daily temperature). |
| Functional data analysis | `mlr3fda` | Developing | Supervised predictions from functional data. |

: Table of extension tasks that can be used with `r mlr3` infrastructure. 'Stable' tasks have an interface that is unlikely to undergo any major changes in the future, these tasks are discussed in this chapter. 'Developing' tasks are in packages that might have breaking API changes. As we have a growing community of contributors, this list is far from exhaustive and many 'experimental' task implementations exist; this list just represents the tasks that have a functioning interface. {#tbl-alltasks}

## Cost-Sensitive Classification {#cost-sens}

We begin by discussing a task that does not require any additional packages or infrastructure, only the tools we have already learnt about from earlier chapters.
In regular classification, the aim is to minimize the misclassification rate, and thus all types of misclassification errors are deemed equally severe.
A more general setting is `r define("cost-sensitive classification", index = "classification!cost-sensitive classification")`, in which costs caused by different kinds of errors are equal.
The objective of cost-sensitive classification is to minimize the expected costs.

As we discuss this task we will use the `r ref("mlr_tasks_german_credit")` task as a running example.

```{r special-001}
task = tsk("german_credit")
task
```

Let us imagine in this task that an insurance company, InsureU, provides fixed loans of \$5K with \$6K to be paid back over one year.
InsureU needs to be predict if a new customer will have good credit to decide if a loan should be given.
This is a deterministic classification problem where InsureU is classifying if a customer will be in class 'Good' or 'Bad'.
Now we can write up the potential costs associated with each prediction and the eventual truth:

```{r special-002}
costs = matrix(c(-1, 0, 5, 0), nrow = 2, dimnames =
  list("Predicted Credit" = c("good", "bad"), Truth = c("good", "bad")))
costs
```

If InsureU thinks a customer has bad credit then there's no profit or loss, the loan is not provided.
If a customer has good credit and returns the loan then InsureU makes a \$1K profit, but if they default then InsureU loses \$5K - note above this is shown as `-1` and `-5` as cost-sensitive classification is a minimisation problem where we assume lower costs are associated with more positive outcomes.

### Cost-sensitive Measure

We will now see how to implement a more nuanced approach to classification errors with `r ref("MeasureClassifCosts")`.
This measure takes one argument, which is a matrix with row and column names corresponding to the class labels. Let's put our insurance example into practice, notice that we have already named the matrix as required for the measure:

```{r special-003}
cost_measure = msr("classif.costs", costs = costs)
cost_measure

learners = lrns(paste0("classif.", c("log_reg", "featureless", "ranger")))
bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 3)))
bmr$aggregate(cost_measure)[, c(4, 7)]
```

In this experiment the logistic regression performs best as it minimises the expected costs (and maximises expected profits); the featureless learner performs the worst.

### Thresholding

As we have discussed in @sec-basics, `r index("thresholding")` is a method to determine at what probability an observation will fall into one category or another.
Currently in our running example, the models above will predict a customer has good credit if the probability of good credit is greater than 0.5.
However, this may not be a sensible approach as the cost of a false positive and false negative is not equal.
A false positive results in a cost of +5 whereas a false negative only results in a cost of 0, hence we would likely prefer a model with a higher false negative rate and lower false positive rate.
This is highlighted in the `"threshold"` `autoplot`:

```{r special-004}
prediction = lrn("classif.log_reg", predict_type = "prob")$train(task)$predict(task)
autoplot(prediction, type = "threshold", measure = cost_measure)
```

As expected, the optimal threshold is greater than 0.5, indicating that to maximise profits, the majority of observations should be predicted to have bad credit.

Instead of having to manually threshold we can also use `r ref("PipeOpTuneThreshold")`, continuing the same example:

```{r special-005}
library(mlr3tuning)
gr = as_learner(po("learner_cv", lrn("classif.log_reg", predict_type = "prob")) %>>% po("tunethreshold", measure = cost_measure))

learners = list(gr, lrn("classif.log_reg"))
bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 3)))
bmr$aggregate(cost_measure)[, c(4, 7)]
```

As expected, our tuned learner performs much better and will even make InsureU a profit not a loss.

## Probabilistic Regression {#sec-probregr}

`r index("Probabilistic regression")` is a simple extension to deterministic regression in which the goal is to predict a full distribution around a single '`r index("point prediction")`'.
By example, instead of just predicting someone's height, $\mu$, we could also predict how confident we are about our prediction, usually estimated as a standard error, $\sigma$.
This 'confidence' could be expressed as a confidence interval, for example $\mu \pm \sigma$, or by assuming a distribution, for example $Norm(\mu = 120, \sigma = 2)$.

### Confidence interval predictions

Predicting a `r index("confidence interval")` is possible using features in `r mlr3`.
By setting `predict_type = "se"` for any `r ref("LearnerRegr")` object, enables predictions of the `r index("point prediction")` (`mean`) as well as measure of confidence for that prediction (`se`).
These can be manually combined to create a confidence interval.
An example is in @fig-probregrconfint where we assume a symmetric Normal 95% confidence interval.

```{r special-006, warning=FALSE}
#| fig-alt: x-label says "Response plus-minus 1.96se" and y-label says "Truth". Plot shows a black line running through the x=y axis. 7 blue points are plotted with error bars. Plot shows that 3 points are lying on the x=y line. Other points are not on the line and their error bars do not overlap with the line either.
#| fig-cap: Scatter plot of predictions (x) vs truth (y) where the straight line is $x=y$. Points lying on $x=y$ indicate very good predictions, points with error bars overlapping $x=y$ indicates truth is within 95% confidence interval.
#| label: fig-probregrconfint
t = tsk("mtcars")
split = partition(t, 0.8)
l = lrn("regr.ranger", predict_type = "se")
l$train(t, split$train)
p = l$predict(t, split$test)
autoplot(p, type = "confidence", quantile = 1.96)
```

This representation provides more detail about our prediction by including our confidence about the predictions.
One of the ways in which this is helpful is in telling us how 'far' our predictions are from the truth.
For example the top prediction in @fig-probregrconfint has wide error bars and is further from the truth than the third prediction which has narrower error bars and is closer to the truth.
Analogously to classification, we might think of the first prediction as "P(Y = 28) = 0.4" versus the third prediction as "P(Y = 24) = 0.8" (numbers not exact).
This level of detail allows more nuanced metrics to be utilised to measure probabilistic predictions, which we will turn to in the next section.

### Probability distribution predictions

In the example above we used the standard error to calculate an approximate symmetric Normal confidence interval.
Instead we could have plugged the estimated `reponse` and `se` into a `r index("probability distribution")` as an estimate of the distribution mean and standard deviation.
By assuming a probability distribution, we can make our predictions even more detailed by not just considering the mean of the distribution (the 'response' in `r mlr3` terminology), but also the probability density function, cumulative distribution function, and all other properties.
`r ref_pkg("distr6")` implements many probability distributions with `r ref_pkg("R6")` and provides a flexible interface that was originally designed to be used with `r mlr3proba`.
Below we demonstrate construction of two distributions (Uniform and Normal) in `r ref_pkg("distr6")` and how to access their methods and properties, below we call the `pdf` (probability density function) and `cdf` (cumulative distribution function) methods and look at the most important properties with `summary`.

```{r special-007}
library(distr6)
response = 5
se = 2

# construct Uniform distribution
unif = dstr("Uniform",
  lower = response - se * sqrt(3),
  upper = response + se * sqrt(3))
# summary of distribution properties
summary(unif)
# call pdf
unif$cdf(1:6)

# construct Normal distribution
norm = dstr("Normal", mean = response, sd = se)
# summary of distribution properties
summary(norm)
# evaluate cdf
norm$pdf(4:8)
```

Full details of the `r ref_pkg("distr6")` interface can be found on the package website (`r link("https://alan-turing-institute.github.io/distr6/")`).
In the above example we manually transformed the `response` and `se` 'predictions' to be compatible with a Uniform distribution construction, but learning these transformations can be time consuming.
`r mlr3proba` offers a pipeline (@sec-pipelines) that incorporates the key steps of distribution prediction for probabilistic regression, which are:

1. Train and predict a learner to make `response` predictions
2. Train and predict a learner to make `se` predictions
3. Select a probability distribution to model these predictions
4. Transform predictions and construct distribution
5. Return prediction

This pipeline, `"probregr"`, is a wrapper around `r ref("mlr3proba::PipeOpProbregr")`, and is showcased in the code below where we specify three inputs: 1) a random forest as the main learner, i.e., the one making the 'usual' regression point prediction; 2) a featureless learner to estimate the standard deviation, which simply predicts the future standard error as the sample standard deviation in the training set; and 3) we assume a Normal distribution for the predicted distribution. We visualise the quality of predictions in @fig-probregrdist.

```{r special-008}
#| fig-alt: Three Normally distributed curves are plotted in blue, red, and green. Each curve has a point of the same colour, representing the true value. The x-label says "x" and the y-label says "f(x)". The point on the green curve is very close to the distribution maximum. The point on the red curve is close to, but not exactly the maximum. The point on the blue curve is about a third of the way down from the top of the curve. The title of the plot says "Logloss = 2.82".
#| fig-cap: Three probability distribution predictions plotted as probability density functions. True values are represented as points. Shows fairly good predictions for the green and red observations as the truth is close to the distribution mean/mode but a less good (but not terrible) prediction for the blue observation. The logloss value is low but could be better (0 is the minimum).
#| label: fig-probregrdist
library(mlr3verse)
library(mlr3proba)
library(ggplot2)

# create our pipeline
l = as_learner(ppl("probregr",
  learner = lrn("regr.ranger"),
  learner_se = lrn("regr.featureless"),
  dist = "Normal")
)

# train and predict
task = tsk("mtcars")
split = partition(task)
l$train(task, split$train)
p = l$predict(task, split$test)
p

plot_probregr(p, 3, "point", "random") +
  ggtitle(sprintf("Logloss = %s", round(p$score(msr("regr.logloss")), 2)))
```

The `$distr` column in our `r ref("PredictionRegr")` contains objects inheriting from `r ref("distr6::Distribution")`, which allows querying of all standard probability distribution methods and properties, as we have seen in the code above.
This example shows that although none of our predictions were perfect -- the truth was not exactly in line with the peak of the predicted distributions -- in each case the truth was at least within an area of reasonable confidence.
In future versions of `r mlr3proba` we will extend functionality further by adding models that can make probabilistic predictions (e.g., with Bayesian methods), more losses (e.g., Brier score), and a dedicated predict type for interval predictions.
We now turn to the next task implemented in `r mlr3proba`, which follows closely from probabilistic regression.

## Survival Analysis {#sec-survival}

Survival analysis is concerned with the prediction or estimation of time-to-event distributions, for example the time until a marathon runner finishes a race.
The predictive problem is unique in that models are trained and tested on data that may include 'censoring'.
Censoring occurs when the event of interest does *not* take place, for example if a marathon runner gives up and does not finish a race.
Instead of throwing away information about censoring events, survival analysis datasets include a status variable that provides information about the 'status' of an observation, so in our marathon runner example we might write their outcome as (4, 1) if they finish the race at 4 hours, otherwise if they give up at 2 hours we would write (2, 1).
The key to understanding survival analysis is that we assume there exists a hypothetical time the marathon runner would have finished if they had not been censored, it is then the job of a survival learner to estimate what the true survival time would have been for a similar runner, assuming they are *not* censored (see @fig-censoring).
Mathematically, this is represented by the hypothetical event time, $Y$, the hypothetical censoring time, $C$, the observed outcome time, $T = min(Y, C)$, the event indicator $\Delta = (T = Y)$, and as usual some features, $X$.
Learners are trained on $(T, \Delta)$ but, critically, make predictions of $Y$ from previously unseen features.
This means that unlike classification and regression, learners are trained on two variables, the time until an event takes place and the status or event indicator.
For convenience, this is often stored using the `r ref("survival::Surv")` object.
An example of this is in the code below, where we randomly generate 6 survival times and 6 event indicators, an outcome with a `+` indicates the outcome is censoring, otherwise it is the event of interest.
```{r special-009}
library(survival)
Surv(runif(6), rbinom(6, 1, 0.5))
```

Readers familiar with survival analysis will recognise that the description above applies specifically to 'right-censoring'.
Currently, this is the only form of censoring available in the `r mlr3` universe hence restricting our discussion to that setting.
For a good introduction to survival analysis see @Collett2014 or for machine learning in survival analysis specifically see @MLSA.

For the remainder of this section we will look at how `r mlr3proba` [@mlr3proba] extends the building blocks of `r mlr3` for survival analysis. We will begin by looking at objects used to construct machine learning tasks for survival analysis, then we will turn to the learners we have implemented to solve these tasks, before looking at measures for evaluating survival analysis predictions, and then finally we will consider how to transform prediction types.

```{r special-010, echo=FALSE}
#| fig-alt:
#| fig-cap: Dead and censored subjects (y-axis) over time (x-axis). Black diamonds indicate true death times and white circles indicate censoring times. Vertical line is the study end time. Subjects 1 and 2 die in the study time. Subject 3 is censored in the study and (unknown) dies within the study time. Subject 4 is censored in the study and (unknown) dies after the study. Subject 5 dies after the end of the study. Figure and caption from @Sonabend2021b.
#| label: fig-censoring
library(ggplot2)
ggplot(data.frame(x = c(1, 2, 5, 7, 10), y = 1:5), aes(x = x, y = y)) +
  geom_blank() +
  geom_vline(xintercept = 8) +
  geom_segment(y = 1, yend = 1, x = 0, xend = 7) +
  geom_segment(y = 2, yend = 2, x = 0, xend = 8) +
  geom_segment(y = 3, yend = 3, x = 0, xend = 4) +
  geom_segment(y = 3, yend = 3, x = 4, xend = 6, linetype = 2) +
  geom_segment(y = 4, yend = 4, x = 0, xend = 1) +
  geom_segment(y = 4, yend = 4, x = 1, xend = 9, linetype = 2) +
  geom_segment(y = 5, yend = 5, x = 0, xend = 8) +
  geom_segment(y = 5, yend = 5, x = 8, xend = 9, linetype = 2) +
  geom_point(x = 4, y = 3, shape = 21, colour = "black", fill = "white", size = 4) +
  geom_point(x = 6, y = 3, shape = 18, size = 4) +
  geom_point(x = 7, y = 1, shape = 18, size = 4) +
  geom_point(x = 8, y = 2, shape = 18, size = 4) +
  geom_point(x = 9, y = 5, shape = 18, size = 4) +
  geom_point(x = 8, y = 5, shape = 21, size = 4, fill = "white") +
  geom_point(x = 1, y = 4, shape = 21, size = 4, fill = "white") +
  geom_point(x = 9, y = 4, shape = 18, size = 4) +
  labs(y = "Subject", x = "Time") +
  scale_x_continuous(labels = as.character(1:10), breaks = 1:10) +
  theme_minimal() + theme(panel.grid.minor = element_blank())
```

### TaskSurv

As we saw in the introduction to this section, survival algorithms require two targets for training, this means the new `r ref("TaskSurv")` object  expects two targets.
The simplest way to create a survival task is to use `r ref("as_task_surv")`, as seen in the following code chunk.

```{r special-011}
library("mlr3verse")
library("mlr3proba")
library("survival")

task = as_task_surv(survival::rats, time = "time",
  event = "status", type = "right", id = "rats")

print(task)
```

The inputs to `r ref("as_task_surv")` follow `r ref("survival::Surv")`, so in theory there is support for left and interval censoring, as well as for competing risks, however no learners have been implemented at the time of writing that allow for these methods.

In construction, `r ref("TaskSurv")` objects coerce the target columns into a survival object.
We have also included a specific `autoplot` for these objects to quickly create Kaplan-Meier plots to visualise the outcome.

```{r special-012, warning=FALSE, message=FALSE}
head(task$truth())
autoplot(task)
```

We have also included a few pre-loaded tasks in `r mlr3proba` that can be used in experiments

```{r special-013}
as.data.table(mlr_tasks)[task_type == "surv"]
```

### LearnerSurv, PredictionSurv and predict types

The interface for `r ref("LearnerSurv")` and `r ref("PredictionSurv")` objects is identical to the regression and classification settings discussed in @sec-basics.
Similarly to these settings, survival learners are constructed with `"lrn"`, a few of the available learners are listed below.

```{r special-014}
mlr_learners$keys("^surv")
```

The primary difference in constructing a survival learner (and the resulting predictions) is the number of possible `predict_type` options that can be passed to `r ref("lrn")`.

Similarly to deterministic regression (@sec-tasks), a survival model can make a single point prediction, and similarly to probabilistic regression (@sec-probregr), a survival model could also make a prediction of a probability distribution.
`r mlr3proba` has a slightly different design interface to `r mlr3` as all possible types of prediction ('predict types') are returned for all survival models.
The reason for this decision is that all these predict types can be transformed to one another and it is therefore computationally simpler to return all at once instead of rerunning models to change predict type.
In survival analysis, the following predictions can be made:

* `response` - Predicted survival time.
* `distr` - Predicted survival distribution, either discrete or continuous.
* `lp` - Linear predictor calculated as the fitted coefficients multiplied by the test data.
* `crank` - Continuous risk ranking.

We will go through each of these prediction types in more detail and with examples to make them less abstract.
We will use the following setup for most of the examples.
In this chunk we are partitioning the `r ref("survival::rats")` and training a `r define("Cox Proportional Hazards")` model on the training set and making predictions for the predict set.
For this model, all predict types except `response` can be computed.

```{r special-015}
t = tsk("rats")
split = partition(t)
p = lrn("surv.coxph")$train(t, split$train)$predict(t, split$test)
p
```

#### `predict_type = "response"`

Counterintuitively for many, the `response` prediction of predicted survival times is actually the least common predict type in survival analysis.
The likely reason for this is due to the presence of censoring.
We rarely observe the true survival time for many observations and therefore it is unlikely any survival model can confidently make predictions for survival times.
This is illustrated in the code below.
We train and predict from a survival support vector machine from `r ref_pkg("survivalsvm")`, we use `type = "regression"` to select the algorithm that optimises survival time predictions.
We then compare the predictions from the model to the true data.
As can be seen from the output, our predictions are all less than the true observed time, which means we know our model definitely underestimated the truth.
However, because each of the true values are censored times, we have absolutely no way of knowing if these predictions are slightly bad or absolutely terrible, i.e., the true survival times could be $105, 92, 92$ or they could be $300, 1000, 200$.
Hence, with no realistic way to evaluate these models, survival time predictions are rarely useful.
It is also worth noting here that the nature of survival analysis means that it is rare to find high-quality datasets without a substantial amount of censoring.

```{r special-016}
library(mlr3extralearners)
# we use a different learner here as few survival models make a response prediction
pred = lrn("surv.svm", type = "regression", gamma.mu = 1e-3)$
  train(t, split$train)$predict(t, split$test)
data.frame(pred = pred$response[1:3], truth = pred$truth[1:3])
```

#### `predict_type = "distr"`

So unlike regression in which deterministic/point predictions are most common, in survival analysis the most common prediction type is actually distribution predictions.
You will therefore find that the majority of survival models in `r mlr3proba` will make distribution predictions by default.
As with probabilistic regression, we implement these predictions using the `r ref_pkg("alan-turing-institute/distr6")` interface, which allows visualisation and evaluation of survival curves.
In the example below we train a Cox PH model on the `rats` dataset and then plot the survival function of the first three predictions and evaluation this function for all three predictions at $t = 40,70,120$.

```{r special-017}
t = tsk("rats")
split = partition(t)
p = lrn("surv.coxph")$train(t, split$train)$predict(t, split$test)
plot(p$distr[1:3], fun = "survival")
p$distr[1:3]$survival(c(40, 70, 120))
```

#### `predict_type = "lp"`

`lp`, often written as $\eta$ in academic writing, is computationally the simplest prediction and has a natural analogue in regression modelling.
Readers familiar with linear regression will know that when fitting a simple linear regression model, $Y = X\beta$, we are actually estimating the values for $\beta$, and the estimated linear predictor is then $X\hat{\beta}$, where $\hat{\beta}$ are our estimated values.
In simple survival models, the linear predictor is the same quantity (but estimated in a slighty more complicated way).
The implementations in `r mlr3proba` are primarily machine-learning focused and few of these models have a simple linear form, which means that `lp` cannot be computed for most of these.

#### `predict_type = "crank"`

The final prediction type, `crank`, is the most common in survival analysis and also the most confusing.
Academic texts will often refer to 'risk' predictions in survival analysis (hence why survival models are often known as 'risk prediction models'), without defining what 'risk' means.
Often risk is defined as $exp(\eta)$ as this is a common quantity found in simple linear survival models.
However, sometimes risk is defined as $exp(-\eta)$, and sometimes it can be an arbitrary quantity that does not have a meaningful interpretation.
To prevent this confusion in `r mlr3proba`, we define the predict type `crank`, which stands for **c**ontinuous **rank**ing.
This is best explained by example.
Continuing from the previous example we output the first three crank predictions.
The output tells us that the first rat is at the higher risk of death (larger values represent higher risk) and the second rat is at the lowest risk of death.
The distance between predictions also tells us that the difference in risk between the first and second rat is smaller than the difference between the second and third.
The key points in this interpretation are: 1) lower values represent lower risk; 2) the difference between values has meaning but should not be over-interpreted; and 3) the actual values are meaningless.
The last point is the most important, comparing these values between samples is not meaningful.
The reason why this predict type interesting, and so common, is because it allows identifying observations at lower/higher risk to each other, which is useful for resource allocation and prioritisation, and clinical trials (e.g., "are people in a treatment arm at lower risk of X than people in the control arm.").

```{r special-018}
p$crank[1:3]
```

### MeasureSurv

Survival models in `r mlr3proba` are evaluated with `r ref("MeasureSurv")` objects, which are constructed in the usual way with `"msr"`; measures currently implemented are listed below.

```{r special-019}
mlr_measures$keys("^surv")
```

Survival measures will evaluate one of the four prediction types listed above.
In general survival measures can be grouped into the following:

1. Discrimination measures - Quantify if a model correctly identifies if one observation is at higher risk than another. Evaluates `crank` predictions.
1. Calibration measures - Quantify if the average prediction is close to the truth (all definitions of calibration are unfortunately vague in a survival context). May evaluate `crank` or `distr` predictions.
1. Scoring rules - Quantify if probabilistic predictions are close to true values. Evalutes `distr` predictions.

```{r special-020}
head(as.data.table(mlr_measures)[task_type == "surv", c("key", "predict_type")])
```

There is a lot of debate in the literature around the 'best' survival measures to use to evaluate models, as a general rule we recommend `"surv.rcll"` to evaluate the quality of `distr` predictions, `"surv.cindex"` to evaluate a model's discrimination, and `"surv.dcal"` to evaluate a model's calibration.

### Composition

Throughout `r mlr3proba` documentation we refer to "native" and "composed" predictions.
We define a 'native' prediction as the prediction made by a model without any post-processing, whereas a 'composed' prediction is one that is returned after post-processing.

#### Internal composition

In `r mlr3proba` we make use of composition internally to return a `"crank"` prediction for every learner.
This is to ensure that we can meaningfully benchmark all models according to at least one critierion.
The package uses the following rules to create `"crank"` predictions:

1. If a model returns a 'risk' prediction then `crank = risk` (we may multiply this by $-1$ to ensure the 'low value low risk' interpretation).
1. Else if a model returns a `response` prediction then we set `crank = -response`.
1. Else if a model returns a `lp` prediction then we set `crank = lp` (or `crank = -lp` if needed).
1. Else if a model returns a `distr` prediction then we set `crank` as the sum of the cumulative hazard function (see @Sonabend2022 for full discussion as to why we picked this method).

#### Explicit composition and pipelines

At the start of this section we mentioned that it is possible to transform prediction types between each other.
In `r mlr3proba` this is possible with 'compositor' pipelines (@sec-pipelines).
There are a number of pipelines implemented in the package but two in particular focus on predict type transformation:

1. `r ref("pipeline_crankcompositor")` - Transforms a `"distr"` prediction to `"crank"`; and
1. `r ref("pipeline_distrcompositor")` - Transforms a `"lp"` prediction to `"distr"`.

In practice, the second pipeline is more common as we internally use a version of the first pipeline whenever we return predictions from survival models.
You may want to use the first pipeline to overwrite the default method of transforming distributions to rankings.
For now we will just look at the second pipeline.

In the example below we load the `rats` dataset, remove factor columns, and then partition the data into training and testing.
We construct the `distrcompositor` pipeline around a survival GLMnet learner which by default can only make predictions for `"lp"` and `"crank"`.
In the pipeline we specify that we will estimate the baseline distribution with a `r define("Kaplan-Meier")` estimator (`estimator = "kaplan") and that we want to assume a proportional hazards form for our estimated distribution.
We then train and predict in the usual way and in our output we can now see a `distr` prediction.

```{r special-021, eval = FALSE}
library(mlr3verse)
library(mlr3extralearners)

task = tsk("rats")
task$select(c("litter", "rx"))
split = partition(task)

learner = lrn("surv.glmnet")

# no distr output
learner$train(task, split$train)$predict(task, split$test)

pipe = as_learner(ppl(
  "distrcompositor",
  learner = learner,
  estimator = "kaplan",
  form = "ph"
))

# now with distr
pipe$train(task, split$train)$predict(task, split$test)
```

Mathematically, we have done the following:

1. Assume our estimated distribution will have the form $h(t) = h_0(t)exp(\eta)$ where $h$ is the hazard function and $h_0$ is the baseline hazard function.
1. Estimate $\hat{\eta}$ prediction using GLMnet
1. Estimate $\hat{h}_0(t)$ with the Kaplan-Meier estimator
1. Put this all together as $h(t) = \hat{h}_0(t)exp(\hat{\eta})$

For more details about prediction types and compositions we recommend @Kalbfleisch2011.


### Putting it all together {#sec-survival-all}

Finally, let's put all the above into practice in a small benchmark experiment.
We first load the `r ref("grace")` dataset and subset to the first 500 rows.
We then select the right-censored logloss measure to evaluate distribution predictions, setup the same pipeline we used in the previous experiment, and load a Cox PH and Kaplan-Meier estimator.
We run our experiment with 3-fold cross-validation and aggregate the results.

```{r special-022, warning=FALSE}
library(mlr3verse)
library(mlr3extralearners)

task = tsk("grace")$filter(1:500)
measures = msrs(paste0("surv.", c("cindex", "dcalib", "rcll")))

pipe = as_learner(ppl(
  "distrcompositor",
  learner = lrn("surv.glmnet"),
  estimator = "kaplan",
  form = "ph"
))
pipe$id = "Coxnet"
learners = c(lrns(c("surv.coxph", "surv.kaplan")), pipe)

bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 3)))
bmr$aggregate(measures)[,
  c("learner_id", "surv.cindex", "surv.dcalib", "surv.rcll")]
```

In this small experiment, Coxnet and Cox PH have the best discrimination, Cox PH has the best calibration, and Coxnet and Cox PH seem equally good at overall predictive accuracy.

## Density Estimation {#sec-density}

Density estimation is the learning task to estimate the unknown distribution from which a univariate dataset is generated, or put more simply to estimate the probability density (or mass) function for a variable.
Density estimation is implemented in `r mlr3proba`, similarly to the previous two tasks that can also predict probability distributions (hence the name "**mlr3proba**bilistic").
Unconditional density estimation, i.e., estimation of a target without any covariates, is viewed as an unsupervised task, which means the 'truth' is never known.
For a good overview to density estimation see *Density estimation for statistics and data analysis* [@Silverman1986].

The package `r mlr3proba` extends `r mlr3` with the following objects for density estimation:

* `r ref("mlr3proba::TaskDens", text = "TaskDens")` to define density tasks
* `r ref("mlr3proba::LearnerDens", text = "LearnerDens")` as the base class for density estimators
* `r ref("mlr3proba::PredictionDens", text = "PredictionDens")` for density predictions
* `r ref("mlr3proba::MeasureDens", text = "MeasureDens")` as specialized class for density performance measures

We will consider each in turn.

### TaskDens

As density estimation is an unsupervised task, there is no target for prediction.
In the code below we construct a density task using `r ref("as_task_dens")` which only takes one argument, a `data.frame` type object with exactly one column (which we will use to estimate the underlying distribution).

```{r special-023}
task = as_task_dens(data.frame(x = rnorm(1000)))
task
```

As with other tasks, we have included a couple tasks that come shipped with `r mlr3proba`:

```{r}
as.data.table(mlr_tasks)[task_type == "dens", c(1:2, 4:5)]
```

### LearnerDens and PredictionDens

Density learners can make one of three possible prediction types.

1. `distr` - probability distribution
1. `pdf` - probability density function
1. `cdf` - cumulative distribution function

All learners will return a `distr` and `pdf` prediction but only some can make `cdf` predictions.
Similarly to survival analysis and probabilistic regression, the `distr` predict type is implemented using `r ref_pkg("alan-turing-institute/distr6")`.

```{r special-024}
learn = lrn("dens.hist")
p = learn$train(task, 1:900)$predict(task, 901:1000)
x = seq.int(-2, 2, 0.01)
ggplot(data.frame(x = x, y = p$distr$pdf(x)), aes(x = x, y = y)) +
  geom_line() + theme_minimal()
```

The `pdf` and `cdf` predict types are simply wrappers around `distr$pdf` and `distr$cdf` respectively, which is best demonstrated by example:

```{r special-025}
learn = lrn("dens.hist")
p = learn$train(task, 1:10)$predict(task, 11:13)
p
cbind(p$distr$pdf(task$data()$x[11:13]), p$pdf[1:3])
```

The reason for returning `pdf` and `cdf` in this way is to enable measures to be used to evaluate the quality of our estimations, which we will return to in the next section.

### MeasureDens

Currently the only measure implemented in `r mlr3proba` for density estimation is logloss, which is defined in the same way as in classification and probabilistic regression, $L(y) = -log(f_Y(y))$, where $\hat{f}_Y$ is our estimated probability density function.
Putting this together with the above we are now ready to train a density learner and to estimate a distribution.
Carrying on from before:

```{r special-026}
meas = msr("dens.logloss")
meas

p$score(meas)
```

### Unconditional density estimation vs. probabilistic regression

Some readers may spot parallels between density estimation and probabilistic regression.
In fact, density estimation is just a special case of probabilistic regression (@sec-probregr), occurring when we make an unconditional prediction.
For example, given average rainfall across the USA, estimating the distribution of rainfall would be unsupervised density estimation.
If we now wanted to estimate the distribution of rainfall next year given rainfall this year, then we would be back to supervised probabilistic regression.
This is illustrated in the code below.
We load the `r ref("precip")` dataset and train, predict, and score a Normal kernel density estimation density learner.
We then create a new dataset based on `r ref("precip")` but with an added variable for rainfall next year.
Now we can specify a target for prediction based on another covariate, thus making this a supervised learning task.
We then use the `probregr` pipeline (@sec-probregr) and train, predict, and score our random forest learner.
Note the key difference in these outputs is that whilst both make distribution predictions, the former (density) *estimates* a single probability distribution for the variable of interest, whereas the latter (regression) *predicts* a distribution for each observation.

```{r special-027}
task = tsk("precip")
split = partition(task)
learn = lrn("dens.kde", kernel = "Norm")
p = learn$train(task, split$train)$predict(task, split$test)
p
p$score(msr("dens.logloss"))

df = task$data()
df$next_year = (df$precip * 1.5) + rnorm(length(df$precip))
task = as_task_regr(df, target = "next_year")
pipe = as_learner(ppl("probregr", lrn("regr.ranger"), dist = "Normal"))
p = pipe$train(task, split$train)$predict(task, split$test)
p
p$score(msr("regr.logloss"))
```

### Putting it all together {#sec-density-all}

Finally, we conduct a small benchmark study on the `r ref("mlr_tasks_faithful")` task using some of the integrated density learners:

```{r special-028, message=FALSE, warning=FALSE, results='hide'}
library(mlr3extralearners)
task = tsk("faithful")
learners = lrns(c("dens.hist", "dens.pen", "dens.kde"))
measure = msr("dens.logloss")
bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 3)))
bmr$aggregate(measure)
autoplot(bmr, measure = measure)
```

The results of this experiment show that the sophisticated Penalized Density Estimator does not outperform the baseline histogram, but that the Kernel Density Estimator has at least consistently better (i.e. lower logloss) results.

## Cluster Analysis {#sec-cluster}

`r index("Cluster analysis")` is another unsupervised task implemented in `r mlr3`.
The objective of cluster analysis is to group data into clusters, where each cluster contains similar observations.
The similarity is based on specified metrics that are task and application-dependent.
Unlike classification where we try to predict a category for each observation, in cluster analysis there is no 'true' label or category to predict.

The package `r mlr3cluster` extends `r mlr3` with the following objects for cluster analysis:

* `r ref("mlr3cluster::TaskClust", text = "TaskClust")` to define clustering tasks
* `r ref("mlr3cluster::LearnerClust", text = "LearnerClust")` as base class for clustering learners
* `r ref("mlr3cluster::PredictionClust", text = "PredictionClust")` as specialized class for `r ref("Prediction")` objects
* `r ref("mlr3cluster::MeasureClust", text = "MeasureClust")` as specialized class for performance measures

We will consider each in turn.

### TaskClust

Similarly to density estimation (@sec-density), there is no target for prediction and so no `truth` field in `r ref("TaskClust")`.
In the code below we construct a cluster task using `r ref("as_task_clust")` which only takes one argument, a `data.frame` type object.

```{r special-029, warning=FALSE, message=FALSE}
library(mlr3verse)
task = as_task_clust(tsk("penguins_simple")$data()[, -1])
task

autoplot(task)
```

Currently we only have one clustering task shipped with `r mlr3cluster`:

```{r special-030}
as.data.table(mlr_tasks)[task_type == "clust", c(1:2, 4:5)]
```

### LearnerClust and PredictionClust

As with density estimation, we refer to `training` and `predicting` for clustering to be consistent with the `r mlr3` interface, but strictly speaking this should be `clustering` and `assigning` (the latter we will return to shortly).
Two `predict_types` are available for clustering learnings:

1. `partition` - which cluster does an observation fall into
1. `prob` - probability of an observation belonging to each cluster

Hence, similarly to classification, prediction types of clustering learners are either deterministic (`partition`) - estimate the cluster an observation falls into - or probabilistic (`prob`) - estimate the probability of an observation falling into each cluster.

Below we construct a C-Means clustering learner with `prob` prediction type, train it on the `r ref("USArrests")` datasets and then return the cluster assignments (`$assignments`) for each observation - as with other tasks, `learner$model` returns the model representation specificed by the underlying implementation.

```{r special-031}
learner = lrn("clust.cmeans", predict_type = "prob", centers = 3)
learner

learner$train(task)
learner$assignments[1:6]
```

As clustering is unsupervised, it rarely makes sense to use `predict` for new data however this is still possible using the `r mlr3` interface.
Continuing our example above:

```{r special-032, warning = FALSE, message = FALSE}
# using same data for estimation (rare use case)
learner$train(task, 1:30)$predict(task, 31:32)

# using same data for estimation (common use case)
prediction = learner$train(task)$predict(task)
autoplot(prediction, task)
```

Whilst two prediction types are possible, there are some learners where 'prediction' can never make sense, for example in `r define("hierarchical clustering")`.
In hierarchical clustering, the goal is to build a hierarchy of nested clusters by either splitting large clusters into smaller ones or merging smaller clusters into bigger ones.
The final result is a tree or dendrogram which can change if a new data point is added.
For consistency, `r mlr3cluster` offers `predict` method for hierarchical clusterers but it simply assigns all points to a specified number of clusters by cutting the resulting tree at a corresponding level.

```{r special-033}
learner = lrn("clust.hclust")
learner$train(task)
learner$predict(task)
autoplot(learner)
```

### MeasureDens

As previously discussed, unsupervised tasks do not have ground truth data to compare to in model evaluation.
However, we can still measure the quality of cluster assignments by quantifying how closely objects within the same cluster are related (cluster cohesion) as well as how distinct different clusters are from each other (cluster separation).

There are a few built-in evaluation metrics available to assess the quality of clustering.

```{r special-034}
mlr_measures$keys("clust")
```

Two common measures are the within sum of squares (WSS), `r ref("mlr_measures_clust.wss")`, and silhouette quality index (SQI), `r ref("mlr_measures_clust.silhouette")`.
WSS calculates the sum of squared differences between observations and centroids, which is a quantification of cluster cohesion (smaller values indicate clusters more compact).
SQI quantifies how well each point belongs to its assigned cluster versus neighboring cluster, where scores closer to `1` indicate well clustered and scores closer to `-1` indicate poorly clustered.

Putting this together with the above we can now score our cluster estimation (note we must pass the `task` to `r ref("msr")`):

```{r special-035}
meas = msrs(c("clust.wss", "clust.silhouette"))

prediction$score(meas, task = task)
```

The very high WSS and middling SQI is 0.6 indicates that our clusters could do with a bit more work.
Often reducing an unsupervised task to a quantitative measure may not be useful (given no ground truth) and instead visualisation may be a more effective tools for assessing the quality of the clusters.

### Visualization

As with other tasks, we use `r mlr3viz` to visualise performance of cluster learners. The two most important plots are principal components analysis (PCA) and silhouette plots.

PCA is a commonly used dimension reduction method in machine learning to reduce the number of variables in a dataset or to visualise the most important 'components', which are linear transformations of the dataset features.
Components are considered more important if they have higher variance (and therefore more predictive power).
In the context of clustering, by plotting observations against the first two components, and then colouring them by cluster, we would expect to see distinct groups of observations if we are clustering well.

Continuing our running example, by plotting a PCA we see that our model has created cleanly separated clusters along the first two principal components.

```{r special-036, message=FALSE, warning=FALSE}
autoplot(prediction, task, type = "pca")
```

Silhouette plots help to visually assess the quality of the analysis and help choose a number of clusters for a given data set.
The average silhouette index is 0.5, which means observations that are scored below this line are considered to be poor estimations whereas observations above are well clustered, the SQI is then jus the average across all observations.
To visualise this, in a silhouette plot, we plot a dotted line at 0.5 and a bar for each data point, coloured by cluster.The red dotted line shows the mean silhouette value and each bar represents a data point.

Continuing our running example, we find that the majority of observations are actually below the average line, and therefore the quality of our prediction is not very good meaning that many observations are likely assigned to the wrong cluster.

```{r special-037, message=FALSE, warning=FALSE}
autoplot(prediction, task, type = "sil")
```

### Putting it all together {#sec-cluster-all}

Finally, we conduct a small benchmark study using the same task and with a few integrated cluster learners:

```{r special-038, message=FALSE, warning=FALSE}
task = as_task_clust(USArrests)
learners = list(
  lrn("clust.featureless"),
  lrn("clust.kmeans"),
  lrn("clust.cmeans", centers = 3L)
)
measures = list(msr("clust.wss"), msr("clust.silhouette"))
bmr = benchmark(benchmark_grid(task, learners, rsmp("cv", folds = 3)))
bmr$aggregate(measures)[, c(4, 7, 8)]
```

The experiment shows that using c-means algorithm with three centers produces a better within sum of squares score than any other learner considered. However, the featureless learner performs the best when considering silhouette measure which takes into the account both cluster cohesion and separation.

## Spatial Analysis {#sec-spatiotemporal}

The final task we will discuss in this book is `r define("spatial analysis")`.
Spatial analysis can be a subset of any other machine learning task (e.g., regression or classification) but is defined by the presence of spatial information in a dataset, usually stored as coordinates that are often named "x" and "y" or "lat" and "lon".

Spatial data has additional complexity due to 'autocorrelation'.
Where `r index("correlation")` is defined as a statistical association *between two* variables, `r define("autocorrelation")` is a statistical association *within one* variable.
In machine learning terms, in a datset with features and observations, correlation occurs when two or more features are statistically associated in some way, whereas autocorrelation occurs when two or more observations are statistically associated across one feature.
Autocorrelation therefore violates one of the fundamental assumptions of ML, that all observations in a dataset are independent [@hastie2001].
Violation of this assumption results in lower confidence about the quality of a trained ML model, meaning we cannot trust any performance estimates.

Autocorrelation is present in spatial data as there is implicit information encoded in coordinates, such as whether two observations (e.g., cities, countries, continents) are close together or far apart.
By example, say we are predicting the number of cases of a disease two months after outbreak in Germany.
Outbreaks radiate outwards from an epicentre and therefore countries closer to Germany will have higher numbers of cases and countries further aware will have lower numbers (Fig @fig-autocorrelation, bottom).
Therefore the 'country' variable has autocorrelation, as countries that are closer (in the figure below this is defined by Euclidean distance) have higher case numbers and vice versa.

```{r special-039, warning=FALSE, echo=FALSE, message=FALSE}
#| label: fig-autocorrelation
#| fig-cap: Heatmaps where darker countries indicate higher number of cases and lighter countries indicate lower number of cases of imaginary Disease X with epicenter in Germany. The top map imagines a world in which there is no spatial autocorrelation and the number of cases of a disease is randomly distributed. The bottom map shows a more accurate world in which the number of cases radiate outwards from the epicentre (Germany).
#| fig-alt: Image shows two separate maps of Europe. Top mat has a random distribution of colours from yellow to red. Bottom mat shows darkest colour (dark orange) at Germany with increasing lightness as the countries are increasingly further away.

library(dplyr)
library(patchwork)

m = map_data("world")
m = m[m$region == "Germany", ]
x1 = mean(m$lon)
x2 = mean(m$lat)
m = map_data("world")
m = m %>% dplyr::filter(long > -15, long < 40, lat > 36, lat <  60)
m$dist = -sqrt((m$lon - x1)^2 + (m$lat - x2)^2)
m$rand = rnorm(nrow(m), mean(m$dist), sd(m$dist))
cou = m %>% group_by(region) %>% summarise(long = mean(long), lat = mean(lat))

p = ggplot(data = m) +
  scale_fill_gradient2(low = "white", mid = "orange", high = "red", midpoint = mean(m$dist)) +
  theme_minimal() +
  theme(legend.position = "n") +
  xlim(-10, 25) + ylim(37, 58)

t = geom_text(aes(label = region, x = long, y = lat), data = cou, size = 2)

p1 = p + geom_map(map = m, aes(long, lat, map_id = region, fill = rand)) + t

p2 = p + geom_map(map = m, aes(long, lat, map_id = region, fill = dist)) + t

p1 / p2
```


Unlike other tasks we have looked at in this chapter, there is no underlying difference to algorithms, instead we provide new resampling methods in `r mlr3spatial` and `r mlr3spatiotempcv` to account for any overconfidence in predictions that might occur from autocorrelation.
To make use of these resamplings we have also implemented special tasks to hold spatial data.

Throughout this section we will use the `r ref("mlr3spatiotempcv::ecuador")` dataset and task.

### `TaskXST`

To make use of spatial resampling methods, we have implemented two extensions of `r ref("TaskClassif")` and `r ref("TaskRegr")` to accommodate spatial data, `r ref("mlr3spatiotempcv::TaskClassifST")` and `r ref("mlr3spatiotempcv::TaskRegrST")` respectively.
Below we only show classification examples but regression follows trivially.

```{r special-040}
library(mlr3spatial)
library(mlr3spatiotempcv)
# create task from `data.frame`
task = as_task_classif_st(ecuador, id = "ecuador_task", target = "slides",
  positive = "TRUE", coordinate_names = c("x", "y"), crs = 32717)

# or create task from 'sf' object
data_sf = sf::st_as_sf(ecuador, coords = c("x", "y"), crs = 32717)
task = as_task_classif_st(data_sf, target = "slides", positive = "TRUE")
task
```

Once a task is created, you can train and predict as normal.

```{r special-041}
lrn("classif.rpart")$train(task)$predict(task)
```

However as discussed above, it is best to use the specialised resampling methods in order to have trustworthy estimates of model performance.

### Spatiotemporal Cross-Validation {#spatiotemp-cv}

Before we look at the spatial resampling methods implemented in `r mlr3spatiotempcv` we will first show what can go wrong if standard resampling methods are used.
Below we benchmark a decision tree on the `r ref("mlr_tasks_ecuador")` task using two different repeated cross-validation resampling methods, the first "NSpCV" is not optimised for spatial data, the second is optimised for spatial data.
The example highlights how "NSpCV" makes it appear as if the decision tree is performing better than it is with significantly higher estimated performance, however this is an overconfident prediction due to the autocorrelation in the data.

```{r special-042}
task = tsk("ecuador")

learner = lrn("classif.rpart", predict_type = "prob")
rsmp_nsp = rsmp("repeated_cv", folds = 3, repeats = 2, id = "NSpCV")
rsmp_sp = rsmp("repeated_spcv_coords", folds = 3, repeats = 2, id = "SpCV")

design = benchmark_grid(task, learner, c(rsmp_nsp, rsmp_sp))
bmr = benchmark(design)
bmr$aggregate(msr("classif.acc"))[, c(5, 7)]
```

In the above example, the non-spatial resampling method means that the train and test data are very similar due to spatial autocorrelation, and so this is almost as bad as evaluating a model on the training data (see @sec-basics).
In contrast, the spatial method has accommodated for autocorrelation and the test data is therefore less correlated (though still some association will remain) with the training data.

Visually this can be seen using the built-in `autoplot` methods.
In @fig-sprsmp we visualise how the task is partitioned according to the spatial resampling method (@fig-sprsmp top) and non-spatial resampling method (@fig-sprsmp bottom).
There is a clear separation between coordinates using the 'correct' spatial resampling method whereas the train and test splits overlap significantly (and are therefore highly correlated) using the non-spatial method.

```{r special-043}
#| layout-nrow: 2
#| label: fig-sprsmp
#| fig-cap: Charts
#| fig-subcap:
#|   - "First"
#|   - "Second"
autoplot(rsmp_sp, task, fold_id = c(1:3), size = 0.7) +
  scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) +
  scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))

autoplot(rsmp_nsp, task, fold_id = c(1:3), size = 0.7) +
  scale_y_continuous(breaks = seq(-3.97, -4, -0.01)) +
  scale_x_continuous(breaks = seq(-79.06, -79.08, -0.02))
```

Now we have seen why spatial resampling methods matter we can take a look at what is implemented in `r mlr3spatiotempcv`.
At the time of writing we have implemented 19 different resampling methods for spatial and spatiotemporal analysis.
These are categorised into:

* Blocking - Create rectangular blocks in 2D/3D space (as opposed to partitioning in 1D space used in non-spatial methods)
* Buffering - Create buffering zones to remove observations between train and test sets
* Spatiotemporal clustering - Clusters based on coordinates (and/or time-points)
* Feature space clustering - Clusted in feature space and not necessarily spatiotemporal
* Partitioning - Grouped by factor variables

The choice of method may depend on specific characteristics of the dataset and there is no easy rule to pick one method over another.
The documentation for each of the implemented methods includes details of each method as well as references to original publications.
Full details of different methods can be found in @Schratz2021.

:::{.callout-tip}
## Spatio*temporal* resampling

We have focused on spatial analysis but referred to "spatiotemporal" and "spatiotemp".
The resampling methods discussed in this section extend trivially to temporal analysis (or spatial and temporal analysis combined) by setting the `"time"` `col_role` in the task (@sec-performance).
See the `r mlr3spatiotempcv` visualisation vignette for specific details about 3D spatiotemporal visualisation `r link("https://mlr3spatiotempcv.mlr-org.com/articles/spatiotemp-viz.html")`.
:::


### Spatial Prediction {#sec-spatial-prediction}

Until now we have looked at resampling to accommodate spatiotemporal *features*, but what if you want to make spatiotemporal *predictions*?
In this case the goal is to make classification or regression predictions for spatial coordinates at a pixel level, usually by creating and labelling raster images.

To enable these predictions we have created a new function, `r ref("predict_spatial")` which users a different interface to allow spatial predictions using any of the following spatial data classes:

* `stars` (from package `r ref_pkg("stars")`)
* `SpatRaster` (from package `r ref_pkg("terra")`)
* `RasterLayer` (from package `r ref_pkg("raster")`)
* `RasterStack` (from package `r ref_pkg("raster")`)

To use a raster image for prediction, it must be wrapped in `r ref("TaskUnsupervised")`.
In the example below we load the `leipzig_points` file for training and coerce this to a spatiotemporal task with `r ref("as_task_classif_st")`, and we load the `leipzig_raster` raster image and coerce this to an unsupervised task.
Both files are shipped with, and can be downloaded from, `r mlr3spatial`.

```{r special-045}
library(sf)

# load sample points
leipzig_vector = sf::read_sf(system.file("extdata",
  "leipzig_points.gpkg", package = "mlr3spatial"),
  stringsAsFactors = TRUE)
# create training data
tsk_leipzig = as_task_classif_st(leipzig_vector, target = "land_cover")

# load raster image
leipzig_raster = terra::rast(system.file("extdata", "leipzig_raster.tif",
  package = "mlr3spatial"))
# create testing data
tsk_predict = as_task_unsupervised(leipzig_raster)
```

Now we can continue as normal to train and predict with a classification learner, in this case a random forest.

```{r special-046}
lrn = lrn("classif.ranger")$train(tsk_leipzig)
pred = predict_spatial(tsk_predict, lrn, format = "terra")
pred
```

In this example we specified creation of a `terra` object, which can be visualised with in-built plotting methods.

```{r special-047, message = FALSE, cache.lazy=FALSE, cache=FALSE}
library(terra, exclude = "resample")
plot(pred, col = c("#440154FF", "#443A83FF", "#31688EFF",
  "#21908CFF", "#35B779FF", "#8FD744FF", "#FDE725FF"))
```

## Conclusion

<!-- TODO -->

### Resources {.unnumbered .unlisted}

<!-- TODO -->

## Exercises

<!-- TODO -->

1. Survival analysis
2. Density estimation
3. Spatiotemporal
4. Clustering
5. Cost-sens class