# Tasks {#sec-appendix-tasks}

{{< include _setup.qmd >}}

```{r include=FALSE}
library(mlr3verse)
library(mlr3proba)
library(mlr3spatiotempcv)
library(ggplot2)
```

The key features of the tasks that we use throughout the book are explained below as well as a plot of the target variable(s).

## Regression Tasks

### `mtcars`

```{r, message=FALSE, warning=FALSE}
tsk("mtcars")
tsk("mtcars")$head()
autoplot(tsk("mtcars"))
```

See more at `?mlr_tasks_mtcars`.

## Classification Tasks

### `german_credit`

```{r, message=FALSE, warning=FALSE}
tsk("german_credit")
tsk("german_credit")$head()
autoplot(tsk("german_credit"))
```

See more at `?mlr_tasks_german_credit`.

### `penguins`

```{r, message=FALSE, warning=FALSE}
tsk("penguins")
tsk("penguins")$head()
autoplot(tsk("penguins"))
```

See more at `?mlr_tasks_penguins`.

### `penguins_simple`

```{r, message=FALSE, warning=FALSE}
tsk("penguins_simple")
tsk("penguins_simple")$head()
autoplot(tsk("penguins_simple"))
```

See more at `?mlr3data::mlr_tasks_penguins_simple`.

### `sonar`

```{r, message=FALSE, warning=FALSE}
tsk("sonar")
tsk("sonar")$head()
autoplot(tsk("sonar"))
```

See more at `?mlr_tasks_sonar`.

### `spam`

```{r, message=FALSE, warning=FALSE}
tsk("spam")
tsk("spam")$head()
autoplot(tsk("spam"))
```

See more at `?mlr_tasks_spam`.

## Survival Tasks

### `rats`

```{r, message=FALSE, warning=FALSE}
tsk("rats")
tsk("rats")$head()
autoplot(tsk("rats"))
```

See more at `?mlr3proba::mlr_tasks_rats`.

## Density Tasks

### `precip`

```{r, message=FALSE, warning=FALSE}
tsk("precip")
tsk("precip")$head()
autoplot(tsk("precip"))
```

See more at `?mlr3proba::mlr_tasks_precip`.

## Spatiotemporal Tasks

### `ecuador`

```{r, message=FALSE, warning=FALSE}
tsk("ecuador")
tsk("ecuador")$head()
autoplot(tsk("ecuador"))
```

See more at `?mlr3spatiotempcv::mlr_tasks_ecuador`.

## Clustering Tasks

### `usarrests`

```{r, message=FALSE, warning=FALSE}
tsk("usarrests")
tsk("usarrests")$head()
autoplot(tsk("usarrests"))
```

See more at `?mlr3cluster::mlr_tasks_usarrests`.
